### 2013-02-01 main.tex
Initial commit

\documentclass[preprint]{elsarticle}

\usepackage{hyperref}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{amsmath}  % Maths
\usepackage{amsfonts} % Maths
\usepackage{amssymb}  % Maths
\usepackage{stmaryrd} % Maths (crochets doubles)

\usepackage{url}     % Mise en forme + liens pour URLs
\usepackage{array}   % Tableaux évolués

\usepackage{comment}

\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{pro}{Property~\ref{#1}}
\newrefformat{pps}{Proposition~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{thm}{Theorem~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\def\pref{\prettyref}

\newdefinition{definition}{Definition}
\newdefinition{property}{Property}
\newdefinition{proposition}{Proposition}
\newdefinition{example*}{Example}
\newdefinition{example}[example*]{Example}

\usepackage{tikz}
\newdimen\pgfex
\newdimen\pgfem
\usetikzlibrary{arrows,shapes,shadows,scopes}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}

\input{macros/macros}
\input{macros/macros-ph}
\input{macros/macros-grn}

% Commandes À FAIRE
\usepackage{color} % Couleurs du texte
\newcommand{\todo}[1]{\textcolor{darkgreen}{[#1]}}

% Id est
\newcommand{\ie}{i.e. }

% Césures
\hyphenation{pa-ra-me-tri-za-tion}
\hyphenation{pa-ra-me-tri-za-tions}

\begin{document}

\begin{frontmatter}

\title{Concretizing the Process Hitting into Biological Regulatory Networks}

\author[irccyn,nii]{Maxime Folschette}
\ead{Maxime.Folschette@irccyn.ec-nantes.fr}
\author[lix]{Lo\"ic Paulev\'e}
\author[nii]{Katsumi Inoue}
\author[irccyn]{Morgan Magni}
\author[irccyn]{Olivier Roux}

\address[irccyn]{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e - B.P. 92101 - 44321 Nantes Cedex 3, France.}
\address[nii]{National Institute of Informatics,\\
2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.}
\address[lix]{LIX, \'Ecole Polytechnique, 91128 Palaiseau Cedex, France.}

\begin{abstract}
\parindent 0.5cm
The Process Hitting (PH) is a recently introduced framework to model concurrent processes.
Its major originality lies in a specific restriction on the causality of actions, which
makes the formal analysis of very large systems tractable.
PH is suitable to model Biological Regulatory Networks (BRNs) with complete or partial
knowledge of cooperations between regulators by defining the most permissive dynamics
with respect to these constraints.

On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.

Given a PH model of a BRN, we first tackle the inference of the underlying Interaction Graph
between components.
Then the inference of corresponding Thomas' models is provided using Answer Set Programming,
which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.

In addition to giving a formal link between different approaches for qualitative BRNs modeling, 
this work emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
cooperations, where Thomas' approach fails because of the combinatorics of parameters.
\end{abstract}

\end{frontmatter}

\input{parts/intro}

\section{Frameworks}\label{sec:frameworks}
\input{parts/ph}
\input{parts/grn}

\input{parts/infer-ig}
\input{parts/infer-param}
\input{parts/example}
\input{parts/discussion}

\bibliographystyle{elsarticle-num}
\bibliography{biblio}

%\input{parts/contributions}

\end{document}

### 2013-02-01 parts/discussion.tex
Initial commit

\section{Conclusion and Discussion}

This work establishes the abstraction relationship between PH and Thomas' approaches for
qualitative BRN modeling.
The PH allows an abstract representation of BRNs dynamics (allowing incomplete knowledge on the
cooperation between components) that can not be exactly represented in Ren\'e Thomas' formalism by a
single instance of BRN parametrization.
This motivates the concretization of PH models into a set of compatible Thomas' models in order to benefit
of the complementary advantages of these two formal frameworks.

We first propose an original inference of the Interaction Graph (IG) from a BRN
having its dynamics specified in the PH framework.
An IG gives a compact abstract representation of the influence of the components between each
others.
Then, based on a prior inference of Ren\'e Thomas' parametrization for BRNs from a PH model, we
delimit the set of compatible Thomas' parametrizations that are compatible with the PH dynamics,
and give arguments for their correctness.
A parametrization is compatible with the PH if its dynamics (in terms of possible transitions) is included in the PH dynamics.
The enumeration of such parametrizations is efficiently tackled using Answer Set Programming.
We illustrate the overall method on simple examples and large biological models.

Several extensions of the presented work are now to be considered.
First, we now plan to explore the inference of Thomas' parameters when the inferred IG involves unsigned interactions, as some particular cases having such IGs are known to have correct
Thomas' parametrization (not respecting some assumptions imposed in \pref{ssec:admissible-K},
however).
Second, the inference of BRN multiplexes \cite{BernotMultiplexes} may be of practical interest 
as they allow to implicitly reduce the possible parametrizations by making cooperations appear
in the IG.
Because of its atomicity, the PH allows to specify a range of cooperations that can not be
completely captured by a single instance of BRN multiplexes, then encouraging the inference of a set
of compatible ones.
Finally, in order to improve the performances in the IG inference, we will consider projection operations on
the PH structure to undo cooperations between components and reduce the cardinality of
configurations to explore by making the interactions independent.

\paragraph{Ack}
This work was partially supported by the Fondation Centrale Initiatives.

### 2013-02-01 parts/example.tex
Initial commit

\section{Examples}\label{sec:examples}

The inference method described in this paper has been implemented as part of
\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related
tools.
Our implementation mainly consists in ASP programs that are solved using Clingo\footnote{Available
at \url{http://potassco.sourceforge.net}}.
The IG and parameters inference can be performed using the command
\texttt{ph2thomas -i model.ph -{}-dot ig.dot}
where \texttt{model.ph} is the PH model in \textsc{Pint} format and \texttt{ig.dot} is an output of the inferred IG in DOT format.
The (possibly partial) inferred parametrization will be returned on the standard output.
The admissible parametrizations enumeration is performed when adding the \texttt{-{}-enumerate}
parameter to the command.

Applied to the example in \pref{fig:runningPH-2} where cooperations have been defined,
our method infers the IG and parametrization given in \pref{fig:runningBRN}.
Regarding the example in \pref{fig:runningPH-1}, the same IG is inferred, as well as for the
parametrization except for the parameters $K_{a,\{b\},\{c\}}$ and $K_{a,\{c\},\{b\}}$ which are
undefined (because of the lack of cooperativity between $b$ and $c$).
In such a case, this partial parametrization allows 36 admissible complete parametrizations, as two
parameters with 3 potential values could not be inferred.
If we constrain these latter parameters so that they contain exactly one element, we obtain only 9
admissible parametrizations.

The current implementation can successfully handle large PH models of BRNs found in the literature
such as an ERBB receptor-regulated G1/S transition model from \cite{Sahin09} which contains 20
components and 15 cooperative sorts, and a T-cells receptor model from \cite{Klamt06} which contains 40
components and 14 cooperative sorts\footnote{Both models are available as examples distributed with \textsc{Pint}.}.
For each model, IG and parameters inferences are performed together in less than a second
on a standard desktop computer.
%\footnote{Using a Dell Inspiron 1720 laptop, with an Intel Core 2 Duo CPU T5550 ($2 \times 1.83\text{GHz}$)
%and 3.9 Gib memory, on an Ubuntu 11.10 64-bits OS}
After removing the cooperations from these models (leaving only raw actions), the inferences allow to
determine 40 parameters out of 195 for the 20 components model, and 77 out of 143 for the 40 components model.
As we thus have an order of magnitude of respectively $10^{31}$ and $10^{73}$ admissible parametrizations,
these models would therefore be more efficiently studied as PH than as BRNs.
We note that the complexity of the method is exponential in the number of regulators of one
component and linear in the number of components.
% We note that despite its smaller size in term of components, the ERBB transmission model takes more time to be computed because the biggest cooperative sorts contain more processes (up to 32 processes) than in the T-cells receptor model (up to 8 processes).

A PH model can be built based on information found in the literature about the local influences between components.
The precision of this knowledge will determine the precision of the modeled activations and inhibitions,
and some information is likely to help in the representation of cooperations.

### 2013-02-01 parts/grn.tex
Initial commit

\subsection{Thomas' modeling}
We concisely present the Thomas' modeling of BRNs dynamics, merely inspired by
\cite{Richard06,BernotSemBRN}.
In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t. the presented
inference), we extend the classical formalism by setting parameters to intervals of values instead of single
values, and briefly discuss this addition.

Thomas' formalism lies on two complementary descriptions of the system. First, the
\emph{Interaction Graph} (IG) models the structure of the system by defining the components'
mutual influences.
The \emph{parametrization} then specifies the levels to which tends a component when a given
configuration of its regulators applies.

The IG is composed of nodes that represent components, and edges labeled with a threshold that stand
for either positive or negative regulations (\pref{def:ig}).
For such a regulation to take place, the expression level of its head component has to be higher than its threshold; otherwise, the opposite influence is expressed.
The uniqueness of these regulations makes the following sections simpler.
%Therefore, for any component $b$, a predecessor $a$ of $b$ such that we have $a \xrightarrow{t} b$
%can be either an activator or an inhibitor of $a$, according to the sign of the interaction and if the expression level of $b$ if above or below the threshold $t$.
We call $\levelsA{a}{b}$ (resp. $\levelsI{a}{b}$) the levels of $a$ where it is an
activator (resp. inhibitor) of $b$ (\pref{def:levels});
$l_a$ denotes the maximum level of $a$.
%For the sake of simplicity and to establish a parallel with PH, if $a$ represents a component, we call $a_i$ its $i^\text{th}$ expression level.

\begin{definition}[Interaction Graph]
\label{def:ig}
An \emph{Interaction Graph} (IG) is a triple $(\Gamma, E_+, E_-)$ where $\Gamma$ is a finite number of \emph{components},
and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.

A regulation from $a$ to $b$ is uniquely referenced:
if $a \xrightarrow{t} b \in E_+$ (resp. $E_-$),
$\nexists a \xrightarrow{t'} b \in E_+ \text{ (resp. $E_-$)}, t' \neq t$
and $\nexists a \xrightarrow{t'} b \in E_-\text{ (resp. $E_+$)}, t' \in \mathbb{N}$.
\end{definition}

\begin{definition}[Effective levels ($\levels$)]\label{def:levels}
Let $(\Gamma,E_+,E_-)$ be an IG and $a, b \in \Gamma$ two of its components:
\begin{itemize}
  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
    $\levelsI{a}{b} \DEF [0; t-1]$;
  \item if $a \xrightarrow{t} b \in E_-$, $\levelsA{a}{b} \DEF [0; t-1]$ and
    $\levelsI{a}{b} \DEF [t; l_a]$;
  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
\end{itemize}
\end{definition}

For all component $a \in \Gamma$, 
$\GRNreg{a} \DEF \{ b\in\Gamma\mid \exists b\xrightarrow t a\in E_+\cup E_- \}$
is the set of its regulators.
We allow any number of levels for the components, without considering the number of outgoing edges,
as the number of processes in a PH sort is not constrained in any way.
%$E_+(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in
%E_+\}$ (resp. $E_-(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in E_-\}$) denotes the set
%of its positive (resp. negative) regulators; and
%$\GRNreg{a} = E_+(a)\cup E_-(a)$ the set of its regulators.

\begin{example*}
\pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E_+,E_-)$ with
$\Gamma = \{a, b, c\}$,
$E_+ = \{b \xrightarrow{1} a, c \xrightarrow{1} a\}$ and
$E_- = \{a \xrightarrow{2} b\}$;
hence $\GRNreg{a} = \{b, c\}$.
%This IG can represent the same behavior as the PH given in \pref{fig:runningPH-2}.
\end{example*}

\begin{figure}[t]
\begin{minipage}{0.4\linewidth}
\centering
\scalebox{1.2}{
\begin{tikzpicture}[grn]
\path[use as bounding box] (-0.3,-0.75) rectangle (2.5,1.5);
\node[inner sep=0] (a) at (2,0) {a};
\node[inner sep=0] (b) at (0,0) {b};
\node[inner sep=0] (c) at (2,1.2) {c};
%\path
%  node[elabel, below=-1em of a] {$0..2$}
%  node[elabel, below=-1em of b] {$0..1$}
%  node[elabel, below=-1em of c] {$0..1$};
\path[->]
  (b) edge[bend right] node[elabel, below=-2pt] {$+1$} (a)
  (c) edge node[elabel, right=-2pt] {$+1$} (a)
  (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b);
\end{tikzpicture}
}
\end{minipage}
\begin{minipage}{0.6\linewidth}
\centering
\begin{align*}
K_{a,\{b,c\},\emptyset} &= [2 ; 2] & K_{b,\{a\},\emptyset} &= [0 ; 1] \\
K_{a,\{b\},\{c\}} &= [1 ; 1] & K_{b,\emptyset,\{a\}} &= [0 ; 0] \\
K_{a,\{c\},\{b\}} &= [1 ; 1] &&\\
K_{a,\emptyset,\{b,c\}} &= [0 ; 0] & K_{c,\emptyset,\emptyset} &= [0 ; 1]
\end{align*}
\end{minipage}
\caption{\label{fig:runningBRN}
(left)
IG example.
%Components are represented by nodes labeled with a name and possible expression levels.
Regulations are represented by the edges labeled with their sign and threshold.
For instance, the edge from $b$ to $a$ is labeled $+1$, which stands for: $b \xrightarrow{1} a \in
E_+$.
%and means that if the level of expression of $b$ is equal to (i.e. above) 1, then $b$ activates $a$,
%otherwise, $b$ inhibits $a$.
(right)
Example parametrization of the left IG.
}
\end{figure}

A \emph{state} $s$ of an IG $(\Gamma, E_+, E_-)$ is an element in $\prod_{a \in \Gamma} [0;l_a]$.
$\GRNget{s}{a}$ refers to the level of component $a$ in $s$.
The specificity of Thomas' approach lies in the use of discrete \emph{parameters} to represent the
focal level interval towards which the component will evolve in each configuration of its regulators
(\pref{def:param}).
Indeed, for each possible state of a BRN, all regulators of a component $a$ can be divided into
\emph{activators} and \emph{inhibitors}, given their type of interaction and expression level,
referred to as the \emph{resources} of $a$ in this state (\pref{def:resources}).

\begin{definition}[Discrete parameter $K_{a,A,B}$ and Parametrization $K$]\label{def:param}
For a given component $a \in \Gamma$ and $A$ (resp. $B$) $\subset \GRNreg{a}$ a set of its activators (resp. inhibitors) such that $A \cup B = \GRNreg{a}$ and $A \cap B = \emptyset$,
the discrete \emph{parameter} $K_{a,A,B} = [i; j]$ is a non-empty interval towards which $a$ will tend
in the states where its activators (resp. inhibitors) are the regulators in set $A$ (resp. $B$).
The complete map $K$ of discrete parameters for $\IG$ is called a \emph{parametrization} of $\IG$.
\end{definition}
%A consequence of this definition is that $0 \leq i_1 \leq i_2 \leq l_a$.
%We also denote: $j < K_{a,A,B} \Leftrightarrow j < i_1$ and $j > K_{a,A,B} \Leftrightarrow j> i_2$.

\begin{definition}[Resources $\GRNres{a}{s}$]\label{def:resources}
For a given state $s$ of a BRN, we define the \emph{activators} $A$ and \emph{inhibitors} $B$ of $a$ in $s$ as $\GRNres{a}{s} = A,B$, where:
\begin{align*}
  A &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsA{b}{a}\} \\
  B &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsI{b}{a}\}
\end{align*}
We also denote: $\GRNallres{a} = \{(A;B) \mid \exists s \in \textstyle\prod_{a \in \Gamma} [0;l_a], \GRNres{a}{s} = A,B\}$
\end{definition}

%\begin{example*}
%\pref{fig:runningBRN}(right) gives a Parametrization of the IG of \pref{fig:runningBRN}(left).
%\end{example*}

At last, \pref{def:dynamics} gives the asynchronous dynamics of a BRN using Thomas' parameters.
From a given state $s$, a transition to another state $s'$ is possible provided that only one component $a$ will evolve of one level towards $K_{a,\GRNres{a}{s}}$.

\begin{definition}[Asynchronous dynamics]\label{def:dynamics}
Let $s$ be a state of a BRN using Thomas' parameters $(\IG, K)$ where $\IG = (\Gamma, E_+, E_-)$.
The state that succeeds to $s$ is given by the indeterministic function $f(s)$:
\begin{align*}
  f(s)  & = s' \Leftrightarrow \exists a \in \Gamma,
    \GRNget{s'}{a} = f^a(s) \wedge
    \forall b \in \Gamma, b \neq a, \GRNget{s}{b} = \GRNget{s'}{b}
    \quad\text{, with}\\
  f^a(s) & =
  \begin{cases}
    \GRNget{s}{a} + 1 & \text{if } \GRNget{s}{a} < K_{a,A,B} \\% \GRNres{a}{s}} \\
    \GRNget{s}{a} & \text{if } \GRNget{s}{a} \in K_{a,A,B} \\ %\GRNres{a}{s}}\\
    \GRNget{s}{a} - 1 & \text{if } \GRNget{s}{a} > K_{a,A,B} % \GRNres{a}{s}}
  \end{cases}
\quad\text{, where $A,B=\GRNres{a}{s}$.}
\end{align*}
\end{definition}

While the use of intervals as parameter values does not add expressivity in boolean
networks, it allows to specify a larger range of dynamics in the general case (w.r.t. the above
definitions).
Indeed, assume that $K_{a,A,B} = [i ; i+2]$;
we aim at obtaining three different parameters $K_{a,A_1,B_1} = i$,  $K_{a,A_2,B_2} = i+1$,
$K_{a,A_3,B_3} = i+2$.
The only possible modification in resources is to add $a$ as a self-regulator.
However, because resources have a boolean definition (a component is either an activator or an inhibitor of
$a$), it is not possible to differentiate the 3 values.
We also remark that the use of intervals makes optional some explicit auto-activations in the IG
(as for $b$ in \pref{fig:runningBRN}, for instance).

\begin{example*}
In the BRN that consists of the IG and parametrization of \pref{fig:runningBRN}, the following
transitions are possible given the semantics defined in \pref{def:dynamics}:
$\GRNetat{a_0, b_1, c_1} \rightarrow \GRNetat{a_1, b_1, c_1} \rightarrow \GRNetat{a_2, b_1, c_1} \rightarrow
\GRNetat{a_2, b_0, c_1} \rightarrow \GRNetat{a_1, b_0, c_1}$,
ending in a steady state,
where $a_i$ is the component $a$ at level $i$.
As $K_{b,\{a\},\emptyset} = [0 ; 1]$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
\end{example*}

### 2013-02-01 parts/infer-ig.tex
Initial commit

\section{Interaction Graph Inference}\label{sec:infer-IG}

In order to infer a complete BRN, one has to find the Interaction Graph (IG) first, as some
constraints on the parametrization rely on it.
Inferring the IG is an abstraction step which consists in determining the global influence of
components on each of its successors.

This section first introduces the notion of focal processes within a PH
(\pref{ssec:focal}) which is used to characterize well-formed PH for IG inference
in \pref{ssec:wf}, and as well used by the parametrization inference presented in \pref{sec:infer-K}.
Finally, the rules for inferring the interactions between components from a PH are
described in \pref{ssec:infer-IG}.
We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
performed.

\subsection{Focal Processes}\label{ssec:focal}

Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
processes} w.r.t. a given context (a set of processes that are potentially present).
When such a context applies, we expect to (always) reach one focal process in a bounded number of
actions.

For $S_a\subseteq L_a$ and a context (set of processes) $\ctx$, let us define as $\PHa(S_a,\ctx)$
the set of actions on the sort $a$ having their hitter in $\ctx$ and target in $S_a$
(\pref{eq:PHa-ctx});
and the digraph $(V, E)$ where arcs are the bounces within the sort $a$ triggered by actions
in $\PHa(S_a,\ctx)$ (\pref{eq:bounce-graph}).
$\focals(a,S_a,\ctx)$ denotes the set of focal processes of sort $a$ in the scope of
$\PHa(S_a,\ctx)$ (\pref{def:focals}).
\begin{align}
\PHa(S_a,\ctx) & \DEF \{ \PHfrappe{b_i}{a_j}{a_k}\in\PHa \mid b_i\in\ctx \wedge a_j\in S_a \}
\label{eq:PHa-ctx}
\\
\begin{split}
E  & \DEF \{(a_j,a_k)\in (S_a \times \PHl_a) \mid 
			\exists\PHfrappe{b_i}{a_j}{a_k}\in \PHa(S_a,\ctx) \}
\\
V & \DEF S_a \cup \{ a_k\in L_a\mid \exists (a_j,a_k)\in E\}
\end{split}
\label{eq:bounce-graph}
\end{align}

\begin{definition}[$\focals(a,S_a,\ctx)$]\label{def:focals}
The set of processes that are focal for processes in $S_a$ in the scope of $\PHa(S_a,\ctx)$
are given by:
%$\focals(a,S_a,\ctx)$ is the set of focal processes of sort $a$ in the context $\ctx$:
\[
\focals(a,S_a,\ctx) \DEF
\begin{cases}
\{ a_i \in V \mid \nexists (a_i,a_j)\in E\} & \text{if the digraph $(V,E)$ is acyclic},\\
\emptyset & \text{otherwise.}\\
\end{cases}
\]
\end{definition}

We note $\PHl(\ctx)$ the set of states $s\in L$ such that $\forall a\in\PHsort(\ctx), \PHget{s}{a}\in\ctx$,
where $\PHsort(\ctx)$ is the set of sorts with processes in $\ctx$.
We say a sequence of actions $h^1,\dots,h^n$ is \emph{bounce-wise} if and only if
$\forall m\in[1;n-1], \PHbounce(h^m)=\PHtarget(h^{m+1})$.
From \pref{def:focals}, it derives that:
\begin{enumerate}
\item if $\focals(a,S_a,\ctx)=\emptyset$, there exists a 
state $s\in \PHl(\ctx\cup S_a)$ such that $\forall n\in\mathbb N$ there
exists a bounce-wise sequence of actions $h^1,\dots,h^{n+1}$ in $\PHa(S_a,\ctx)$ 
with $\PHtarget(h^1)\in s$.
\item if $\focals(a,S_a,\ctx)\neq\emptyset$, for all
state $s\in \PHl(\ctx\cup S_a)$,
for any bounce-wise sequence of actions $h^1,\dots,h^n$ in $\PHa(S_a,\ctx)$ where $\PHtarget(h^1)\in
s$,
either
 $\PHbounce(h^n) \in \focals(a,S_a,\ctx)$,
or
$\exists h^{n+1}\in \PHa(a,\ctx)$ such that $\PHbounce(h^n) = \PHtarget(h^{n+1})$.
Moreover $n\leq|\PHa(S_a,\ctx)|$ (i.e. no cycle of actions possible).
\end{enumerate}

It is worth noticing that those bounce-wise sequences of actions may not be successively playable in
a state $s\in L(\ctx\cup S_a)$.
Indeed, nothing impose that the hitters of actions are present in $s$.
In the general case, the playability of those bounce-wise sequences, referred to as \emph{focals
reachability} may be hard to prove.
However, in the scope of this paper, the particular contexts used with $\focals$ ensure this property.
Notably, the rest of this section uses only \emph{strict} contexts (\pref{def:strict-ctx}) which
allow at most one hitter per sort in the bounce-wise sequences (and thus are present in $s$).

\begin{comment}
\begin{property}[$\focals$ reachability]\label{pro:focals-reach}
$\focals(a,S_a,\ctx)$ is reachable if and only if 
$\forall s\in L(\ctx\cup S_a)$,
if $\exists h\in \PHa(S_a,\ctx)$ with $\PHtarget(h)\in s$,
there exists a (possibly empty) sequence of actions
$h^1,\dots,h^n \in \PHa(\ctx\cup S_a)$
	such that $h^1,\dots,h^n,h$ are successively playable in $s$;
where $\PHa(\ctx) \DEF \{ \PHfrappe{b_i}{c_j}{c_k}\in\PHa \mid c\neq a \wedge
			b\in \PHsort(\ctx) \Rightarrow b_i\in \ctx \wedge c\in\PHsort(\ctx) \Rightarrow
			c_i\in\ctx \}$.
\end{property}
\end{comment}

\begin{definition}[Strict context for $S_a$]\label{def:strict-ctx}
A context (set of processes) $\ctx$ is strict for $S_a\subseteq L_a$ if and only if
$\{b_i,b_j\} \subset \ctx \wedge b\neq a \Rightarrow i=j$.
\end{definition}

In other words, assuming focals reachability, if $\focals(a,S_a,\ctx)$ is empty, there exists a
sequence of actions that may be played an unbound number of times (cycle);
if it is non-empty, it is ensured that any state in $\PHl(\ctx\cup S_a)$ converges, in a bounded
number of steps, either to a process in $S_a$ that is not hit by processes in $\ctx$, or to a process in
$L_a\setminus S_a$.

\begin{example*}
In the PH of \pref{fig:runningPH-1}, we obtain:
\begin{align*}
\focals(a,L_a,\{b_0,c_0\}) &= \{ a_0 \}
&
\focals(a,L_a,\{b_1,c_1\}) &= \{ a_2 \}
\\
\focals(a,L_a,\{b_1,c_0\}) &= \emptyset
&
\focals(a,\{a_1\},\{b_1,c_0\}) &= \{ a_0, a_2 \}
\end{align*}
\end{example*}

\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}

The inference of an IG from a PH assumes that the PH defines two types of sorts:
the sorts corresponding to BRN components, and the cooperative sorts.
This leads to the characterization of a \emph{well-formed} PH for IG inference.

The identification of sorts modeling components relies on the observation that their processes
represent (ordered) qualitative levels.
Hence an action on such a sort cannot make it bounce to a process at a distance more than one.
The set of sorts satisfying such a condition is referred to as $\Gamma$
(\pref{eq:PH-components}).
Therefore, in the rest of this paper, $\Gamma$ denotes the set of components of the BRN to infer.

\begin{equation}
\Gamma \DEF \{a \in \PHs \mid \nexists \PHfrappe{b_i}{a_j}{a_k} \in \PHa, |j - k| > 1\} \\
\label{eq:PH-components}
\end{equation}

Any sort that does not act as a component should then be treated as a cooperative sort.
As explained in \pref{ssec:PH}, the role of a cooperative sort $\upsilon$ is to compute the current
state of set of cooperating processes.
Hence, for each sub-state $\sigma$ formed by the sorts hitting $\upsilon$, $\upsilon$ should
converge to a focal process.
This is expressed by \pref{pro:wf-cooperative-sort}, where
the set of sorts having an action on a given sort $a$ is given by 
$\PHdirectpredec{a}$ (\pref{eq:ph_direct_predec})
and $\PHproc(\sigma)$ is the set of processes that compose the sub-state $\sigma$.

\begin{equation}
\forall a \in \PHs, \PHdirectpredec{a} \DEF \{b \in \PHs \mid \exists \PHfrappe{b_i}{a_j}{a_k}\in\PHa \}
\label{eq:ph_direct_predec}
\end{equation}

\begin{property}[Well-formed cooperative sort]\label{pro:wf-cooperative-sort}
A sort $\upsilon\in\PHs$ is a well-formed cooperative sort if and only if
each configuration $\sigma$ of its predecessors leads $\upsilon$ to a unique focal process,
denoted by $\upsilon(\sigma)$:
\[
\forall \sigma \in {\textstyle\prod_{
a\in\PHdirectpredec{\upsilon} \wedge a\neq \upsilon}}
\PHl_{a},
\focals(\upsilon,\PHl_\upsilon,\PHproc(\sigma)\cup \PHl_\upsilon) = \{ \upsilon(\sigma) \}\]
\end{property}

Such a property allows a large variety of definitions of a cooperative sort, but
for the sake of simplicity, does not allow the existence of multiple focal processes.
While this may be easily extended to (the condition becomes 
$\focals(\upsilon,\PHl_\upsilon, \PHproc(\sigma)\cup \PHl_\upsilon)\neq\emptyset$), it makes some
hereafter equations a bit more complex to read as they should handle a set of focal processes instead
of a unique focal process.

Finally, \pref{pro:wf-ph} sums up the conditions for a Process Hitting to be suitable for IG
inference.
In addition of having either component sorts or well-formed cooperative sorts, we also require that
there is no cycle between cooperative sorts, and that
sorts being never hit (\ie{} serving as an invariant environment) are components.

\begin{property}[Well-formed Process Hitting for IG inference]\label{pro:wf-ph}
A PH is well-formed for IG inference if and only if the following conditions are verified:
\begin{itemize}
\item 
each sort $a\in\PHs$ either belongs to $\Gamma$, or is a well-formed cooperative sort;
\item 
there is no cycle between cooperative sorts
(the digraph $(\Sigma,\{(a,b)\in(\Sigma\times\Sigma)\mid \exists \PHfrappe{a_i}{b_j}{b_k}\in\PHa
\wedge a\neq b\wedge \{a,b\}\cap\Gamma=\emptyset \})$ is
acyclic);
\item 
sorts having no action hitting them belong to $\Gamma$
($\{ a \in \Sigma\mid \nexists \PHfrappe{b_i}{a_j}{a_k}\in\PHa\} \subset \Gamma$).
\end{itemize}
\end{property}

\begin{example*}
In the PH of \pref{fig:runningPH-2}, $bc$ is a well-formed cooperative sort as defined in \pref{pro:wf-cooperative-sort}, because:
\begin{align*}
\focals(bc, \PHl_{bc}, \{b_0, c_0\} \cup \PHl_{bc}) = \{bc_{00}\} && \focals(bc, \PHl_{bc}, \{b_0, c_1\} \cup \PHl_{bc}) = \{bc_{01}\} \\
\focals(bc, \PHl_{bc}, \{b_1, c_0\} \cup \PHl_{bc}) = \{bc_{10}\} && \focals(bc, \PHl_{bc}, \{b_1, c_1\} \cup \PHl_{bc}) = \{bc_{11}\}
\end{align*}
Hence, both \pref{fig:runningPH-1} and \pref{fig:runningPH-2} are well-formed PH for IG inference
with $\Gamma = \{a,b,c\}$.
\end{example*}

\subsection{Interaction Inference}\label{ssec:infer-IG}

At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
($\PHs \setminus \Gamma$) that will not appear in the IG. 
We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
by considering direct actions and possible intermediate cooperative sorts.
The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
(\pref{eq:regulators}).
\begin{align}
\begin{split}
\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
(c^k)_{k \in [0;n]} \in \PHs^{n+1}, \\
                                   & \quad \quad c^0 = b \wedge c^n = a \\
                                   & \quad \quad \wedge \forall k \in \llbracket 0 ; n-1 \rrbracket,
								   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
\end{split}
\label{eq:ph_predec}
\\
\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
\label{eq:regulators}
\end{align}

Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
If $g=\{b\}$, we simple note $\ctx_b(\sigma)$.
This set is composed of the active processes of sorts in $g$, and the focal process (assumed
unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
$\upsilon$ are defined in $\sigma$.
When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval}.
Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
always terminates.
\begin{align}
\forall g\subset \Gamma,
	\ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
\label{eq:ctx-sigma}
\\
\upsilon(\sigma) & \DEF
\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
	\upsilon'\in\PHdirectpredec{\upsilon} \wedge
	\upsilon'\in\PHs\setminus\Gamma })
\label{eq:cooperative-eval}
\end{align}

We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
the level of $b$ makes possible the increase (decrease) of the level of $a$.
This is analogous to standard IG inferences from discrete maps \cite{RiCo07}.

This reasoning can be straightforwardly applied to a PH when inferring the influence of $b$ on $a$
when $b\neq a$ (\pref{eq:edges-inference-b}).
Let us define $\gamma(b\rightarrow a)$ as the set of components cooperating with $b$ to hit $a$,
including $b$ and $a$ (\pref{eq:cooperating-with-b}).
Given a configuration $\sigma\in\prod_{c\in\gamma(b\rightarrow a)} L_c$, 
$\focals(a,\{a_i\},\ctx_b(\sigma))$ gives the bounces that a given process $a_i$ can make in the
context $\ctx_b(\sigma)$.
We note $\sigma\{b_i\}$ the configuration $\sigma$ where the process of sort $b$ has been replaced
by $b_i$.
If there exists $b_i,b_{i+1}\in L_b$ such that one bounce in 
$\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_i\}))$
has a lower (resp. higher) level that one bounce in
$\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_{i+1}\}))$, then
$b$ as positive (resp. negative) influence on $a$ with a maximum threshold $l=i+1$.
\begin{equation}
\gamma(b\rightarrow a)  \DEF \{ a, b \} \cup \{ c \in \Gamma \mid 
			\exists \upsilon\in\PHs\setminus\Gamma,
				\upsilon\in\PHpredec{a} \wedge \{b,c\}\subset\PHpredec{\upsilon} \}
\label{eq:cooperating-with-b}
\end{equation}

Then, we infer that $a$ has a self-influence if its current level can have an impact on its own
evolution at a given configuration $\sigma$.
We consider here a configuration $\sigma$ of a group $g$ of sorts having a cooperation on $a$.
This set of sort groups is given by $X(a)$ (\pref{eq:influence-groups}) which returns the set of
connected components (noted $\mathcal C$) of the graph linking two regulators $b,c$ of $a$ if there
is a cooperative sort hitting $a$ regulated by both of them.
Given $a_i,a_{i+1}\in L_a$, we pick $a_j\in\focals(a,\{a_i\},\ctx_g(\sigma\{a_i\}))$ and
$a_k\in\focals(a,\{a_{i+1}\},\ctx_g(\sigma\{a_{i+1}\}))$.
If $k=j+1$, we can not conclude as there is no difference in the evolution of both levels.
If $k\neq j+1$ and $k-j\neq 0$, then $a_i$ and $a_{i+1}$ have divergent evolutions: we infer an
influence of sign of $k-j$ at threshold $i+1$.
We note that some aspects of this inference are arbitrary and may impact the number of parameters to
infer in the next section.
In particular, in some cases, the use of intervals for Thomas' parameters drops the requirement of
inferring a self-activation.
%Future work may propose alternative definitions of self-influences inference in order to range over
%different parametrization configurations.
\begin{equation}
X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
				\exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
					\{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
\label{eq:influence-groups}
\end{equation}

\pref{pps:inference-edges} details the inference of all existing influences between components occurring
with a threshold $l$.
These influences are split into positive and negative ones, and represent possible edges in the final IG.
We do not consider the cases where a component has no visible influence on another.
\begin{proposition}[Edges inference]\label{pps:inference-edges}
We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any
$a\in\Gamma$ by:
\begin{align}
\begin{split}
\forall b\in\PHpredecgene{a}, b\neq a, \\
b\xrightarrow l a \in \hat{E}_s & \Longleftrightarrow
 \exists \sigma\in\textstyle\prod_{c\in\gamma(b\rightarrow a)} L_c, \exists b_i,b_{i+1}\in \PHl_b,\\
& \qquad\qquad
        \exists a_{j}\in\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_i\})), \\
& \qquad\qquad
        \exists a_{k}\in\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_{i+1}\})), \\
& \qquad\qquad\qquad
                        s = \f{sign}(k - j) \wedge l = i+1
\end{split}
\label{eq:edges-inference-b}
\\
\begin{split}
a \xrightarrow l a \in \hat{E}_s & \Longleftrightarrow
\exists g \in X(a), \sigma \in L_a\times \textstyle\prod_{b\in g} L_b,
			\exists a_i,a_{i+1}\in \PHl_a, \\
& \qquad\qquad
        \exists a_{j}\in\focals(a, \{a_i\}, \ctx_g(\sigma\{a_i\})), \\
& \qquad\qquad
        \exists a_{k}\in\focals(a, \{a_{i+1}\},  \ctx_g(\sigma\{a_{i+1}\})), \\
& \qquad\qquad\quad
			k \neq j+1
				\wedge s = \f{sign}(k - j) \wedge l = i+1
\end{split}
\label{eq:edges-inference-a}
\end{align}
where $s \in \{ +, - \}$, $\bar s = + \EQDEF s = -$, $\bar s = - \EQDEF s = +$,
$\f{sign}(n) = + \EQDEF n > 0$,
$\f{sign}(n) = - \EQDEF n < 0$,
and $\f{sign}(0) \DEF 0$.
\end{proposition}

We are now able to infer the edges of the final IG by considering positive and negative influences
(\pref{pps:inference-IG}).
We infer a positive (resp. negative) edge if there exists a corresponding influence with the same
sign. If an influence is both positive and negative, we infer an unsigned edge. In the end, the
threshold of each edge is the minimum threshold for which the influence has been found. As unsigned
edges represent ambiguous interactions, no threshold is inferred.
\begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
We infer $\IG = (\Gamma,E_+,E_-,E_\pm)$ from \pref{pps:inference-edges} as follows:
\begin{align*}
E_+ &= \{a \xrightarrow{t} b \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
  \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
E_- &= \{a \xrightarrow{t} b \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
E_\pm &= \{a \rightarrow b \mid \exists a \xrightarrow{t} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t'} b \in \hat{E}_-\} \\
\end{align*}
\end{proposition}

\begin{example*}
The IG inference from the PH of \pref{fig:runningPH-2} gives
$\hat{E}_+ = \{b \xrightarrow{1} a, c \xrightarrow{1} a\}$ and 
$\hat{E}_- = \{a \xrightarrow{2} b\}$, corresponding to the IG of \pref{fig:runningBRN}.
No self-influence are inferred ($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
\end{example*}

### 2013-02-01 parts/infer-param.tex
Initial commit

\section{Parametrization inference}\label{sec:infer-K}

Given the IG inferred from a PH as presented in the previous section, one can find the discrete parameters that model the behavior of the studied PH using the method presented in the following.
It relies on an exhaustive enumeration of all predecessors of each component in order to find attractor processes and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations.
The last step consists of the enumeration of all compatible complete parametrizations given this
set of inferred parameters, the PH dynamics and some biological constraints on parameters.

\subsection{Parameters inference}

This subsection presents some results related to the inference of independent discrete parameters from a given PH. These results are equivalent to those presented in \cite{PMR10-TCSB}, with notation adapted to be shared with the previous section.
In addition, we introduce the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
which implies that the inferred IG does not contain any unsigned interactions, and thus can be seen as the
regular IG $(\Gamma, E_+, E_-)$,
and that any processes in $\levelsA{b}{a}$ (resp. $\levelsI{b}{a}$) share the same behavior
regarding $a$.
\begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
A PH is well-formed for parameter inference if and only if
it is well-formed for IG inference, and
the IG $(\Gamma, E_+, E_-, E_\pm)$ inferred by \pref{pps:inference-IG}
verifies $E_\pm=\emptyset$ and if the following property holds:
\begin{align}
\begin{split}
\forall b\in \GRNreg{a} &,
	\forall (i,j\in\levelsA{b}{a} \vee i,j\in\levelsI{b}{a}), \\
& \qquad	\forall c, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
& \qquad \qquad
			\PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
				\PHfrappe{b_j}{c_k}{c_l}\in\PHa
\end{split}
\label{eq:wf-levels}
\end{align}
\end{property}

Let $K_{a,A,B}$ be the parameter we want to infer for a given component $a \in \Gamma$
%and $A,B \in \GRNallres{a}$ a configuration of resources of $a$ (activators and inhibitors).
and $A \subset \GRNreg{a}$ (resp. $B \subset \GRNreg{a}$) a set of its activators (resp. inhibitors).
This inference, as for the IG inference, relies on the search of focal processes of the component for the given configuration of its regulators.

For each sort $b \in \GRNreg{a}$, we define a context $C^b_{a,A,B}$ in \pref{eq:param_context} that contains all processes representing the influence of the regulators in the configuration $A,B$.
The context of a cooperative sort $\upsilon$ that regulates $a$ is given in
\pref{eq:param_context_coop} as the set of focal processes matching the current configuration.
$C_{a,A,B}$ refers to the union of all these contexts (\pref{eq:K-ctx}).
\begin{align}
\label{eq:param_context}
\forall b\in\Gamma,~
C_{a,A,B}^b & \DEF \begin{cases}
\levelsA{b}{a} & \text{if $b \in A$,}\\
\levelsI{b}{a} & \text{if $b \in B$,}\\
L_b		& \text{otherwise;}\\
\end{cases}
\\
\label{eq:param_context_coop}
\forall \upsilon \in \PHpredec{a}\setminus\Gamma,~
C_{a,A,B}^\upsilon & \DEF \{
\upsilon(\sigma) \mid \sigma \in \textstyle\prod_{c\in\PHdirectpredec{\upsilon}}C_{a,A,B}^c \}
\\
C_{a,A,B} & \DEF \textstyle\bigcup_{b\in\PHpredec{a}} C^b_{a,A,B}
\label{eq:K-ctx}
\end{align}

The parameter $K_{a,A,B}$ specifies to which values $a$ eventually evolves as long as the context
$C_{a,A,B}$ holds, which is precisely the definition of the $\focals$ function
(\pref{def:focals} in \pref{ssec:focal}),
where the focals reachability property can be derived from \pref{pro:wf-ph-K} and
\pref{eq:param_context_coop}.
Hence $K_{a,A,B} = \focals(a,C^a_{a,A,B},C_{a,A,B})$ if this latter is a non-empty interval
(\pref{pps:param_K}).

\begin{proposition}[Parameter inference]
\label{pps:param_K}
Let $(\PHs, \PHl, \PHh)$ be a Process Hitting well-formed for parameter inference, and $\IG = (\Gamma,
E_+, E_-)$ the inferred IG.
Let $A$ (resp. $B$) $\subseteq \Gamma$ be the set of regulators that activate (resp. inhibit) a sort
$a$.
%If $\focals(a,C_{a,A,B})$ is a non-empty interval, then $K_{a,A,B} = \focals(a, C_{a,A,B})$.
If $\focals(a,C^a_{a,A,B},C_{a,A,B})=[a_i;a_j]$ is a non-empty interval, 
	then $K_{a,A,B} = [i;j]$.
\end{proposition}

\begin{example*}
Applied to the PH in \pref{fig:runningPH-1}, we obtain, for instance, 
$K_{b,\{a\},\emptyset} = [0 ; 1]$ and
$K_{a,\{b,c\},\emptyset} = [2 ; 2]$,
while $K_{a,\{b\},\{c\}}$ can not be inferred.
For the PH in \pref{fig:runningPH-2}, this latter is evaluated to $[1;1]$.
\end{example*}

Given the \pref{pps:param_K}, we see that in some cases, the inference of the targeted parameter is impossible.
This can be due to a lack of cooperation between regulators: when two regulators independently hit a component, their actions can have opposite effects, leading to either an indeterministic evolution or to oscillations.
Such an indeterminism is not possible in a BRN as in a given configuration of regulators, a component can only have an interval attractor, and eventually reaches a steady-state.
In order to avoid such inconclusive cases, one has to ensure that no such behavior is allowed by
either removing undesired actions or using cooperative sorts to prevent opposite influences between
regulators.

\subsection{Admissible parametrizations}\label{ssec:admissible-K}

When building a BRN, one has to find the parametrization that best describes the desired behavior of the studied system.
Complexity is inherent to this process as the number of possible parametrizations for a given IG is exponential w.r.t. the number of components.
However, the method of parameters inference presented in this section gives some information about necessary parameters given a certain dynamics described by a PH.
This information thus drops the number of possible parametrizations, allowing to find the desired behavior more easily.

We first delimit the validity of a parameter (\pref{pro:K-valid}) in order to ensure that any
transition in the resulting BRN is allowed by the studied PH.
This is verified by the existence of a hit making the concerned component bounce into the direction
of the value of the parameter in the matching context.
Thus, assuming \pref{pro:wf-ph-K} holds, any transition in the inferred BRN corresponds to at least
one transition in the PH, proving the correctness of our inference.
We remark that any parameter inferred by \pref{pps:param_K} satisfies this property.

\begin{property}[Parameter validity]\label{pro:K-valid}
A parameter $K_{a,A,B}$ is valid w.r.t. the PH iff the following equation is verified:
\begin{align*}
\forall a_i\in C^a_{a,A,B},
		a_i \notin K_{a,A,B} \Longrightarrow (
  \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, & c_k \in C^c_{a,A,B} \\
 \wedge a_i < K_{a,A,B} \Rightarrow j > i 
 & \wedge  a_i > K_{a,A,B} \Rightarrow j <i )
\end{align*}
\end{property}

Then, we use some additional biological constraints on Thomas' parameters given in
\cite{BernotSemBRN}, that we sum up in the following three properties:

\begin{property}[Extreme values assumption]
Let $\IG = (\Gamma, E_+, E_-)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} iff:
\label{prop:param_enum_extreme}
\[
  \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Rightarrow 0 \in K_{b,\emptyset,\GRNreg{b}} \wedge l_b \in K_{b,\GRNreg{b},\emptyset}
\]
\end{property}

\begin{property}[Activity assumption]
\label{prop:param_enum_activity}
Let $\IG = (\Gamma, E_+, E_-)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} iff:
\begin{align*}
  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A,B} <_{[]} K_{b,A \cup \{b\},B \setminus \{b\}}
\\
  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A \setminus \{b\},B \cup \{b\}} <_{[]} K_{b,A,B}
\end{align*}
\end{property}

\begin{property}[Monotonicity assumption]
\label{prop:param_enum_monotonicity}
Let $\IG = (\Gamma, E_+, E_-)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} iff:
\[
  \forall b \in \Gamma, \forall (A;B), (A';B') \in \GRNallres{b},
  A \subset A' \wedge B' \subset B \Rightarrow K_{b,A,B} \leq_{[]} K_{b,A',B'}
\]
\end{property}

\begin{comment}
\begin{definition}[Admissible parametrization \& Admissible parametrization with respect to inferred parameters]
\label{def:param_enum_inf}
Let $\PH = (\PHs, \PHl, \PHh)$ be a PH so that IG inference is possible, and $\IG = (\Gamma, E_+,
E_-)$ the inferred IG.
A parametrization $K$ on $\IG$ is said to be \emph{admissible} iff it respects
the extreme values assumption, the activity assumption and the monotonicity assumption.
A parametrization $K$ on $\IG$ is said to be \emph{admissible with respect to the
inferred parameters} iff it is admissible and that all parameters that can be inferred regarding
\pref{pps:param_K} are equal to their inferred value.
\end{definition}

\todo{utilité de “Admissible parametrization” seul ?}
\end{comment}

\subsection{Answer Set Programming implementation concepts}

\newcommand{\ti}[1]{\texttt{\textit{#1}}}
\newcommand{\aspil}[1]{\texttt{#1}}
\newcommand{\asp}[1]{\begin{itemize} \item[] \aspil{#1} \end{itemize}}

Answer Set Programming (ASP) \cite{Baral03} has been chosen to address the enumeration of all admissible parametrizations.
The motivations are following:
\begin{itemize}
  \item ASP efficiently tackles the inherent complexity of the models we use, thus allowing a fast execution of the formal tools defined in this paper,
  \item it is convenient to enumerate a large set of possible answers,
  \item it allows us to easily constrain the answers according to some properties.
\end{itemize}
We now synthesize some key points to better make the reader understand our ASP implementation with the enumeration example.

All information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
For functional purposes, we assign a unique label to each couple $A,B$ of activators and inhibitors of a given component, and in the following we note $K^p_{a,A,B}$ the parameter of component $a$ whose regulators $A,B$ are assigned to the label $p$.
Then, to state the existence of a parameter $K^\ti{p}_{\ti{a},A,B}$, we use an atom named \aspil{param\_label} in the following fact:
\asp{param\_label(\ti{a}, \ti{p}).}

Defining a set in ASP is equivalent to defining the rule for belonging to this set.
For example, we define an atom \aspil{param\_act} that describes the set of active regulators of a given a parameter.
%of component \ti{a} and label \ti{p} (\ie the set $A$ of a parameter $K^\ti{p}_{\ti{a},A,B}$).
Describing the activators of $K^\ti{p}_{\ti{a},\{\ti{b},\ti{c}\},\{\ti{d}\}}$ gives:
\asp{param\_act(\ti{a}, \ti{p}, \ti{b}).
\item[] param\_act(\ti{a}, \ti{p}, \ti{c}).}
The absence of such a fact involving \ti{d} with label \ti{p} indicates that \ti{d} is an inhibitor in the configuration of regulators related to this parameter.

Rules allow more detailed declarations than facts as they have a body (right-hand part below) containing constraints and allowing to use variables, while facts only have a head (left-hand part).
For instance, in order to define the set of expression levels of a component, we declare:
\asp{component\_levels(X, 0..M) :- component(X, M).}
where the \aspil{component(X, M)} atom stands for the existence of a component \aspil{X} with a maximum level \aspil{M}.
Considering this declaration, any possible answer for the atom \aspil{component\_levels} will be found by binding all possible values of its two terms with all existing \aspil{component} facts: the existence of an answer \aspil{component\_levels(\ti{a}, \ti{k})} will depend on the existence of a term \ti{a}, which is bound with \aspil{X}, and an integer~\ti{k}, constrained by: $0 \leq \ti{k} \leq \aspil{M}$.

Cardinalities are convenient to enumerate all possible parametrizations by creating multiple answer sets.
A cardinality (denoted hereafter with curly brackets) gives any number of possible answers for some atoms between a lower and upper bounds.
For example,
\asp{1 \{ param(X, P, I) : component\_levels(X, I) \} :-
\item[] ~~~~~~param\_label(X, P), not infered\_param(X, P).}
where \aspil{param(X, P, I)} stands for: $\aspil{I} \in K^\aspil{P}_{\aspil{X},A,B}$,
means that any parameter of component \aspil{X} and label \aspil{P} must contain at least one level value (\aspil{I}) in the possible expression levels of \aspil{X}.
Indeed, the lower bound is 1, forcing at least one element in the parameter, but no upper bound is specified, allowing up to any number of answers.
The body (right-hand side) of the rule also checks for the existence of a parameter of \aspil{X} with label \aspil{P}, and constrains that the parametrization inference was not conclusive for the considered parameter (\aspil{not} stands for negation by failure: \aspil{not L} becomes true if \aspil{L} is not true).
Such a constraint gives multiple results as any set of \aspil{param} atoms satisfying the cardinality will lead to a new global set of answers.
In this way, we enumerate all possible parametrizations which respects the results of parameters
inference, but completely disregarding the notion of admissible parametrizations given in
\pref{ssec:admissible-K}.

We rely on integrity constraints to filter only admissible parametrizations.
An integrity constraint is a rule with no head, that makes an answer set unsatisfiable if its body turns out to be true.
Hence, if we suppose that:
\begin{itemize}
  \item the \aspil{less\_active(\ti{a}, \ti{p}, \ti{q})} atom means that $K^\ti{p}_{\ti{a},A,B}$ stands for a configuration with less activating regulators than $K^\ti{q}_{\ti{a},A',B'}$ (\ie $A \subset A'$),
  \item the \aspil{param\_inf(\ti{a}, \ti{p}, \ti{q})} atom means: $K^\ti{p}_{\ti{a},A,B} \leq_{[]} K^\ti{q}_{\ti{a},A',B'}$,
\end{itemize}
the monotonicity assumption is formulated as the following integrity constraint:
\asp{:- less\_active(X, P, Q), not param\_inf(X, P, Q).}
which removes all parametrization results where parameters $K^\aspil{P}_{\aspil{X},A,B}$ and $K^\aspil{Q}_{\aspil{X},A',B'}$ exist such that $A \subset A'$ and $K^\aspil{Q}_{\aspil{X},A',B'} <_{[]} K^\aspil{P}_{\aspil{X},A,B}$, thus violating the monotonicity assumption.
Of course, other assumptions can be formulated in the same way.

This subsection succinctly described how we write ASP programs to represent a model and solve all steps of Thomas' modeling inference.
It finds a particularly interesting application in the enumeration of parameters: all possible parametrizations are generated in separate answer sets, and integrity constraints are formulated to remove those that do not fit the assumptions of admissible parametrizations,
thus reducing the number of interesting parametrizations to be considered in the end.

### 2013-02-01 parts/intro.tex
Initial commit

\section{Introduction}
As regulatory phenomena play a crucial role in biological systems, they need to be studied accurately.
Biological Regulatory Networks (BRNs) consist in sets of either positive or negative mutual effects between the components.
With the purpose of analyzing these systems, they are often modeled as graphs which make it possible to determine the possible evolutions of all the interacting components of the system.
Indeed, besides continuous models of physicists, often designed through systems of ordinary
differential equations, a discrete modeling approach was initiated by René Thomas in 1973
\cite{Thomas73}.

In this approach, the different levels of a component, such as concentration or expression levels, are abstractly represented by (positive) integer values and transitions between these levels may be considered as instantaneous.
Hence, qualitative state graphs may be derived from which we are able to formally find out all the possible behaviors expressed as sequences of transitions between these states.
Nevertheless, these dynamics can be precisely established only with regard to some discrete parameters,
hereafter called ``Thomas' parameters'',
which stand for kinds of ``focal points'', i.e. the evolutionary tendency from each state and depending on the set of the other currently interacting components.
%resources in this very state, that is, the set of the other currently interacting components.
%Hereafter, we refer to these discrete parameters as Thomas' parameters.

Thomas' modeling has motivated numerous works around the link between the Interaction Graph (IG)
(summarizing the global influences between components) and the possible dynamics (e.g.,
\cite{RiCo07,RRT08}),
model reduction (e.g., \cite{Naldi09}), formal checking of dynamics (e.g., \cite{Richard06,Naldi07}), 
and the incorporation of time (e.g., \cite{Siebert06,Ahmad08}) and probability
(e.g., \cite{Twardziok10-CMSB}) dimensions, to name but a few.
While the formal checking of dynamical properties is often limited to small networks because of the
state graph explosion, the main drawback of this framework is the difficulty to specify Thomas'
parameters, especially for large networks.

In order to address the formal checking of dynamical properties within very large BRNs, we recently
introduced in \cite{PMR10-TCSB} a new formalism, named the \emph{``Process Hitting''} (PH), to model
concurrent systems having components with a few qualitative levels.
A PH describes, in an atomic manner, the possible evolutions of a process (representing one
component at one level) triggered by the hit of at most one other process in the system.
This framework can be seen as a special class of formalisms like Petri Nets or Communicating Finite
State Machines, where the causality between actions is restricted.
Thanks to the particular structure of interactions within a PH, very efficient static analysis
methods have been developed to over- and under-approximate reachability properties making tractable
the formal analysis of BRNs with hundreds of components \cite{PMR12-MSCS}.

PH is suitable to model BRNs with different levels of abstraction in the specification of
cooperations (associated influences) between components.
This allows to model BRNs with a partial knowledge on precise evolution functions for components
by capturing the largest (the most general) dynamics.

The objectives of the work presented in this paper are the following.
Firstly, we show that starting from one PH model, it is possible to find back the underlying IG.
We perform an exhaustive search for the possible interactions on one component from all the
others, consistently with the knowledge of the dynamics that these interactions lead to and that are
expressed in PH.
The second phase of our work concerns the Thomas' parameters inference.
It consists in determining the nesting set (possibly too large) of the parameters which necessarily
lead to the satisfaction of the known cooperating constraints.
The resulting BRN dynamics is ensured to respect the PH dynamics, \ie no spurious transitions are
made possible by the inference.
Answer Set Programming (ASP) \cite{Baral03} turns out to be effective for these enumerative searches.

The outcome of this work is twofold.
The first benefit is that such an approach makes it possible to refine the construction of
BRNs with a partial and progressively brought knowledge in PH, while being able to export such
models in the Thomas' framework.
This work thus strengthens the formal link between both modelings.
The second feature of our method is that it can be applied on very large BRNs.

Finally, it must be noticed that we are not interested in this paper in the derivation of one
PH from a BRN (which was previously described in \cite{PMR10-TCSB}) but, on the contrary, to finding out
a set of BRNs from one PH.

Our work is related to the approach of \cite{Khalis09} which relies on temporal logic, and \cite{20646302,DBLP:conf/ipcat/CorblinFTCT12} which also uses constraint programming. Both aim at determining a class of models which are consistent with available partial data on the regulatory structure and dynamical properties.
%These classes are built in order to infer properties common to some studied models.
%In our approach, we intend to focus on the Thomas' parameters inference
Our method is based on a model rather than on constraints, which allows to define some properties on the system structure (such as cooperations).
Furthermore, we claim that we are able to deal with larger biological networks.

\paragraph{Outline}
\pref{sec:frameworks} recalls the PH and Thomas frameworks;
\pref{sec:infer-IG} defines the IG inference from PH;
\pref{sec:infer-K} details the enumeration of Thomas parametrizations compatible with a PH
and discuss its implementation in ASP.
\pref{sec:examples} illustrates the applicability of our method on simple examples
and large biological models.

\paragraph{Notations}
$[i;j]$ is the set of integers $\{ i, i+1, \dots, j \}$;
we note $[i_1;j_1] \leq_{[]} [i_2;j_2] \EQDEF (i_1\leq i_2\wedge j_1\leq j_2)$
and $[i_1;j_1] <_{[]} [i_2;j_2] \EQDEF 
%	[i_1;j_1] \leq_{[]} [i_2;j_2] \wedge (i_1\neq i_2 \vee j_1\neq j_2)$.
(i_1<i_2 \wedge j_1\leq j_2) \vee (i_1\leq i_2\wedge j_1 < j_2)$.
Given an integer $k$, $k<[i;j] \EQDEF k <i$ and $k>[i;j]\EQDEF k > j$.

### 2013-02-01 parts/ph.tex
Initial commit

\subsection{The Process Hitting framework}
\label{ssec:PH}

We recall here the definition and semantics of the Process Hitting (PH), and its usage to model
cooperation between concurrent components.
Two examples of PH modeling a BRN at different abstraction levels are given.
They serve as running examples in the rest of this article.

\medskip

A PH (\pref{def:PH}) gathers a finite number of concurrent \emph{processes}
grouped into a finite set of \emph{sorts}.
A process belongs to a unique sort and is noted $a_i$ where $a$ is the
sort and $i$ the identifier of the process within the sort $a$.
At any time, one and only one process of each sort is present; a state of the PH thus corresponds to the set of such processes.

The concurrent interactions between processes are defined by a set of
\emph{actions}.
Actions describe the replacement of a process by another of the same sort
conditioned by the presence of at most one other process in the current
state of the PH.
An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$ where $a_i,b_j,b_k$ are processes
of sorts $a$ and $b$.
It is required that $b_j\neq b_k$ and that $a=b\Rightarrow a_i=b_j$.
An action $h=\PHfrappe{a_i}{b_j}{b_k}$ is read as ``$a_i$ \emph{hits} $b_j$ to
make it bounce to $b_k$'', and
$a_i,b_j,b_k$ are called respectively \emph{hitter}, \emph{target} and
\emph{bounce} of the action, and can be referred to as
$\PHhitter(h), \PHtarget(h), \PHbounce(h)$, respectively.

\begin{definition}[Process Hitting]\label{def:PH}
A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
\begin{itemize}
\item $\PHs \DEF \{a,b,\dots\}$ is the finite set of \emph{sorts};
\item $\PHl \DEF \prod_{a\in\PHs} \PHl_a$ is the set of states with $\PHl_a = \{a_0,\dots,a_{l_a}\}$
the finite set of \emph{processes} of sort $a\in\Sigma$ and $l_a$ a positive integer with
	$a\neq b\Rightarrow \forall(a_i,b_j)\in\PHl_a\times\PHl_b,a_i\neq b_j$;
\item $\PHa \DEF \{ \PHfrappe{a_i}{b_j}{b_k}, \dots \mid
					(a,b)\in\PHs^2 \wedge (a_i,b_j,b_k)\in \PHl_a\times\PHl_b\times\PHl_b$ \\
	\hspace*{2cm} $\wedge b_j\neq b_k \wedge a=b\Rightarrow a_i=b_j\}$
			is the finite set of \emph{actions}.
\end{itemize}
$\PHproc$ denotes the set of all processes ($\PHproc \DEF \{ a_i\mid a\in\PHs \wedge a_i\in\PHl_a\}$).
\end{definition}

\noindent
The sort of a process $a_i$ is referred to as $\PHsort(a_i)=a$ and the set of
sorts present in an action $h\in\PHa$ as 
$\PHsort(h) = \{\PHsort(\PHhitter(h)),\PHsort(\PHtarget(h))\}$.
Given a state $s\in \PHl$, the process of sort $a\in\PHs$ present in $s$ is
denoted by $\PHget{s}{a}$, that is the $a$-coordinate of the state $s$.
If $a_i\in \PHl_a$, we define the notation $a_i\in s \EQDEF \PHget{s}{a}=a_i$.

An action $h=\PHfrappe{a_i}{b_j}{b_k} \in\PHa$ is \emph{playable} in $s\in L$
if and only if $\PHget{s}{a}=a_i$ and $\PHget{s}{b}=b_j$.
In such a case, $(s\play h)$ stands for the state resulting from the play of
the action $h$ in $s$, that is 
$\PHget{(s\play h)}{b} = b_k$ and 
$\forall c\in\PHs, c\neq b, \PHget{(s\play h)}{c} = \PHget{s}{c}$.
For the sake of clarity, $((s\play h)\play h')$, $h'\in\PHa$ is abbreviated as
$(s\play h\play h')$.

\begin{example*}
\pref{fig:runningPH-1} represents a PH $(\PHs,\PHl,\PHa)$ with
$\PHs = \{a,b,c\}$,
$\PHl_a = \{a_0,a_1,a_2\}$,
$\PHl_b = \{b_0, b_1\}$,
$\PHl_c = \{c_0, c_1\}$, and
\begin{align*}
\PHa & = \{
	\PHfrappe{a_2}{b_1}{b_0},
&&  \PHfrappe{b_0}{a_2}{a_1},
&&	\PHfrappe{c_0}{a_2}{a_1},\\
&&& \PHfrappe{b_0}{a_1}{a_0},
&&	\PHfrappe{c_0}{a_1}{a_0},\\
&&& \PHfrappe{b_1}{a_0}{a_1},
&&	\PHfrappe{c_1}{a_0}{a_1},\\
&&& \PHfrappe{b_1}{a_1}{a_2},
&&	\PHfrappe{c_1}{a_1}{a_2} \}\enspace.
\end{align*}
The action $h=\PHfrappe{b_1}{a_1}{a_2}$ is playable in the state
$s = \state{b_1,a_1,c_0}$; and $s\play h=\state{b_1,a_2,c_0}$.

\begin{figure}[t]
\centering
\scalebox{1.3}{
\begin{tikzpicture}
\TSort{(0,0)}{a}{3}{r}
\TSort{(-3,0.5)}{b}{2}{l}
\TSort{(3,0.5)}{c}{2}{r}

\THit{b_1}{very thick}{a_0}{.west}{a_1}
\THit{b_1}{very thick}{a_1}{.north west}{a_2}
\THit{b_0}{}{a_2}{.west}{a_1}
\THit{b_0}{}{a_1}{.west}{a_0}

\path[bounce, bend left=60]
\TBounce{a_1}{very thick}{a_2}{.south}
\TBounce{a_0}{very thick}{a_1}{.south}
;
\path[bounce, bend right=60]
\TBounce{a_2}{}{a_1}{.north}
\TBounce{a_1}{}{a_0}{.north}
;

\THit{c_1}{very thick}{a_0}{.east}{a_1}
\THit{c_1}{very thick}{a_1}{.north east}{a_2}
\THit{c_0}{}{a_2}{.east}{a_1}
\THit{c_0}{}{a_1}{.east}{a_0}

\path[bounce, bend right=60]
\TBounce{a_1}{very thick}{a_2}{.south east}
\TBounce{a_0}{very thick}{a_1}{.south east}
;
\path[bounce, bend left=60]
\TBounce{a_2}{}{a_1}{.north}
\TBounce{a_1}{}{a_0}{.north east}
;

\THit{a_2}{bend right}{b_1}{.north east}{b_0}
\path[bounce, bend left=80]
\TBounce{b_1}{out=100,in=140}{b_0}{.north}
;

\end{tikzpicture}
}
\caption{\label{fig:runningPH-1}
A Process Hitting (PH) example.
Sorts are represented by labeled boxes, and processes by circles (ticks are
the identifiers of the processes within the sort, for instance, $a_0$ is the
process ticked $0$ in the box $a$).
An action (for instance $\PHfrappe{b_1}{a_1}{a_2}$) is represented by a pair of
directed arcs, having the hit part ($b_1$ to $a_1$) in plain line and the bounce
part ($a_1$ to $a_2$) in dotted line.
Actions involving $b_1$ or $c_1$ are in thick lines.
%The current state is represented by the grayed processes:
%$\state{a_0,b_1,c_0,d_0}$.
}
\end{figure}

This PH example actually models a BRN where the component $a$ has three qualitative
levels and components $b$ and $c$ are boolean.
In this BRN, $b$ and $c$ activate $a$, while $a$ inhibits $b$.
The inhibition of $b$ by $a$ is only effective when $a$ is at level $2$;
in the other cases, $b$ cannot evolve in any direction.
The activation of $a$ by $b$ ($c$) is encoded by the actions making the level of $a$ increase (resp.
decrease) when $b$ ($c$) is present (resp. absent).
It is worth noticing that the activation of $a$ by $b$ ($c$) is independent from $c$ ($b$).
This may express a lack of knowledge on the cooperation between these two regulators:
we thus model an over-approximation of the possible actions.
\end{example*}

\paragraph{Modeling cooperation}
As described in \cite{PMR10-TCSB}, the cooperation between processes to make another bounce can be
expressed in PH by building a \emph{cooperative sort}.
\pref{fig:PH-cooperativity} shows an example of cooperation between processes $b_1$ and $c_1$ to
make $a_1$ bounce to $a_2$:
a cooperative sort $bc$ is defined with 4 processes (one for each sub-state of the presence of
processes $b_1$ and $c_1$).
For the sake of clarity, the $bc$ processes are indexed using the sub-state they represent.
Hence, $bc_{01}$ represents the sub-state $\state{b_0,c_1}$, and so on.
Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the status of the sorts $b$
and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
Then, it is the process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of the
independent hits from $b_1$ and $c_1$.

\begin{figure}[p]
\centering
\scalebox{1.3}{
\begin{tikzpicture}
\TSort{(0,0)}{b}{2}{t}
\TSort{(0,-3.8)}{c}{2}{b}
\TSort{(4.5,-3)}{a}{3}{r}

\TSetTick{bc}{0}{00}
\TSetTick{bc}{1}{01}
\TSetTick{bc}{2}{10}
\TSetTick{bc}{3}{11}
% \TSetSortLbcel{bc}{$\neg a\wedge b$}
\TSort{(-0.5,-2)}{bc}{4}{b}

\THit{b_1}{very thick,bend right}{bc_0}{.north}{bc_2}
\THit{b_1}{very thick,bend right}{bc_1}{.north}{bc_3}
\THit{b_0}{}{bc_2}{.north west}{bc_0}
\THit{b_0}{}{bc_3}{.north west}{bc_1}

\THit{c_0}{}{bc_1}{.south}{bc_0}
\THit{c_0}{}{bc_3}{.south}{bc_2}
\THit{c_1}{very thick}{bc_0}{.south}{bc_1}
\THit{c_1}{very thick}{bc_2}{.south}{bc_3}

\path[bounce, bend right=25]
\TBounce{bc_2}{}{bc_0}{.north east}
\TBounce{bc_3}{}{bc_1}{.north east}
;
\path[bounce, bend left=80, distance=30]
\TBounce{bc_0}{very thick}{bc_2}{.north}
\TBounce{bc_1}{very thick}{bc_3}{.north}
;
\path[bounce, bend right]
\TBounce{bc_0}{very thick}{bc_1}{.west}
\TBounce{bc_2}{very thick}{bc_3}{.west}
;
\path[bounce, bend left]
\TBounce{bc_3}{}{bc_2}{.east}
\TBounce{bc_1}{}{bc_0}{.east}
;

\THit{bc_3}{}{a_1}{.west}{a_2}
\path[bounce, bend left=40]
\TBounce{a_1}{}{a_2}{.south west}
;

\end{tikzpicture}
}

\caption{\label{fig:PH-cooperativity}
A PH modeling a cooperativity between $b_1$ and $c_1$ to make
$a_1$ bounce to $a_2$.
Actions involving $b_1$ or $c_1$ are in thick lines.
}
\end{figure}
\begin{figure}[p]
\centering
\scalebox{1.3}{
\begin{tikzpicture}
\path[use as bounding box] (-4,-1.9) rectangle (4.5,3.9);

\TSort{(0,0)}{a}{3}{l}
\TSort{(3, 3)}{b}{2}{t}
\TSort{(3,-1)}{c}{2}{b}

\TSetTick{bc}{0}{00}
\TSetTick{bc}{1}{01}
\TSetTick{bc}{2}{10}
\TSetTick{bc}{3}{11}
% \TSetSortLbcel{bc}{$\neg a\wedge b$}
\TSort{(-3,-0.5)}{bc}{4}{l}

\THit{bc_3}{}{a_1}{.north west}{a_2}
\THit{bc_0}{}{a_1}{.south west}{a_0}
\path[bounce]
\TBounce{a_1}{bend left}{a_2}{.south west}
\TBounce{a_1}{bend right}{a_0}{.north west}
;

\THit{b_0}{}{a_2}{.east}{a_1}
\THit{b_1}{}{a_0}{.north east}{a_1}
\path[bounce]
\TBounce{a_2}{bend left}{a_1}{.north east}
\TBounce{a_0}{bend right=20}{a_1}{.south}
;

\THit{c_0}{bend right}{a_2}{.south east}{a_1}
\THit{c_1}{bend right}{a_0}{.east}{a_1}
\path[bounce]
\TBounce{a_2}{bend left=20}{a_1}{.north}
\TBounce{a_0}{bend right=30}{a_1}{.south east}
;

\path[dashed,hit]
	(2,-1.3) edge[bend left=10] (-2.3,-0.7)
	(2.2, 3.3) edge[bend right=10] (-2.3,3)
;

\THit{a_2}{bend left,out=40,in=80}{b_1}{.north west}{b_0}
\path[bounce, bend right]
\TBounce{b_1}{}{b_0}{.east}
;

\end{tikzpicture}
}

\caption{\label{fig:runningPH-2}
PH resulting from the refinement of the one in \pref{fig:runningPH-1} by the
specification of several cooperations.
The actions from $b$ and $c$ to the cooperative sort $bc$ are identical to those defined in
\pref{fig:PH-cooperativity} and are represented here by a single dashed arc.
}
\end{figure}

We note that cooperative sorts are standard PH sorts and do not involve any
special treatment regarding the semantics of related actions.

When the number of cooperating processes is large, it is possible to chain several cooperative sorts
to prevent the combinatoric explosion of the number of processes created within cooperative sorts.
For instance, if $b_1$, $c_1$, and $d_1$ cooperate, one can create a cooperative sort $bc$ with 4
processes reflecting the presence of $b_1$ and $c_1$, and a cooperative sort $bcd$ with 4 processes
reflecting the presence of $bc_{11}$ and $d_1$.  Such constructions are helpful in PH
as the static analysis of dynamics developed in \cite{PMR12-MSCS} does not suffer from the number of
sorts, but on the number of processes within a single sort.

While the construction of cooperation in PH allows to encode any boolean functions
between cooperating processes \cite{PMR10-TCSB}, it is worth noticing they introduce a temporal
shift in their application.
This allows the existence of interleaving of actions leading to a cooperative sort representing a
past sub-state of the presence of the cooperative processes.
The resulting behavior is then an over-approximation
of the realization of an instantaneous cooperation.

\begin{example*}
The PH in \pref{fig:runningPH-2} results from the refinement of the PH in \pref{fig:runningPH-1}
where several cooperations have been specified.
In particular, the bounce to $a_2$ is the result of a cooperation between $b_1$ and $c_1$; and the
bounce to $a_0$ of a cooperation between $b_0$ and $c_0$.
Hence, this PH expresses a BRN where $a$ requires both $b$ and $c$ active to reach its
highest level, and $a$ does not become inactive unless both $b$ and $c$ are inactive.
\end{example*}

### 2013-02-04 parts/ph.tex
Add TODOs and REWRITEs

@@
 \paragraph{Modeling cooperation}
+\rewrite{Detail the modeling of cooperations}
 As described in \cite{PMR10-TCSB}, the cooperation between processes to make another bounce can be
@@
 \end{example*}
+
+\rewrite{Explain the efficiency of PH for reachability matters}

### 2013-04-15 parts/ph.tex
minor adds

@@
 \rewrite{Detail the modeling of cooperations}
-As described in \cite{PMR10-TCSB}, the cooperation between processes to make another bounce can be
+As described in \cite{PMR10-TCSB}, the cooperation between several processes to make another process bounce can be
 expressed in PH by building a \emph{cooperative sort}.
@@
 processes $b_1$ and $c_1$).
-For the sake of clarity, the $bc$ processes are indexed using the sub-state they represent.
+For the sake of clarity, the processes of $bc$ are indexed using the sub-state they represent.
 Hence, $bc_{01}$ represents the sub-state $\state{b_0,c_1}$, and so on.
@@
 and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
-Then, it is the process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of the
+Then, it is process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of
 independent hits from $b_1$ and $c_1$.

### 2013-04-17 parts/grn.tex
Add TODOs and REWRITEs
minor adds
add unsigned edges in BRN defs

@@
 \subsection{Thomas' modeling}
-We concisely present the Thomas' modeling of BRNs dynamics, merely inspired by
+
+\todo{Find a new semantics that could tackle unsigned edges}
+\newcommand{\uns}{\pm}
+\newcommand{\sN}{\mathbb{N}}
+
+We concisely present Thomas' modeling of BRNs dynamics, merely inspired by
 \cite{Richard06,BernotSemBRN}.
 In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t. the presented
-inference), we extend the classical formalism by setting parameters to intervals of values instead of single
-values, and briefly discuss this addition.
+inference),
+we propose the notion of unsigned edge modeling an interaction whose nature (activation or inhibition) is undefined,
+and we extend the classical parametrization formalism by setting parameters to intervals of values instead of single values.
+We briefly discuss these additions.
 
@@
 \emph{Interaction Graph} (IG) models the structure of the system by defining the components'
-mutual influences.
-The \emph{parametrization} then specifies the levels to which tends a component when a given
+mutual influences and the conditions of these influences.
+The \emph{parametrization} then specifies the levels towards which tends a component when a given
 configuration of its regulators applies.
@@
 %can be either an activator or an inhibitor of $a$, according to the sign of the interaction and if the expression level of $b$ if above or below the threshold $t$.
-We call $\levelsA{a}{b}$ (resp. $\levelsI{a}{b}$) the levels of $a$ where it is an
-activator (resp. inhibitor) of $b$ (\pref{def:levels});
+%We call $\levelsA{a}{b}$ (resp. $\levelsI{a}{b}$) the levels of $a$ where it is an
+%activator (resp. inhibitor) of $b$ (\pref{def:levels});
+We call $\levels{a}{b}$ the levels of $a$ above the threshold of $\GRNedge{a}{s}{t}{b}$;
 $l_a$ denotes the maximum level of $a$.
@@
 
-\begin{definition}[Interaction Graph]
+\begin{definition}[\new{Interaction Graph}]
 \label{def:ig}
-An \emph{Interaction Graph} (IG) is a triple $(\Gamma, E_+, E_-)$ where $\Gamma$ is a finite number of \emph{components},
-and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
-is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.
+%An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E_+, E_-, E_\uns)$ where $\Gamma$ is a finite number of \emph{components},
+%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
+%is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.
+An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E)$ where $\Gamma$ is a finite number of \emph{components},
+and $E \subset \{ \GRNedge{a}{s}{t}{b} %\in \Gamma \times S \times \sN \times \Gamma 
+\mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\}$
+%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
+is the set of 
+%positive (resp. negative)
+\emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}.
 
 A regulation from $a$ to $b$ is uniquely referenced:
-if $a \xrightarrow{t} b \in E_+$ (resp. $E_-$),
-$\nexists a \xrightarrow{t'} b \in E_+ \text{ (resp. $E_-$)}, t' \neq t$
-and $\nexists a \xrightarrow{t'} b \in E_-\text{ (resp. $E_+$)}, t' \in \mathbb{N}$.
+$\forall \GRNedge{a}{s}{t}{b} \in E$, 
+$\exists a \xrightarrow{s', t'} b \in E \Rightarrow s = s' \wedge t = t'$.
+%if $a \xrightarrow{t} b \in E_+$ (resp. $E_-$),
+%$\nexists a \xrightarrow{t'} b \in E_+ \text{ (resp. $E_-$)}, t' \neq t$
+%and $\nexists a \xrightarrow{t'} b \in E_-\text{ (resp. $E_+$)}, t' \in \mathbb{N}$.
 \end{definition}
 
-\begin{definition}[Effective levels ($\levels$)]\label{def:levels}
-Let $(\Gamma,E_+,E_-)$ be an IG and $a, b \in \Gamma$ two of its components:
+\begin{definition}[\new{Effective levels ($\levelsl$)}]\label{def:levels}
+Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components:
 \begin{itemize}
-  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
-    $\levelsI{a}{b} \DEF [0; t-1]$;
-  \item if $a \xrightarrow{t} b \in E_-$, $\levelsA{a}{b} \DEF [0; t-1]$ and
-    $\levelsI{a}{b} \DEF [t; l_a]$;
-  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
+%  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
+%    $\levelsI{a}{b} \DEF [0; t-1]$;
+%  \item if $a \xrightarrow{t} b \in E_-$, $\levelsA{a}{b} \DEF [0; t-1]$ and
+%    $\levelsI{a}{b} \DEF [t; l_a]$;
+%  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
+  \item if $\GRNedge{a}{s}{t}{b} \in E$, $\levels{a}{b} \DEF [t; l_a]$;
+  \item otherwise, $\levels{a}{b} \DEF \emptyset$.
 \end{itemize}
@@
 
-For all component $a \in \Gamma$, 
-$\GRNreg{a} \DEF \{ b\in\Gamma\mid \exists b\xrightarrow t a\in E_+\cup E_- \}$
+Given this definition, we denote:
+$E_s = \{ \GRNedge{a}{s}{t}{b} \in E \}$
+for $s \in \{ +, -, \uns \}$.
+Furthermore,
+for all component $a \in \Gamma$, 
+$\GRNreg{a} \DEF \{ b\in\Gamma\mid \exists \GRNedge{b}{s}{t}{a} \in E \}$
 is the set of its regulators.
@@
 \begin{example*}
-\pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E_+,E_-)$ with
+\pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E)$ with
 $\Gamma = \{a, b, c\}$,
-$E_+ = \{b \xrightarrow{1} a, c \xrightarrow{1} a\}$ and
-$E_- = \{a \xrightarrow{2} b\}$;
+$E_+ = \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}\}$,
+$E_- = \{\GRNedgef{a}{-}{2}{b}\}$ and
+$E_\uns = \emptyset$;
 hence $\GRNreg{a} = \{b, c\}$.
@@
 
-A \emph{state} $s$ of an IG $(\Gamma, E_+, E_-)$ is an element in $\prod_{a \in \Gamma} [0;l_a]$.
+A \emph{state} $s$ of an IG $(\Gamma, E)$ is an element in $\prod_{a \in \Gamma} [0;l_a]$.
 $\GRNget{s}{a}$ refers to the level of component $a$ in $s$.
@@
 
-\begin{definition}[Discrete parameter $K_{a,A,B}$ and Parametrization $K$]\label{def:param}
-For a given component $a \in \Gamma$ and $A$ (resp. $B$) $\subset \GRNreg{a}$ a set of its activators (resp. inhibitors) such that $A \cup B = \GRNreg{a}$ and $A \cap B = \emptyset$,
-the discrete \emph{parameter} $K_{a,A,B} = [i; j]$ is a non-empty interval towards which $a$ will tend
-in the states where its activators (resp. inhibitors) are the regulators in set $A$ (resp. $B$).
+\begin{definition}[Resources $\GRNres{a}{s}$]\label{def:resources}
+For a given state $s$ of a BRN, we define the set of \emph{resources} of a component $a$ in $s$ as $\GRNres{a}{s}$, where:
+%\begin{align*}
+%  A &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsA{b}{a}\} \\
+%  B &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsI{b}{a}\}
+%\end{align*}
+$$\GRNres{a}{s} = \{ b \in \GRNreg{a} \mid \GRNget{s}{b} \in \levels{b}{a} \}$$
+%We also denote: $\GRNallres{a} = \{(A;B) \mid \exists s \in \textstyle\prod_{a \in \Gamma} [0;l_a], \GRNres{a}{s} = A,B\}$
+We also denote: $\GRNallres{a} = \{ \omega \subset \GRNreg{a} \}$
+\end{definition}
+
+\begin{definition}[Discrete parameter $K_{a,\omega}$ and Parametrization $K$]\label{def:param}
+For a given component $a \in \Gamma$ and $\omega \subset \GRNreg{a}$ a set of regulators of $a$,
+the discrete \emph{parameter} $K_{a,\omega} = [i; j]$ is a non-empty interval towards which $a$ will tend
+in the states where its resources are exactly the regulators in $\omega$.
 The complete map $K$ of discrete parameters for $\IG$ is called a \emph{parametrization} of $\IG$.
+%For a given component $a \in \Gamma$ and $A$ (resp. $B$) $\subset \GRNreg{a}$ a set of its activators (resp. inhibitors) such that $A \cup B = \GRNreg{a}$ and $A \cap B = \emptyset$,
+%the discrete \emph{parameter} $K_{a,A,B} = [i; j]$ is a non-empty interval towards which $a$ will tend
+%in the states where its activators (resp. inhibitors) are the regulators in set $A$ (resp. $B$).
+%The complete map $K$ of discrete parameters for $\IG$ is called a \emph{parametrization} of $\IG$.
 \end{definition}
@@
 %We also denote: $j < K_{a,A,B} \Leftrightarrow j < i_1$ and $j > K_{a,A,B} \Leftrightarrow j> i_2$.
-
-\begin{definition}[Resources $\GRNres{a}{s}$]\label{def:resources}
-For a given state $s$ of a BRN, we define the \emph{activators} $A$ and \emph{inhibitors} $B$ of $a$ in $s$ as $\GRNres{a}{s} = A,B$, where:
-\begin{align*}
-  A &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsA{b}{a}\} \\
-  B &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsI{b}{a}\}
-\end{align*}
-We also denote: $\GRNallres{a} = \{(A;B) \mid \exists s \in \textstyle\prod_{a \in \Gamma} [0;l_a], \GRNres{a}{s} = A,B\}$
-\end{definition}
 
@@
 At last, \pref{def:dynamics} gives the asynchronous dynamics of a BRN using Thomas' parameters.
-From a given state $s$, a transition to another state $s'$ is possible provided that only one component $a$ will evolve of one level towards $K_{a,\GRNres{a}{s}}$.
+From a given state $s$, a transition to another state $s'$ is possible provided that only one component $a$ will evolve of one level towards $K_{a,\GRNres{a}{s}}$,
+as stated by the definition of the transition relation $\GRNtrans{s}{s'}$.
 
-\begin{definition}[Asynchronous dynamics]\label{def:dynamics}
+\begin{definition}[\new{Asynchronous dynamics}]\label{def:dynamics}
+The dynamics of a BRN using Thomas' parameters is defined by the transition relation $\GRN{}{} \in \prod_{a \in \Gamma} [0;l_a] \times \prod_{a \in \Gamma} [0;l_a]$:
+$\forall s, s' \in \prod_{a \in \Gamma} [0;l_a]$,
+\begin{align*}
+\GRNtrans{s}{s'} &\Leftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
+  & \qquad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
+\end{align*}
+with: $f^a(s) = 
+  \begin{cases}
+    +1 & \text{if } \GRNget{s}{a} < K_{a, \GRNres{a}{s}} \\
+    -1 & \text{if } \GRNget{s}{a} > K_{a, \GRNres{a}{s}} \\
+  \end{cases}$
+\begin{comment}
 Let $s$ be a state of a BRN using Thomas' parameters $(\IG, K)$ where $\IG = (\Gamma, E_+, E_-)$.
@@
 \end{align*}
+\end{comment}
 \end{definition}
@@
 (as for $b$ in \pref{fig:runningBRN}, for instance).
+\rewrite{Detail the reason of using interval parameters}
 

### 2013-04-17 parts/infer-ig.tex
Add TODOs and REWRITEs
minor adds
add unsigned edges to IG inference

@@
 
+%\todo{Try to use projections in order to perform the IG inference}
+
 At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
@@
 \begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
-We infer $\IG = (\Gamma,E_+,E_-,E_\pm)$ from \pref{pps:inference-edges} as follows:
+We infer $\IG = (\Gamma,E)$ from \pref{pps:inference-edges} as follows:
 \begin{align*}
-E_+ &= \{a \xrightarrow{t} b \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
+E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
   \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
-E_- &= \{a \xrightarrow{t} b \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
+E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
   \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
-E_\pm &= \{a \rightarrow b \mid \exists a \xrightarrow{t} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t'} b \in \hat{E}_-\} \\
+E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_-
+  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\} \\
 \end{align*}

### 2013-04-17 parts/infer-param.tex
Add TODOs and REWRITEs
add unsigned edges to parameters inference

@@
 
+\todo{Try to address the inference of parameters despite unsigned edges}
+
 This subsection presents some results related to the inference of independent discrete parameters from a given PH.
@@
 regarding $a$.
-\begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
+\begin{property}[\new{Well-formed PH for parameter inference}]\label{pro:wf-ph-K}
 A PH is well-formed for parameter inference if and only if
@@
 \forall b\in \GRNreg{a} &,
-	\forall (i,j\in\levelsA{b}{a} \vee i,j\in\levelsI{b}{a}), \\
+	\forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
 & \qquad	\forall c, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
@@
 
-Let $K_{a,A,B}$ be the parameter we want to infer for a given component $a \in \Gamma$
+Let $K_{a,\omega}$ be the parameter we want to infer for a given component $a \in \Gamma$
 %and $A,B \in \GRNallres{a}$ a configuration of resources of $a$ (activators and inhibitors).
-and $A \subset \GRNreg{a}$ (resp. $B \subset \GRNreg{a}$) a set of its activators (resp. inhibitors).
+and $\omega \subset \GRNreg{a}$ a set of its regulators.
 This inference, as for the IG inference, relies on the search of focal processes of the component for the given configuration of its regulators.
 
-For each sort $b \in \GRNreg{a}$, we define a context $C^b_{a,A,B}$ in \pref{eq:param_context} that contains all processes representing the influence of the regulators in the configuration $A,B$.
+For each sort $b \in \GRNreg{a}$, we define a context $C^b_{a,\omega}$ in \pref{eq:param_context} that contains all processes representing the influence of the resources in the configuration modeled by $\omega$.
 The context of a cooperative sort $\upsilon$ that regulates $a$ is given in
 \pref{eq:param_context_coop} as the set of focal processes matching the current configuration.
-$C_{a,A,B}$ refers to the union of all these contexts (\pref{eq:K-ctx}).
+$C_{a,\omega}$ refers to the union of all these contexts (\pref{eq:K-ctx}).
 \begin{align}
@@
 \forall b\in\Gamma,~
-C_{a,A,B}^b & \DEF \begin{cases}
-\levelsA{b}{a} & \text{if $b \in A$,}\\
-\levelsI{b}{a} & \text{if $b \in B$,}\\
+C_{a,\omega}^b & \DEF \begin{cases}
+  \levels{b}{a} & \text{if $b \in \omega$,}\\
+  \ulevels{b}{a} & \text{if $b \notin \omega$,}\\
   L_b            & \text{otherwise;}\\
@@
 \forall \upsilon \in \PHpredec{a}\setminus\Gamma,~
-C_{a,A,B}^\upsilon & \DEF \{
-\upsilon(\sigma) \mid \sigma \in \textstyle\prod_{c\in\PHdirectpredec{\upsilon}}C_{a,A,B}^c \}
+C_{a,\omega}^\upsilon & \DEF \{
+\upsilon(\sigma) \mid \sigma \in \textstyle\prod_{c\in\PHdirectpredec{\upsilon}}C_{a,\omega}^c \}
 \\
-C_{a,A,B} & \DEF \textstyle\bigcup_{b\in\PHpredec{a}} C^b_{a,A,B}
+C_{a,\omega} & \DEF \textstyle\bigcup_{b\in\PHpredec{a}} C^b_{a,\omega}
 \label{eq:K-ctx}
@@
 
-The parameter $K_{a,A,B}$ specifies to which values $a$ eventually evolves as long as the context
-$C_{a,A,B}$ holds, which is precisely the definition of the $\focals$ function
+The parameter $K_{a,\omega}$ specifies to which values $a$ eventually evolves as long as the context
+$C_{a,\omega}$ holds, which is precisely the definition of the $\focals$ function
 (\pref{def:focals} in \pref{ssec:focal}),
@@
 \pref{eq:param_context_coop}.
-Hence $K_{a,A,B} = \focals(a,C^a_{a,A,B},C_{a,A,B})$ if this latter is a non-empty interval
+Hence $K_{a,\omega} = \focals(a,C^a_{a,\omega},C_{a,\omega})$ if this latter is a non-empty interval
 (\pref{pps:param_K}).
@@
 \label{pps:param_K}
-Let $(\PHs, \PHl, \PHh)$ be a Process Hitting well-formed for parameter inference, and $\IG = (\Gamma,
-E_+, E_-)$ the inferred IG.
+Let $(\PHs, \PHl, \PHh)$ be a Process Hitting well-formed for parameter inference, and $\IG = (\Gamma, E)$ the inferred IG.
 Let $A$ (resp. $B$) $\subseteq \Gamma$ be the set of regulators that activate (resp. inhibit) a sort
@@
 %If $\focals(a,C_{a,A,B})$ is a non-empty interval, then $K_{a,A,B} = \focals(a, C_{a,A,B})$.
-If $\focals(a,C^a_{a,A,B},C_{a,A,B})=[a_i;a_j]$ is a non-empty interval, 
-	then $K_{a,A,B} = [i;j]$.
+If $\focals(a,C^a_{a,\omega},C_{a,\omega})=[a_i;a_j]$ is a non-empty interval, 
+	then $K_{a,\omega} = [i;j]$.
 \end{proposition}
@@
 \begin{property}[Parameter validity]\label{pro:K-valid}
-A parameter $K_{a,A,B}$ is valid w.r.t. the PH iff the following equation is verified:
+A parameter $K_{a,\omega}$ is valid w.r.t. the PH iff the following equation is verified:
 \begin{align*}
-\forall a_i\in C^a_{a,A,B},
-		a_i \notin K_{a,A,B} \Longrightarrow (
-  \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, & c_k \in C^c_{a,A,B} \\
- \wedge a_i < K_{a,A,B} \Rightarrow j > i 
- & \wedge  a_i > K_{a,A,B} \Rightarrow j <i )
+\forall a_i\in C^a_{a,\omega},
+		a_i \notin K_{a,\omega} \Longrightarrow (
+  \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, & c_k \in C^c_{a,\omega} \\
+ \wedge a_i < K_{a,\omega} \Rightarrow j > i 
+ & \wedge  a_i > K_{a,\omega} \Rightarrow j <i )
 \end{align*}
@@
 \begin{property}[Extreme values assumption]
-Let $\IG = (\Gamma, E_+, E_-)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} iff:
+Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} iff:
 \label{prop:param_enum_extreme}
 \[
-  \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Rightarrow 0 \in K_{b,\emptyset,\GRNreg{b}} \wedge l_b \in K_{b,\GRNreg{b},\emptyset}
+  \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Longrightarrow \exists \omega \subset \GRNreg{b}, 0 \in K_{b,\omega} \wedge \exists \omega' \subset \GRNreg{b}, l_b \in K_{b,\omega'}
 \]
@@
 \label{prop:param_enum_activity}
-Let $\IG = (\Gamma, E_+, E_-)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} iff:
+Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} iff:
 \begin{align*}
-  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A,B} <_{[]} K_{b,A \cup \{b\},B \setminus \{b\}}
-\\
-  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A \setminus \{b\},B \cup \{b\}} <_{[]} K_{b,A,B}
+%  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A,B} <_{[]} K_{b,A \cup \{b\},B \setminus \{b\}}
+%\\
+%  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A \setminus \{b\},B \cup \{b\}} <_{[]} K_{b,A,B}
+  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists \omega \subset \GRNreg{b}, K_{b,\omega} \neq K_{b,\omega \cup \{ a \}}
 \end{align*}
@@
 \label{prop:param_enum_monotonicity}
-Let $\IG = (\Gamma, E_+, E_-)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} iff:
-\[
-  \forall b \in \Gamma, \forall (A;B), (A';B') \in \GRNallres{b},
-  A \subset A' \wedge B' \subset B \Rightarrow K_{b,A,B} \leq_{[]} K_{b,A',B'}
-\]
+Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} iff:
+%\[
+%  \forall b \in \Gamma, \forall (A;B), (A';B') \in \GRNallres{b},
+%  A \subset A' \wedge B' \subset B \Rightarrow K_{b,A,B} \leq_{[]} K_{b,A',B'}
+%\]
+$\forall b \in \Gamma$,
+$\forall X^+ \subset \{ a \in \Gamma \mid \GRNedge{a}{+}{t}{b} \in E_+ \}$,
+$\forall X^- \subset \{ a \in \Gamma \mid \GRNedge{a}{-}{t}{b} \in E_- \}$,
+%A \subset A' \wedge B' \subset B \Rightarrow
+$K_{b,\omega \cup X^-} \leq_{[]} K_{b,\omega \cup X^+}$
 \end{property}
@@
 \subsection{Answer Set Programming implementation concepts}
+
+\rewrite{This whole subsection may be completed, or turned into a new section}
+
+\todo{Rewrite implementation with the new semantics}
 

### 2013-04-17 parts/grn.tex
add unsigned edges to parameters inference
more explicit \new labels

@@
 
-\begin{definition}[\new{Interaction Graph}]
+\new{Added unsigned edges ($\uns$) that stand for an unknown king of regulation.
+The definition of $\levels{a}{b}$ now gives the levels for which $a$ regulates $b$ (i.e. $a > t$), independently from the regulation sign.}
+
+\begin{definition}[Interaction Graph]
 \label{def:ig}
@@
 
-\begin{definition}[\new{Effective levels ($\levelsl$)}]\label{def:levels}
-Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components:
-\begin{itemize}
+\begin{definition}[Effective levels ($\levelsl$)]\label{def:levels}
+Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components.
+If $\GRNedge{a}{s}{t}{b} \in E$, we define: $\levels{a}{b} \DEF [t; l_a]$ and $\ulevels{a}{b} \DEF [0; t-1]$.
+%%%
+%\begin{itemize}
 %  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
@@
 %  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
-  \item if $\GRNedge{a}{s}{t}{b} \in E$, $\levels{a}{b} \DEF [t; l_a]$;
-  \item otherwise, $\levels{a}{b} \DEF \emptyset$.
-\end{itemize}
+%%%
+%  \item if $\GRNedge{a}{s}{t}{b} \in E$, $\levels{a}{b} \DEF [t; l_a]$;
+%  \item otherwise, $\levels{a}{b} \DEF \emptyset$.
+%\end{itemize}
 \end{definition}
@@
 
+\new{No more $A,B$ but sets $\omega$ of resources instead. Definitions of $K$ were adapted.}
+
 \begin{definition}[Resources $\GRNres{a}{s}$]\label{def:resources}
@@
 
-\begin{definition}[\new{Asynchronous dynamics}]\label{def:dynamics}
-The dynamics of a BRN using Thomas' parameters is defined by the transition relation $\GRN{}{} \in \prod_{a \in \Gamma} [0;l_a] \times \prod_{a \in \Gamma} [0;l_a]$:
+\begin{definition}[Asynchronous dynamics]\label{def:dynamics}
+The dynamics of a BRN using Thomas' parameters is defined by the transition relation $\GRNtrans{}{} \in \prod_{a \in \Gamma} [0;l_a] \times \prod_{a \in \Gamma} [0;l_a]$:
 $\forall s, s' \in \prod_{a \in \Gamma} [0;l_a]$,

### 2013-04-17 parts/infer-param.tex
more explicit \new labels

@@
 regarding $a$.
-\begin{property}[\new{Well-formed PH for parameter inference}]\label{pro:wf-ph-K}
+
+\new{Changed $\levelsA{a}{b}$ and $\levelsI{a}{b}$ into $\levels{a}{b}$ in \pref{eq:wf-levels} and \pref{eq:param_context}.}
+
+\begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
 A PH is well-formed for parameter inference if and only if
@@
 We remark that any parameter inferred by \pref{pps:param_K} satisfies this property.
+
+\new{Update all assumptions.}
 

### 2013-04-17 parts/grn.tex
update notations in BRN defs

@@
 \todo{Find a new semantics that could tackle unsigned edges}
-\newcommand{\uns}{\pm}
-\newcommand{\sN}{\mathbb{N}}
 
@@
 
+Given this definition, we denote as a shortcut:
+$E_s = \{ \GRNedge{a}{s}{t}{b} \in E \}$
+for $s \in \{ +, -, \uns \}$.
+Furthermore, for all component $a \in \Gamma$, 
+we denote $\GRNreg{a}$ the set of its regulators as defined in \pref{eq:grn-regulators}.
+\[
+  \label{eq:grn-regulators}
+  \GRNreg{a} \DEF \{ b\in\Gamma\mid \exists \GRNedge{b}{s}{t}{a} \in E \}
+\]
+We allow any number of levels for the components, without considering the number of outgoing edges,
+as the number of processes in a PH sort is not constrained in any way.
+\rewrite{Detail}
+%$E_+(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in
+%E_+\}$ (resp. $E_-(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in E_-\}$) denotes the set
+%of its positive (resp. negative) regulators; and
+%$\GRNreg{a} = E_+(a)\cup E_-(a)$ the set of its regulators.
+
 \begin{definition}[Effective levels ($\levelsl$)]\label{def:levels}
-Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components.
-If $\GRNedge{a}{s}{t}{b} \in E$, we define: $\levels{a}{b} \DEF [t; l_a]$ and $\ulevels{a}{b} \DEF [0; t-1]$.
+%Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components.
+Let $b \in \Gamma$ and $a \in \GRNreg{a}$;
+we define: $\levels{a}{b} \DEF [t; l_a]$ and $\ulevels{a}{b} \DEF [0; t-1]$.
 %%%
@@
 
-Given this definition, we denote:
-$E_s = \{ \GRNedge{a}{s}{t}{b} \in E \}$
-for $s \in \{ +, -, \uns \}$.
-Furthermore,
-for all component $a \in \Gamma$, 
-$\GRNreg{a} \DEF \{ b\in\Gamma\mid \exists \GRNedge{b}{s}{t}{a} \in E \}$
-is the set of its regulators.
-We allow any number of levels for the components, without considering the number of outgoing edges,
-as the number of processes in a PH sort is not constrained in any way.
-%$E_+(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in
-%E_+\}$ (resp. $E_-(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in E_-\}$) denotes the set
-%of its positive (resp. negative) regulators; and
-%$\GRNreg{a} = E_+(a)\cup E_-(a)$ the set of its regulators.
-
 \begin{example*}
 \pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E)$ with
-$\Gamma = \{a, b, c\}$,
-$E_+ = \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}\}$,
-$E_- = \{\GRNedgef{a}{-}{2}{b}\}$ and
-$E_\uns = \emptyset$;
-hence $\GRNreg{a} = \{b, c\}$.
+$\Gamma = \{a, b, c\}$ and
+\begin{align*}
+  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}\} &
+  E_\uns &= \emptyset \\
+  E_- &= \{\GRNedgef{a}{-}{2}{b}\}
+\end{align*}
+Hence:
+\begin{align*}
+  \GRNreg{a} &= \{ b, c \} &
+  \GRNreg{b} &= \{ a \} \\
+  \GRNreg{c} &= \emptyset
+\end{align*}
 %This IG can represent the same behavior as the PH given in \pref{fig:runningPH-2}.
@@
 \begin{align*}
-K_{a,\{b,c\},\emptyset} &= [2 ; 2] & K_{b,\{a\},\emptyset} &= [0 ; 1] \\
-K_{a,\{b\},\{c\}} &= [1 ; 1] & K_{b,\emptyset,\{a\}} &= [0 ; 0] \\
-K_{a,\{c\},\{b\}} &= [1 ; 1] &&\\
-K_{a,\emptyset,\{b,c\}} &= [0 ; 0] & K_{c,\emptyset,\emptyset} &= [0 ; 1]
+K_{a,\{b,c\}} &= [2 ; 2] & K_{b,\emptyset} &= [0 ; 1] \\
+K_{a,\{b\}} &= [1 ; 1] & K_{b,\{a\}} &= [0 ; 0] \\
+K_{a,\{c\}} &= [1 ; 1] &&\\
+K_{a,\emptyset} &= [0 ; 0] & K_{c,\emptyset} &= [0 ; 1]
 \end{align*}
@@
 
-A \emph{state} $s$ of an IG $(\Gamma, E)$ is an element in $\prod_{a \in \Gamma} [0;l_a]$.
+A \emph{state} $s$ of an IG $(\Gamma, E)$ is an element in $\GRNstates \DEF \prod_{a \in \Gamma} [0;l_a]$.
 $\GRNget{s}{a}$ refers to the level of component $a$ in $s$.
+\rewrite{Functional definition of a state? (More rigorous)}
 The specificity of Thomas' approach lies in the use of discrete \emph{parameters} to represent the
@@
 \begin{definition}[Resources $\GRNres{a}{s}$]\label{def:resources}
-For a given state $s$ of a BRN, we define the set of \emph{resources} of a component $a$ in $s$ as $\GRNres{a}{s}$, where:
+For a given component $a \in \Gamma$ and a state $s \in \GRNstates$,
+the set of regulators of $a$ whose level in $s$ is above the required threshold to regulate $a$
+is called the set of \emph{resources} of $a$ in $s$ and is noted $\GRNres{a}{s}$:
 %\begin{align*}
@@
 %We also denote: $\GRNallres{a} = \{(A;B) \mid \exists s \in \textstyle\prod_{a \in \Gamma} [0;l_a], \GRNres{a}{s} = A,B\}$
-We also denote: $\GRNallres{a} = \{ \omega \subset \GRNreg{a} \}$
+%We also denote: $\GRNallres{a} = \{ \omega \subset \GRNreg{a} \}$
 \end{definition}
@@
 
+\new{Use a transition relation instead of an indeterministic function.}
+
 \begin{definition}[Asynchronous dynamics]\label{def:dynamics}
-The dynamics of a BRN using Thomas' parameters is defined by the transition relation $\GRNtrans{}{} \in \prod_{a \in \Gamma} [0;l_a] \times \prod_{a \in \Gamma} [0;l_a]$:
-$\forall s, s' \in \prod_{a \in \Gamma} [0;l_a]$,
+The dynamics of a BRN using Thomas' parameters is given by the transition relation $\GRNtrans{}{} \in \GRNstates \times \GRNstates$ defined by:
 \begin{align*}
-\GRNtrans{s}{s'} &\Leftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
+  \forall s, s' \in \GRNstates, \GRNtrans{s}{s'} &\Longleftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
     & \qquad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
@@
 definitions).
-Indeed, assume that $K_{a,A,B} = [i ; i+2]$;
-we aim at obtaining three different parameters $K_{a,A_1,B_1} = i$,  $K_{a,A_2,B_2} = i+1$,
-$K_{a,A_3,B_3} = i+2$.
+Indeed, assume that $K_{a,\omega} = [i ; i+2]$;
+we aim at obtaining three different parameters $K_{a,\omega_1} = i$,  $K_{a,\omega_2} = i+1$, $K_{a,\omega_3} = i+2$.
 The only possible modification in resources is to add $a$ as a self-regulator.
-However, because resources have a boolean definition (a component is either an activator or an inhibitor of
-$a$), it is not possible to differentiate the 3 values.
+However, because resources have a boolean definition,
+%(a component is either an activator or an inhibitor of $a$),
+it is not possible to differentiate the 3 values.
 We also remark that the use of intervals makes optional some explicit auto-activations in the IG
@@
 transitions are possible given the semantics defined in \pref{def:dynamics}:
-$\GRNetat{a_0, b_1, c_1} \rightarrow \GRNetat{a_1, b_1, c_1} \rightarrow \GRNetat{a_2, b_1, c_1} \rightarrow
-\GRNetat{a_2, b_0, c_1} \rightarrow \GRNetat{a_1, b_0, c_1}$,
+$\GRNstate{a_0, b_1, c_1} \rightarrow \GRNstate{a_1, b_1, c_1} \rightarrow \GRNstate{a_2, b_1, c_1} \rightarrow
+\GRNstate{a_2, b_0, c_1} \rightarrow \GRNstate{a_1, b_0, c_1}$,
 ending in a steady state,
-where $a_i$ is the component $a$ at level $i$.
-As $K_{b,\{a\},\emptyset} = [0 ; 1]$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
+where $a_i$ denotes the component $a$ at level $i$.
+As $K_{b,\emptyset} = [0 ; 1]$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
 \end{example*}

### 2013-04-17 parts/infer-param.tex
finish updating notations in parameters inference

@@
 which implies that the inferred IG does not contain any unsigned interactions, and thus can be seen as the
-regular IG $(\Gamma, E_+, E_-)$,
-and that any processes in $\levelsA{b}{a}$ (resp. $\levelsI{b}{a}$) share the same behavior
+regular IG $(\Gamma, E)$,
+and that any processes in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
 regarding $a$.
@@
 Applied to the PH in \pref{fig:runningPH-1}, we obtain, for instance, 
-$K_{b,\{a\},\emptyset} = [0 ; 1]$ and
-$K_{a,\{b,c\},\emptyset} = [2 ; 2]$,
-while $K_{a,\{b\},\{c\}}$ can not be inferred.
+$K_{b,\emptyset} = [0 ; 1]$ and
+$K_{a,\{b,c\}} = [2 ; 2]$,
+while $K_{a,\{b\}}$ can not be inferred.
 For the PH in \pref{fig:runningPH-2}, this latter is evaluated to $[1;1]$.

### 2013-04-18 parts/infer-param.tex
fix property of well-formed PH for parameter inference

@@
 it is well-formed for IG inference, and
-the IG $(\Gamma, E_+, E_-, E_\pm)$ inferred by \pref{pps:inference-IG}
-verifies $E_\pm=\emptyset$ and if the following property holds:
+the IG $(\Gamma, E)$ inferred by \pref{pps:inference-IG}
+verifies the following property:
 \begin{align}
@@
 	\forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
-& \qquad	\forall c, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
+& \qquad \forall c \in \Gamma, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
 & \qquad \qquad

### 2013-04-23 parts/grn.tex
several adds and remove some labels

@@
 \subsection{Thomas' modeling}
-
-\todo{Find a new semantics that could tackle unsigned edges}
 
@@
 %For the sake of simplicity and to establish a parallel with PH, if $a$ represents a component, we call $a_i$ its $i^\text{th}$ expression level.
-
-\new{Added unsigned edges ($\uns$) that stand for an unknown king of regulation.
-The definition of $\levels{a}{b}$ now gives the levels for which $a$ regulates $b$ (i.e. $a > t$), independently from the regulation sign.}
+\rewrite{Rewrite + reference + add some glue}
+%\new{Added unsigned edges ($\uns$) that stand for an unknown king of regulation.
+%The definition of $\levels{a}{b}$ now gives the levels for which $a$ regulates $b$ (i.e. $a > t$), independently from the regulation sign.}
 
@@
 %is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.
-An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E)$ where $\Gamma$ is a finite number of \emph{components},
-and $E \subset \{ \GRNedge{a}{s}{t}{b} %\in \Gamma \times S \times \sN \times \Gamma 
-\mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\}$
+An \emph{Interaction Graph} (IG) is a couple $(\Gamma, E)$ where $\Gamma$ is the finite set of \emph{components},
+and $E$
 %and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
-is the set of 
+is the finite set of 
 %positive (resp. negative)
-\emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}.
-
+\emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}:
+$$E \DEF \{ \GRNedge{a}{s}{t}{b}, \ldots \mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\}$$
 A regulation from $a$ to $b$ is uniquely referenced:
-$\forall \GRNedge{a}{s}{t}{b} \in E$, 
-$\exists a \xrightarrow{s', t'} b \in E \Rightarrow s = s' \wedge t = t'$.
+$$\forall \GRNedge{a}{s}{t}{b} \in E, \forall \GRNedge{a}{s'}{t'}{b} \in E, s = s' \wedge t = t' \enspace.$$
 %if $a \xrightarrow{t} b \in E_+$ (resp. $E_-$),
@@
 Given this definition, we denote as a shortcut:
-$E_s = \{ \GRNedge{a}{s}{t}{b} \in E \}$
+$E_s \DEF \{ \GRNedge{a}{s}{t}{b} \in E \}$
 for $s \in \{ +, -, \uns \}$.
@@
 \]
-We allow any number of levels for the components, without considering the number of outgoing edges,
+The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$, \ie the number of components that $a$ regulates.
+Indeed, each discrete expression level in $[0 ; l_a]$ stands for an arbitrary set of concentration levels for which $a$ has a certain influence on several other components it regulates.
+Allowing more expression levels makes several levels useless as they represent the same influence of $a$ than some other levels;
+on the contrary, reducing the maximum level requires that several thresholds on outgoing edges of $a$ share the same value, which is biologically unlikely.
+However, these considerations are usually taken into account when one has to directly build a model in this framework, but are not mandatory.
+In the scope of this paper, we allow any number of levels for components, without considering the number of outgoing edges,
 as the number of processes in a PH sort is not constrained in any way.
-\rewrite{Detail}
+
 %$E_+(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in
@@
 \pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E)$ with
-$\Gamma = \{a, b, c\}$ and
+$\Gamma = \{a, b, c\}$ and:
 \begin{align*}
@@
 Regulations are represented by the edges labeled with their sign and threshold.
-For instance, the edge from $b$ to $a$ is labeled $+1$, which stands for: $b \xrightarrow{1} a \in
-E_+$.
+For instance, the edge from $b$ to $a$ is labeled $+1$, which stands for: $\GRNedgef{b}{+}{1}{a}$.
 %and means that if the level of expression of $b$ is equal to (i.e. above) 1, then $b$ activates $a$,
@@
 referred to as the \emph{resources} of $a$ in this state (\pref{def:resources}).
+%
+%\new{No more $A,B$ but sets $\omega$ of resources instead. Definitions of $K$ were adapted.}
 
-\new{No more $A,B$ but sets $\omega$ of resources instead. Definitions of $K$ were adapted.}
-
-\begin{definition}[Resources $\GRNres{a}{s}$]\label{def:resources}
+\begin{definition}[Resources ($\GRNreslabel$)]\label{def:resources}
 For a given component $a \in \Gamma$ and a state $s \in \GRNstates$,
@@
 %\end{align*}
-$$\GRNres{a}{s} = \{ b \in \GRNreg{a} \mid \GRNget{s}{b} \in \levels{b}{a} \}$$
+$$\GRNres{a}{s} \DEF \{ b \in \GRNreg{a} \mid \GRNget{s}{b} \in \levels{b}{a} \}$$
 %We also denote: $\GRNallres{a} = \{(A;B) \mid \exists s \in \textstyle\prod_{a \in \Gamma} [0;l_a], \GRNres{a}{s} = A,B\}$
@@
 as stated by the definition of the transition relation $\GRNtrans{s}{s'}$.
+%
+%\new{Use a transition relation instead of an indeterministic function.}
 
-\new{Use a transition relation instead of an indeterministic function.}
-
-\begin{definition}[Asynchronous dynamics]\label{def:dynamics}
+\begin{definition}[Asynchronous dynamics ($\GRNtrans{}{}$)]\label{def:dynamics}
 The dynamics of a BRN using Thomas' parameters is given by the transition relation $\GRNtrans{}{} \in \GRNstates \times \GRNstates$ defined by:
@@
 
-While the use of intervals as parameter values does not add expressivity in boolean
-networks, it allows to specify a larger range of dynamics in the general case (w.r.t. the above
-definitions).
+While the use of intervals as parameter values does not add expressivity in boolean networks,
+it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t. the above definitions).
 Indeed, assume that $K_{a,\omega} = [i ; i+2]$;
-we aim at obtaining three different parameters $K_{a,\omega_1} = i$,  $K_{a,\omega_2} = i+1$, $K_{a,\omega_3} = i+2$.
-The only possible modification in resources is to add $a$ as a self-regulator.
-However, because resources have a boolean definition,
+we aim at obtaining the same dynamics with three different integer parameters instead: $K_{a,\omega_1} = i$,  $K_{a,\omega_2} = i+1$, $K_{a,\omega_3} = i+2$.
+The only possible modification in the resources of $a$ is to add a self-regulation $\GRNedge{a}{s}{t}{a}$.
+However, because resources have a boolean definition
+(each component can be only below or above the threshold related to its regulation on $a$),
 %(a component is either an activator or an inhibitor of $a$),
-it is not possible to differentiate the 3 values.
-We also remark that the use of intervals makes optional some explicit auto-activations in the IG
+it is not possible to differentiate the 3 sets of resources $\omega_1$, $\omega_2$ and $\omega_3$
+by relying only on the regulation of $a$ on itself.
+Finally, with the same reasoning, we also remark that the use of intervals makes optional some explicit auto-activations in the IG
 (as for $b$ in \pref{fig:runningBRN}, for instance).
-\rewrite{Detail the reason of using interval parameters}
 

### 2013-04-23 parts/grn.tex
formatting and remove comments

@@
 The uniqueness of these regulations makes the following sections simpler.
-%Therefore, for any component $b$, a predecessor $a$ of $b$ such that we have $a \xrightarrow{t} b$
-%can be either an activator or an inhibitor of $a$, according to the sign of the interaction and if the expression level of $b$ if above or below the threshold $t$.
-%We call $\levelsA{a}{b}$ (resp. $\levelsI{a}{b}$) the levels of $a$ where it is an
-%activator (resp. inhibitor) of $b$ (\pref{def:levels});
 We call $\levels{a}{b}$ the levels of $a$ above the threshold of $\GRNedge{a}{s}{t}{b}$;
 $l_a$ denotes the maximum level of $a$.
-%For the sake of simplicity and to establish a parallel with PH, if $a$ represents a component, we call $a_i$ its $i^\text{th}$ expression level.
+\rewrite{Discuss unsigned edges}
 \rewrite{Rewrite + reference + add some glue}
-%\new{Added unsigned edges ($\uns$) that stand for an unknown king of regulation.
-%The definition of $\levels{a}{b}$ now gives the levels for which $a$ regulates $b$ (i.e. $a > t$), independently from the regulation sign.}
 
@@
 \label{def:ig}
-%An \emph{Interaction Graph} (IG) is a quadruplet $(\Gamma, E_+, E_-, E_\uns)$ where $\Gamma$ is a finite number of \emph{components},
-%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
-%is the set of positive (resp. negative) \emph{regulations} between two nodes, labeled with a \emph{threshold}.
-An \emph{Interaction Graph} (IG) is a couple $(\Gamma, E)$ where $\Gamma$ is the finite set of \emph{components},
-and $E$
-%and $E_+$ (resp. $E_-$) $\subset \{a \xrightarrow{t} b \mid a, b \in \Gamma \wedge t \in [1; l_a]\}$
-is the finite set of 
-%positive (resp. negative)
-\emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}:
-$$E \DEF \{ \GRNedge{a}{s}{t}{b}, \ldots \mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\}$$
-A regulation from $a$ to $b$ is uniquely referenced:
+An \emph{Interaction Graph} (IG) is a couple $\IG = (\Gamma, E)$ with $\Gamma$ the finite set of \emph{components},
+and $E$ the finite set of \emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}:
+$$E \DEF \{ \GRNedge{a}{s}{t}{b}, \ldots \mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\} \enspace,$$
+where a regulation from $a$ to $b$ is uniquely referenced:
 $$\forall \GRNedge{a}{s}{t}{b} \in E, \forall \GRNedge{a}{s'}{t'}{b} \in E, s = s' \wedge t = t' \enspace.$$
-%if $a \xrightarrow{t} b \in E_+$ (resp. $E_-$),
-%$\nexists a \xrightarrow{t'} b \in E_+ \text{ (resp. $E_-$)}, t' \neq t$
-%and $\nexists a \xrightarrow{t'} b \in E_-\text{ (resp. $E_+$)}, t' \in \mathbb{N}$.
 \end{definition}
@@
 Given this definition, we denote as a shortcut:
-$E_s \DEF \{ \GRNedge{a}{s}{t}{b} \in E \}$
-for $s \in \{ +, -, \uns \}$.
-Furthermore, for all component $a \in \Gamma$, 
-we denote $\GRNreg{a}$ the set of its regulators as defined in \pref{eq:grn-regulators}.
-\[
+$E_s \DEF \{ \GRNedge{a}{s}{t}{b} \in E \}$ for $s \in \{ +, -, \uns \}$.
+Furthermore, for all component $a \in \Gamma$, we denote $\GRNreg{a}$ the set of its regulators as defined in \pref{eq:grn-regulators}.
+\begin{align}
 \label{eq:grn-regulators}
   \GRNreg{a} \DEF \{ b\in\Gamma\mid \exists \GRNedge{b}{s}{t}{a} \in E \}
-\]
+\end{align}
 The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$, \ie the number of components that $a$ regulates.
@@
 
-%$E_+(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in
-%E_+\}$ (resp. $E_-(a) = \{b \in \Gamma \mid \exists b \xrightarrow{t} a \in E_-\}$) denotes the set
-%of its positive (resp. negative) regulators; and
-%$\GRNreg{a} = E_+(a)\cup E_-(a)$ the set of its regulators.
-
 \begin{definition}[Effective levels ($\levelsl$)]\label{def:levels}
-%Let $(\Gamma,E)$ be an IG and $a, b \in \Gamma$ two of its components.
 Let $b \in \Gamma$ and $a \in \GRNreg{a}$;
 we define: $\levels{a}{b} \DEF [t; l_a]$ and $\ulevels{a}{b} \DEF [0; t-1]$.
-%%%
-%\begin{itemize}
-%  \item if $a \xrightarrow{t} b \in E_+$, $\levelsA{a}{b} \DEF [t; l_a]$ and
-%    $\levelsI{a}{b} \DEF [0; t-1]$;
-%  \item if $a \xrightarrow{t} b \in E_-$, $\levelsA{a}{b} \DEF [0; t-1]$ and
-%    $\levelsI{a}{b} \DEF [t; l_a]$;
-%  \item otherwise, $\levelsA{a}{b} \DEF \levelsI{a}{b} \DEF \emptyset$.
-%%%
-%  \item if $\GRNedge{a}{s}{t}{b} \in E$, $\levels{a}{b} \DEF [t; l_a]$;
-%  \item otherwise, $\levels{a}{b} \DEF \emptyset$.
-%\end{itemize}
 \end{definition}
@@
 \end{align*}
-%This IG can represent the same behavior as the PH given in \pref{fig:runningPH-2}.
 \end{example*}
@@
 %  node[elabel, below=-1em of b] {$0..1$}
-%  node[elabel, below=-1em of c] {$0..1$};
+%  node[elabel, above=-.8em of c] {$0..1$};
 \path[->]
@@
 \begin{align*}
-K_{a,\{b,c\}} &= [2 ; 2] & K_{b,\emptyset} &= [0 ; 1] \\
+K_{a,\emptyset} &= [0 ; 0] & K_{b,\emptyset} &= [0 ; 1] \\
 K_{a,\{b\}} &= [1 ; 1] & K_{b,\{a\}} &= [0 ; 0] \\
 K_{a,\{c\}} &= [1 ; 1] &&\\
-K_{a,\emptyset} &= [0 ; 0] & K_{c,\emptyset} &= [0 ; 1]
+K_{a,\{b,c\}} &= [2 ; 2] & K_{c,\emptyset} &= [0 ; 1]
 \end{align*}
@@
 IG example.
-%Components are represented by nodes labeled with a name and possible expression levels.
-Regulations are represented by the edges labeled with their sign and threshold.
+Components are represented by nodes labeled with a name
+and regulations by edges labeled with their sign and threshold.
 For instance, the edge from $b$ to $a$ is labeled $+1$, which stands for: $\GRNedgef{b}{+}{1}{a}$.
-%and means that if the level of expression of $b$ is equal to (i.e. above) 1, then $b$ activates $a$,
-%otherwise, $b$ inhibits $a$.
+This means that if the level of expression of $b$ is equal to (\ie above) 1, then $b$ activates $a$;
+otherwise, $b$ inhibits $a$.
 (right)
-Example parametrization of the left IG.
+Example parametrization on the left IG.
 }
@@
 referred to as the \emph{resources} of $a$ in this state (\pref{def:resources}).
-%
-%\new{No more $A,B$ but sets $\omega$ of resources instead. Definitions of $K$ were adapted.}
 
@@
 For a given component $a \in \Gamma$ and a state $s \in \GRNstates$,
-the set of regulators of $a$ whose level in $s$ is above the required threshold to regulate $a$
+the set of regulators of $a$ whose level in $s$ is above the related threshold %to regulate $a$
 is called the set of \emph{resources} of $a$ in $s$ and is noted $\GRNres{a}{s}$:
-%\begin{align*}
-%  A &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsA{b}{a}\} \\
-%  B &= \{b \in \Gamma \mid \GRNget{s}{b} \in \levelsI{b}{a}\}
-%\end{align*}
 $$\GRNres{a}{s} \DEF \{ b \in \GRNreg{a} \mid \GRNget{s}{b} \in \levels{b}{a} \}$$
-%We also denote: $\GRNallres{a} = \{(A;B) \mid \exists s \in \textstyle\prod_{a \in \Gamma} [0;l_a], \GRNres{a}{s} = A,B\}$
-%We also denote: $\GRNallres{a} = \{ \omega \subset \GRNreg{a} \}$
 \end{definition}
 
-\begin{definition}[Discrete parameter $K_{a,\omega}$ and Parametrization $K$]\label{def:param}
-For a given component $a \in \Gamma$ and $\omega \subset \GRNreg{a}$ a set of regulators of $a$,
-the discrete \emph{parameter} $K_{a,\omega} = [i; j]$ is a non-empty interval towards which $a$ will tend
+\begin{definition}[Parameter $K_{a,\omega}$ and Parametrization $K$]\label{def:param}
+For a given component $a \in \Gamma$, and $\omega \subset \GRNreg{a}$ a set of regulators of $a$,
+the \emph{parameter} $K_{a,\omega} = [i; j]$, with $0 \leq i \leq j \leq l_a$, is a non-empty interval towards which $a$ will tend
 in the states where its resources are exactly the regulators in $\omega$.
-The complete map $K$ of discrete parameters for $\IG$ is called a \emph{parametrization} of $\IG$.
-%For a given component $a \in \Gamma$ and $A$ (resp. $B$) $\subset \GRNreg{a}$ a set of its activators (resp. inhibitors) such that $A \cup B = \GRNreg{a}$ and $A \cap B = \emptyset$,
-%the discrete \emph{parameter} $K_{a,A,B} = [i; j]$ is a non-empty interval towards which $a$ will tend
-%in the states where its activators (resp. inhibitors) are the regulators in set $A$ (resp. $B$).
-%The complete map $K$ of discrete parameters for $\IG$ is called a \emph{parametrization} of $\IG$.
+The complete map $K$ of parameters on $\IG$ is called a \emph{parametrization} on $\IG$.
 \end{definition}
-%A consequence of this definition is that $0 \leq i_1 \leq i_2 \leq l_a$.
-%We also denote: $j < K_{a,A,B} \Leftrightarrow j < i_1$ and $j > K_{a,A,B} \Leftrightarrow j> i_2$.
 
-%\begin{example*}
-%\pref{fig:runningBRN}(right) gives a Parametrization of the IG of \pref{fig:runningBRN}(left).
-%\end{example*}
+\begin{example*}
+\pref{fig:runningBRN}(right) gives a Parametrization on the IG of \pref{fig:runningBRN}(left).
+\end{example*}
 
@@
 as stated by the definition of the transition relation $\GRNtrans{s}{s'}$.
-%
-%\new{Use a transition relation instead of an indeterministic function.}
 
@@
   \forall s, s' \in \GRNstates, \GRNtrans{s}{s'} &\Longleftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
-    & \qquad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
+    & \qquad\quad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
 \end{align*}
@@
   \end{cases}$
-\begin{comment}
-Let $s$ be a state of a BRN using Thomas' parameters $(\IG, K)$ where $\IG = (\Gamma, E_+, E_-)$.
-The state that succeeds to $s$ is given by the indeterministic function $f(s)$:
-\begin{align*}
-  f(s)  & = s' \Leftrightarrow \exists a \in \Gamma,
-    \GRNget{s'}{a} = f^a(s) \wedge
-    \forall b \in \Gamma, b \neq a, \GRNget{s}{b} = \GRNget{s'}{b}
-    \quad\text{, with}\\
-  f^a(s) & =
-  \begin{cases}
-    \GRNget{s}{a} + 1 & \text{if } \GRNget{s}{a} < K_{a,A,B} \\% \GRNres{a}{s}} \\
-    \GRNget{s}{a} & \text{if } \GRNget{s}{a} \in K_{a,A,B} \\ %\GRNres{a}{s}}\\
-    \GRNget{s}{a} - 1 & \text{if } \GRNget{s}{a} > K_{a,A,B} % \GRNres{a}{s}}
-  \end{cases}
-\quad\text{, where $A,B=\GRNres{a}{s}$.}
-\end{align*}
-\end{comment}
 \end{definition}
 
-While the use of intervals as parameter values does not add expressivity in boolean networks,
+While the use of intervals as parameter values instead of simple integers does not add expressivity in boolean networks,
 it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t. the above definitions).
@@
 (each component can be only below or above the threshold related to its regulation on $a$),
-%(a component is either an activator or an inhibitor of $a$),
 it is not possible to differentiate the 3 sets of resources $\omega_1$, $\omega_2$ and $\omega_3$

### 2013-04-23 parts/infer-ig.tex
formatting
more explicit \new labels
several adds and remove some labels
formatting and remove comments

@@
 allow at most one hitter per sort in the bounce-wise sequences (and thus are present in $s$).
-
-\begin{comment}
-\begin{property}[$\focals$ reachability]\label{pro:focals-reach}
-$\focals(a,S_a,\ctx)$ is reachable if and only if 
-$\forall s\in L(\ctx\cup S_a)$,
-if $\exists h\in \PHa(S_a,\ctx)$ with $\PHtarget(h)\in s$,
-there exists a (possibly empty) sequence of actions
-$h^1,\dots,h^n \in \PHa(\ctx\cup S_a)$
-	such that $h^1,\dots,h^n,h$ are successively playable in $s$;
-where $\PHa(\ctx) \DEF \{ \PHfrappe{b_i}{c_j}{c_k}\in\PHa \mid c\neq a \wedge
-			b\in \PHsort(\ctx) \Rightarrow b_i\in \ctx \wedge c\in\PHsort(\ctx) \Rightarrow
-			c_i\in\ctx \}$.
-\end{property}
-\end{comment}
 
@@
 \subsection{Interaction Inference}\label{ssec:infer-IG}
-
-%\todo{Try to use projections in order to perform the IG inference}
 
@@
   \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
-E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_-
-  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\} \\
+E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_- \\
+  & \qquad\qquad\qquad \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\}
 \end{align*}

### 2013-04-23 parts/infer-param.tex
several adds and remove some labels
formatting and remove comments

@@
 
-\todo{Try to address the inference of parameters despite unsigned edges}
-
 This subsection presents some results related to the inference of independent discrete parameters from a given PH.
@@
 
-\new{Changed $\levelsA{a}{b}$ and $\levelsI{a}{b}$ into $\levels{a}{b}$ in \pref{eq:wf-levels} and \pref{eq:param_context}.}
-
 \begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
@@
 verifies the following property:
-\begin{align}
+\begin{align*}
   \begin{split}
-\forall b\in \GRNreg{a} &,
+  \forall a \in \Gamma, \forall b\in \GRNreg{a}&,
           \forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
-& \qquad \forall c \in \Gamma, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
-& \qquad \qquad
+  & \quad \forall c \in \Gamma, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
+  & \qquad
                           \PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
@@
   \end{split}
-\label{eq:wf-levels}
-\end{align}
+\end{align*}
 \end{property}
@@
 Let $K_{a,\omega}$ be the parameter we want to infer for a given component $a \in \Gamma$
-%and $A,B \in \GRNallres{a}$ a configuration of resources of $a$ (activators and inhibitors).
 and $\omega \subset \GRNreg{a}$ a set of its regulators.
@@
 $a$.
-%If $\focals(a,C_{a,A,B})$ is a non-empty interval, then $K_{a,A,B} = \focals(a, C_{a,A,B})$.
-If $\focals(a,C^a_{a,\omega},C_{a,\omega})=[a_i;a_j]$ is a non-empty interval, 
-	then $K_{a,\omega} = [i;j]$.
+If $\focals(a,C^a_{a,\omega},C_{a,\omega})=[a_i;a_j]$ is a non-empty interval, then $K_{a,\omega} = [i;j]$.
 \end{proposition}
@@
 
-\new{Update all assumptions.}
-
 \begin{property}[Parameter validity]\label{pro:K-valid}
-A parameter $K_{a,\omega}$ is valid w.r.t. the PH iff the following equation is verified:
+A parameter $K_{a,\omega}$ is valid w.r.t. the PH if and only if the following equation is verified:
 \begin{align*}
-\forall a_i\in C^a_{a,\omega},
-		a_i \notin K_{a,\omega} \Longrightarrow (
-  \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, & c_k \in C^c_{a,\omega} \\
- \wedge a_i < K_{a,\omega} \Rightarrow j > i 
- & \wedge  a_i > K_{a,\omega} \Rightarrow j <i )
+  \forall a_i\in C^a_{a,\omega}, a_i \notin K_{a,\omega} \Longrightarrow
+    (& \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, c_k \in C^c_{a,\omega} \\
+     & \wedge a_i < K_{a,\omega} \Rightarrow j > i \wedge  a_i > K_{a,\omega} \Rightarrow j <i )
 \end{align*}
@@
 \begin{property}[Extreme values assumption]
-Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} iff:
+Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} if and only if:
 \label{prop:param_enum_extreme}
-\[
+\begin{align*}
   \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Longrightarrow \exists \omega \subset \GRNreg{b}, 0 \in K_{b,\omega} \wedge \exists \omega' \subset \GRNreg{b}, l_b \in K_{b,\omega'}
-\]
+\end{align*}
 \end{property}
@@
 \label{prop:param_enum_activity}
-Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} iff:
+Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} if and only if:
 \begin{align*}
-%  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A,B} <_{[]} K_{b,A \cup \{b\},B \setminus \{b\}}
-%\\
-%  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists (A;B) \in \GRNallres{a}, K_{b,A \setminus \{b\},B \cup \{b\}} <_{[]} K_{b,A,B}
   \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists \omega \subset \GRNreg{b}, K_{b,\omega} \neq K_{b,\omega \cup \{ a \}}
@@
 \label{prop:param_enum_monotonicity}
-Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} iff:
-%\[
-%  \forall b \in \Gamma, \forall (A;B), (A';B') \in \GRNallres{b},
-%  A \subset A' \wedge B' \subset B \Rightarrow K_{b,A,B} \leq_{[]} K_{b,A',B'}
-%\]
-$\forall b \in \Gamma$,
-$\forall X^+ \subset \{ a \in \Gamma \mid \GRNedge{a}{+}{t}{b} \in E_+ \}$,
-$\forall X^- \subset \{ a \in \Gamma \mid \GRNedge{a}{-}{t}{b} \in E_- \}$,
-%A \subset A' \wedge B' \subset B \Rightarrow
-$K_{b,\omega \cup X^-} \leq_{[]} K_{b,\omega \cup X^+}$
+Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} if and only if:
+\begin{align*}
+  \forall b \in \Gamma,
+  \forall A^+ \subset \{ a \in \Gamma \mid \GRNedge{a}{+}{t}{b} \in E_+ \}&,
+  \forall A^- \subset \{ a \in \Gamma \mid \GRNedge{a}{-}{t}{b} \in E_- \},\\
+  K_{b,\omega \cup A^-} & \leq_{[]} K_{b,\omega \cup A^+}
+\end{align*}
 \end{property}
-
-\begin{comment}
-\begin{definition}[Admissible parametrization \& Admissible parametrization with respect to inferred parameters]
-\label{def:param_enum_inf}
-Let $\PH = (\PHs, \PHl, \PHh)$ be a PH so that IG inference is possible, and $\IG = (\Gamma, E_+,
-E_-)$ the inferred IG.
-A parametrization $K$ on $\IG$ is said to be \emph{admissible} iff it respects
-the extreme values assumption, the activity assumption and the monotonicity assumption.
-A parametrization $K$ on $\IG$ is said to be \emph{admissible with respect to the
-inferred parameters} iff it is admissible and that all parameters that can be inferred regarding
-\pref{pps:param_K} are equal to their inferred value.
-\end{definition}
-
-\todo{utilité de “Admissible parametrization” seul ?}
-\end{comment}
 

### 2013-04-23 parts/grn.tex
notations

@@
 and $E$ the finite set of \emph{regulations} between two nodes, labeled with a \emph{sign} and a \emph{threshold}:
-$$E \DEF \{ \GRNedge{a}{s}{t}{b}, \ldots \mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in [1; l_a]\} \enspace,$$
+$$E \DEF \{ \GRNedge{a}{s}{t}{b}, \ldots \mid a, b \in \Gamma \wedge s \in \{ +, -, \uns \} \wedge t \in \segm{1}{l_a}\} \enspace,$$
 where a regulation from $a$ to $b$ is uniquely referenced:
@@
 The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$, \ie the number of components that $a$ regulates.
-Indeed, each discrete expression level in $[0 ; l_a]$ stands for an arbitrary set of concentration levels for which $a$ has a certain influence on several other components it regulates.
+Indeed, each discrete expression level in $\segm{0}{l_a}$ stands for an arbitrary set of concentration levels for which $a$ has a certain influence on several other components it regulates.
 Allowing more expression levels makes several levels useless as they represent the same influence of $a$ than some other levels;
@@
 Let $b \in \Gamma$ and $a \in \GRNreg{a}$;
-we define: $\levels{a}{b} \DEF [t; l_a]$ and $\ulevels{a}{b} \DEF [0; t-1]$.
+we define: $\levels{a}{b} \DEF \segm{t}{l_a}$ and $\ulevels{a}{b} \DEF \segm{0}{t-1}$.
 \end{definition}
@@
 \begin{align*}
-K_{a,\emptyset} &= [0 ; 0] & K_{b,\emptyset} &= [0 ; 1] \\
-K_{a,\{b\}} &= [1 ; 1] & K_{b,\{a\}} &= [0 ; 0] \\
-K_{a,\{c\}} &= [1 ; 1] &&\\
-K_{a,\{b,c\}} &= [2 ; 2] & K_{c,\emptyset} &= [0 ; 1]
+K_{a,\emptyset} &= \segm{0}{0} & K_{b,\emptyset} &= \segm{0}{1} \\
+K_{a,\{b\}} &= \segm{1}{1} & K_{b,\{a\}} &= \segm{0}{0} \\
+K_{a,\{c\}} &= \segm{1}{1} &&\\
+K_{a,\{b,c\}} &= \segm{2}{2} & K_{c,\emptyset} &= \segm{0}{1}
 \end{align*}
@@
 
-A \emph{state} $s$ of an IG $(\Gamma, E)$ is an element in $\GRNstates \DEF \prod_{a \in \Gamma} [0;l_a]$.
+A \emph{state} $s$ of an IG $(\Gamma, E)$ is an element in $\GRNstates \DEF \prod_{a \in \Gamma} \segm{0}{l_a}$.
 $\GRNget{s}{a}$ refers to the level of component $a$ in $s$.
@@
 For a given component $a \in \Gamma$, and $\omega \subset \GRNreg{a}$ a set of regulators of $a$,
-the \emph{parameter} $K_{a,\omega} = [i; j]$, with $0 \leq i \leq j \leq l_a$, is a non-empty interval towards which $a$ will tend
+the \emph{parameter} $K_{a,\omega} = \segm{i}{j}$, with $0 \leq i \leq j \leq l_a$, is a non-empty interval towards which $a$ will tend
 in the states where its resources are exactly the regulators in $\omega$.
@@
 it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t. the above definitions).
-Indeed, assume that $K_{a,\omega} = [i ; i+2]$;
+Indeed, assume that $K_{a,\omega} = \segm{i}{i+2}$;
 we aim at obtaining the same dynamics with three different integer parameters instead: $K_{a,\omega_1} = i$,  $K_{a,\omega_2} = i+1$, $K_{a,\omega_3} = i+2$.
@@
 where $a_i$ denotes the component $a$ at level $i$.
-As $K_{b,\emptyset} = [0 ; 1]$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
+As $K_{b,\emptyset} = \segm{0}{1}$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
 \end{example*}

### 2013-04-23 parts/infer-ig.tex
notations

@@
 We say a sequence of actions $h^1,\dots,h^n$ is \emph{bounce-wise} if and only if
-$\forall m\in[1;n-1], \PHbounce(h^m)=\PHtarget(h^{m+1})$.
+$\forall m\in\segm{1}{n-1}, \PHbounce(h^m)=\PHtarget(h^{m+1})$.
 From \pref{def:focals}, it derives that:
@@
 \forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
-(c^k)_{k \in [0;n]} \in \PHs^{n+1}, \\
+(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
                                    & \quad \quad c^0 = b \wedge c^n = a \\
-                                   & \quad \quad \wedge \forall k \in \llbracket 0 ; n-1 \rrbracket,
+                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
 								   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
@@
 When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
-the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval}.
+the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
+where $\uplus$ stands for a union of states (which are sets of processes).
 Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$

### 2013-04-23 parts/infer-param.tex
notations

@@
 $a$.
-If $\focals(a,C^a_{a,\omega},C_{a,\omega})=[a_i;a_j]$ is a non-empty interval, then $K_{a,\omega} = [i;j]$.
+If $\focals(a,C^a_{a,\omega},C_{a,\omega})=\segm{a_i}{a_j}$ is a non-empty interval, then $K_{a,\omega} = \segm{i}{j}$.
 \end{proposition}
@@
 Applied to the PH in \pref{fig:runningPH-1}, we obtain, for instance, 
-$K_{b,\emptyset} = [0 ; 1]$ and
-$K_{a,\{b,c\}} = [2 ; 2]$,
+$K_{b,\emptyset} = \segm{0}{1}$ and
+$K_{a,\{b,c\}} = \segm{2}{2}$,
 while $K_{a,\{b\}}$ can not be inferred.
-For the PH in \pref{fig:runningPH-2}, this latter is evaluated to $[1;1]$.
+For the PH in \pref{fig:runningPH-2}, this latter is evaluated to $\segm{1}{1}$.
 \end{example*}
@@
   \forall A^- \subset \{ a \in \Gamma \mid \GRNedge{a}{-}{t}{b} \in E_- \},\\
-  K_{b,\omega \cup A^-} & \leq_{[]} K_{b,\omega \cup A^+}
+  K_{b,\omega \cup A^-} & \leqsegm K_{b,\omega \cup A^+}
 \end{align*}
@@
   \item the \aspil{less\_active(\ti{a}, \ti{p}, \ti{q})} atom means that $K^\ti{p}_{\ti{a},A,B}$ stands for a configuration with less activating regulators than $K^\ti{q}_{\ti{a},A',B'}$ (\ie $A \subset A'$),
-  \item the \aspil{param\_inf(\ti{a}, \ti{p}, \ti{q})} atom means: $K^\ti{p}_{\ti{a},A,B} \leq_{[]} K^\ti{q}_{\ti{a},A',B'}$,
+  \item the \aspil{param\_inf(\ti{a}, \ti{p}, \ti{q})} atom means: $K^\ti{p}_{\ti{a},A,B} \leqsegm K^\ti{q}_{\ti{a},A',B'}$,
 \end{itemize}
@@
 \asp{:- less\_active(X, P, Q), not param\_inf(X, P, Q).}
-which removes all parametrization results where parameters $K^\aspil{P}_{\aspil{X},A,B}$ and $K^\aspil{Q}_{\aspil{X},A',B'}$ exist such that $A \subset A'$ and $K^\aspil{Q}_{\aspil{X},A',B'} <_{[]} K^\aspil{P}_{\aspil{X},A,B}$, thus violating the monotonicity assumption.
+which removes all parametrization results where parameters $K^\aspil{P}_{\aspil{X},A,B}$ and $K^\aspil{Q}_{\aspil{X},A',B'}$ exist such that $A \subset A'$ and $K^\aspil{Q}_{\aspil{X},A',B'} \ltsegm K^\aspil{P}_{\aspil{X},A,B}$, thus violating the monotonicity assumption.
 Of course, other assumptions can be formulated in the same way.

### 2013-04-24 parts/grn.tex
discussions on interval parameters & unsigned edges

@@
 \cite{Richard06,BernotSemBRN}.
-In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t. the presented
+In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t.~the presented
 inference),
@@
 
-The IG is composed of nodes that represent components, and edges labeled with a threshold that stand
-for either positive or negative regulations (\pref{def:ig}).
-For such a regulation to take place, the expression level of its head component has to be higher than its threshold; otherwise, the opposite influence is expressed.
-The uniqueness of these regulations makes the following sections simpler.
-We call $\levels{a}{b}$ the levels of $a$ above the threshold of $\GRNedge{a}{s}{t}{b}$;
-$l_a$ denotes the maximum level of $a$.
-\rewrite{Discuss unsigned edges}
-\rewrite{Rewrite + reference + add some glue}
+The IG is composed of nodes that represent components,
+and edges labeled with a sign and a threshold that stand for regulations between these components (\pref{def:ig}).
+The activity, concentration rate or presence of each component in a given state of the system is modeled by a discrete value called \emph{expression level}.
+The maximum level of a component $a$ is denoted $l_a$.
+The sign of an edge denotes the kind of regulation it models: it can be positive ($+$), negative ($-$) or unsigned ($\uns$),
+the latter meaning that the nature (activation or inhibition) of the regulation is unknown.
+For a regulation to take place (activation or inhibition), the expression level of its head component has to be higher than its threshold;
+otherwise, the opposite influence is expressed.
+The uniqueness of each regulation is forced, in order to make the following sections simpler.
+The purpose of unsigned edges is discussed at the end of the current section.
 
@@
 $E_s \DEF \{ \GRNedge{a}{s}{t}{b} \in E \}$ for $s \in \{ +, -, \uns \}$.
-Furthermore, for all component $a \in \Gamma$, we denote $\GRNreg{a}$ the set of its regulators as defined in \pref{eq:grn-regulators}.
+Furthermore, for all component $b \in \Gamma$, we denote $\GRNreg{b}$ the set of its regulators as defined in \pref{eq:grn-regulators}.
 \begin{align}
 \label{eq:grn-regulators}
-  \GRNreg{a} \DEF \{ b\in\Gamma\mid \exists \GRNedge{b}{s}{t}{a} \in E \}
+  \GRNreg{b} \DEF \{ b\in\Gamma\mid \exists \GRNedge{a}{s}{t}{b} \in E \}
 \end{align}
-The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$, \ie the number of components that $a$ regulates.
-Indeed, each discrete expression level in $\segm{0}{l_a}$ stands for an arbitrary set of concentration levels for which $a$ has a certain influence on several other components it regulates.
-Allowing more expression levels makes several levels useless as they represent the same influence of $a$ than some other levels;
-on the contrary, reducing the maximum level requires that several thresholds on outgoing edges of $a$ share the same value, which is biologically unlikely.
-However, these considerations are usually taken into account when one has to directly build a model in this framework, but are not mandatory.
-In the scope of this paper, we allow any number of levels for components, without considering the number of outgoing edges,
-as the number of processes in a PH sort is not constrained in any way.
+Then, for all component $a \in \GRNreg{b}$ regulating $b$,
+we denote $\levels{a}{b}$ (resp.~$\ulevels{a}{b}$) the levels of $a$ above (resp.~below) the threshold $t$ of the (unique) edge $\GRNedge{a}{s}{t}{b}$ (\pref{def:levels}).
+If the expression level of $a$ belongs to the interval $\levels{a}{b}$, then $a$ is expected to have the influence corresponding to the sign $s$ on $b$;
+if its expression level belongs to $\ulevels{a}{b}$, the opposite influence is expected.
 
 \begin{definition}[Effective levels ($\levelsl$)]\label{def:levels}
-Let $b \in \Gamma$ and $a \in \GRNreg{a}$;
-we define: $\levels{a}{b} \DEF \segm{t}{l_a}$ and $\ulevels{a}{b} \DEF \segm{0}{t-1}$.
+Let $b \in \Gamma$ and $a \in \GRNreg{a}$; we define:
+$$\levels{a}{b} \DEF \segm{t}{l_a} \quad \text{and} \quad \ulevels{a}{b} \DEF \segm{0}{t-1} \enspace.$$
 \end{definition}
@@
   \GRNreg{c} &= \emptyset
+\end{align*}
+We also have especially:
+\begin{align*}
+  \levels{a}{b} &= \segm{2}{2} & \ulevels{a}{b} &= \segm{0}{1}
 \end{align*}
@@
 For instance, the edge from $b$ to $a$ is labeled $+1$, which stands for: $\GRNedgef{b}{+}{1}{a}$.
-This means that if the level of expression of $b$ is equal to (\ie above) 1, then $b$ activates $a$;
+This means that if the expression level of $b$ is equal to (\ie above) 1, then $b$ activates $a$;
 otherwise, $b$ inhibits $a$.
@@
 \end{definition}
+The purpose of interval parameters is discussed at the end of this section.
 
@@
 
-While the use of intervals as parameter values instead of simple integers does not add expressivity in boolean networks,
+\begin{example*}
+In the BRN that consists of the IG and parametrization of \pref{fig:runningBRN}, the following
+transitions are possible given the semantics defined in \pref{def:dynamics}:
+$$\GRNstate{a_0, b_1, c_1} \rightarrow \GRNstate{a_1, b_1, c_1} \rightarrow \GRNstate{a_2, b_1, c_1} \rightarrow
+\GRNstate{a_2, b_0, c_1} \rightarrow \GRNstate{a_1, b_0, c_1} \enspace,$$
+where $a_i$ denotes the component $a$ at level $i$.
+This sequence of states ends in a steady state: no evolution is possible in $\GRNstate{a_1, b_0, c_1}$.
+Due to the semantics of interval parameters and as $K_{b,\emptyset} = \segm{0}{1}$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
+\end{example*}
+
+\paragraph{Discussion on maximum level restrictions}
+The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$ in the IG, \ie the number of components that $a$ regulates.
+Indeed, each discrete expression level in $\segm{0}{l_a}$ stands for an arbitrary set of concentration levels for which $a$ has a certain influence on several other components it regulates.
+Allowing more expression levels makes several levels useless as they represent the same influence of $a$ than some other levels;
+on the contrary, reducing the maximum level requires that several thresholds on outgoing edges of $a$ share the same value, which is biologically unlikely.
+However, these considerations are usually taken into account when one has to directly build a model in this framework, but are not mandatory.
+In the scope of this paper, we allow any number of levels for components, without considering the number of outgoing edges,
+as the number of processes in a PH sort is not constrained in any way.
+
+\paragraph{Discussion on unsigned edges}
+Unsigned edges allow to model regulations whom trend (activation or inhibition) is unknown.
+It has to be noted that using unsigned edges in addition to the regular positive and negative ones does not add expressivity to the framework.
+Indeed, signs have no impact on the parametrization (\pref{def:param}) or the dynamics (\pref{def:dynamics}).
+They do have an impact later in this paper, when admissible parametrizations are enumerated (\pref{ssec:admissible-K}).
+However, signs in an IG are useful to understand the trend of a regulation.
+PH actions are very atomistic and the inference of the underlying BRN proposed in \pref{sec:infer-IG}
+allows to find the global trends of all regulations between components,
+which may turn out to be difficult to perform by hand.
+As signs have no direct influence on the semantics, unsigned edges car prove convenient to model an IG with partial knowledge on some regulations.
+
+\paragraph{Discussion on interval parameters}
+While the use of intervals as parameter values ($K_{a,\omega} = \segm{i}{j}$) instead of simple integers ($K_{a,\omega} = i$) does not add expressivity in boolean networks,
 it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t. the above definitions).
@@
 (as for $b$ in \pref{fig:runningBRN}, for instance).
-
-\begin{example*}
-In the BRN that consists of the IG and parametrization of \pref{fig:runningBRN}, the following
-transitions are possible given the semantics defined in \pref{def:dynamics}:
-$\GRNstate{a_0, b_1, c_1} \rightarrow \GRNstate{a_1, b_1, c_1} \rightarrow \GRNstate{a_2, b_1, c_1} \rightarrow
-\GRNstate{a_2, b_0, c_1} \rightarrow \GRNstate{a_1, b_0, c_1}$,
-ending in a steady state,
-where $a_i$ denotes the component $a$ at level $i$.
-As $K_{b,\emptyset} = \segm{0}{1}$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
-\end{example*}

### 2013-04-24 parts/ph.tex
several adds and remove some labels
discussions on interval parameters & unsigned edges

@@
 	$a\neq b\Rightarrow \forall(a_i,b_j)\in\PHl_a\times\PHl_b,a_i\neq b_j$;
-\item $\PHa \DEF \{ \PHfrappe{a_i}{b_j}{b_k}, \dots \mid
+\item $\PHa \DEF \{ \PHfrappe{a_i}{b_j}{b_k}, \ldots \mid
 					(a,b)\in\PHs^2 \wedge (a_i,b_j,b_k)\in \PHl_a\times\PHl_b\times\PHl_b$ \\
@@
 part ($a_1$ to $a_2$) in dotted line.
-Actions involving $b_1$ or $c_1$ are in thick lines.
+Here, actions involving $b_1$ or $c_1$ are in thick lines.
 %The current state is represented by the grayed processes:
@@
 \paragraph{Modeling cooperation}
-\rewrite{Detail the modeling of cooperations}
+\rewrite{Detail the modeling of cooperations?}
 As described in \cite{PMR10-TCSB}, the cooperation between several processes to make another process bounce can be
@@
 
-\rewrite{Explain the efficiency of PH for reachability matters}
+\rewrite{Explain the efficiency of PH for reachability?}

### 2013-04-25 parts/intro.tex
notations
adds after Morgan's review

@@
 
+\todo{Emphasis on the new contribution since CMSB.}
+
 \paragraph{Outline}
@@
 \paragraph{Notations}
-$[i;j]$ is the set of integers $\{ i, i+1, \dots, j \}$;
-we note $[i_1;j_1] \leq_{[]} [i_2;j_2] \EQDEF (i_1\leq i_2\wedge j_1\leq j_2)$
-and $[i_1;j_1] <_{[]} [i_2;j_2] \EQDEF 
+$\segm{i}{j}$ is the set of integers $\{ i, i+1, \dots, j \}$.
+We note: $$\segm{i_1}{j_1} \leqsegm \segm{i_2}{j_2} \EQDEF (i_1\leq i_2\wedge j_1\leq j_2) \enspace,$$
+%and $$[i_1;j_1] \ltsegm [i_2;j_2] \EQDEF 
 %	[i_1;j_1] \leq_{[]} [i_2;j_2] \wedge (i_1\neq i_2 \vee j_1\neq j_2)$.
-(i_1<i_2 \wedge j_1\leq j_2) \vee (i_1\leq i_2\wedge j_1 < j_2)$.
-Given an integer $k$, $k<[i;j] \EQDEF k <i$ and $k>[i;j]\EQDEF k > j$.
+%(i_1<i_2 \wedge j_1\leq j_2) \vee (i_1\leq i_2\wedge j_1 < j_2)$$.
+and, given an integer $k$: $k < \segm{i}{j} \EQDEF k < i$ and $k > \segm{i}{j} \EQDEF k > j$.

### 2013-04-25 main.tex
Add TODOs and REWRITEs
add unsigned edges in BRN defs
more explicit \new labels
affiliation of Morgan
additions after Morgan's review

@@
 % Commandes À FAIRE
-\usepackage{color} % Couleurs du texte
-\newcommand{\todo}[1]{\textcolor{darkgreen}{[#1]}}
+%\usepackage{color} % Couleurs du texte
+\newcommand{\rewrite}[1]{\textcolor{blue}{\textbf{[REWRITE: #1]}}}
+\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
+\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
+%\newcommand{\new}[1]{#1}
 
@@
 
-\title{Concretizing the Process Hitting into Biological Regulatory Networks}
+\title{\todo{New title} Concretizing the Process Hitting into Biological Regulatory Networks}
 
@@
 \ead{Maxime.Folschette@irccyn.ec-nantes.fr}
-\author[lix]{Lo\"ic Paulev\'e}
+\author[eth]{Lo\"ic Paulev\'e}
 \author[nii]{Katsumi Inoue}
-\author[irccyn]{Morgan Magni}
+\author[irccyn,nii]{Morgan Magnin}
 \author[irccyn]{Olivier Roux}
@@
 2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.}
-\address[lix]{LIX, \'Ecole Polytechnique, 91128 Palaiseau Cedex, France.}
+\address[eth]{ETH Zürich, Institut für Automatik,\\
+ETL I 28, Physikstrasse 3, 8092 Zurich, Switzerland.}
 
@@
 cooperations, where Thomas' approach fails because of the combinatorics of parameters.
+
+\todo{Emphasis on the new contribution since CMSB.}
 \end{abstract}

### 2013-04-25 parts/grn.tex
additions after Morgan's review

@@
 \begin{example*}
-\pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E)$ with
-$\Gamma = \{a, b, c\}$ and:
+\pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E)$ where
+$\Gamma = \{a, b, c\}$, with $l_a = 2$ and $l_b = l_c = 1$,
+and:
 \begin{align*}
@@
 \end{align*}
+\todo{Find an example with an unsigned edge?}
 \end{example*}
@@
 Example parametrization on the left IG.
+\todo{Find an example with an unsigned edge?}
 }
@@
 At last, \pref{def:dynamics} gives the asynchronous dynamics of a BRN using Thomas' parameters.
-From a given state $s$, a transition to another state $s'$ is possible provided that only one component $a$ will evolve of one level towards $K_{a,\GRNres{a}{s}}$,
+From a given state $s$, a transition to another state $s'$ is possible provided that exactly one component $a$ evolves of one level towards $K_{a,\GRNres{a}{s}}$,
 as stated by the definition of the transition relation $\GRNtrans{s}{s'}$.
+However, $a$ cannot evolve if its expression level already belongs to $K_{a,\GRNres{a}{s}}$.
 
 \begin{definition}[Asynchronous dynamics ($\GRNtrans{}{}$)]\label{def:dynamics}
-The dynamics of a BRN using Thomas' parameters is given by the transition relation $\GRNtrans{}{} \in \GRNstates \times \GRNstates$ defined by:
+The dynamics of a BRN using Thomas' parameters is given by the transition relation $\GRNtrans{}{}\ \in \GRNstates \times \GRNstates$ defined by:
 \begin{align*}
@@
 as the number of processes in a PH sort is not constrained in any way.
+For instance, in the example of \pref{fig:runningBRN}, it would be more relevant to have $l_a = 1$, as $a$ has only one outgoing edge;
+however, the value $l_a = 2$ makes the example more interesting.
 
@@
 which may turn out to be difficult to perform by hand.
-As signs have no direct influence on the semantics, unsigned edges car prove convenient to model an IG with partial knowledge on some regulations.
+As signs have no direct influence on the semantics, unsigned edges can be convenient to model an IG with partial knowledge on some regulations.
 

### 2013-04-25 parts/infer-param.tex
discussions on interval parameters & unsigned edges
start refactoring section 4.3

@@
 Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} if and only if:
-\label{prop:param_enum_extreme}
+\label{pro:param_enum_extreme}
 \begin{align*}
@@
 \begin{property}[Activity assumption]
-\label{prop:param_enum_activity}
+\label{pro:param_enum_activity}
 Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} if and only if:
@@
 \begin{property}[Monotonicity assumption]
-\label{prop:param_enum_monotonicity}
+\label{pro:param_enum_monotonicity}
 Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} if and only if:
@@
 
+\newcommand{\atom}{\mathbf}
+\newcommand{\la}{\leftarrow}
+\newcommand{\var}[1]{#1}
+\newcommand{\nota}{\neg}
+
 Answer Set Programming (ASP) \cite{Baral03} has been chosen to address the enumeration of all admissible parametrizations.
@@
   \item it is convenient to enumerate a large set of possible answers,
-  \item it allows us to easily constrain the answers according to some properties.
+  \item it allows to easily constrain the answers according to some properties.
 \end{itemize}
@@
 
-All information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
-For functional purposes, we assign a unique label to each couple $A,B$ of activators and inhibitors of a given component, and in the following we note $K^p_{a,A,B}$ the parameter of component $a$ whose regulators $A,B$ are assigned to the label $p$.
-Then, to state the existence of a parameter $K^\ti{p}_{\ti{a},A,B}$, we use an atom named \aspil{param\_label} in the following fact:
-\asp{param\_label(\ti{a}, \ti{p}).}
+ASP is a logic programming paradigm based on a set of rules.
+A rule takes the form of:
+$$head \la body.$$
+where $head$ is an atom and $body$ is a series of atoms or negations of atoms.
+Solving an ASP program means finding an answer set, \ie a minimal set of atoms that fulfills all the rules.
+Several answer sets can be a solution to the same ASP program and a solver can be directed to enumerate them all.
+
+A rule with no $body$ part is called a fact, and its $head$ atom has to be present in all answer sets.
+For instance, all information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
+For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
+and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
+Then, to state the existence of a parameter $K^p_{a,\omega}$, an atom named $\atom{param\_label}$ is used in the following fact:
+$$\atom{param\_label}(a, p).$$
 
 Defining a set in ASP is equivalent to defining the rule for belonging to this set.
-For example, we define an atom \aspil{param\_act} that describes the set of active regulators of a given a parameter.
+For example, we define an atom $param\_act$ that describes the set of active regulators of a given a parameter.
 %of component \ti{a} and label \ti{p} (\ie the set $A$ of a parameter $K^\ti{p}_{\ti{a},A,B}$).
-Describing the activators of $K^\ti{p}_{\ti{a},\{\ti{b},\ti{c}\},\{\ti{d}\}}$ gives:
-\asp{param\_act(\ti{a}, \ti{p}, \ti{b}).
-\item[] param\_act(\ti{a}, \ti{p}, \ti{c}).}
-The absence of such a fact involving \ti{d} with label \ti{p} indicates that \ti{d} is an inhibitor in the configuration of regulators related to this parameter.
+Describing the activators of $K^p_{a,\{b,c\}}$ gives:
+\begin{align*}
+  \atom{param\_act}(a, p, b).&&
+  \atom{param\_act}(a, p, c).
+\end{align*}
+The absence of such a fact involving another potential regulator $d$ of $a$ with label $p$ indicates that $d$ is not in the set of resources related to this parameter.
+\rewrite{Find another example.}
 
@@
 For instance, in order to define the set of expression levels of a component, we declare:
-\asp{component\_levels(X, 0..M) :- component(X, M).}
-where the \aspil{component(X, M)} atom stands for the existence of a component \aspil{X} with a maximum level \aspil{M}.
-Considering this declaration, any possible answer for the atom \aspil{component\_levels} will be found by binding all possible values of its two terms with all existing \aspil{component} facts: the existence of an answer \aspil{component\_levels(\ti{a}, \ti{k})} will depend on the existence of a term \ti{a}, which is bound with \aspil{X}, and an integer~\ti{k}, constrained by: $0 \leq \ti{k} \leq \aspil{M}$.
+$$\atom{component\_levels}(\var{X}, 0..\var{M}) \la \atom{component}(\var{X}, \var{M}).$$
+where the atom $\atom{component}(X, M)$ stands for the existence of a component $X$ with a maximum level $M$.
+Considering this declaration, any possible answer for the atom $\atom{component\_levels}$ will be found by binding all possible values of its two terms with all existing $\atom{component}$ facts:
+the existence of an answer $\atom{component\_levels}(a, k)$ will depend on the existence of a term $a$, which is bound with $\var{X}$, and an integer $k$, constrained by: $0 \leq k \leq \var{M}$.
 
@@
 For example,
-\asp{1 \{ param(X, P, I) : component\_levels(X, I) \} :-
-\item[] ~~~~~~param\_label(X, P), not infered\_param(X, P).}
-where \aspil{param(X, P, I)} stands for: $\aspil{I} \in K^\aspil{P}_{\aspil{X},A,B}$,
-means that any parameter of component \aspil{X} and label \aspil{P} must contain at least one level value (\aspil{I}) in the possible expression levels of \aspil{X}.
-Indeed, the lower bound is 1, forcing at least one element in the parameter, but no upper bound is specified, allowing up to any number of answers.
-The body (right-hand side) of the rule also checks for the existence of a parameter of \aspil{X} with label \aspil{P}, and constrains that the parametrization inference was not conclusive for the considered parameter (\aspil{not} stands for negation by failure: \aspil{not L} becomes true if \aspil{L} is not true).
-Such a constraint gives multiple results as any set of \aspil{param} atoms satisfying the cardinality will lead to a new global set of answers.
-In this way, we enumerate all possible parametrizations which respects the results of parameters
-inference, but completely disregarding the notion of admissible parametrizations given in
-\pref{ssec:admissible-K}.
+\begin{align*}
+  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
+  & \qquad\qquad \atom{param\_label}(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
+\end{align*}
+where $\atom{param}(\var{X}, \var{P}, \var{I})$ stands for: $\var{I} \in K^\var{P}_{\var{X},\omega}$,
+means that any parameter of component $\var{X}$ and label $\var{P}$ must contain at least one level value ($\var{I}$) in the possible expression levels of $\var{X}$.
+Indeed, the lower bound is $1$, forcing at least one element in the parameter, but no upper bound is specified, allowing up to any number of answers.
+The body (right-hand side) of the rule also checks for the existence of a parameter of $\var{X}$ with label $\var{P}$,
+and constrains that the parametrization inference was not conclusive for the considered parameter (“$\nota$” stands for negation by failure: $\nota \atom{L}$ becomes true if $\atom{L}$ is not true).
+Such a constraint gives multiple results as any set of $\atom{param}$ atoms satisfying the cardinality will lead to a new global set of answers.
+In this way, we enumerate all possible parametrizations which respects the results of parameters inference,
+but completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K}.
 
@@
 \begin{itemize}
-  \item the \aspil{less\_active(\ti{a}, \ti{p}, \ti{q})} atom means that $K^\ti{p}_{\ti{a},A,B}$ stands for a configuration with less activating regulators than $K^\ti{q}_{\ti{a},A',B'}$ (\ie $A \subset A'$),
-  \item the \aspil{param\_inf(\ti{a}, \ti{p}, \ti{q})} atom means: $K^\ti{p}_{\ti{a},A,B} \leqsegm K^\ti{q}_{\ti{a},A',B'}$,
+  \item the $\atom{less\_active}(a, p, q)$ atom means that $K^p_{a,\omega}$ stands for a configuration with less activating regulators than $K^q_{a,\omega'}$, % (\ie $A \subset A'$),
+  \item the $\atom{param\_inf}(a, p, q)$ atom means: $K^p_{a,\omega} \leqsegm K^q_{a,\omega'}$,
 \end{itemize}
-the monotonicity assumption is formulated as the following integrity constraint:
-\asp{:- less\_active(X, P, Q), not param\_inf(X, P, Q).}
-which removes all parametrization results where parameters $K^\aspil{P}_{\aspil{X},A,B}$ and $K^\aspil{Q}_{\aspil{X},A',B'}$ exist such that $A \subset A'$ and $K^\aspil{Q}_{\aspil{X},A',B'} \ltsegm K^\aspil{P}_{\aspil{X},A,B}$, thus violating the monotonicity assumption.
+then the monotonicity assumption (\pref{pro:param_enum_monotonicity}) is formulated as the following integrity constraint:
+$$\la \atom{less\_active}(\var{X}, \var{P}, \var{Q}), \nota \atom{param\_inf}(\var{X}, \var{P}, \var{Q}).$$
+which removes all parametrization results where parameters $K^\var{P}_{\var{X},\omega}$ and $K^\var{Q}_{\var{X},\omega'}$ exist such that $\var{X}$ is less activated by $\omega$ than $\omega'$ %$A \subset A'$
+but $K^\var{Q}_{\var{X},\omega'} \ltsegm K^\var{P}_{\var{X},\omega}$,
+thus violating the monotonicity assumption.
 Of course, other assumptions can be formulated in the same way.

### 2013-04-26 parts/grn.tex
continue refactoring section 4.3

@@
 
-\begin{example*}
+\begin{example}
 \pref{fig:runningBRN}(left) represents an Interaction Graph $(\Gamma,E)$ where
@@
 \todo{Find an example with an unsigned edge?}
-\end{example*}
+\end{example}
 
@@
 
-\begin{example*}
+\begin{example}
 \pref{fig:runningBRN}(right) gives a Parametrization on the IG of \pref{fig:runningBRN}(left).
-\end{example*}
+\end{example}
 
@@
 
-\begin{example*}
+\begin{example}
 In the BRN that consists of the IG and parametrization of \pref{fig:runningBRN}, the following
@@
 Due to the semantics of interval parameters and as $K_{b,\emptyset} = \segm{0}{1}$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
-\end{example*}
+\end{example}
 

### 2013-04-26 parts/infer-ig.tex
continue refactoring section 4.3

@@
 
-\begin{example*}
+\begin{example}
 In the PH of \pref{fig:runningPH-1}, we obtain:
@@
 \end{align*}
-\end{example*}
+\end{example}
 
@@
 
-\begin{example*}
+\begin{example}
 In the PH of \pref{fig:runningPH-2}, $bc$ is a well-formed cooperative sort as defined in \pref{pro:wf-cooperative-sort}, because:
@@
 with $\Gamma = \{a,b,c\}$.
-\end{example*}
+\end{example}
 
@@
 
-\begin{example*}
+\begin{example}
 The IG inference from the PH of \pref{fig:runningPH-2} gives
@@
 No self-influence are inferred ($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
-\end{example*}
+\end{example}

### 2013-04-26 parts/infer-param.tex
continue refactoring section 4.3

@@
 
-\begin{example*}
+\begin{example}
+\label{ex:infer-param-runningPH-1}
 Applied to the PH in \pref{fig:runningPH-1}, we obtain, for instance, 
@@
 $K_{a,\{b,c\}} = \segm{2}{2}$,
-while $K_{a,\{b\}}$ can not be inferred.
-For the PH in \pref{fig:runningPH-2}, this latter is evaluated to $\segm{1}{1}$.
-\end{example*}
+while $K_{a,\{b\}}$ and $K_{a,\{c\}}$ can not be inferred.
+\end{example}
+
+\begin{example}
+For the PH in \pref{fig:runningPH-2}, we obtain
+$K_{b,\emptyset} = \segm{0}{1}$,
+$K_{a,\{b,c\}} = \segm{2}{2}$,
+$K_{a,\{b\}} = \segm{1}{1}$,
+and $K_{a,\{c\}} = \segm{1}{1}$.
+For this PH, all parameters can be inferred, and the obtained parametrization is the one of \pref{fig:runningBRN}(right).
+\end{example}
 
@@
 
+\newcommand{\paramlabel}{\atom{param\_label}}
+\newcommand{\paramres}{\atom{param\_resource}}
+
 Answer Set Programming (ASP) \cite{Baral03} has been chosen to address the enumeration of all admissible parametrizations.
@@
 A rule takes the form of:
-$$head \la body.$$
-where $head$ is an atom and $body$ is a series of atoms or negations of atoms.
-Solving an ASP program means finding an answer set, \ie a minimal set of atoms that fulfills all the rules.
-Several answer sets can be a solution to the same ASP program and a solver can be directed to enumerate them all.
+$$\underbrace{{\ }\atom{H}_{\ }}_{head} \la \underbrace{\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m}_{body}.$$
+where the $head$ is an atom ($\atom{H}$) and the $body$ is a series of atoms ($\atom{A}_i$) and negations of atoms ($\nota \atom{B}_i$).
+Such a rule states that if all atoms $\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n$ are true
+and all atoms $\atom{B}_1, \atom{B}_2, \dots, \atom{B}_m$ are not true (negation by failure), then $\atom{H}$ has to be true.
+Solving an ASP program means finding an \emph{answer set}, which is a minimal set of true atoms that respects all the rules.
+Several answer sets can be solution to the same ASP program and the solver can be directed to enumerate them all.
+\todo{Distinguish atoms and predicates.}
 
-A rule with no $body$ part is called a fact, and its $head$ atom has to be present in all answer sets.
+A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
 For instance, all information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
+In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
 For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
 and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
-Then, to state the existence of a parameter $K^p_{a,\omega}$, an atom named $\atom{param\_label}$ is used in the following fact:
-$$\atom{param\_label}(a, p).$$
+In other words, $\paramlabel(a, p)$ states that a parameter $K^p_{a,\omega}$ exists.
+Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 7 parameters, and to state the existence of each of them, the following facts are used:
+\begin{align*}
+  &\paramlabel(a,1).
+  &\paramlabel(b,1).\\
+  &\paramlabel(a,2).
+  &\paramlabel(b,2).\\
+  &\paramlabel(a,3).\\
+  &\paramlabel(a,4).
+  &\paramlabel(c,1).
+\end{align*}
 
-Defining a set in ASP is equivalent to defining the rule for belonging to this set.
-For example, we define an atom $param\_act$ that describes the set of active regulators of a given a parameter.
-%of component \ti{a} and label \ti{p} (\ie the set $A$ of a parameter $K^\ti{p}_{\ti{a},A,B}$).
-Describing the activators of $K^p_{a,\{b,c\}}$ gives:
-\begin{align*}
-  \atom{param\_act}(a, p, b).&&
-  \atom{param\_act}(a, p, c).
-\end{align*}
-The absence of such a fact involving another potential regulator $d$ of $a$ with label $p$ indicates that $d$ is not in the set of resources related to this parameter.
-\rewrite{Find another example.}
+A set (such as sets of resources $\omega$) can be defined in ASP with a series of atoms. \rewrite{Find a more precise sentence.}
+For example, for a given parameter $K^p_{a,\omega}$, expressing $x \in \omega$ is done by the atom: $\paramres(a, p, x)$.
+Then, the sets of resources of \pref{ex:infer-param-runningPH-1} is defined by:
+$$\begin{array}{>{\hspace{.8em}}l|>{\phantom{\raisebox{.6em}{xx}}}l}
+  \hspace{-.8em}\text{Parameter} & \text{Corresponding ASP fact} \\
+\hline
+  K^0_{a,\emptyset} & \\
+  K^1_{a,\{b\}} & \paramres(a,1,b). \\
+  K^2_{a,\{c\}} & \paramres(a,2,c). \\
+  K^3_{a,\{b,c\}} & \paramres(a,3,b). \quad \paramres(a,3,c). \phantom{\raisebox{-.8em}{xx}} \\ \hline
+  K^0_{b,\emptyset} & \\
+  K^1_{b,\{a\}} & \paramres(b,1,a). \phantom{\raisebox{-1em}{xx}} \\ \hline
+  K^0_{c,\emptyset} &
+\end{array}$$
+The absence of an atom $\paramres(a,1,c)$, for example, states that $c$ is not a resource for the parameter $K^1_{a,\{b\}}$ (indeed, $c \notin \{b\}$).
+
+\todo{Continue refactoring.}
+\rewrite{Explain variables}
 
@@
   & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
-  & \qquad\qquad \atom{param\_label}(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
+  & \qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
 \end{align*}

### 2013-04-26 parts/ph.tex
continue refactoring section 4.3

@@
 
-\begin{example*}
+\begin{example}
 \pref{fig:runningPH-1} represents a PH $(\PHs,\PHl,\PHa)$ with
@@
 we thus model an over-approximation of the possible actions.
-\end{example*}
+\end{example}
 
@@
 
-\begin{example*}
+\begin{example}
 The PH in \pref{fig:runningPH-2} results from the refinement of the PH in \pref{fig:runningPH-1}
@@
 highest level, and $a$ does not become inactive unless both $b$ and $c$ are inactive.
-\end{example*}
+\end{example}
 

### 2013-04-29 parts/infer-param.tex
continue refactoring section 4.3

@@
 \newcommand{\paramres}{\atom{param\_resource}}
+\newcommand{\component}{\atom{component}}
+\newcommand{\componentlevels}{\atom{component\_levels}}
 
@@
 
+\rewrite{Facts : component}
+
+\rewrite{Rules + variables : component\_levels}
+
+\rewrite{Cardinalities : énumération de param + param\_label et infered\_param}
+
 A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
-For instance, all information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
+For instance, the information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
+In particular, the predicate $\component$ allows to define all components belonging to the inferred IG.
+Thus, an atom $\component(a, m)$ states that $a$ is a component of the IG and $l_a = m$.
+Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
+\begin{align*}
+  &\component(a, 2).
+  &\component(c, 1). \\
+  &\component(b, 1).
+\end{align*}
+
+To define the sets of possible expression levels of each component (\ie the set $\segm{0}{l_a}$ for each $a \in \Gamma$),
+one can use atoms of the form $\componentlevels(a, k)$ to state that $k \in \segm{0}{l_a}$.
+Variables are then useful in order to enumerate each possible $k$ for all component $a$.
+A variable starts with a capital letter, and during the solving,
+any rule containing variables will be duplicated in order to replace each variable by all the possible values it could represent.
+\rewrite{Reformulate}
+For example, the following rule, which contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$), enumerates all possible expression levels of all components in the system:
+\begin{align*}
+  \componentlevels(A, K) \la \component(\var{A}, \var{M}), 0 \leq K \leq M.
+\end{align*}
+The “$\leq$” notation stand for a shortcut in ASP which has the same meaning as the mathematical operator.
+Regarding \pref{ex:infer-param-runningPH-1}, the previous rule will make the following set of atoms belong to all answer sets:
+\begin{align*}
+  &\{&\componentlevels(b, 0)
+  &&\componentlevels(b, 1) \\
+  &&\componentlevels(c, 0)
+  &&\componentlevels(c, 1) \\
+  &&\componentlevels(a, 0)
+  &&\componentlevels(a, 1) \\
+  &&\componentlevels(a, 2) &&&&\}
+\end{align*}
+
+Cardinality rules are convenient to enumerate all possible parametrizations by creating multiple answer sets.
+The head of a cardinality rule specifies a set of atoms $H$ and two integers $min$ and $max$, and is denoted:
+\begin{align*}
+  min\ \{\ H\ \}\ max \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
+\end{align*}
+Given such a rule, every answer $S$ sets have to verify:
+\begin{align*}
+  min \leq |S \cap H| \leq max
+\end{align*}
+In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
+The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} : \atom{Q}$,
+which is a shorthand for “all atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
+\rewrite{Reformulate}
+
+Given a set of atoms, it creates as many answer sets as there are possible 
+constrains gives any number of possible answers for some atoms between a lower and upper bounds.
+For example,
+\begin{align*}
+  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
+  & \qquad\qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
+\end{align*}
+
+---------------
+
 In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.

### 2013-05-02 parts/infer-param.tex
finish refactoring section 4.3

--- 2013-04-26 parts/infer-param.tex
@@
 
-\rewrite{This whole subsection may be completed, or turned into a new section}
-
-\todo{Rewrite implementation with the new semantics}
+\rewrite{Should this subsection be turned into a new section?}
 
@@
 \newcommand{\atom}{\mathbf}
+%\newcommand{\predicate}{\mathbf}
+\newcommand{\predicate}[1]{#1}
 \newcommand{\la}{\leftarrow}
@@
 
-\newcommand{\paramlabel}{\atom{param\_label}}
-\newcommand{\paramres}{\atom{param\_resource}}
+\newcommand{\paramlabel}{\predicate{param\_label}}
+\newcommand{\paramres}{\predicate{param\_resource}}
+\newcommand{\component}{\predicate{component}}
+\newcommand{\componentlevels}{\predicate{component\_levels}}
+\newcommand{\param}{\predicate{param}}
+\newcommand{\inferedparam}{\predicate{infered\_param}}
+\newcommand{\lessactive}{\predicate{less\_active}}
+\newcommand{\paraminf}{\predicate{param\_inf}}
 
-Answer Set Programming (ASP) \cite{Baral03} has been chosen to address the enumeration of all admissible parametrizations.
+Answer Set Programming (ASP) is a logic programming paradigm \cite{Baral03},
+which has been chosen to address the enumeration of all admissible parametrizations.
 The motivations are following:
 \begin{itemize}
-  \item ASP efficiently tackles the inherent complexity of the models we use, thus allowing a fast execution of the formal tools defined in this paper,
-  \item it is convenient to enumerate a large set of possible answers,
-  \item it allows to easily constrain the answers according to some properties.
+  \item ASP efficiently tackles the inherent complexity of the models used, thus allowing a fast execution of the formal tools defined in this paper,
+  \item ASP is convenient to enumerate a large set of possible answers,
+  \item and it allows to easily constrain the answers according to some properties.
 \end{itemize}
@@
 
-ASP is a logic programming paradigm based on a set of rules.
-A rule takes the form of:
-$$\underbrace{{\ }\atom{H}_{\ }}_{head} \la \underbrace{\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m}_{body}.$$
-where the $head$ is an atom ($\atom{H}$) and the $body$ is a series of atoms ($\atom{A}_i$) and negations of atoms ($\nota \atom{B}_i$).
+\subsubsection{Simple rules}\label{sssec:simple_rules}
+ASP is based on a set of rules of of the form:
+\begin{align*}
+  \underbrace{{\ }\atom{H}_{\ }}_{head} \la \underbrace{\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m}_{body}.
+\end{align*}
+where the $body$ is a series of atoms ($\atom{A}_i$) and negations of atoms ($\nota \atom{B}_i$).
+In the case of \emph{simple rules} (as opposed to the \emph{cardinality rules} of \pref{sssec:cardinality_rules}), the $head$ is also an atom ($\atom{H}$).
 Such a rule states that if all atoms $\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n$ are true
 and all atoms $\atom{B}_1, \atom{B}_2, \dots, \atom{B}_m$ are not true (negation by failure), then $\atom{H}$ has to be true.
-Solving an ASP program means finding an \emph{answer set}, which is a minimal set of true atoms that respects all the rules.
+Solving an ASP program means finding an \emph{answer set}, that is a minimal set of true atoms that respect all the rules.
 Several answer sets can be solution to the same ASP program and the solver can be directed to enumerate them all.
-\todo{Distinguish atoms and predicates.}
+
+An atom is composed of a predicate and a series of arguments (possibly empty).
+For example, the following atom:
+\begin{align*}
+  \predicate{p}(x_1, x_2, \dots, x_r)
+\end{align*}
+is composed of the predicate $\predicate{p}$ and $r$ arguments: $x_1, x_2, \dots, x_r$.
+Each argument can be either an value or a variable,
+a value being a representation of a piece of data (component name, expression level, …)
+and a variable can is used to represent any existing value which respects the rules.
+In this paper, a variable is always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, …)
+while values are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …).
 
 A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
-For instance, all information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
+For instance, the information describing the studied model (the original PH model and the inferred IG and parameters) are expressed in ASP using facts.
+In particular, the predicate $\component$ allows to define all components belonging to the inferred IG.
+Thus, an atom $\component(a, m)$ states that $a$ is a component of the IG and that $l_a = m$.
+
+\begin{example}
+Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
+\begin{align*}
+  &\component(a, 2). \\
+  &\component(b, 1). \\
+  &\component(c, 1).
+\end{align*}
+These three atoms are built with the predicate $\component$.
+Furthermore, $a$, $b$, $c$, $1$ and $2$ are values.
+\end{example}
+
+To describe the sets of all expression levels of each component (\ie the set $\segm{0}{l_a}$ for each $a \in \Gamma$),
+one can use atoms of the form $\componentlevels(a, k)$ to state that $k \in \segm{0}{l_a}$.
+Variables here come in handy to enumerate each possible value $k$ for each component $a$:
+during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible values it could represent.
+The following rule, for example, contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$) and enumerates the set of possible expression levels of each component in the system:
+\begin{align*}
+  \componentlevels(A, K) \la \component(\var{A}, \var{M}), 0 \leq K \leq M.
+\end{align*}
+where the notation “$\leq$” stands for a shortcut in ASP which has the same meaning as the mathematical operator.
+
+\begin{example}
+Regarding \pref{ex:infer-param-runningPH-1}, the previous rule will make the following set of atoms belong to every answer set:
+\begin{align*}
+  &\{&\componentlevels(b, 0)
+  &&\componentlevels(b, 1) \\
+  &&\componentlevels(c, 0)
+  &&\componentlevels(c, 1) \\
+  &&\componentlevels(a, 0)
+  &&\componentlevels(a, 1) \\
+  &&\componentlevels(a, 2) &&&&\}
+\end{align*}
+\end{example}
+
+\subsubsection{Cardinality rules}
+\label{sssec:cardinality_rules}
+As an extension of simple rules, \emph{cardinality rules} turn out to be convenient to enumerate a set of answer sets.
+The head of a cardinality rule specifies a set of atoms $H$ and two integers $min$ and $max$, and is denoted:
+\begin{align*}
+  min\ \{\ H\ \}\ max \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
+\end{align*}
+Given such a rule, as many answer sets as possible are created, so that each answer $S$ set verifies:
+\begin{align*}
+  min \leq |S \cap H| \leq max
+\end{align*}
+and every atom $\atom{H}_i \in S \cap H$ respects the simple rule:
+\begin{align*}
+  \atom{H}_i \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
+\end{align*}
+In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
+The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} : \atom{Q}$,
+which is a shorthand for “the set of atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
+\rewrite{Reformulate?}
+
+Cardinality rules turn out to be convenient to enumerate all possible parametrizations by creating multiple answer sets.
+For functional purposes, a unique label is assigned to every possible set of resources of a given component.
+Thus, we denote $\omega_p$ the set of resources of a given component $a$ labeled by $p$,
+and naturally, $K_{a,\omega_p}$ is the related parameter.
+%and in the following we note $K_{a,\omega_p}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
+We note that labeling the sets of resources of a component is obviously equivalent to labeling its parameters.
+Then, suppose that:
+\begin{itemize}
+  \item $\paramlabel(a, p)$ states that $p$ is a valid label for a set of resources of component $a$ (and therefore $K_{a,\omega_p}$ is a valid parameter);
+  \item $\param(a, p, i)$ states that: $i \in K_{a, \omega_p}$;
+  \item $\inferedparam(a, p)$ states that the parameter inference of $K_{a, \omega_p}$ was conclusive (\pref{pps:param_K}).
+\end{itemize}
+It is thereby possible to enumerate the possible values of all parameters for which \pref{pps:param_K} was not conclusive, with the following cardinality rule:
+\begin{align*}
+  & 1\ \{\ \param(\var{A}, \var{P}, \var{I}) : \componentlevels(\var{A}, \var{I})\ \}\ \infty\ \la \\
+  & \qquad\qquad\qquad \paramlabel(\var{A}, \var{P}), \nota \inferedparam(\var{A}, \var{P}).
+\end{align*}
+Indeed, this rule applies to any possible parameter $\var{P}$ of any component $\var{A}$ ($\paramlabel$) whose value is still unknown ($\nota \inferedparam$),
+and states that any expression level $\var{I}$ of this component ($\componentlevels$) can belong to the value of the parameter ($\param$).
+Furthermore, the lower bound is $1$, which forces each enumerated parameter to contain at least one value,
+but no upper bound is specified ($\infty$) for the size of each parameter (which is bounded by the number of possible expression levels of each gene).
+%each parameter contains at least $1$ value, as stated by the lower bound, and has no upper bound ($\infty$) but the number of expression levels of its component.
+In other worlds, this cardinality rule creates as many answer sets as there are \emph{candidate} parametrizations
+so that if $K_{a, \omega_p}$ could not be inferred by \pref{pps:param_K}, then
+$K_{a, \omega_p} \subset \segm{0}{l_a} \wedge K_{a, \omega_p} \neq \emptyset$
+(thus completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K} or the fact that parameters have to be intervals).
+
+%so that
+%inferred parameters with \pref{pps:param_K} keep their value and
+%parameters that could not be inferred take any possible value amongst the expression levels of the related component: $K_{a, \omega_p} \subset \segm{0}{l_a} \wedge K_{a, \omega_p} \neq \emptyset$.
+
+\subsubsection{Constraints}\label{sssec:constraints}
+%Finally, constraints can be used to filter answer sets that model parametrizations which are impossible or do not respect the assumptions of \pref{ssec:admissible-K}.
+Finally, a constraint is a rule with no $head$ part:
+\begin{align*}
+  \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
+\end{align*}
+A constraint is satisfied only if its $body$ is not satisfied,
+which thus allows to invalidate answer sets containing some unwanted combinations of atoms.
+In the scope of parameters enumeration, for example, constraints are especially useful to filter parametrizations
+that contain non-interval parameters or that do not respect the assumptions of \pref{ssec:admissible-K}.
+Indeed, suppose that:
+\begin{itemize}
+  \item $\lessactive(a, p, q)$ states that $\omega_p$ is a set of resources of $a$ with (loosely) less activators and more inhibitors than $\omega_q$;
+  \item $\paraminf(a, p, q)$ states that: $K_{a,\omega_p} \leqsegm K_{a,\omega_q}$.
+\end{itemize}
+Then, the monotonicity assumption (\pref{pro:param_enum_monotonicity}) is formulated as the following constraint:
+\begin{align*}
+  \la \lessactive(\var{A}, \var{P}, \var{Q}), \nota \paraminf(\var{A}, \var{P}, \var{Q}).
+\end{align*}
+Indeed, this constraint removes all parametrization results where parameters $K_{\var{A},\omega_\var{P}}$ and $K_{\var{Z},\omega_\var{Q}}$ exist
+such that $\var{A}$ is less activated by the set of resources $\omega_\var{P}$ than $\omega_\var{Q}$,
+but $K_{\var{A},\omega_\var{Q}} \ltsegm K_{\var{A},\omega_\var{P}}$,
+thus violating the monotonicity assumption.
+Of course, other assumptions can be formulated in the same way.
+
+\begin{example}
+In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{b\}}$ and $K_{a,\{c\}}$ could not be inferred by \pref{pps:param_K}.
+The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values
+(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
+\begin{align*}
+  (K_{a,\{b\}} ; K_{a,\{c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
+\end{align*}
+We note that $(\segm{1}{1} ; \segm{1}{1})$ belongs to the enumerated set of admissible parametrizations.
+Therefore, this enumeration allows, from the model in \pref{fig:runningPH-1}, to find the behavior of the model completed with a cooperative sort described in \pref{fig:runningPH-2}.
+\end{example}
+
+\begin{comment}
+---------------
+
 In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
@@
 Of course, other assumptions can be formulated in the same way.
+\end{comment}
 
-This subsection succinctly described how we write ASP programs to represent a model and solve all steps of Thomas' modeling inference.
+This subsection succinctly described how ASP programs come in handy to represent a model and solve all steps of Thomas' modeling inference.
 It finds a particularly interesting application in the enumeration of parameters: all possible parametrizations are generated in separate answer sets, and integrity constraints are formulated to remove those that do not fit the assumptions of admissible parametrizations,

### 2013-05-03 parts/grn.tex
fixes after review from Olivier

@@
 
-We concisely present Thomas' modeling of BRNs dynamics, merely inspired by
+We concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
 \cite{Richard06,BernotSemBRN}.
@@
 inference),
-we propose the notion of unsigned edge modeling an interaction whose nature (activation or inhibition) is undefined,
+we propose the notion of unsigned edge modeling an interaction whose nature (activation or inhibition) is undetermined,
 and we extend the classical parametrization formalism by setting parameters to intervals of values instead of single values.
@@
 
-The IG is composed of nodes that represent components,
-and edges labeled with a sign and a threshold that stand for regulations between these components (\pref{def:ig}).
-The activity, concentration rate or presence of each component in a given state of the system is modeled by a discrete value called \emph{expression level}.
-The maximum level of a component $a$ is denoted $l_a$.
+The IG is composed of nodes ($a$, $b$, $c$, …) that represent components,
+and edges ($\GRNedge{a}{s}{t}{b}$, …) labeled with a sign ($s$) and a threshold ($t$) that stand for regulations between these components (\pref{def:ig}).
+The activity, concentration rate or presence of each component in a given state of the system is modeled by an abstract discrete value called \emph{expression level}.
+The maximum expression level of a component $a$ is denoted $l_a$.
 The sign of an edge denotes the kind of regulation it models: it can be positive ($+$), negative ($-$) or unsigned ($\uns$),
-the latter meaning that the nature (activation or inhibition) of the regulation is unknown.
-For a regulation to take place (activation or inhibition), the expression level of its head component has to be higher than its threshold;
-otherwise, the opposite influence is expressed.
+the latter meaning that the nature of the regulation can not be reduced to a simple activation or inhibition.
+Regarding the dynamics, an edge $\GRNedge{a}{s}{t}{b}$ states that $a$ influences the evolution of $b$ in a certain way when its expression level is above or equal to the threshold $t$;
+when its expression level is strictly below this threshold, another kind of influence is expressed, which usually consists of the opposite influence.
+%For a regulation to take place (activation or inhibition), the expression level of its head component has to be higher than its threshold;
+%otherwise, the opposite influence is expressed.
 The uniqueness of each regulation is forced, in order to make the following sections simpler.
@@
 \end{align}
-Then, for all component $a \in \GRNreg{b}$ regulating $b$,
-we denote $\levels{a}{b}$ (resp.~$\ulevels{a}{b}$) the levels of $a$ above (resp.~below) the threshold $t$ of the (unique) edge $\GRNedge{a}{s}{t}{b}$ (\pref{def:levels}).
-If the expression level of $a$ belongs to the interval $\levels{a}{b}$, then $a$ is expected to have the influence corresponding to the sign $s$ on $b$;
-if its expression level belongs to $\ulevels{a}{b}$, the opposite influence is expected.
+Then, for all component $a$ regulating $b$,
+\ie if $\GRNedge{a}{s}{t}{b} \in E$,
+we denote $\levels{a}{b}$ (resp.~$\ulevels{a}{b}$) the interval of expression levels of $a$ above (resp.~below) the threshold $t$ (\pref{def:levels}).
+For all expression levels of $a$ that belong to $\levels{a}{b}$, $a$ is expected to have the influence corresponding to the sign $s$ on $b$;
+for all expression levels belonging to $\ulevels{a}{b}$, the opposite influence is expected.
+This definition holds because of the uniqueness of the edge $\GRNedge{a}{s}{t}{b}$.
 
 \begin{definition}[Effective levels ($\levelsl$)]\label{def:levels}
-Let $b \in \Gamma$ and $a \in \GRNreg{a}$; we define:
+If $\GRNedge{a}{s}{t}{b} \in E$, we define:
+%Let $b \in \Gamma$ and $a \in \GRNreg{a}$; we define:
 $$\levels{a}{b} \DEF \segm{t}{l_a} \quad \text{and} \quad \ulevels{a}{b} \DEF \segm{0}{t-1} \enspace.$$
@@
 $\GRNget{s}{a}$ refers to the level of component $a$ in $s$.
-\rewrite{Functional definition of a state? (More rigorous)}
-The specificity of Thomas' approach lies in the use of discrete \emph{parameters} to represent the
-focal level interval towards which the component will evolve in each configuration of its regulators
-(\pref{def:param}).
-Indeed, for each possible state of a BRN, all regulators of a component $a$ can be divided into
-\emph{activators} and \emph{inhibitors}, given their type of interaction and expression level,
-referred to as the \emph{resources} of $a$ in this state (\pref{def:resources}).
+%\rewrite{Functional definition of a state? (More rigorous)}
+For each possible state, the set of \emph{resources} of a given component
+is the set of regulators of this component whose expression level is above the threshold of the regulation (\pref{def:resources}).
+In other words, in every state $s$, a regulator $b$ of a component $a$
+is either a resource (if $\GRNget{s}{b} \in \levels{b}{a}$)
+or not a resource (if $\GRNget{s}{b} \in \ulevels{b}{a}$).
+%For each possible state, all regulators of a component $a$ can be divided into
+%\emph{activators} and \emph{inhibitors}, given their type of interaction and expression level,
+%referred to as the \emph{resources} of $a$ in this state (\pref{def:resources}).
+Relying on this observation, the specificity of Thomas' approach lies in the use of discrete \emph{parameters} to represent the
+focal level interval towards which the component will evolve in every configuration of the resources (\pref{def:param}).
 
@@
 For a given component $a \in \Gamma$, and $\omega \subset \GRNreg{a}$ a set of regulators of $a$,
-the \emph{parameter} $K_{a,\omega} = \segm{i}{j}$, with $0 \leq i \leq j \leq l_a$, is a non-empty interval towards which $a$ will tend
-in the states where its resources are exactly the regulators in $\omega$.
+the \emph{parameter} $K_{a,\omega} = \segm{i}{j}$ is a non-empty interval with with $0 \leq i \leq j \leq l_a$.
 The complete map $K$ of parameters on $\IG$ is called a \emph{parametrization} on $\IG$.
 \end{definition}
-The purpose of interval parameters is discussed at the end of this section.
-
-\begin{example}
-\pref{fig:runningBRN}(right) gives a Parametrization on the IG of \pref{fig:runningBRN}(left).
-\end{example}
-
-At last, \pref{def:dynamics} gives the asynchronous dynamics of a BRN using Thomas' parameters.
-From a given state $s$, a transition to another state $s'$ is possible provided that exactly one component $a$ evolves of one level towards $K_{a,\GRNres{a}{s}}$,
-as stated by the definition of the transition relation $\GRNtrans{s}{s'}$.
-However, $a$ cannot evolve if its expression level already belongs to $K_{a,\GRNres{a}{s}}$.
+An IG and a parametrization make up a complete BRN.
+The purpose of intervals as parameters is discussed at the end of this section.
+Regarding the dynamics, an interval parameter $K_{a,\omega}$ is the set of values towards which $a$ will tend
+in the states where its resources are exactly the regulators in $\omega$.
+%as defined in the asynchronous dynamics of \pref{def:dynamics}.
+%At last, \pref{def:dynamics} gives the asynchronous dynamics of a BRN using Thomas' parameters.
+Indeed, from a given state $s$, a transition to another state $s'$ is possible provided that
+exactly one component $a$ evolves of one expression level towards the closest value of the interval $K_{a,\GRNres{a}{s}}$,
+as stated by the definition of the transition relation $\GRNtrans{s}{s'}$ (\pref{def:dynamics}).
+However, $a$ cannot evolve if its expression level already belongs to the interval $K_{a,\GRNres{a}{s}}$.
 
@@
 \end{definition}
+The notations “$<$” and “$>$” comparing an integer and an interval were defined at the end of \pref{sec:intro}.
 
 \begin{example}
-In the BRN that consists of the IG and parametrization of \pref{fig:runningBRN}, the following
-transitions are possible given the semantics defined in \pref{def:dynamics}:
+\pref{fig:runningBRN}(right) gives a Parametrization on the IG of \pref{fig:runningBRN}(left).
+In this BRN, the following transitions are possible given the semantics defined in \pref{def:dynamics}:
 $$\GRNstate{a_0, b_1, c_1} \rightarrow \GRNstate{a_1, b_1, c_1} \rightarrow \GRNstate{a_2, b_1, c_1} \rightarrow
@@
 \paragraph{Discussion on unsigned edges}
-Unsigned edges allow to model regulations whom trend (activation or inhibition) is unknown.
+Unsigned edges allow to model regulations whose trend is more complex than a simple activation or inhibition.
 It has to be noted that using unsigned edges in addition to the regular positive and negative ones does not add expressivity to the framework.
@@
 PH actions are very atomistic and the inference of the underlying BRN proposed in \pref{sec:infer-IG}
-allows to find the global trends of all regulations between components,
+allows to find the global trends of regulations between components,
 which may turn out to be difficult to perform by hand.
@@
 While the use of intervals as parameter values ($K_{a,\omega} = \segm{i}{j}$) instead of simple integers ($K_{a,\omega} = i$) does not add expressivity in boolean networks,
-it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t. the above definitions).
+it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t.~the above definitions).
+Indeed, it is not always possible to create the same dynamics of a BRN with interval parameters by using only integer parameters.
+Consider for example an interval parameter with three values: $K_{a,\omega} = \segm{i}{i+2}$,
+and let $s$ be its corresponding state: $\GRNres{a}{s} = \omega$.
+From \pref{def:dynamics}, if $\GRNget{s}{a} \in K_{a,\omega}$, then $a$ can not evolve in $s$; therefore, $a$ has three steady-states ($i$, $i+1$ and $i+2$).
+This behavior can not be reproduced with integer intervals, because it is impossible to distinguish three different sets of resources corresponding to the three involved states.
+At most two steady-states can be modeled by using a self-regulation $\GRNedge{a}{s}{t}{a}$, allowing to distinguish two cases ($\GRNget{s}{a} < t$ and $\GRNget{s}{a} \geq t$), which is not enough.
+\begin{comment}
 Indeed, assume that $K_{a,\omega} = \segm{i}{i+2}$;
@@
 by relying only on the regulation of $a$ on itself.
+\end{comment}
 Finally, with the same reasoning, we also remark that the use of intervals makes optional some explicit auto-activations in the IG

### 2013-05-03 parts/infer-param.tex
fixes after review from Olivier

@@
 \label{ex:infer-param-runningPH-1}
-Applied to the PH in \pref{fig:runningPH-1}, we obtain, for instance, 
+Applied to the PH in \pref{fig:runningPH-1}, we obtain, in particular,
 $K_{b,\emptyset} = \segm{0}{1}$ and
@@
 \begin{example}
-For the PH in \pref{fig:runningPH-2}, we obtain
+Regarding the PH in \pref{fig:runningPH-2}, we obtain
 $K_{b,\emptyset} = \segm{0}{1}$,

### 2013-05-03 parts/intro.tex
start refactoring section 4.3
fixes after review from Olivier

@@
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 As regulatory phenomena play a crucial role in biological systems, they need to be studied accurately.
@@
 $\segm{i}{j}$ is the set of integers $\{ i, i+1, \dots, j \}$.
-We note: $$\segm{i_1}{j_1} \leqsegm \segm{i_2}{j_2} \EQDEF (i_1\leq i_2\wedge j_1\leq j_2) \enspace,$$
-%and $$[i_1;j_1] \ltsegm [i_2;j_2] \EQDEF 
-%	[i_1;j_1] \leq_{[]} [i_2;j_2] \wedge (i_1\neq i_2 \vee j_1\neq j_2)$.
-%(i_1<i_2 \wedge j_1\leq j_2) \vee (i_1\leq i_2\wedge j_1 < j_2)$$.
-and, given an integer $k$: $k < \segm{i}{j} \EQDEF k < i$ and $k > \segm{i}{j} \EQDEF k > j$.
+Given an integer $k$, we note: $k < \segm{i}{j} \EQDEF k < i$ and $k > \segm{i}{j} \EQDEF k > j$.
+We also define:
+\begin{align*}
+  \segm{i_1}{j_1} \leqsegm \segm{i_2}{j_2} &\EQDEF (i_1\leq i_2\wedge j_1\leq j_2) \enspace\text{, and:}\\
+  \segm{i_1}{j_1} \ltsegm \segm{i_2}{j_2} &\EQDEF (i_1<i_2 \wedge j_1\leq j_2) \vee (i_1\leq i_2\wedge j_1 < j_2) \enspace.
+\end{align*}

### 2013-05-03 parts/grn.tex
discussion on edge a -2 b

@@
 as the number of processes in a PH sort is not constrained in any way.
-For instance, in the example of \pref{fig:runningBRN}, it would be more relevant to have $l_a = 1$, as $a$ has only one outgoing edge;
-however, the value $l_a = 2$ makes the example more interesting.
+For instance, in the example of \pref{fig:runningBRN},
+it would be more relevant to have $l_a = 1$ and $\GRNedgef{a}{-}{1}{b} \in E$, as $a$ has only one outgoing edge,
+but the value $l_a = 2$ makes the example more interesting, therefore $\GRNedgef{a}{-}{2}{b} \in E$ can be arbitrarily chosen.
 

### 2013-05-03 main.tex
continue refactoring section 4.3
finish refactoring section 4.3
new contributions + titles

@@
 \newrefformat{ssec}{Subsect.~\ref{#1}}
+\newrefformat{sssec}{Subsect.~\ref{#1}}
 \newrefformat{suppl}{Appendix~\ref{#1}}
 \newrefformat{eq}{Eq.~\eqref{#1}}
+\newrefformat{ex}{Example~\ref{#1}}
 \def\pref{\prettyref}
@@
 \newdefinition{proposition}{Proposition}
-\newdefinition{example*}{Example}
-\newdefinition{example}[example*]{Example}
+\newdefinition{example}{Example}
 
@@
 
-\title{\todo{New title} Concretizing the Process Hitting into Biological Regulatory Networks}
+\title{
+\todo{CMSB title:Concretizing the Process Hitting into Biological Regulatory Networks}
+\\\rewrite{Find a new title:}
+\begin{itemize}
+  \item Abstracting BRNs from PH models
+  \item Abstracting the BRNs of a PH model
+  \item Extracting BRNs from PH models
+  \item ???
+\end{itemize}}
 
@@
 between components.
+To handle wider class of models, we introduce unsigned edges to model regulations that are
+more complex than simple activations or inhibitions.
 Then the inference of corresponding Thomas' models is provided using Answer Set Programming,
@@
 cooperations, where Thomas' approach fails because of the combinatorics of parameters.
-
-\todo{Emphasis on the new contribution since CMSB.}
 \end{abstract}

### 2013-05-03 parts/intro.tex
new contributions + titles

@@
 expressed in PH.
-The second phase of our work concerns the Thomas' parameters inference.
+The second phase of our work concerns the inference of Thomas' parameters.
 It consists in determining the nesting set (possibly too large) of the parameters which necessarily
@@
 Answer Set Programming (ASP) \cite{Baral03} turns out to be effective for these enumerative searches.
+
+%A major contribution from [CMSB] is the use of unsigned edges to model a wider class of models.
+While the work presented in~\cite{FPIMR12-CMSB} could not handle regulations that were not strictly activations or inhibitions,
+the results presented in this paper allow to study a wider class of models by using interactions of a new type: \emph{unsigned} regulations.
+The benefit is that inferring an unsigned edge does not prevent the second phase of the inference,
+and the method is thus conclusive on all boolean models.
+\rewrite{Should we cite [CMSB]?}
 
@@
 
-\todo{Emphasis on the new contribution since CMSB.}
-
 \paragraph{Outline}

### 2013-05-28 parts/grn.tex
discussion on interval parameters

@@
 \paragraph{Discussion on interval parameters}
-While the use of intervals as parameter values ($K_{a,\omega} = \segm{i}{j}$) instead of simple integers ($K_{a,\omega} = i$) does not add expressivity in boolean networks,
-it allows to specify a larger range of dynamics in the general case of multivalued networks (w.r.t.~the above definitions).
-Indeed, it is not always possible to create the same dynamics of a BRN with interval parameters by using only integer parameters.
-Consider for example an interval parameter with three values: $K_{a,\omega} = \segm{i}{i+2}$,
-and let $s$ be its corresponding state: $\GRNres{a}{s} = \omega$.
-From \pref{def:dynamics}, if $\GRNget{s}{a} \in K_{a,\omega}$, then $a$ can not evolve in $s$; therefore, $a$ has three steady-states ($i$, $i+1$ and $i+2$).
-This behavior can not be reproduced with integer intervals, because it is impossible to distinguish three different sets of resources corresponding to the three involved states.
+Whereas most of the literature use single value parameters (e.g., $K_{a,\omega} = i$), we introduce
+in this paper intervals as parameter values (e.g., $K_{a,\omega} = \segm{i}{j}$) for specifying
+the discrete functions for components.
+These interval parameters are mainly introduced for convenience, and it is worth noticing that they
+do not add extra expressivity for modelling multivalued networks.
+Although, interval parameters are more expressive that single value parameters for multivalued networks.
+First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists
+a positive edge from $a$ to $a$ in the interaction graph.
+\todo{be sure to have introduced the inteference of IG from discrete maps}
+Then, let us consider the case where $K_{a,\omega} = \segm{i}{i+2}$ (interval of three values), and
+let $s$ be a state such that $\GRNres{a}{s} = \omega$.
+From \pref{def:dynamics}, if $\GRNget{s}{a} \in K_{a,\omega}$, then $a$ can not evolve in $s$;
+therefore, $a$ has three (local) fixed points ($i$, $i+1$ and $i+2$).
+This behavior can not be reproduced with single value parameters, because it is impossible to distinguish three different sets of resources corresponding to the three involved states.
 At most two steady-states can be modeled by using a self-regulation $\GRNedge{a}{s}{t}{a}$, allowing to distinguish two cases ($\GRNget{s}{a} < t$ and $\GRNget{s}{a} \geq t$), which is not enough.
-\begin{comment}
-Indeed, assume that $K_{a,\omega} = \segm{i}{i+2}$;
-we aim at obtaining the same dynamics with three different integer parameters instead: $K_{a,\omega_1} = i$,  $K_{a,\omega_2} = i+1$, $K_{a,\omega_3} = i+2$.
-The only possible modification in the resources of $a$ is to add a self-regulation $\GRNedge{a}{s}{t}{a}$.
-However, because resources have a boolean definition
-(each component can be only below or above the threshold related to its regulation on $a$),
-it is not possible to differentiate the 3 sets of resources $\omega_1$, $\omega_2$ and $\omega_3$
-by relying only on the regulation of $a$ on itself.
-\end{comment}
-Finally, with the same reasoning, we also remark that the use of intervals makes optional some explicit auto-activations in the IG
-(as for $b$ in \pref{fig:runningBRN}, for instance).

### 2013-06-06 parts/grn.tex
typo

@@
 a positive edge from $a$ to $a$ in the interaction graph.
-\todo{be sure to have introduced the inteference of IG from discrete maps}
+\todo{be sure to have introduced the inference of IG from discrete maps}
 Then, let us consider the case where $K_{a,\omega} = \segm{i}{i+2}$ (interval of three values), and

### 2013-06-10 parts/grn.tex
typos

@@
 These interval parameters are mainly introduced for convenience, and it is worth noticing that they
-do not add extra expressivity for modelling multivalued networks.
-Although, interval parameters are more expressive that single value parameters for multivalued networks.
+do not add extra expressivity for modeling multivalued networks.
+Although, interval parameters are more expressive than single value parameters for multivalued networks.
 First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists

### 2013-06-10 parts/asp.tex
new section for ASP

--- 2013-05-02 parts/infer-param.tex
@@
-\section{Parametrization inference}\label{sec:infer-K}
+\section{Answer Set Programming implementation concepts}
 
-Given the IG inferred from a PH as presented in the previous section, one can find the discrete parameters that model the behavior of the studied PH using the method presented in the following.
-It relies on an exhaustive enumeration of all predecessors of each component in order to find attractor processes and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations.
-The last step consists of the enumeration of all compatible complete parametrizations given this
-set of inferred parameters, the PH dynamics and some biological constraints on parameters.
-
-\subsection{Parameters inference}
-
-This subsection presents some results related to the inference of independent discrete parameters from a given PH.
-These results are equivalent to those presented in \cite{PMR10-TCSB}, with notation adapted to be shared with the previous section.
-In addition, we introduce the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
-which implies that the inferred IG does not contain any unsigned interactions, and thus can be seen as the
-regular IG $(\Gamma, E)$,
-and that any processes in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
-regarding $a$.
-
-\begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
-A PH is well-formed for parameter inference if and only if
-it is well-formed for IG inference, and
-the IG $(\Gamma, E)$ inferred by \pref{pps:inference-IG}
-verifies the following property:
-\begin{align*}
-  \begin{split}
-  \forall a \in \Gamma, \forall b\in \GRNreg{a}&,
-          \forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
-  & \quad \forall c \in \Gamma, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
-  & \qquad
-                          \PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
-                                  \PHfrappe{b_j}{c_k}{c_l}\in\PHa
-  \end{split}
-\end{align*}
-\end{property}
-
-Let $K_{a,\omega}$ be the parameter we want to infer for a given component $a \in \Gamma$
-and $\omega \subset \GRNreg{a}$ a set of its regulators.
-This inference, as for the IG inference, relies on the search of focal processes of the component for the given configuration of its regulators.
-
-For each sort $b \in \GRNreg{a}$, we define a context $C^b_{a,\omega}$ in \pref{eq:param_context} that contains all processes representing the influence of the resources in the configuration modeled by $\omega$.
-The context of a cooperative sort $\upsilon$ that regulates $a$ is given in
-\pref{eq:param_context_coop} as the set of focal processes matching the current configuration.
-$C_{a,\omega}$ refers to the union of all these contexts (\pref{eq:K-ctx}).
-\begin{align}
-  \label{eq:param_context}
-  \forall b\in\Gamma,~
-  C_{a,\omega}^b & \DEF \begin{cases}
-    \levels{b}{a} & \text{if $b \in \omega$,}\\
-    \ulevels{b}{a} & \text{if $b \notin \omega$,}\\
-    L_b            & \text{otherwise;}\\
-  \end{cases}
-  \\
-  \label{eq:param_context_coop}
-  \forall \upsilon \in \PHpredec{a}\setminus\Gamma,~
-  C_{a,\omega}^\upsilon & \DEF \{
-  \upsilon(\sigma) \mid \sigma \in \textstyle\prod_{c\in\PHdirectpredec{\upsilon}}C_{a,\omega}^c \}
-  \\
-  C_{a,\omega} & \DEF \textstyle\bigcup_{b\in\PHpredec{a}} C^b_{a,\omega}
-  \label{eq:K-ctx}
-\end{align}
-
-The parameter $K_{a,\omega}$ specifies to which values $a$ eventually evolves as long as the context
-$C_{a,\omega}$ holds, which is precisely the definition of the $\focals$ function
-(\pref{def:focals} in \pref{ssec:focal}),
-where the focals reachability property can be derived from \pref{pro:wf-ph-K} and
-\pref{eq:param_context_coop}.
-Hence $K_{a,\omega} = \focals(a,C^a_{a,\omega},C_{a,\omega})$ if this latter is a non-empty interval
-(\pref{pps:param_K}).
-
-\begin{proposition}[Parameter inference]
-\label{pps:param_K}
-Let $(\PHs, \PHl, \PHh)$ be a Process Hitting well-formed for parameter inference, and $\IG = (\Gamma, E)$ the inferred IG.
-Let $A$ (resp. $B$) $\subseteq \Gamma$ be the set of regulators that activate (resp. inhibit) a sort
-$a$.
-If $\focals(a,C^a_{a,\omega},C_{a,\omega})=\segm{a_i}{a_j}$ is a non-empty interval, then $K_{a,\omega} = \segm{i}{j}$.
-\end{proposition}
-
-\begin{example}
-\label{ex:infer-param-runningPH-1}
-Applied to the PH in \pref{fig:runningPH-1}, we obtain, for instance, 
-$K_{b,\emptyset} = \segm{0}{1}$ and
-$K_{a,\{b,c\}} = \segm{2}{2}$,
-while $K_{a,\{b\}}$ and $K_{a,\{c\}}$ can not be inferred.
-\end{example}
-
-\begin{example}
-For the PH in \pref{fig:runningPH-2}, we obtain
-$K_{b,\emptyset} = \segm{0}{1}$,
-$K_{a,\{b,c\}} = \segm{2}{2}$,
-$K_{a,\{b\}} = \segm{1}{1}$,
-and $K_{a,\{c\}} = \segm{1}{1}$.
-For this PH, all parameters can be inferred, and the obtained parametrization is the one of \pref{fig:runningBRN}(right).
-\end{example}
-
-Given the \pref{pps:param_K}, we see that in some cases, the inference of the targeted parameter is impossible.
-This can be due to a lack of cooperation between regulators: when two regulators independently hit a component, their actions can have opposite effects, leading to either an indeterministic evolution or to oscillations.
-Such an indeterminism is not possible in a BRN as in a given configuration of regulators, a component can only have an interval attractor, and eventually reaches a steady-state.
-In order to avoid such inconclusive cases, one has to ensure that no such behavior is allowed by
-either removing undesired actions or using cooperative sorts to prevent opposite influences between
-regulators.
-
-\subsection{Admissible parametrizations}\label{ssec:admissible-K}
-
-When building a BRN, one has to find the parametrization that best describes the desired behavior of the studied system.
-Complexity is inherent to this process as the number of possible parametrizations for a given IG is exponential w.r.t. the number of components.
-However, the method of parameters inference presented in this section gives some information about necessary parameters given a certain dynamics described by a PH.
-This information thus drops the number of possible parametrizations, allowing to find the desired behavior more easily.
-
-We first delimit the validity of a parameter (\pref{pro:K-valid}) in order to ensure that any
-transition in the resulting BRN is allowed by the studied PH.
-This is verified by the existence of a hit making the concerned component bounce into the direction
-of the value of the parameter in the matching context.
-Thus, assuming \pref{pro:wf-ph-K} holds, any transition in the inferred BRN corresponds to at least
-one transition in the PH, proving the correctness of our inference.
-We remark that any parameter inferred by \pref{pps:param_K} satisfies this property.
-
-\begin{property}[Parameter validity]\label{pro:K-valid}
-A parameter $K_{a,\omega}$ is valid w.r.t. the PH if and only if the following equation is verified:
-\begin{align*}
-  \forall a_i\in C^a_{a,\omega}, a_i \notin K_{a,\omega} \Longrightarrow
-    (& \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, c_k \in C^c_{a,\omega} \\
-     & \wedge a_i < K_{a,\omega} \Rightarrow j > i \wedge  a_i > K_{a,\omega} \Rightarrow j <i )
-\end{align*}
-\end{property}
-
-Then, we use some additional biological constraints on Thomas' parameters given in
-\cite{BernotSemBRN}, that we sum up in the following three properties:
-
-\begin{property}[Extreme values assumption]
-Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} if and only if:
-\label{pro:param_enum_extreme}
-\begin{align*}
-  \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Longrightarrow \exists \omega \subset \GRNreg{b}, 0 \in K_{b,\omega} \wedge \exists \omega' \subset \GRNreg{b}, l_b \in K_{b,\omega'}
-\end{align*}
-\end{property}
-
-\begin{property}[Activity assumption]
-\label{pro:param_enum_activity}
-Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{activity assumption} if and only if:
-\begin{align*}
-  \forall b \in \Gamma, \forall a \in \GRNreg{b}, \exists \omega \subset \GRNreg{b}, K_{b,\omega} \neq K_{b,\omega \cup \{ a \}}
-\end{align*}
-\end{property}
-
-\begin{property}[Monotonicity assumption]
-\label{pro:param_enum_monotonicity}
-Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{monotonicity assumption} if and only if:
-\begin{align*}
-  \forall b \in \Gamma,
-  \forall A^+ \subset \{ a \in \Gamma \mid \GRNedge{a}{+}{t}{b} \in E_+ \}&,
-  \forall A^- \subset \{ a \in \Gamma \mid \GRNedge{a}{-}{t}{b} \in E_- \},\\
-  K_{b,\omega \cup A^-} & \leqsegm K_{b,\omega \cup A^+}
-\end{align*}
-\end{property}
-
-\subsection{Answer Set Programming implementation concepts}
-
-\rewrite{Should this subsection be turned into a new section?}
+\rewrite{Check and correctly glue this section}
 
@@
 
-\newcommand{\atom}{\mathbf}
+%\newcommand{\atom}{\mathbf}
+\newcommand{\atom}[1]{#1}
 %\newcommand{\predicate}{\mathbf}
@@
 
-\subsubsection{Simple rules}\label{sssec:simple_rules}
+\subsection{Simple rules}\label{sssec:simple_rules}
 ASP is based on a set of rules of of the form:
@@
 
-\subsubsection{Cardinality rules}
+\subsection{Cardinality rules}
 \label{sssec:cardinality_rules}
@@
 
-\subsubsection{Constraints}\label{sssec:constraints}
+\subsection{Constraints}\label{sssec:constraints}
 %Finally, constraints can be used to filter answer sets that model parametrizations which are impossible or do not respect the assumptions of \pref{ssec:admissible-K}.

### 2013-06-10 parts/infer-param.tex
normal shape for atoms
new section for ASP

--- 2013-04-29 parts/infer-param.tex
@@
 \end{property}
-
-\subsection{Answer Set Programming implementation concepts}
-
-\rewrite{This whole subsection may be completed, or turned into a new section}
-
-\todo{Rewrite implementation with the new semantics}
-
-\newcommand{\ti}[1]{\texttt{\textit{#1}}}
-\newcommand{\aspil}[1]{\texttt{#1}}
-\newcommand{\asp}[1]{\begin{itemize} \item[] \aspil{#1} \end{itemize}}
-
-\newcommand{\atom}{\mathbf}
-\newcommand{\la}{\leftarrow}
-\newcommand{\var}[1]{#1}
-\newcommand{\nota}{\neg}
-
-\newcommand{\paramlabel}{\atom{param\_label}}
-\newcommand{\paramres}{\atom{param\_resource}}
-\newcommand{\component}{\atom{component}}
-\newcommand{\componentlevels}{\atom{component\_levels}}
-
-Answer Set Programming (ASP) \cite{Baral03} has been chosen to address the enumeration of all admissible parametrizations.
-The motivations are following:
-\begin{itemize}
-  \item ASP efficiently tackles the inherent complexity of the models we use, thus allowing a fast execution of the formal tools defined in this paper,
-  \item it is convenient to enumerate a large set of possible answers,
-  \item it allows to easily constrain the answers according to some properties.
-\end{itemize}
-We now synthesize some key points to better make the reader understand our ASP implementation with the enumeration example.
-
-ASP is a logic programming paradigm based on a set of rules.
-A rule takes the form of:
-  $$\underbrace{{\ }\atom{H}_{\ }}_{head} \la \underbrace{\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m}_{body}.$$
-where the $head$ is an atom ($\atom{H}$) and the $body$ is a series of atoms ($\atom{A}_i$) and negations of atoms ($\nota \atom{B}_i$).
-Such a rule states that if all atoms $\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n$ are true
-and all atoms $\atom{B}_1, \atom{B}_2, \dots, \atom{B}_m$ are not true (negation by failure), then $\atom{H}$ has to be true.
-Solving an ASP program means finding an \emph{answer set}, which is a minimal set of true atoms that respects all the rules.
-Several answer sets can be solution to the same ASP program and the solver can be directed to enumerate them all.
-\todo{Distinguish atoms and predicates.}
-
-\rewrite{Facts : component}
-
-\rewrite{Rules + variables : component\_levels}
-
-\rewrite{Cardinalities : énumération de param + param\_label et infered\_param}
-
-A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
-For instance, the information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
-In particular, the predicate $\component$ allows to define all components belonging to the inferred IG.
-Thus, an atom $\component(a, m)$ states that $a$ is a component of the IG and $l_a = m$.
-Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
-\begin{align*}
-  &\component(a, 2).
-  &\component(c, 1). \\
-  &\component(b, 1).
-\end{align*}
-
-To define the sets of possible expression levels of each component (\ie the set $\segm{0}{l_a}$ for each $a \in \Gamma$),
-one can use atoms of the form $\componentlevels(a, k)$ to state that $k \in \segm{0}{l_a}$.
-Variables are then useful in order to enumerate each possible $k$ for all component $a$.
-A variable starts with a capital letter, and during the solving,
-any rule containing variables will be duplicated in order to replace each variable by all the possible values it could represent.
-\rewrite{Reformulate}
-For example, the following rule, which contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$), enumerates all possible expression levels of all components in the system:
-\begin{align*}
-  \componentlevels(A, K) \la \component(\var{A}, \var{M}), 0 \leq K \leq M.
-\end{align*}
-The “$\leq$” notation stand for a shortcut in ASP which has the same meaning as the mathematical operator.
-Regarding \pref{ex:infer-param-runningPH-1}, the previous rule will make the following set of atoms belong to all answer sets:
-\begin{align*}
-  &\{&\componentlevels(b, 0)
-  &&\componentlevels(b, 1) \\
-  &&\componentlevels(c, 0)
-  &&\componentlevels(c, 1) \\
-  &&\componentlevels(a, 0)
-  &&\componentlevels(a, 1) \\
-  &&\componentlevels(a, 2) &&&&\}
-\end{align*}
-
-Cardinality rules are convenient to enumerate all possible parametrizations by creating multiple answer sets.
-The head of a cardinality rule specifies a set of atoms $H$ and two integers $min$ and $max$, and is denoted:
-\begin{align*}
-  min\ \{\ H\ \}\ max \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
-\end{align*}
-Given such a rule, every answer $S$ sets have to verify:
-\begin{align*}
-  min \leq |S \cap H| \leq max
-\end{align*}
-In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
-The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} : \atom{Q}$,
-which is a shorthand for “all atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
-\rewrite{Reformulate}
-
-Given a set of atoms, it creates as many answer sets as there are possible 
-constrains gives any number of possible answers for some atoms between a lower and upper bounds.
-For example,
-\begin{align*}
-  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
-  & \qquad\qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
-\end{align*}
-
----------------
-
-In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
-For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
-and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
-In other words, $\paramlabel(a, p)$ states that a parameter $K^p_{a,\omega}$ exists.
-Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 7 parameters, and to state the existence of each of them, the following facts are used:
-\begin{align*}
-  &\paramlabel(a,1).
-  &\paramlabel(b,1).\\
-  &\paramlabel(a,2).
-  &\paramlabel(b,2).\\
-  &\paramlabel(a,3).\\
-  &\paramlabel(a,4).
-  &\paramlabel(c,1).
-\end{align*}
-
-A set (such as sets of resources $\omega$) can be defined in ASP with a series of atoms. \rewrite{Find a more precise sentence.}
-For example, for a given parameter $K^p_{a,\omega}$, expressing $x \in \omega$ is done by the atom: $\paramres(a, p, x)$.
-Then, the sets of resources of \pref{ex:infer-param-runningPH-1} is defined by:
-$$\begin{array}{>{\hspace{.8em}}l|>{\phantom{\raisebox{.6em}{xx}}}l}
-  \hspace{-.8em}\text{Parameter} & \text{Corresponding ASP fact} \\
-\hline
-  K^0_{a,\emptyset} & \\
-  K^1_{a,\{b\}} & \paramres(a,1,b). \\
-  K^2_{a,\{c\}} & \paramres(a,2,c). \\
-  K^3_{a,\{b,c\}} & \paramres(a,3,b). \quad \paramres(a,3,c). \phantom{\raisebox{-.8em}{xx}} \\ \hline
-  K^0_{b,\emptyset} & \\
-  K^1_{b,\{a\}} & \paramres(b,1,a). \phantom{\raisebox{-1em}{xx}} \\ \hline
-  K^0_{c,\emptyset} &
-\end{array}$$
-The absence of an atom $\paramres(a,1,c)$, for example, states that $c$ is not a resource for the parameter $K^1_{a,\{b\}}$ (indeed, $c \notin \{b\}$).
-
-\todo{Continue refactoring.}
-\rewrite{Explain variables}
-
-Rules allow more detailed declarations than facts as they have a body (right-hand part below) containing constraints and allowing to use variables, while facts only have a head (left-hand part).
-For instance, in order to define the set of expression levels of a component, we declare:
-$$\atom{component\_levels}(\var{X}, 0..\var{M}) \la \atom{component}(\var{X}, \var{M}).$$
-where the atom $\atom{component}(X, M)$ stands for the existence of a component $X$ with a maximum level $M$.
-Considering this declaration, any possible answer for the atom $\atom{component\_levels}$ will be found by binding all possible values of its two terms with all existing $\atom{component}$ facts:
-the existence of an answer $\atom{component\_levels}(a, k)$ will depend on the existence of a term $a$, which is bound with $\var{X}$, and an integer $k$, constrained by: $0 \leq k \leq \var{M}$.
-
-Cardinalities are convenient to enumerate all possible parametrizations by creating multiple answer sets.
-A cardinality (denoted hereafter with curly brackets) gives any number of possible answers for some atoms between a lower and upper bounds.
-For example,
-\begin{align*}
-  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
-  & \qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
-\end{align*}
-where $\atom{param}(\var{X}, \var{P}, \var{I})$ stands for: $\var{I} \in K^\var{P}_{\var{X},\omega}$,
-means that any parameter of component $\var{X}$ and label $\var{P}$ must contain at least one level value ($\var{I}$) in the possible expression levels of $\var{X}$.
-Indeed, the lower bound is $1$, forcing at least one element in the parameter, but no upper bound is specified, allowing up to any number of answers.
-The body (right-hand side) of the rule also checks for the existence of a parameter of $\var{X}$ with label $\var{P}$,
-and constrains that the parametrization inference was not conclusive for the considered parameter (“$\nota$” stands for negation by failure: $\nota \atom{L}$ becomes true if $\atom{L}$ is not true).
-Such a constraint gives multiple results as any set of $\atom{param}$ atoms satisfying the cardinality will lead to a new global set of answers.
-In this way, we enumerate all possible parametrizations which respects the results of parameters inference,
-but completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K}.
-
-We rely on integrity constraints to filter only admissible parametrizations.
-An integrity constraint is a rule with no head, that makes an answer set unsatisfiable if its body turns out to be true.
-Hence, if we suppose that:
-\begin{itemize}
-  \item the $\atom{less\_active}(a, p, q)$ atom means that $K^p_{a,\omega}$ stands for a configuration with less activating regulators than $K^q_{a,\omega'}$, % (\ie $A \subset A'$),
-  \item the $\atom{param\_inf}(a, p, q)$ atom means: $K^p_{a,\omega} \leqsegm K^q_{a,\omega'}$,
-\end{itemize}
-then the monotonicity assumption (\pref{pro:param_enum_monotonicity}) is formulated as the following integrity constraint:
-$$\la \atom{less\_active}(\var{X}, \var{P}, \var{Q}), \nota \atom{param\_inf}(\var{X}, \var{P}, \var{Q}).$$
-which removes all parametrization results where parameters $K^\var{P}_{\var{X},\omega}$ and $K^\var{Q}_{\var{X},\omega'}$ exist such that $\var{X}$ is less activated by $\omega$ than $\omega'$ %$A \subset A'$
-but $K^\var{Q}_{\var{X},\omega'} \ltsegm K^\var{P}_{\var{X},\omega}$,
-thus violating the monotonicity assumption.
-Of course, other assumptions can be formulated in the same way.
-
-This subsection succinctly described how we write ASP programs to represent a model and solve all steps of Thomas' modeling inference.
-It finds a particularly interesting application in the enumeration of parameters: all possible parametrizations are generated in separate answer sets, and integrity constraints are formulated to remove those that do not fit the assumptions of admissible parametrizations,
-thus reducing the number of interesting parametrizations to be considered in the end.

### 2013-06-14 parts/intro.tex
several TODOs

@@
 \section{Introduction}\label{sec:intro}
+
+\rewrite{Include Olivier's additions}
+
+\rewrite{Emphasize on the results on IGs (multistationarity, cyclic attractors) and parameters (?) and its complementarity with PH (reachability, fixed points)}
+
+\rewrite{Idea: IG abstracts PH abstracts BRN}
+
 As regulatory phenomena play a crucial role in biological systems, they need to be studied accurately.

### 2013-06-17 parts/infer-ig.tex
new TODOs in IG inference
several TODOs
fix old inference (remove gamma) and add new inference (new section)

@@
 Finally, the rules for inferring the interactions between components from a PH are
-described in \pref{ssec:infer-IG}.
+described in two complementary subsections with a different perspective in the inference.
+\pref{ssec:infer-IG-structure} proposes an inference based on the structure of the original PH model
+while \pref{ssec:infer-IG-dynamics} uses a dynamics point of view to infer the Interaction Graph.
 We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
@@
 
-\subsection{Interaction Inference}\label{ssec:infer-IG}
+\subsection{Structure-based Interaction Inference}\label{ssec:infer-IG-structure}
 
@@
 
+This first inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
 We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
 the level of $b$ makes possible the increase (decrease) of the level of $a$.
-This is analogous to standard IG inferences from discrete maps \cite{RiCo07}.
+To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
+Such groups are given by the set $X(a)$ that is the set of connected components in the graph linking two regulators
+$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
+\begin{equation}
+X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
+        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
+        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
+\label{eq:influence-groups}
+\end{equation}
 
-This reasoning can be straightforwardly applied to a PH when inferring the influence of $b$ on $a$
-when $b\neq a$ (\pref{eq:edges-inference-b}).
-Let us define $\gamma(b\rightarrow a)$ as the set of components cooperating with $b$ to hit $a$,
-including $b$ and $a$ (\pref{eq:cooperating-with-b}).
-Given a configuration $\sigma\in\prod_{c\in\gamma(b\rightarrow a)} L_c$, 
+This reasoning can then be straightforwardly applied when $b \neq a$ (\pref{eq:edges-inference-b}).
+Indeed, to study the influence of any component $b$ on $a$, it is sufficient to study the (unique) group of influence
+$g \in X(a)$ which $b$ belongs to. %, that is, the (unique) set in $X_a$ containing $b$, denoted $X_a(b)$.
+%The reasoning mentioned in the previous paragraph can then be straightforwardly applied when $b \neq a$ (\pref{eq:edges-inference-b}),
+%by considering this 
+Indeed, given a configuration $\sigma\in\prod_{c\in g \cup \{ a \}} L_c$, 
 $\focals(a,\{a_i\},\ctx_b(\sigma))$ gives the bounces that a given process $a_i$ can make in the
@@
 $\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_{i+1}\}))$, then
-$b$ as positive (resp. negative) influence on $a$ with a maximum threshold $l=i+1$.
-\begin{equation}
-\gamma(b\rightarrow a)  \DEF \{ a, b \} \cup \{ c \in \Gamma \mid 
-			\exists \upsilon\in\PHs\setminus\Gamma,
-				\upsilon\in\PHpredec{a} \wedge \{b,c\}\subset\PHpredec{\upsilon} \}
-\label{eq:cooperating-with-b}
-\end{equation}
+$b$ as positive (resp. negative) influence on $a$ with a maximum threshold $t=i+1$.
 
-Then, we infer that $a$ has a self-influence if its current level can have an impact on its own
-evolution at a given configuration $\sigma$.
-We consider here a configuration $\sigma$ of a group $g$ of sorts having a cooperation on $a$.
-This set of sort groups is given by $X(a)$ (\pref{eq:influence-groups}) which returns the set of
-connected components (noted $\mathcal C$) of the graph linking two regulators $b,c$ of $a$ if there
-is a cooperative sort hitting $a$ regulated by both of them.
+Finally, we infer that $a$ has a self-influence if its current level can have an impact on its own
+evolution in a given configuration $\sigma$ of any group of influence on $a$.
 Given $a_i,a_{i+1}\in L_a$, we pick $a_j\in\focals(a,\{a_i\},\ctx_g(\sigma\{a_i\}))$ and
@@
 inferring a self-activation.
-%Future work may propose alternative definitions of self-influences inference in order to range over
-%different parametrization configurations.
-\begin{equation}
-X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
-				\exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
-					\{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
-\label{eq:influence-groups}
-\end{equation}
 
 \pref{pps:inference-edges} details the inference of all existing influences between components occurring
-with a threshold $l$.
+with a threshold $t$.
 These influences are split into positive and negative ones, and represent possible edges in the final IG.
@@
 \begin{split}
-\forall b\in\PHpredecgene{a}, b\neq a, \\
-b\xrightarrow l a \in \hat{E}_s & \Longleftrightarrow
- \exists \sigma\in\textstyle\prod_{c\in\gamma(b\rightarrow a)} L_c, \exists b_i,b_{i+1}\in \PHl_b,\\
+\forall b\in\PHpredecgene{a}, b &\neq a, \\
+b\xrightarrow t a \in \hat{E}_s & \Longleftrightarrow
+  \exists g \in X(a), b \in g, 
+  \exists \sigma\in\textstyle\prod_{c\in g \cup \{ a \}} L_c, \exists b_i,b_{i+1}\in \PHl_b,\\
 & \qquad\qquad
@@
 & \qquad\qquad\qquad
-                        s = \f{sign}(k - j) \wedge l = i+1
+                        s = \f{sign}(k - j) \wedge t = i+1
 \end{split}
@@
 \begin{split}
-a \xrightarrow l a \in \hat{E}_s & \Longleftrightarrow
-\exists g \in X(a), \sigma \in L_a\times \textstyle\prod_{b\in g} L_b,
+a \xrightarrow t a \in \hat{E}_s & \Longleftrightarrow
+\exists g \in X(a), \sigma \in \textstyle\prod_{b\in g \cup \{ a \}} L_b,
       \exists a_i,a_{i+1}\in \PHl_a, \\
@@
       k \neq j+1
-				\wedge s = \f{sign}(k - j) \wedge l = i+1
+        \wedge s = \f{sign}(k - j) \wedge t = i+1
 \end{split}
@@
 \end{align}
-where $s \in \{ +, - \}$, $\bar s = + \EQDEF s = -$, $\bar s = - \EQDEF s = +$,
+where $s \in \{ +, - \}$,
 $\f{sign}(n) = + \EQDEF n > 0$,
@@
 sign. If an influence is both positive and negative, we infer an unsigned edge. In the end, the
-threshold of each edge is the minimum threshold for which the influence has been found. As unsigned
-edges represent ambiguous interactions, no threshold is inferred.
+threshold of each edge is the minimum threshold for which an influence has been found.
+%As unsigned edges represent ambiguous interactions, no threshold is inferred.
 \begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
-We infer $\IG = (\Gamma,E)$ from \pref{pps:inference-edges} as follows:
+We infer $\IG = (\Gamma,E)$ using \pref{pps:inference-edges} as follows:
 \begin{align*}
@@
 $\hat{E}_- = \{a \xrightarrow{2} b\}$, corresponding to the IG of \pref{fig:runningBRN}.
-No self-influence are inferred ($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
+No self-influence are inferred. %($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
+\rewrite{Check this example}
 \end{example}
+
+\todo{Add a new subsection about the alternative inference (cf.~A.~Richard)}
+
+\todo{Insist on the absence of non-functional edges in the inferred IG (contrary to biologists' IGs)}
+
+\subsection{Dynamics-based Interaction Inference}\label{ssec:infer-IG-dynamics}
+
+This second inference is complementary to the one presented in the previous subsection.
+It is based on an approach considering the global dynamics of a component rather than independent dynamics of each regulator;
+in that, it is analogous to standard IG inferences from discrete maps \cite{RiCo07}.
+Contrary to the previously presented method, we focus on a configuration (or: sub-state) $\sigma$
+on the whole set $\reg{a}$ of predecessors of a component $a$ (rather than on a subset of cooperating regulators).
+
+\rewriteil{$B_a(\sigma)$ = Ensemble des processus vers lesquels on bondit}
+
+\begin{align}
+\begin{split}\label{eq:edges-inference-richard-pos}
+  \forall b\in\PHpredecgene{a}, \qquad\qquad\\
+  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
+    \exists a_j \in B_a(\sigma\{b_{t}\}), \exists a_k \in B_a(\sigma\{b_{t+1}\}),\\
+    &\qquad a_j \leq \PHget{\sigma}{a} < a_k \warnul{\vee a_j < \PHget{\sigma}{a} \leq a_k \quad^*}
+\end{split}\\
+\begin{split}\label{eq:edges-inference-richard-neg}
+  \forall b\in\PHpredecgene{a}, \qquad\qquad\\
+  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
+    \exists a_j \in B_a(\sigma\{b_{t}\}), \exists a_k \in B_a(\sigma\{b_{t+1}\}),\\
+    &\qquad a_j > \PHget{\sigma}{a} \geq a_k \warnul{\vee a_j \geq \PHget{\sigma}{a} > a_k \quad^*}
+\end{split}
+\end{align}
+
+\todo{\underline{$^*$Garder les conditions larges ?}}

### 2013-06-18 parts/infer-ig.tex
finish new inference

@@
 Finally, we infer that $a$ has a self-influence if its current level can have an impact on its own
-evolution in a given configuration $\sigma$ of any group of influence on $a$.
+evolution in a given configuration $\sigma$ of any group of influence on $a$ (\pref{eq:edges-inference-a}).
 Given $a_i,a_{i+1}\in L_a$, we pick $a_j\in\focals(a,\{a_i\},\ctx_g(\sigma\{a_i\}))$ and
@@
 We do not consider the cases where a component has no visible influence on another.
-\begin{proposition}[Edges inference]\label{pps:inference-edges}
+\begin{proposition}[Influences inference]\label{pps:inference-edges}
 We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any
@@
   \exists g \in X(a), b \in g, 
-  \exists \sigma\in\textstyle\prod_{c\in g \cup \{ a \}} L_c, \exists b_i,b_{i+1}\in \PHl_b,\\
+  \exists \sigma\in\textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \exists b_i,b_{i+1}\in \PHl_b,\\
 & \qquad\qquad
@@
 a \xrightarrow t a \in \hat{E}_s & \Longleftrightarrow
-\exists g \in X(a), \sigma \in \textstyle\prod_{b\in g \cup \{ a \}} L_b,
+\exists g \in X(a), \sigma \in \textstyle\prod_{b\in g \cup \{ a \}} \PHl_b,
       \exists a_i,a_{i+1}\in \PHl_a, \\
@@
 
-\todo{Add a new subsection about the alternative inference (cf.~A.~Richard)}
-
 \todo{Insist on the absence of non-functional edges in the inferred IG (contrary to biologists' IGs)}
@@
 This second inference is complementary to the one presented in the previous subsection.
-It is based on an approach considering the global dynamics of a component rather than independent dynamics of each regulator;
-in that, it is analogous to standard IG inferences from discrete maps \cite{RiCo07}.
+It is based on an approach considering the global dynamics of a component rather than independent dynamics of each regulator,
+%in that, it is analogous to standard IG inferences from discrete maps \cite{RiCo07}.
+and is inspired from the definition of Interaction Graph built from a Transition Graph as described in~\cite{Richard2010378}.
 Contrary to the previously presented method, we focus on a configuration (or: sub-state) $\sigma$
@@
 
-\rewriteil{$B_a(\sigma)$ = Ensemble des processus vers lesquels on bondit}
+For a given configuration $\sigma$ of a component $a$ and its regulators,
+$\irB_a(\sigma)$ is the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
+This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
+and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
+If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
+%
+\begin{align}
+\begin{split}\label{eq:possible-actions}
+  \forall \sigma &\in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \sigma' = \ctx_{\reg{a}}(\sigma),\\
+  & \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
+  & \qquad\qquad\qquad\qquad\qquad \PHget{\sigma'}{b} = b_i \wedge \PHget{\sigma'}{a} = a_j \}\\
+\end{split}
+\end{align}
+%
+\begin{align}
+\begin{split}\label{eq:possible-bounces}
+  \forall \sigma &\in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c,
+%    \sigma' = \ctx_{\reg{a}}(\sigma),\\
+  \irB_a(\sigma) \DEF 
+  \begin{cases}
+    \irF_a(\sigma)
+      & \text{ if } \irF_a(\sigma) \neq \emptyset\\
+    \{ \PHget{\sigma}{a} \}
+      & \text{ if } \irF_a(\sigma) = \emptyset
+  \end{cases}
+\end{split}
+\end{align}
 
+The main idea of the inference detailed in \pref{pps:inference-edges-dynamics} is close to the one of \pref{ssec:infer-IG-structure}:
+the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
+there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop) (\pref{eq:edges-inference-richard-pos} \& \eqref{eq:edges-inference-richard-neg}).
+the main difference lies in the fact that this inference focuses on local evolutions (the possibility for $a$ to bounce away from its original state)
+rather than global evolutions (by considering focal processes denoting the overall influence).
+Therefore, an influence is inferred if the possible bounces of $a$ in $\sigma$ depend on the level of $b$.
+%
+\begin{proposition}[Influences inference]\label{pps:inference-edges-dynamics}
+We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
 \begin{align}
 \begin{split}\label{eq:edges-inference-richard-pos}
-  \forall b\in\PHpredecgene{a}, \qquad\qquad\\
+  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
   b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
-    \exists a_j \in B_a(\sigma\{b_{t}\}), \exists a_k \in B_a(\sigma\{b_{t+1}\}),\\
+    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
     &\qquad a_j \leq \PHget{\sigma}{a} < a_k \warnul{\vee a_j < \PHget{\sigma}{a} \leq a_k \quad^*}
@@
 \begin{split}\label{eq:edges-inference-richard-neg}
-  \forall b\in\PHpredecgene{a}, \qquad\qquad\\
+  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
   b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
-    \exists a_j \in B_a(\sigma\{b_{t}\}), \exists a_k \in B_a(\sigma\{b_{t+1}\}),\\
+    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
     &\qquad a_j > \PHget{\sigma}{a} \geq a_k \warnul{\vee a_j \geq \PHget{\sigma}{a} > a_k \quad^*}
@@
 \end{align}
+\end{proposition}
+%
+\todo{\underline{$^*$Keep these large conditions?}}
 
-\todo{\underline{$^*$Garder les conditions larges ?}}
+Finally it is possible to infer the final IG by using the same reasoning as \pref{pps:inference-IG}:
+a positive (negative) edge in inferred if there exist positives (negative) influences only, otherwise an unsigned edge in inferred.
+We point out that the main use of this alternative inference is the fact that the conditions used to infer each influence
+in \pref{eq:edges-inference-richard-pos} \& \eqref{eq:edges-inference-richard-neg}
+allows to apply some results about the links between the structure of an IG and the dynamics of the model.
+These results consist mainly in Thomas' conjectures, that focus on the presence of positive and negative circuits in the IG, and state that:
+\begin{itemize}
+  \item the presence of a positive circuit in the IG is a necessary condition for several fixed points,
+  \item the presence of a negative circuit in the IG is a necessary condition for sustained oscillations.
+\end{itemize}
+The sign of a circuit is defined by the number of negative edges it contains;
+given the nature of unsigned edges, the presence of such an edge in a circuit would count for the presence of both a positive and negative circuit.
+These conjectures have been demonstrated in~\cite{Richard2010378} \towrite{Other references! (multi-stationarity \& boolean models)}.
+
+\todo{Example}

### 2013-07-01 parts/infer-ig.tex
create new new inference 2 (groups of regulators and conditions)

@@
 Finally, the rules for inferring the interactions between components from a PH are
-described in two complementary subsections with a different perspective in the inference.
-\pref{ssec:infer-IG-structure} proposes an inference based on the structure of the original PH model
-while \pref{ssec:infer-IG-dynamics} uses a dynamics point of view to infer the Interaction Graph.
+described in \pref{ssec:infer-IG}.
 We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
@@
 
+\subsection{Interaction Inference}\label{ssec:infer-IG}
+
+At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
+($\PHs \setminus \Gamma$) that will not appear in the IG. 
+We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
+by considering direct actions and possible intermediate cooperative sorts.
+The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
+(\pref{eq:regulators}).
+\begin{align}
+\begin{split}
+\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
+(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
+                                   & \quad \quad c^0 = b \wedge c^n = a \\
+                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
+                   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
+\end{split}
+\label{eq:ph_predec}
+\\
+\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
+\label{eq:regulators}
+\end{align}
+
+Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
+refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
+If $g=\{b\}$, we simple note $\ctx_b(\sigma)$.
+This set is composed of the active processes of sorts in $g$, and the focal process (assumed
+unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
+The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
+relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
+$\upsilon$ are defined in $\sigma$.
+When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
+the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
+where $\uplus$ stands for a union of states (which are sets of processes).
+Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
+always terminates.
+\begin{align}
+\forall g\subset \Gamma,
+  \ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
+\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
+\label{eq:ctx-sigma}
+\\
+\upsilon(\sigma) & \DEF
+\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
+  \upsilon'\in\PHdirectpredec{\upsilon} \wedge
+  \upsilon'\in\PHs\setminus\Gamma })
+\label{eq:cooperative-eval}
+\end{align}
+
+This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
+We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
+the level of $b$ makes possible the increase (decrease) of the level of $a$,
+and is directly inspired from the works of~\cite{Richard2010378}.
+To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
+Such groups are given by the set $X(a)$ that is the set of connected components in the graph linking two regulators
+$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
+\begin{equation}
+X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
+        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
+        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
+\label{eq:influence-groups}
+\end{equation}
+
+To infer influence of a component $b$ on a component $a$,
+we observe of the direction of evolution of $a$ for different active processes of $b$.
+The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
+This reasoning can also be applied in the case where $a = b$.
+
+%To infer influences on a component $a$, we compare several configurations of groups of components of $a$.
+%Indeed, to study the influence of any component $b$ on $a$, it is sufficient to study the (unique) group of influence
+%$g \in X(a)$ which $b$ belongs to
+%(that is, the unique set of components regulating $a$ through common cooperative sorts containing $b$).
+%This reasoning relies on the observation of the direction of evolution of $a$ for different active processes of $b$.
+%and we can infer a positive influence if raising the level of $b$ changes the evolution of $a$
+
+Formally, for a given group $g$ of regulators of a component $a$
+(that is, a minimal set of components regulating $a$ through common cooperative sorts),
+and a configuration $\ctx$ on $g$,
+$\irB_a(\sigma)$ is the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
+%This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
+%and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
+If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
+%
+\begin{align}
+\begin{split}\label{eq:possible-bounces}
+  &\forall g \in X(a), \forall \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c,
+%    \sigma' = \ctx_{\reg{a}}(\sigma),\\
+  \irB_a(\sigma) \DEF 
+  \begin{cases}
+    \irF_a(\sigma)
+      & \text{ if } \irF_a(\sigma) \neq \emptyset\\
+    \{ \PHget{\sigma}{a} \}
+      & \text{ if } \irF_a(\sigma) = \emptyset
+  \end{cases}\\
+%\end{split}\\
+%\begin{split}
+  &\text{where: } \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
+  & \qquad\qquad\qquad\qquad\qquad\qquad \PHget{\sigma}{a} = a_j \wedge \PHget{(\ctx_{g}(\sigma))}{b} = b_i \}\\
+\end{split}
+\end{align}
+
+\pref{pps:inference-edges} details the inference of all existing influences between components occurring
+with a threshold $t$.
+The main idea of this inference is that
+the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
+there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
+(\pref{eq:edges-inference-pos} \& \eqref{eq:edges-inference-neg}).
+Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
+We ignore the cases where a component has no visible influence on another.
+%
+\begin{proposition}[Influences inference]\label{pps:inference-edges}
+We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
+\begin{align}
+\begin{split}\label{eq:edges-inference-pos}
+  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
+  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
+    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
+    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \quad^*}\\
+    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \vee a_j < \PHget{\sigma}{a} \leq a_k \quad^{**}}\\
+    &\qquad \warnul{a_j < a_k \quad^{***}}
+\end{split}\\
+\begin{split}\label{eq:edges-inference-neg}
+  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
+  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
+    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
+    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \quad^*}\\
+    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \vee a_j \geq \PHget{\sigma}{a} > a_k \quad^{**}}\\
+    &\qquad \warnul{a_j > a_k \quad^{***}}
+\end{split}
+\end{align}
+\end{proposition}
+%
+\todo{\underline{$^*$Strict inference}}\\
+\todo{\underline{$^{**}$Large inference}}\\
+\todo{\underline{$^{***}$Non-unitary-style inference}}
+
+We are now able to infer the edges of the final IG by considering positive and negative influences
+(\pref{pps:inference-IG}).
+We infer a positive (resp. negative) edge if there only exist corresponding influences with the same sign.
+If an influence is both positive and negative, we infer an unsigned edge.
+In the end, the threshold of each edge is the minimum threshold for which an influence has been found.
+%
+\begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
+We infer $\IG = (\Gamma,E)$ using \pref{pps:inference-edges} as follows:
+\begin{align*}
+E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
+  \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
+E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
+  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
+E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_- \\
+  & \qquad\qquad\qquad \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\}
+\end{align*}
+\end{proposition}
+
+We point out that one of the benefits of this alternative inference is the fact that the conditions used to infer each influence
+in \pref{eq:edges-inference-pos} \& \eqref{eq:edges-inference-neg}
+allow to apply some results about the links between the structure of an IG and the dynamics of the model.
+These results consist mainly in Thomas' conjectures, that focus on the presence of positive and negative circuits in the IG, and state that:
+\begin{itemize}
+  \item the presence of a positive circuit in the IG is a necessary condition for several fixed points,
+  \item the presence of a negative circuit in the IG is a necessary condition for sustained oscillations.
+\end{itemize}
+The sign of a circuit is defined by the number of negative edges it contains;
+given the nature of unsigned edges, the presence of such an edge in a circuit would count for the presence of both a positive and negative circuit.
+These conjectures have been demonstrated in~\cite{Richard2010378} \towrite{Other references! (multi-stationarity \& boolean models)}.
+
+\todo{Insist on the absence of non-functional edges in the inferred IG (contrary to biologists' IGs)}
+
+\begin{example}
+The IG inference from the PH of \pref{fig:runningPH-2} gives
+$\hat{E}_+ = \{b \xrightarrow{1} a, c \xrightarrow{1} a\}$ and 
+$\hat{E}_- = \{a \xrightarrow{2} b\}$, corresponding to the IG of \pref{fig:runningBRN}.
+No self-influence are inferred. %($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
+\rewrite{Check this example}
+\end{example}
+
+\begin{comment}
 \subsection{Structure-based Interaction Inference}\label{ssec:infer-IG-structure}
@@
 \todo{Example}
+\end{comment}

### 2013-07-04 parts/infer-ig.tex
fix mistakes in IG inference

@@
 there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
-(\pref{eq:edges-inference-pos} \& \eqref{eq:edges-inference-neg}).
+(\pref{eq:edges-inference}).
 Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
+Furthermore, studying the influences of the groups of regulators of $a$
+allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
+Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
 We ignore the cases where a component has no visible influence on another.
@@
 \begin{align}
-\begin{split}\label{eq:edges-inference-pos}
-  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
-  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
-    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
-    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \quad^*}\\
-    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \vee a_j < \PHget{\sigma}{a} \leq a_k \quad^{**}}\\
-    &\qquad \warnul{a_j < a_k \quad^{***}}
+\begin{split}\label{eq:edges-inference}
+  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
+  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
+  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+    &\qquad s = \f{sign}(a_k - a_j)
 \end{split}\\
-\begin{split}\label{eq:edges-inference-neg}
-  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
-  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
-    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
-    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \quad^*}\\
-    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \vee a_j \geq \PHget{\sigma}{a} > a_k \quad^{**}}\\
-    &\qquad \warnul{a_j > a_k \quad^{***}}
+\begin{split}\label{eq:edges-inference-auto}
+  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
+  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+    &\qquad s = \f{sign}(a_k - a_j)
+\end{split}\\
+\begin{split}\label{eq:edges-inference-noreg}
+  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
+    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
+    &\qquad s = \f{sign}(a_k - a_j)
 \end{split}
+%\begin{split}\label{eq:edges-inference-pos}
+%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
+%  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
+%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j < a_k
+%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \quad^*}\\
+%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \vee a_j < \PHget{\sigma}{a} \leq a_k \quad^{**}}\\
+%%    &\qquad \warnul{a_j < a_k \quad^{***}}
+%\end{split}\\
+%\begin{split}\label{eq:edges-inference-neg}
+%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
+%  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
+%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j > a_k
+%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \quad^*}\\
+%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \vee a_j \geq \PHget{\sigma}{a} > a_k \quad^{**}}\\
+%    %&\qquad \warnul{a_j > a_k \quad^{***}}
+%\end{split}
 \end{align}
+where $\f{sign}(x) = \begin{cases}\frac{|x|}{x} & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
 \end{proposition}
 %
-\todo{\underline{$^*$Strict inference}}\\
-\todo{\underline{$^{**}$Large inference}}\\
-\todo{\underline{$^{***}$Non-unitary-style inference}}
+%\todo{\underline{$^*$Strict inference}}\\
+%\todo{\underline{$^{**}$Large inference}}\\
+%\todo{\underline{$^{***}$Non-unitary-style inference}}
 
@@
 
-We point out that one of the benefits of this alternative inference is the fact that the conditions used to infer each influence
-in \pref{eq:edges-inference-pos} \& \eqref{eq:edges-inference-neg}
+We point out that one of the benefits of this inference is the fact that the conditions used to infer each influence
+in \pref{pps:inference-edges}
 allow to apply some results about the links between the structure of an IG and the dynamics of the model.

### 2013-07-04 parts/infer-ig.tex
finally, switch back and shrink

@@
 Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
-Furthermore, studying the influences of the groups of regulators of $a$
-allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
-Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
+%Furthermore, studying the influences of the groups of regulators of $a$
+%allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
+%Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
 We ignore the cases where a component has no visible influence on another.
@@
 \begin{split}\label{eq:edges-inference}
-  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
+  \forall b\in\PHpredecgene{a} \cup \{ a \}, \forall s \in \{ +, - \}, \\
   b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
@@
 \end{split}\\
-\begin{split}\label{eq:edges-inference-auto}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
-\end{split}\\
-\begin{split}\label{eq:edges-inference-noreg}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
-    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
-\end{split}
+%
+%%% Tentative de réduire le nombre d'auto-arcs superflus
+%
+%\begin{split}\label{eq:edges-inference-auto}
+%  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+%  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
+%  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+%    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+%    &\qquad s = \f{sign}(a_k - a_j)
+%\end{split}\\
+%\begin{split}\label{eq:edges-inference-noreg}
+%  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+%  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
+%    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
+%    &\qquad s = \f{sign}(a_k - a_j)
+%\end{split}
+%
+%%% Ancienne version de l'inférence
+%
 %\begin{split}\label{eq:edges-inference-pos}

### 2013-07-05 main.tex
new title
typo in title
new section for ASP
several TODOs
fix old inference (remove gamma) and add new inference (new section)
finish new inference
avancées sur l'intro et l'abstract

--- 2013-04-25 main.tex
@@
 %\usepackage{color} % Couleurs du texte
-\newcommand{\rewrite}[1]{\textcolor{blue}{\textbf{[REWRITE: #1]}}}
+\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
+\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
+\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
+\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
+\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
 \newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
 \newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
+\def\updated#1{{\color{magenta}#1}}
 %\newcommand{\new}[1]{#1}
@@
 
-\title{\todo{New title} Concretizing the Process Hitting into Biological Regulatory Networks}
+\title{Constructing Biological Regulatory Networks from Process Hitting models}
+%\\\rewriteil{Or possibly: “Complementarity between PH, IG and Thomas modeling”}
 
@@
 2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.}
-\address[eth]{ETH Zürich, Institut für Automatik,\\
-ETL I 28, Physikstrasse 3, 8092 Zurich, Switzerland.}
+\address[eth]{BISON group, Automatic Control Laboratory, ETH Zürich\\
+Physikstrasse 3, 8092 Zurich, Switzerland.}
 
@@
 
-Given a PH model of a BRN, we first tackle the inference of the underlying Interaction Graph
-between components.
-Then the inference of corresponding Thomas' models is provided using Answer Set Programming,
-which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
-
-In addition to giving a formal link between different approaches for qualitative BRNs modeling, 
-this work emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
+This paper establishes formal links between these two formalisms, permitting to appreciate their
+complementary features.
+First, the inference of the Interaction Graph (IG) from a PH models allows a compact representation of
+the regulations that are effective for the dynamics.
+From this IG, well-known results can be applied to derive global dynamical properties for the
+modelled BRN.
+Second, the inference of all Thomas models that are compatible with a given PH is tackled.
+As the PH allows partial specification of cooperation between components, this inference
+emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
 cooperations, where Thomas' approach fails because of the combinatorics of parameters.
 
-\todo{Emphasis on the new contribution since CMSB.}
+The inference of corresponding Thomas' models is provided using Answer Set Programming,
+which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
+\todo{application}
 \end{abstract}
@@
 \input{parts/infer-param}
+\input{parts/asp}
 \input{parts/example}

### 2013-07-05 parts/infer-ig.tex
less spurious auto-edges in IG inference
avancées sur l'intro et l'abstract

@@
-\section{Interaction Graph Inference}\label{sec:infer-IG}
+% vim:spell spelllang=en:
+\section{Interaction Graph Inference from Process Hitting}\label{sec:infer-IG}
 
-In order to infer a complete BRN, one has to find the Interaction Graph (IG) first, as some
-constraints on the parametrization rely on it.
-Inferring the IG is an abstraction step which consists in determining the global influence of
-components on each of its successors.
+\updated{%
+The Interaction Graph (IG) is an abstract representation of the direct qualitative influences,
+positives and/or negatives, between the components of the system.
+As discussed in \pref{sec:intro}, the IG allows to efficiently characterise global dynamical
+properties for the concrete system, such as the capability for multi-stationarity or oscillation.}
+
+\updated{%
+In a typical biological network modelling process, a prior IG is generally the starting point for
+the formal system specification.
+However, it is very common that the prior IG actually refers interaction that reveal to be 
+non-effective with respect to the dynamics.
+Hence, deriving the IG directly from the dynamical models lead to more concise IGs, enhancing the
+conclusiveness of static analyses based upon this abstract representation.}
+
+\updated{%
+In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
+modelling.}
+
+%In order to infer a complete BRN, one has to find the Interaction Graph (IG) first, as some
+%constraints on the parametrization rely on it.
+%Inferring the IG is an abstraction step which consists in determining the global influence of
+%components on each of its successors.
 
@@
     &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
+    &\qquad s = \f{sign}(k - a)
 \end{split}\\
@@
 \end{align}
-where $\f{sign}(x) = \begin{cases}\frac{|x|}{x} & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
+where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
 \end{proposition}
@@
 %\todo{\underline{$^{***}$Non-unitary-style inference}}
+
+\todo{remark for reducing the number of self-regulation}
 
@@
 \end{proposition}
-
-We point out that one of the benefits of this inference is the fact that the conditions used to infer each influence
-in \pref{pps:inference-edges}
-allow to apply some results about the links between the structure of an IG and the dynamics of the model.
-These results consist mainly in Thomas' conjectures, that focus on the presence of positive and negative circuits in the IG, and state that:
-\begin{itemize}
-  \item the presence of a positive circuit in the IG is a necessary condition for several fixed points,
-  \item the presence of a negative circuit in the IG is a necessary condition for sustained oscillations.
-\end{itemize}
-The sign of a circuit is defined by the number of negative edges it contains;
-given the nature of unsigned edges, the presence of such an edge in a circuit would count for the presence of both a positive and negative circuit.
-These conjectures have been demonstrated in~\cite{Richard2010378} \towrite{Other references! (multi-stationarity \& boolean models)}.
-
-\todo{Insist on the absence of non-functional edges in the inferred IG (contrary to biologists' IGs)}
 

### 2013-07-05 parts/intro.tex
avancées sur l'intro et l'abstract

--- 2013-05-03 parts/intro.tex
@@
+% vim:spell spelllang=en:
 \section{Introduction}\label{sec:intro}
+
+\rewrite{Include Olivier's additions}
+
 As regulatory phenomena play a crucial role in biological systems, they need to be studied accurately.
@@
 (e.g., \cite{Twardziok10-CMSB}) dimensions, to name but a few.
+
 While the formal checking of dynamical properties is often limited to small networks because of the
-state graph explosion, the main drawback of this framework is the difficulty to specify Thomas'
-parameters, especially for large networks.
+state graph explosion, \updated{an other major} drawback of this framework is the difficulty to
+specify Thomas' parameters, especially for large networks.
 
-In order to address the formal checking of dynamical properties within very large BRNs, we recently
-introduced in \cite{PMR10-TCSB} a new formalism, named the \emph{``Process Hitting''} (PH), to model
+In order to address the formal analysis of dynamical properties within very large BRNs, we recently
+introduced in \cite{PMR10-TCSB} a new formalism, named the \emph{Process Hitting} (PH), to model
 concurrent systems having components with a few qualitative levels.
-A PH describes, in an atomic manner, the possible evolutions of a process (representing one
+A PH model describes, in an atomic manner, the possible evolutions of a process (representing one
 component at one level) triggered by the hit of at most one other process in the system.
 This framework can be seen as a special class of formalisms like Petri Nets or Communicating Finite
-State Machines, where the causality between actions is restricted.
+State Machines, \updated{where the arity and the effect of synchronisations between components is
+restricted.}
 Thanks to the particular structure of interactions within a PH, very efficient static analysis
 methods have been developed to over- and under-approximate reachability properties making tractable
-the formal analysis of BRNs with hundreds of components \cite{PMR12-MSCS}.
+the formal analysis of the qualitative dynamics of BRNs with hundreds or thousands of components,
+\updated{which was impossible with other state-of-the-art approaches \cite{PMR12-MSCS,PAK13-CAV,FPMR13-CS2Bio}}.
 
-PH is suitable to model BRNs with different levels of abstraction in the specification of
+The PH is suitable to model BRNs with different levels of abstraction in the specification of
 cooperations (associated influences) between components.
@@
 
-The objectives of the work presented in this paper are the following.
-Firstly, we show that starting from one PH model, it is possible to find back the underlying IG.
-We perform an exhaustive search for the possible interactions on one component from all the
-others, consistently with the knowledge of the dynamics that these interactions lead to and that are
-expressed in PH.
-The second phase of our work concerns the Thomas' parameters inference.
-It consists in determining the nesting set (possibly too large) of the parameters which necessarily
-lead to the satisfaction of the known cooperating constraints.
-The resulting BRN dynamics is ensured to respect the PH dynamics, \ie no spurious transitions are
-made possible by the inference.
-Answer Set Programming (ASP) \cite{Baral03} turns out to be effective for these enumerative searches.
+\medskip
 
-The outcome of this work is twofold.
-The first benefit is that such an approach makes it possible to refine the construction of
-BRNs with a partial and progressively brought knowledge in PH, while being able to export such
-models in the Thomas' framework.
-This work thus strengthens the formal link between both modelings.
-The second feature of our method is that it can be applied on very large BRNs.
+\updated{%
+The aim of this paper is to formally estiblish links between several layers suited for specifying the
+qualitative dynamics of BRNs, namely:
+the Interaction Graph (IG), referencing the sign of direct regulations between components;
+the Process Hitting (PH), which permits to specify the qualitatif dynamics of the system with
+various degrees of knowledge for joint regulations;
+and Thomas' parameters, which completely specify the functions governing the evolution of
+components.}
 
-Finally, it must be noticed that we are not interested in this paper in the derivation of one
-PH from a BRN (which was previously described in \cite{PMR10-TCSB}) but, on the contrary, to finding out
-a set of BRNs from one PH.
+\updated{%
+Firstly, we derive rules that, given a PH model, infer the IG corresponding to the dynamics of the
+encoded BRN.
+The obtained IG relates only components that have actual influence in the dynamics.
+This typically results into IGs that contain less interaction than an hypothetical starting prior
+IG, as the specification of component evolutions may reveal non-functional regulations.
+Based on the derived IG, various static analysis allows to conclude on global properties
+of the system dynamics (see \cite{PR11-SASB} for a short survey).
+The most known are the Thomas' conjectures (that have been proved since \todo{refs}),
+which relate the absence of
+positive cycle to the impossibility for multi-stationarity (distinct attractors),
+and the absence of negative cycle to the impossibility for oscillatory behaviours.
+}
+\todo{check if the inferred IG contains the IG of compatible Thomas}
+
+\updated{%
+Second, we tackle the characterisation of Thomas' parameters that are compatible with a given PH
+model, \ie that fully specify the evolution of components while respecting the cooperations
+specified in PH.
+More formally, the resulting BRN dynamics is ensured to be included in the PH dynamics: any
+transition in the (asynchronous) Thomas model exists in the PH model.
+In practice, the number of compatible Thomas parameters can be huge for large networks where only a
+few cooperations are specified.
+This highlights the fact that the PH is very suited to analyse dynamics of large-scale networks
+where the knowledge on cooperations is often incomplete:
+instead of dealing with a possibly non-tractable number of Thomas models, one can already capture
+precise dynamical features using a single PH models which gathers all the plausible dynamics.
+}
+
+\updated{%
+Finally, we discuss on an implementation of these inference schemes using Answer Set
+Programming (ASP) \cite{Baral03},  which turns out to be effective for these enumerative searches.
+}
+
+%The outcome of this work is twofold.
+%The first benefit is that such an approach makes it possible to refine the construction of
+%BRNs with a partial and progressively brought knowledge in PH, while being able to export such
+%models in the Thomas' framework.
+%This work thus strengthens the formal link between both modelings.
+%The second feature of our method is that it can be applied on very large BRNs.
 
@@
 Furthermore, we claim that we are able to deal with larger biological networks.
+Finally, it must be noticed that we are not interested in this paper in the derivation of one PH
+from a BRN (which was previously described in \cite{PMR10-TCSB}) but, on the contrary, to finding
+out a set of BRNs from one PH.
 
-\todo{Emphasis on the new contribution since CMSB.}
+The contributions presented in this paper significantly extends and improves the preliminary results
+presented in \cite{FPIMR12-CMSB}.
+In addition of improving the efficiency and accuracy of the IG inference, we have added support for
+unsigned regulations, that are regulations being positive or negative depending on a particular
+context.
+This allows to apply our methodology to a wider class of models.
+\todo{asp improvement? more detailed application?}
 
@@
 and discuss its implementation in ASP.
+\todo{add ref to ASP section}
 \pref{sec:examples} illustrates the applicability of our method on simple examples

### 2013-07-05 parts/ph.tex
avancées sur l'intro et l'abstract

@@
 \paragraph{Modeling cooperation}
-\rewrite{Detail the modeling of cooperations?}
+%\rewrite{Detail the modeling of cooperations?}
 As described in \cite{PMR10-TCSB}, the cooperation between several processes to make another process bounce can be
@@
 
-\rewrite{Explain the efficiency of PH for reachability?}
+%\rewrite{Explain the efficiency of PH for reachability?}

### 2013-07-05 parts/discussion.tex
several TODOs
classical interaction graph inference

@@
 
-\paragraph{Ack}
-This work was partially supported by the Fondation Centrale Initiatives.
+\paragraph{Acknowledgement}
+This work was partially supported by the Fondation Centrale Initiatives and
+the French National Agency for Research (ANR-10-BLANC-0218 BioTempo project).
+The work of LP was supported by the Swiss SystemsX.ch project.

### 2013-07-05 parts/grn.tex
several TODOs
classical interaction graph inference

@@
+% vim:spell spelllang=en
 \subsection{Thomas' modeling}
@@
 \end{definition}
+\rewrite{Discuss on the widespread use of discrete parameters + binary resources}
 An IG and a parametrization make up a complete BRN.
@@
 \begin{align*}
-  \forall s, s' \in \GRNstates, \GRNtrans{s}{s'} &\Longleftrightarrow \exists a \in \Gamma, \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + f^a(s) \\
+  \forall s, s' \in \GRNstates, \GRNtrans{s}{s'} &\Longleftrightarrow \exists a \in \Gamma,
+  \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + \delta^a(s) \\
     & \qquad\quad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
 \end{align*}
-with: $f^a(s) = 
+with: $\delta^a(s) = 
   \begin{cases}
@@
 \end{example}
+
+\updated{
+\paragraph{Interaction Graph of the dynamics}
+From the dynamics specification of a BRN, one can infer back the IG that contain only the functional
+regulations.
+Let us define
+$f^a(s) = s[a]$ if $s[a]\in K_{a, \GRNres{a}{s}}$, 
+$f^a(s) = s[a]+1$ if $s[a] < K_{a, \GRNres{a}{s}}$, 
+and 
+$f^a(s) = s[a]-1$ if $s[a] > K_{a, \GRNres{a}{s}}$.
+A positive (resp. negative) interaction from $b$ to $a$ is inferred if there exists a state $s$ such
+that increasing $b$ in $s$ would increase (resp. decrease) $f^a$:
+$f^a(s\{b_i\}) < \text{ (resp. $>$) } f^a(s\{b_{i+1}\})$, 
+where $i < l_b$ and $s\{b_i\}$ denotes the state $s$ where the component $b$ is assigned to $b_i$.
+Then, such an IG can be used to derives global properties on the dynamics
+(e.g., \cite{Richard2010378,PR11-SASB}).
+}
 
@@
 First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists
-a positive edge from $a$ to $a$ in the interaction graph.
-\todo{be sure to have introduced the inference of IG from discrete maps}
+a positive edge from $a$ to $a$ in the interaction graph of the dynamics.
 Then, let us consider the case where $K_{a,\omega} = \segm{i}{i+2}$ (interval of three values), and

### 2013-07-05 parts/intro.tex
classical interaction graph inference

@@
 \updated{%
-The aim of this paper is to formally estiblish links between several layers suited for specifying the
+The aim of this paper is to formally establish links between several layers suited for specifying the
 qualitative dynamics of BRNs, namely:
 the Interaction Graph (IG), referencing the sign of direct regulations between components;
-the Process Hitting (PH), which permits to specify the qualitatif dynamics of the system with
+the Process Hitting (PH), which permits to specify the qualitative dynamics of the system with
 various degrees of knowledge for joint regulations;
@@
 of the system dynamics (see \cite{PR11-SASB} for a short survey).
-The most known are the Thomas' conjectures (that have been proved since \todo{refs}),
+The most known are the Thomas' conjectures (that have been proved since, e.g.
+\cite{RRT08,RiCo07,Richard2010378}
+for Boolean and discrete networks),
 which relate the absence of
@@
 }
+\todo{describe application}
 

### 2013-07-05 parts/infer-ig.tex
typos + suppr urls from bib

@@
 the formal system specification.
-However, it is very common that the prior IG actually refers interaction that reveal to be 
+However, it is common that the prior IG actually refers to interactions that reveal to be 
 non-effective with respect to the dynamics.
@@
 
-\todo{remark for reducing the number of self-regulation}
-
 We are now able to infer the edges of the final IG by considering positive and negative influences

### 2013-07-05 parts/example.tex
Add TODOs and REWRITEs
begin example

@@
 \section{Examples}\label{sec:examples}
+
+\rewrite{Reformulate / glue.}
+
+\subsection{Implementation}
+
+The inference method described in this paper has been implemented as part of
+\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related
+tools.
+Our implementation mainly consists in ASP programs that are solved using Clingo\footnote{Available
+at \url{http://potassco.sourceforge.net}}.
+The IG and parameters inference can be performed using the command
+\texttt{ph2thomas -i model.ph -{}-dot ig.dot}
+where \texttt{model.ph} is the PH model in \textsc{Pint} format and \texttt{ig.dot} is an output of the inferred IG in DOT format.
+The (possibly partial) inferred parametrization will be returned on the standard output.
+The admissible parametrizations enumeration is performed when adding the \texttt{-{}-enumerate}
+parameter to the command.
+
+\todo{Check this paragraph}
+Applied to the example in \pref{fig:runningPH-2} where cooperations have been defined,
+our method infers the IG and parametrization given in \pref{fig:runningBRN}.
+Regarding the example in \pref{fig:runningPH-1}, the same IG is inferred, as well as for the
+parametrization except for the parameters $K_{a,\{b\},\{c\}}$ and $K_{a,\{c\},\{b\}}$ which are
+undefined (because of the lack of cooperativity between $b$ and $c$).
+In such a case, this partial parametrization allows 36 admissible complete parametrizations, as two
+parameters with 3 potential values could not be inferred.
+If we constrain these latter parameters so that they contain exactly one element, we obtain only 9
+admissible parametrizations.
+
+\subsection{Biological application}
+
+The system studied is extracted from a model of ERBB receptor-regulated G1/S transition detailed in~\cite{Sahin09}.
+This model represents a chain of reactions that lead to the activation of protein pRB which is responsible for regulating the cell division,
+and is therefore essential to prevent cancer development.
+Its IG is composed of 20 components and 52 edges.
+
+Three models are created from the original IG, with different levels of precision regarding the cooperation between components.
+Model (1) represents a translation of the raw IG into Process Hitting, that is,
+without any knowledge of the Boolean rules (and therefore the cooperations) of the components.
+Model (2) implements some of the rules based on the results of several knockdown experiments.
+Given the results of the experiments, the introduced cooperative sorts to model the most important cooperations,
+which were emphasized by the impossibility to reach the production of pRB without the knocked protein.
+Model (3) is the totally refined model with all cooperations implemented given the Boolean functions for all components.
+Therefore, those three models can be considered as successive refinements of the original and most general one.
+The results of the IG and parameters inferences are detailed in \pref{tb:egfr20}
+and discussed in the following alongside with details about their construction.
+
+\begin{table}[ht]
+~\hfill%\begin{center}
+  \begin{tabular}{r|l|l|m{2cm}|m{2.5cm}|m{1.5cm}}
+    \textbf{Model} & \textbf{$|E|$} & \textbf{$|K|$} & \textbf{Inferred parameters} & \textbf{Possible\newline models} & \textbf{Fixed points}
+  \\\hline\hline
+    (1) & $52$ & $196$ & $20$ & $2^{176}\simeq 9.6\cdot10^{52}$ & $0$   % v1_0.ph
+  \\\hline
+    (2) & $51$ & $192$ & $98$ & $2^{94}\simeq 2.0\cdot10^{28}$ & $0$    % v2_1.ph
+  \\\hline
+    (3) & $51$ & $192$ & $192$ & $1$ & $3$                              % ori.ph
+  \\\hline
+  \end{tabular}
+\hfill~%\end{center}
+  \caption{Results of the IG and parameters inference on three models
+  derived the ERBB receptor-regulated G1/S transition model of~\cite{Sahin09}
+  with different precisions in the definition of the cooperations.
+  Model (1) contains no cooperations between the components.
+  Several cooperations were included in model (2) under the form of 14 cooperative sorts.
+  Model (3) models all the Boolean functions between components using 22 cooperative sorts.}
+  \label{tb:egfr20}
+\end{table}
+
+Model (1) encompasses only sole interactions between components, that is,
+independent activations or inhibitions of a component on another given the regulations specified in the original IG.
+Therefore, the IG inferred from model (1) is exactly the IG ised to create the model, and the inferred parameters
+are the parameters for the extreme cases (all activators present and all inhibitors absent).
+This first model therefore abstracts a large number of Thomas models as a lot of parameters are left undecided.
+
+In order to build model (2), 14 cooperative sorts were added in order to model the Boolean functions of several components
+(consisting of AND and OR operands).
+To do so, the following components were noticed due to their importance in the chain of reactions:
+CDK4, CDK6, CycD1, ER$\alpha$ and c-MYC.
+Indeed, knockdown experiments have been conducted in~\cite{Sahin09}
+and the results showed that knocking down these components lead to an important decrease in the production of pRB.
+We therefore concluded that these components were involved in other components' Boolean functions
+in a way that their knockdown was sufficient to prevent its production (which is typical of AND operands).
+In order to reproduce such requirements, the boolean functions of their successors,
+that is CDK4, CDK6, prB, p21, p27, IGF1R, MYC, CycD1 and CycE1,
+were modeled as cooperative sorts, if needed.
+We note that the added cooperations allowed to infer about half the parameters.
+However, the number of possible Thomas models that can be inferred from this PH is still significant
+as the remaining unknown parameters leave a big number of possibilities.
+Furthermore, the inferred IG contains one edge less than the original IG. This is due to the fact that
+one of the Boolean function could in fact be simplified in a way that one of the components did not appear anymore.
+No edge have therefore been inferred in this case.
+
+Finally, model (3) was build using all the Boolean functions provided in~\cite{Sahin09}.
+These functions take the form of 22 cooperative sorts into the model in order to match the desired behavior of the system.
+As all cooperations are fully defined in this model, all the parameters are inferred and only one Thomas model can be derived.
+We note also that this model is the only one containing at least one fixed point.
+
+\subsection{Computation times and results}
+
+The current implementation can successfully handle large PH models of BRNs found in the literature
+such as the ERBB receptor-regulated G1/S transition model from~\cite{Sahin09} presented in the last subsection,
+and a T cell receptor model from~\cite{Klamt06} which contains 40 components and 14 cooperative
+sorts\footnote{All models mentioned in this section are available as examples distributed with \textsc{Pint}.}.
+For each model, IG and parameters inferences are performed together in less than a second
+on a standard desktop computer.
+
+Bigger models related to the aforementioned systems were also tested with our implementation:
+an other T cell receptor model described in~\cite{SaezRodriguez2007} containing 94 components
+and an EGFR/ErbB signaling model described in~\cite{Samaga2009} which containing 104 components.
+\todo{Mention how these models were created?}
+
+The composition of all models and the results of the inferences are summarized in \pref{tb:computation}.
+We note that due to a very high number of parameters, no parameters inference could be performed on the $94$ and $104$ component models.
+Therefore, these models would therefore be more efficiently studied as PH than as BRNs.
+finally, we note that the complexity of the method is exponential in the number of regulators of one
+component and linear in the number of components.
+
+\begin{table}
+\begin{center}
+  \begin{tabular}{r|c|c|c|c}
+    \textbf{Model} & \textbf{Components} & \textbf{\# CS} & \textbf{$\Delta t$ IG} & \textbf{$\Delta t$ K}
+\\\hline\hline
+    ERBB-G1/S & $20$ & $22$ & <1s & <1s
+\\\hline
+    T cell receptor & $40$ & $14$ & <1s & <1s
+\\\hline
+    T cell receptor & $94$ & $39$ & 10s & 
+\\\hline
+    EGFR/ErbB signaling & $104$ & $89$ & 3min 20s & 
+  \end{tabular}
+\end{center}
+  \caption{Other models \rewrite{Caption}}
+  \label{tb:computation}
+\end{table}
+
+\begin{comment}
+\subsection{\warnul{Old version of this section}}
 
@@
 and some information is likely to help in the representation of cooperations.
+\end{comment}

### 2013-07-08 parts/infer-param.tex
TODO on well-formed PH for parameters inference

@@
 In addition, we introduce the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
-which implies that the inferred IG does not contain any unsigned interactions, and thus can be seen as the
-regular IG $(\Gamma, E)$,
-and that any processes in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
+which implies that %the inferred IG does not contain any unsigned interactions, and thus can be seen as the
+%regular IG $(\Gamma, E)$,
+%and that 
+any processes in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
 regarding $a$.
 
+\todo{Revoir !}
 \begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}

### 2013-07-09 parts/asp.tex
add running example

@@
 \begin{align*}
-  &\{&\componentlevels(b, 0)
-  &&\componentlevels(b, 1) \\
-  &&\componentlevels(c, 0)
-  &&\componentlevels(c, 1) \\
-  &&\componentlevels(a, 0)
-  &&\componentlevels(a, 1) \\
-  &&\componentlevels(a, 2) &&&&\}
+  &\{&&\componentlevels(b, 0),
+  &&\componentlevels(b, 1), \\
+  &&&\componentlevels(c, 0),
+  &&\componentlevels(c, 1), \\
+  &&&\componentlevels(a, 0),
+  &&\componentlevels(a, 1), \\
+  &&&\componentlevels(a, 2) &&&&\}
 \end{align*}
@@
 \begin{example}
-In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{b\}}$ and $K_{a,\{c\}}$ could not be inferred by \pref{pps:param_K}.
-The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values
-(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
+\towrite{Check this example.}
+In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred by \pref{pps:param_K}.
+A cardinality rule allows to produce 36 parametrizations, in which these parameters can take all possible values:
+%The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
+%(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
 \begin{align*}
-  (K_{a,\{b\}} ; K_{a,\{c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
+  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
 \end{align*}
-We note that $(\segm{1}{1} ; \segm{1}{1})$ belongs to the enumerated set of admissible parametrizations.
-Therefore, this enumeration allows, from the model in \pref{fig:runningPH-1}, to find the behavior of the model completed with a cooperative sort described in \pref{fig:runningPH-2}.
+and all the other parameters keep their inferred values.
+These parametrizations take the form of sets of $\paraminf$ atoms.
+
+Then, due to the fact that $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ were already inferred,
+the monotonicity assumption removes all parametrizations with: $0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
+In the end, the remaining possible values for the parameters that were not inferred are:
+\begin{align*}
+  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{1}{1} , \segm{2}{2} , \segm{1}{2} \}^2
+\end{align*}
+We thus obtain the results of \pref{ex:enum-param-runningPH-1}.
 \end{example}

### 2013-07-09 parts/infer-ig.tex
add running example

@@
 Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
-%Furthermore, studying the influences of the groups of regulators of $a$
-%allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
-%Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
+Furthermore, studying the influences of the groups of regulators of $a$
+allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
+Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
 We ignore the cases where a component has no visible influence on another.
@@
 \begin{split}\label{eq:edges-inference}
-  \forall b\in\PHpredecgene{a} \cup \{ a \}, \forall s \in \{ +, - \}, \\
+%  \forall b\in\PHpredecgene{a} \cup \{ a \}, \forall s \in \{ +, - \}, \\
+  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
   b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
@@
 %
-%\begin{split}\label{eq:edges-inference-auto}
-%  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-%  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
-%  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-%    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-%    &\qquad s = \f{sign}(a_k - a_j)
-%\end{split}\\
-%\begin{split}\label{eq:edges-inference-noreg}
-%  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-%  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
-%    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
-%    &\qquad s = \f{sign}(a_k - a_j)
-%\end{split}
+\begin{split}\label{eq:edges-inference-auto}
+  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
+  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+    &\qquad s = \f{sign}(a_k - a_j)
+\end{split}\\
+\begin{split}\label{eq:edges-inference-noreg}
+  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
+    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
+    &\qquad s = \f{sign}(a_k - a_j)
+\end{split}
 %
@@
 \begin{example}
-The IG inference from the PH of \pref{fig:runningPH-2} gives
-$\hat{E}_+ = \{b \xrightarrow{1} a, c \xrightarrow{1} a\}$ and 
-$\hat{E}_- = \{a \xrightarrow{2} b\}$, corresponding to the IG of \pref{fig:runningBRN}.
-No self-influence are inferred. %($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
-\rewrite{Check this example}
+The IG inference of the PH of \pref{fig:runningPH-1} gives the
+IG in \pref{fig:BRN-inf1}, containing the following edges:
+\begin{align*}
+  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{b}{+}{1}{b}, \GRNedgef{c}{+}{1}{c}\}\\
+  E_- &= \{\GRNedgef{a}{-}{2}{b}\} \qquad\qquad\qquad\qquad\qquad
+  E_\uns = \emptyset
+\end{align*}
+This IG is close to the one in \pref{fig:runningBRN} but not equivalent,
+as each component has an additional auto-action.
+The auto-actions on $b$ and $c$ are the consequence of a global stability
+in some configurations: $c$ never evolves, and neither does $b$ when $a_2$ is not active.
+The auto-action on $a$ is mainly caused by its multi-valued nature.
+
+The inference of the refined PH of \pref{fig:runningPH-2} gives the same IG.
+
+\begin{figure}[t]
+\centering
+\scalebox{1.2}{
+\begin{tikzpicture}[grn]
+  \path[use as bounding box] (-1.3,-0.75) rectangle (3.5,1.5);
+  \node[inner sep=0] (a) at (2,0) {a};
+  \node[inner sep=0] (b) at (0,0) {b};
+  \node[inner sep=0] (c) at (2,1.2) {c};
+  \path[->]
+    (b) edge[bend right] node[elabel, below=-2pt] {$+1$} (a)
+    (c) edge node[elabel, right=-2pt] {$+1$} (a)
+    (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b)
+    (b) edge[in=-15+180, out=15+180, loop] node[elabel, left=-2pt] {+1} (b)
+    (c) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (c)
+    (a) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (a);
+\end{tikzpicture}
+}
+\caption{\label{fig:BRN-inf1}
+  Result of the IG inference performed on the PH of \pref{fig:runningPH-2}.
+}
+\end{figure}
+
 \end{example}

### 2013-07-09 parts/infer-param.tex
fix property 3: well-formed for parameters inference
add running example

@@
 %and that 
-any processes in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
+any process in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
 regarding $a$.
 
-\todo{Revoir !}
 \begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
@@
   \begin{split}
-  \forall a \in \Gamma, \forall b\in \GRNreg{a}&,
+  \forall a \in \Gamma&, \forall b\in \GRNreg{a},
           \forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
-  & \quad \forall c \in \Gamma, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a} \wedge b\in\PHdirectpredec{c})), \\
+  & \quad \forall c \in \PHs, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a}\setminus \Gamma \wedge b\in\PHdirectpredec{c})), \\
   & \qquad
@@
 \label{ex:infer-param-runningPH-1}
-Applied to the PH in \pref{fig:runningPH-1}, we obtain, in particular,
-$K_{b,\emptyset} = \segm{0}{1}$ and
-$K_{a,\{b,c\}} = \segm{2}{2}$,
-while $K_{a,\{b\}}$ and $K_{a,\{c\}}$ can not be inferred.
+
+Applied to the PH in \pref{fig:runningPH-1}, we infer the following parameters:
+\begin{align*}
+K_{a, \emptyset} &= \segm{0}{0} &
+K_{b, \emptyset} &= \segm{0}{0} \\
+K_{a, \{a\}} &= \segm{0}{0} &
+K_{b, \{a\}} &= \segm{0}{0} \\
+K_{a, \{c\}} &= \segm{1}{1} &
+K_{b, \{b\}} &= \segm{1}{1} \\
+K_{a, \{b\}} &= \segm{1}{1} &
+K_{b, \{a,b\}} &= \segm{0}{0} \\
+K_{a, \{b,c\}} &= \segm{1}{1} &
+K_{c, \emptyset} &= \segm{0}{0} \\
+K_{a, \{a,b,c\}} &= \segm{2}{2} &
+K_{c, \{c\}} &= \segm{1}{1}
+%K_{a, \{a,b\}} = ?
+%K_{a, \{a,c\}} = ?
+\end{align*}
+$K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ cannot be inferred,
+which is a direct consequence of the lack of cooperation between $b$ and $c$ on $a$.
 \end{example}
@@
 \begin{example}
-Regarding the PH in \pref{fig:runningPH-2}, we obtain
-$K_{b,\emptyset} = \segm{0}{1}$,
-$K_{a,\{b,c\}} = \segm{2}{2}$,
-$K_{a,\{b\}} = \segm{1}{1}$,
-and $K_{a,\{c\}} = \segm{1}{1}$.
-For this PH, all parameters can be inferred, and the obtained parametrization is the one of \pref{fig:runningBRN}(right).
+Regarding the refined PH of \pref{fig:runningPH-2}, all parameters can be inferred.
+We obtain the same value for the inferred parameters of \pref{ex:infer-param-runningPH-1},
+together with the following results:
+\begin{align*}
+  K_{a,\{a,b\}} &= 1 &
+  K_{a,\{a,c\}} &= 1
+\end{align*}
 \end{example}
@@
 \end{property}
+
+\begin{example}\label{ex:enum-param-runningPH-1}
+The parametrization inferred in \pref{ex:infer-param-runningPH-1} was partial because $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred.
+It is however possible to enumerate all complete and admissible parametrizations
+compatible with both the inferred parameters, and the properties of this subsection.
+This enumeration gives 9 different parametrizations which correspond to the 3 possible values
+for each of the two parameters that could not be inferred:
+\begin{align*}
+  K_{a,\{a,b\}} &\in \{ \segm{1}{1}, \segm{1}{2}, \segm{2}{2} \} \\
+  K_{a,\{a,c\}} &\in \{ \segm{1}{1}, \segm{1}{2}, \segm{2}{2} \}
+\end{align*}
+We note that for all solutions, $0 \notin K_{a,\{a,b\}} \wedge 0 \notin K_{a,\{a,c\}}$.
+This is due to the monotonicity assumption (\pref{pro:param_enum_monotonicity}) which especially states that:
+\begin{align*}
+  K_{a,\{b\}} \leqsegm K_{a,\{a,b\}} \wedge
+  K_{a,\{c\}} \leqsegm K_{a,\{a,c\}}
+\end{align*}
+
+Finally, we note that $\segm{1}{1}$ belongs to the possible values for both parameters.
+Therefore this enumeration allows, from the model in \pref{fig:runningPH-1},
+to find the behavior of the model refined with a cooperative sort described in \pref{fig:runningPH-2}.
+\end{example}

### 2013-07-09 parts/infer-ig.tex
add example with unsigned edge

@@
 
+\begin{example}
+If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
+then two unsigned edges towards $b$ are inferred instead of the previous signed edges:
+\begin{align*}
+  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{c}{+}{1}{c}\}\\
+  E_- &= \emptyset \qquad\qquad\qquad\qquad
+  E_\uns = \{\GRNedgef{a}{\uns}{2}{b}, \GRNedgef{b}{\uns}{1}{b}\}
+\end{align*}
+This is due to the fact that the actions $\PHfrappe{a_2}{b_1}{b_0}$ and $\PHfrappe{a_2}{b_0}{b_1}$
+introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
+\end{example}
+
 \begin{comment}

### 2013-07-09 parts/intro.tex
check TODOs in Introduction

@@
 \section{Introduction}\label{sec:intro}
-
-\rewrite{Include Olivier's additions}
 
@@
 hereafter called ``Thomas' parameters'',
-which stand for kinds of ``focal points'', i.e. the evolutionary tendency from each state and depending on the set of the other currently interacting components.
+which stand for kinds of ``focal points'', \ie the evolutionary tendency from each state and depending on the set of the other currently interacting components.
 %resources in this very state, that is, the set of the other currently interacting components.
@@
 of the system dynamics (see \cite{PR11-SASB} for a short survey).
-The most known are the Thomas' conjectures (that have been proved since, e.g.
-\cite{RRT08,RiCo07,Richard2010378}
+The most known are the Thomas' conjectures (that have been proved since,
+e.g.~\cite{RRT08,RiCo07,Richard2010378}
 for Boolean and discrete networks),
@@
 Programming (ASP) \cite{Baral03},  which turns out to be effective for these enumerative searches.
+}%
+\updatedmax{%
+The whole method is also applied to the biological model of ERBB receptor-regulated G1/S transition,
+containing 20 components,
+which allows to detail how joint actions can be defined inside a PH model in order
+to refine its dynamics, and and study a more precise class of underlying Thomas models.
 }
-\todo{describe application}
+%\todo{describe application}
 
@@
 
-The contributions presented in this paper significantly extends and improves the preliminary results
+\updated{%
+The contributions presented in this paper significantly extend and improve the preliminary results
 presented in \cite{FPIMR12-CMSB}.
@@
 This allows to apply our methodology to a wider class of models.
-\todo{asp improvement? more detailed application?}
+}%
+\updatedmax{%
+Furthermore, some parts have been completely rewritten in order to explain more precisely the efficiency of the ASP implementation,
+and give a new detailed biological application of our work.
+}%
 
@@
 \pref{sec:infer-IG} defines the IG inference from PH;
-\pref{sec:infer-K} details the enumeration of Thomas parametrizations compatible with a PH
-and discuss its implementation in ASP.
-\todo{add ref to ASP section}
-\pref{sec:examples} illustrates the applicability of our method on simple examples
-and large biological models.
+\pref{sec:infer-K} details the parameters inference and the enumeration of Thomas parametrizations compatible with a PH;
+\pref{sec:impl} discusses the implementation of the latter in ASP.
+\pref{sec:examples} illustrates our method on a biological model
+and discusses its applicability on large models.
 

### 2013-07-09 parts/grn.tex
add running example
check TODOs in Frameworks

@@
 We briefly discuss these additions.
+
+\medskip
 
@@
 \end{align*}
-\todo{Find an example with an unsigned edge?}
 \end{example}
@@
   \node[inner sep=0] (c) at (2,1.2) {c};
-%\path
-%  node[elabel, below=-1em of a] {$0..2$}
-%  node[elabel, below=-1em of b] {$0..1$}
-%  node[elabel, above=-.8em of c] {$0..1$};
   \path[->]
@@
   Example parametrization on the left IG.
-\todo{Find an example with an unsigned edge?}
 }
@@
 $\GRNget{s}{a}$ refers to the level of component $a$ in $s$.
-%\rewrite{Functional definition of a state? (More rigorous)}
 For each possible state, the set of \emph{resources} of a given component
@@
 where $i < l_b$ and $s\{b_i\}$ denotes the state $s$ where the component $b$ is assigned to $b_i$.
-Then, such an IG can be used to derives global properties on the dynamics
+Then, such an IG can be used to derive global properties on the dynamics
 (e.g., \cite{Richard2010378,PR11-SASB}).
@@
 These interval parameters are mainly introduced for convenience, and it is worth noticing that they
-do not add extra expressivity for modeling multivalued networks.
+do not add extra expressivity for modeling multivalued networks in the general case.
 Although, interval parameters are more expressive than single value parameters for multivalued networks.
+\rewrite{“However, interval parameters are more expressive than single value parameters for Thomas models.”?}
 First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists

### 2013-07-09 parts/ph.tex
check TODOs in Frameworks

--- 2013-04-26 parts/ph.tex
@@
 In this BRN, $b$ and $c$ activate $a$, while $a$ inhibits $b$.
-The inhibition of $b$ by $a$ is only effective when $a$ is at level $2$;
+The inhibition of $b$ by $a$ is only effective when $a$ is at level $2$ (noted: $a_2$);
 in the other cases, $b$ cannot evolve in any direction.
@@
 \paragraph{Modeling cooperation}
-\rewrite{Detail the modeling of cooperations?}
 As described in \cite{PMR10-TCSB}, the cooperation between several processes to make another process bounce can be
@@
 \end{example}
-
-\rewrite{Explain the efficiency of PH for reachability?}

### 2013-07-09 parts/infer-ig.tex
proofread and check TODOs in IG and Parametrization Inference

@@
 The Interaction Graph (IG) is an abstract representation of the direct qualitative influences,
-positives and/or negatives, between the components of the system.
+positive and/or negative, between the components of the system.
 As discussed in \pref{sec:intro}, the IG allows to efficiently characterise global dynamical
@@
 modelling.}
-
+%
 %In order to infer a complete BRN, one has to find the Interaction Graph (IG) first, as some
@@
 %components on each of its successors.
-
-This section first introduces the notion of focal processes within a PH
-(\pref{ssec:focal}) which is used to characterize well-formed PH for IG inference
-in \pref{ssec:wf}, and as well used by the parametrization inference presented in \pref{sec:infer-K}.
-Finally, the rules for inferring the interactions between components from a PH are
+%
+%
+This section first introduces the notion of focal processes within a PH (\pref{ssec:focal})
+which is used to characterize well-formed PH for IG inference (\pref{ssec:wf}).
+%and as well used by the parametrization inference presented in \pref{sec:infer-K}.
+The rules for inferring the interactions between components from a PH are
 described in \pref{ssec:infer-IG}.

### 2013-07-09 parts/infer-param.tex
proofread and check TODOs in IG and Parametrization Inference

@@
 
-Given the IG inferred from a PH as presented in the previous section, one can find the discrete parameters that model the behavior of the studied PH using the method presented in the following.
-It relies on an exhaustive enumeration of all predecessors of each component in order to find attractor processes and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations.
-The last step consists of the enumeration of all compatible complete parametrizations given this
+Given the IG inferred from a PH as presented in the previous section, one can find the discrete parameters
+that model the behavior of the studied PH using the method presented in the following.
+It relies on an exhaustive enumeration of all predecessors of each component in order to find attractor processes
+and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations (\pref{ssec:infer-K}).
+The last step consists of the enumeration of all compatible complete parametrizations (\pref{ssec:admissible-K}) given this
 set of inferred parameters, the PH dynamics and some biological constraints on parameters.
 
-\subsection{Parameters inference}
+\subsection{Parameters inference}\label{ssec:infer-K}
 

### 2013-07-09 parts/asp.tex
check TODOs in Introduction
proofread and check TODOs in ASP Implementation

@@
-\section{Answer Set Programming implementation concepts}
+\section{Answer Set Programming implementation concepts}\label{sec:impl}
 
@@
   \item ASP efficiently tackles the inherent complexity of the models used, thus allowing a fast execution of the formal tools defined in this paper,
-  \item ASP is convenient to enumerate a large set of possible answers,
+  \item it is convenient to enumerate a large set of possible answers,
   \item and it allows to easily constrain the answers according to some properties.
@@
 is composed of the predicate $\predicate{p}$ and $r$ arguments: $x_1, x_2, \dots, x_r$.
-Each argument can be either an value or a variable,
-a value being a representation of a piece of data (component name, expression level, …)
-and a variable can is used to represent any existing value which respects the rules.
-In this paper, a variable is always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, …)
-while values are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …).
+Each argument can be either an constant or a variable, %value -> constant
+a constant being a representation of a piece of data (component name, expression level, …)
+and a variable is used to represent any possible existing constant which respects the rules.
+In this paper, a variable is always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, $\var{Q}$, …)
+while constants are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …).
 
@@
 These three atoms are built with the predicate $\component$.
-Furthermore, $a$, $b$, $c$, $1$ and $2$ are values.
+Furthermore, $a$, $b$, $c$, $1$ and $2$ are constants.
 \end{example}
@@
 one can use atoms of the form $\componentlevels(a, k)$ to state that $k \in \segm{0}{l_a}$.
-Variables here come in handy to enumerate each possible value $k$ for each component $a$:
-during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible values it could represent.
+Variables here come in handy to enumerate each possible constant $k$ for each component $a$:
+during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible constants it could represent.
 The following rule, for example, contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$) and enumerates the set of possible expression levels of each component in the system:
@@
 \end{align*}
-Given such a rule, as many answer sets as possible are created, so that each answer $S$ set verifies:
+Given such a rule, as many answer sets as possible are created, so that each answer set $S$ verifies:
 \begin{align*}
@@
 In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
-The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} : \atom{Q}$,
+The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} \mid \atom{Q}$,
 which is a shorthand for “the set of atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
-\rewrite{Reformulate?}
+%\rewrite{Reformulate?}
 
@@
 \begin{align*}
-  & 1\ \{\ \param(\var{A}, \var{P}, \var{I}) : \componentlevels(\var{A}, \var{I})\ \}\ \infty\ \la \\
+  & 1\ \{\ \param(\var{A}, \var{P}, \var{I}) \mid \componentlevels(\var{A}, \var{I})\ \}\ \infty\ \la \\
   & \qquad\qquad\qquad \paramlabel(\var{A}, \var{P}), \nota \inferedparam(\var{A}, \var{P}).
@@
 Furthermore, the lower bound is $1$, which forces each enumerated parameter to contain at least one value,
-but no upper bound is specified ($\infty$) for the size of each parameter (which is bounded by the number of possible expression levels of each gene).
+but no upper bound is specified ($\infty$) for the size of each parameter
+(which is already bounded by the number of possible expression levels of the related component).
 %each parameter contains at least $1$ value, as stated by the lower bound, and has no upper bound ($\infty$) but the number of expression levels of its component.
@@
 
+\begin{example}\label{ex:cardinality}
+In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred by \pref{pps:param_K}.
+The previous cardinality rule allows to produce 49 parametrizations, in which these two parameters can take all possible values:
+%The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
+%(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
+\begin{align*}
+  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2}, \{ 0, 2 \} \}^2
+\end{align*}
+and all the other parameters keep their inferred values.
+%These parametrizations take the form of sets of $\param$ atoms.
+Note that $\{ 0, 2 \}$ belongs to the set of candidate parametrizations
+as no rule specifying that a parameter has to be an interval has been defined yet.
+\end{example}
+
 \subsection{Constraints}\label{sssec:constraints}
@@
 In the scope of parameters enumeration, for example, constraints are especially useful to filter parametrizations
-that contain non-interval parameters or that do not respect the assumptions of \pref{ssec:admissible-K}.
+that contain non-interval parameters, or that do not respect the assumptions of \pref{ssec:admissible-K}.
 Indeed, suppose that:
@@
 \end{align*}
-Indeed, this constraint removes all parametrization results where parameters $K_{\var{A},\omega_\var{P}}$ and $K_{\var{Z},\omega_\var{Q}}$ exist
-such that $\var{A}$ is less activated by the set of resources $\omega_\var{P}$ than $\omega_\var{Q}$,
+Indeed, this constraint removes all parametrization results where parameters $K_{\var{A},\omega_\var{P}}$ and $K_{\var{A},\omega_\var{Q}}$ exist
+such that $\var{A}$ is less activated by the set of resources $\omega_\var{P}$ than it is by $\omega_\var{Q}$,
 but $K_{\var{A},\omega_\var{Q}} \ltsegm K_{\var{A},\omega_\var{P}}$,
@@
 \begin{example}
-\towrite{Check this example.}
-In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred by \pref{pps:param_K}.
-A cardinality rule allows to produce 36 parametrizations, in which these parameters can take all possible values:
-%The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
-%(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
+
+The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
+For example, a constraint was written in order to obtain only interval parameter
+(thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
+Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
 \begin{align*}
@@
 \end{align*}
-and all the other parameters keep their inferred values.
-These parametrizations take the form of sets of $\paraminf$ atoms.
 
-Then, due to the fact that $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ were already inferred,
-the monotonicity assumption removes all parametrizations with: $0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
+Then, due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
+the monotonicity assumption (\pref{pro:param_enum_monotonicity}) removes all parametrizations with: $0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
 In the end, the remaining possible values for the parameters that were not inferred are:
@@
 \end{align*}
+All these candidates respect the other properties of \pref{ssec:admissible-K} and no more candidates are filtered.
+%The other properties do not filter other values for these parameters;
 We thus obtain the results of \pref{ex:enum-param-runningPH-1}.
@@
 
-This subsection succinctly described how ASP programs come in handy to represent a model and solve all steps of Thomas' modeling inference.
-It finds a particularly interesting application in the enumeration of parameters: all possible parametrizations are generated in separate answer sets, and integrity constraints are formulated to remove those that do not fit the assumptions of admissible parametrizations,
-thus reducing the number of interesting parametrizations to be considered in the end.
+\medskip
+
+This subsection succinctly described how ASP programs come in handy to represent a model and solve complex problems on it.
+%all steps of Thomas' modeling inference.
+It finds a particularly interesting application in the enumeration of parameters:
+all possible parametrizations are generated in separate answer sets,
+and integrity constraints are formulated to remove those that do not fit the assumptions of admissible parametrizations,
+thus reducing the number of candidate parametrizations to be considered in the end.
+However, all steps of the inference presented in this paper (\pref{sec:infer-IG} \& \ref{sec:infer-K})
+were implemented in and benefited from this programming paradigm, although in significantly different ways.

### 2013-07-09 parts/example.tex
start proofreading Examples

@@
 
-\rewrite{Reformulate / glue.}
+\todo{Proofread this section}
 
-\subsection{Implementation}
+This section details the implementation and usage of the piece of software developed for this work.
+\pref{ssec:appli} gives a practical and detailed application of our results on a biological model taken from the literature,
+while \pref{ssec:cpu} gives data related to the computation times of our software when applied to several large biological models.
 
-The inference method described in this paper has been implemented as part of
-\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related
-tools.
-Our implementation mainly consists in ASP programs that are solved using Clingo\footnote{Available
-at \url{http://potassco.sourceforge.net}}.
-The IG and parameters inference can be performed using the command
-\texttt{ph2thomas -i model.ph -{}-dot ig.dot}
-where \texttt{model.ph} is the PH model in \textsc{Pint} format and \texttt{ig.dot} is an output of the inferred IG in DOT format.
+\medskip
+
+The inference methods described in this paper have been implemented as part of
+\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related tools.
+Our implementation mainly consists in ASP programs that are solved using
+Clingo\footnote{Available at \url{http://potassco.sourceforge.net}}.
+The IG and parameters inference can be performed using the following command:\\
+  \hspace*{1.2em}\texttt{ph2thomas -i model.ph -{}-dot ig.dot}\\
+where \texttt{model.ph} is the PH model in \textsc{Pint} format,
+and \texttt{ig.dot} is an output of the inferred IG in DOT format.
 The (possibly partial) inferred parametrization will be returned on the standard output.
-The admissible parametrizations enumeration is performed when adding the \texttt{-{}-enumerate}
-parameter to the command.
+The admissible parametrizations enumeration is performed when adding the \texttt{-{}-full-enumerate} parameter to the command,
+or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
 
+\begin{comment}
 \todo{Check this paragraph}
@@
 admissible parametrizations.
+\end{comment}
 
-\subsection{Biological application}
+\subsection{Biological application}\label{ssec:appli}
 
-The system studied is extracted from a model of ERBB receptor-regulated G1/S transition detailed in~\cite{Sahin09}.
-This model represents a chain of reactions that lead to the activation of protein pRB which is responsible for regulating the cell division,
-and is therefore essential to prevent cancer development.
-Its IG is composed of 20 components and 52 edges.
+This subsection focuses on the study of a model of ERBB receptor-regulated G1/S transition detailed in~\cite{Sahin09}.
+This model is represented by an IG containing 20 components and 52 edges.
+A protein named EGF, having no ingoing edge, can be considered as the only input.
+Then, a chain of reactions lead to the activation of protein pRB which is responsible for regulating the cell division,
+therefore making it essential to prevent cancer development.
 
 Three models are created from the original IG, with different levels of precision regarding the cooperation between components.
-Model (1) represents a translation of the raw IG into Process Hitting, that is,
+Model (1) represents a translation of the original IG into Process Hitting, that is,
 without any knowledge of the Boolean rules (and therefore the cooperations) of the components.
 Model (2) implements some of the rules based on the results of several knockdown experiments.
-Given the results of the experiments, the introduced cooperative sorts to model the most important cooperations,
-which were emphasized by the impossibility to reach the production of pRB without the knocked protein.
 Model (3) is the totally refined model with all cooperations implemented given the Boolean functions for all components.
@@
 
+\todo{Intégrer}
+Given the results of the experiments, the introduced cooperative sorts to model the most important cooperations,
+which were emphasized by the impossibility to reach the production of pRB without the knocked protein.
+
 In order to build model (2), 14 cooperative sorts were added in order to model the Boolean functions of several components
@@
 
-\subsection{Computation times and results}
+\subsection{Computation times and results}\label{ssec:cpu}
 

### 2013-07-09 parts/example.tex
finish proofreading of Examples

@@
 
+\updatedmax{%
 This section details the implementation and usage of the piece of software developed for this work.
@@
 while \pref{ssec:cpu} gives data related to the computation times of our software when applied to several large biological models.
+}
 
@@
 
+\updatedmax{%
 \subsection{Biological application}\label{ssec:appli}
@@
 independent activations or inhibitions of a component on another given the regulations specified in the original IG.
-Therefore, the IG inferred from model (1) is exactly the IG ised to create the model, and the inferred parameters
-are the parameters for the extreme cases (all activators present and all inhibitors absent).
+Therefore, the IG inferred from model (1) is the same as the IG used to create the model,
+with one additional positive auto-edge on the only input EGF (which is due to its absence of regulators).
+The only parameters that could be inferred are the parameters for the extreme cases
+(all activators present and all inhibitors absent, and the reverse).
 This first model therefore abstracts a large number of Thomas models as a lot of parameters are left undecided.
-
-\todo{Intégrer}
-Given the results of the experiments, the introduced cooperative sorts to model the most important cooperations,
-which were emphasized by the impossibility to reach the production of pRB without the knocked protein.
 
@@
 We therefore concluded that these components were involved in other components' Boolean functions
-in a way that their knockdown was sufficient to prevent its production (which is typical of AND operands).
-In order to reproduce such requirements, the boolean functions of their successors,
+in a way that the knockdown of the former was sufficient to prevent the production of the latter (which is typical of AND operands).
+In order to reproduce such requirements, the Boolean functions of their successors,
 that is CDK4, CDK6, prB, p21, p27, IGF1R, MYC, CycD1 and CycE1,
 were modeled as cooperative sorts, if needed.
+In theory, 9 cooperative sorts would have sufficed, but the chaining of cooperative sorts described
+in \pref{ssec:PH} was used to reduce the number of processes in each cooperative sort.
 We note that the added cooperations allowed to infer about half the parameters.
 However, the number of possible Thomas models that can be inferred from this PH is still significant
-as the remaining unknown parameters leave a big number of possibilities.
+because of the numerous remaining unknown parameters.
 Furthermore, the inferred IG contains one edge less than the original IG. This is due to the fact that
-one of the Boolean function could in fact be simplified in a way that one of the components did not appear anymore.
-No edge have therefore been inferred in this case.
+one of the Boolean functions could in fact be simplified in a way that this component did not appear anymore.
+No edge have therefore been inferred by our method in this case.
 
@@
 As all cooperations are fully defined in this model, all the parameters are inferred and only one Thomas model can be derived.
-We note also that this model is the only one containing at least one fixed point.
+We note also that this PH model is the only one containing at least one fixed point.
 
-\subsection{Computation times and results}\label{ssec:cpu}
+\subsection{Computation times on several large models}\label{ssec:cpu}
 
@@
 an other T cell receptor model described in~\cite{SaezRodriguez2007} containing 94 components
-and an EGFR/ErbB signaling model described in~\cite{Samaga2009} which containing 104 components.
+and an EGFR/ErbB signaling model described in~\cite{Samaga2009} containing 104 components.
 \todo{Mention how these models were created?}
@@
 We note that due to a very high number of parameters, no parameters inference could be performed on the $94$ and $104$ component models.
-Therefore, these models would therefore be more efficiently studied as PH than as BRNs.
-finally, we note that the complexity of the method is exponential in the number of regulators of one
+These models would therefore be more efficiently studied as PH than as BRNs.
+Finally, we note that the complexity of the method is exponential in the number of regulators of one
 component and linear in the number of components.
@@
 \\\hline\hline
-    ERBB-G1/S & $20$ & $22$ & <1s & <1s
+    ERBB-G1/S \cite{Sahin09} & $20$ & $22$ & <1s & <1s
 \\\hline
-    T cell receptor & $40$ & $14$ & <1s & <1s
+    T cell receptor \cite{Klamt06} & $40$ & $14$ & <1s & <1s
 \\\hline
-    T cell receptor & $94$ & $39$ & 10s & 
+    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & 10s & ---
 \\\hline
-    EGFR/ErbB signaling & $104$ & $89$ & 3min 20s & 
+    EGFR/ErbB signaling \cite{Samaga2009} & $104$ & $89$ & 3min 20s & ---
   \end{tabular}
 \end{center}
-  \caption{Other models \rewrite{Caption}}
+  \caption{Other models \rewrite{Caption}
+  Computation times and several pieces of data related to the IG and parametrization inferences of fours biological models.
+  The third column gives the number of cooperative sorts used in the models to model joint actions,
+  and the fourth (resp.~fifth) column gives the computation times of the IG inference (resp.~the parametrization inference).
+  }
   \label{tb:computation}
 \end{table}
+}%
 

### 2013-07-09 parts/discussion.tex
proofread Conclusion

@@
 \section{Conclusion and Discussion}
+
+\todo{Complete/extend}
 
@@
 The enumeration of such parametrizations is efficiently tackled using Answer Set Programming.
-We illustrate the overall method on simple examples and large biological models.
+We illustrate the overall method on several large biological models.
 
 Several extensions of the presented work are now to be considered.
-First, we now plan to explore the inference of Thomas' parameters when the inferred IG involves unsigned interactions, as some particular cases having such IGs are known to have correct
-Thomas' parametrization (not respecting some assumptions imposed in \pref{ssec:admissible-K},
-however).
-Second, the inference of BRN multiplexes \cite{BernotMultiplexes} may be of practical interest 
+First, the inference of BRN multiplexes \cite{BernotMultiplexes} may be of practical interest 
 as they allow to implicitly reduce the possible parametrizations by making cooperations appear
@@
 of compatible ones.
-Finally, in order to improve the performances in the IG inference, we will consider projection operations on
+Second, in order to improve the performances in the IG inference, we will consider projection operations on
 the PH structure to undo cooperations between components and reduce the cardinality of

### 2013-07-09 parts/grn.tex
TODOs

@@
 \end{definition}
-\rewrite{Discuss on the widespread use of discrete parameters + binary resources}
+\todo{Discuss on the widespread use of discrete parameters + binary resources}
 An IG and a parametrization make up a complete BRN.
@@
 Although, interval parameters are more expressive than single value parameters for multivalued networks.
-\rewrite{“However, interval parameters are more expressive than single value parameters for Thomas models.”?}
+\todo{“However, interval parameters are more expressive than single value parameters for Thomas models.”?}
 First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists

### 2013-07-18 parts/grn.tex
--todo

@@
 do not add extra expressivity for modeling multivalued networks in the general case.
-Although, interval parameters are more expressive than single value parameters for multivalued networks.
-\todo{“However, interval parameters are more expressive than single value parameters for Thomas models.”?}
+However, interval parameters are more expressive than Thomas' parameters with single value.
 First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists

### 2013-07-22 parts/intro.tex
--todo
Olivier's additions in Introduction

@@
 components.}
+\updatedmax{%
+Another motivation for the work depicted in this paper is that it constitutes an important step
+to go further in the study of large biological regulatory networks systems beyond simple analysis:
+it makes it possible to partly control some behaviors.
+Indeed, when one wants to modify some behaviors
+(e.g.~so as to avoid the reachability of some states or to create stable states)
+we intend to be able to get upstream the results in accordance at the regulatory level.
+The outcome of the present work is that we could make such changes on the PH model
+and then recover afterwards the BRN corresponding to the transformed behavioral system.
+In a more general framework and in an interesting way,
+we should be able to synthesize families of BRN having some featuring behavioral properties.
+}
 
@@
 }
-\todo{check if the inferred IG contains the IG of compatible Thomas}
 
@@
 which allows to detail how joint actions can be defined inside a PH model in order
-to refine its dynamics, and and study a more precise class of underlying Thomas models.
+to refine its dynamics, and study a more precise class of underlying Thomas models.
 }

### 2013-07-23 parts/example.tex
TODOs
cleanup

@@
 
-\todo{Proofread this section}
+%\todo{Proofread this section}
 
-\updatedmax{%
 This section details the implementation and usage of the piece of software developed for this work.
@@
 while \pref{ssec:cpu} gives data related to the computation times of our software when applied to several large biological models.
-}
 
@@
 
-\updatedmax{%
 \subsection{Biological application}\label{ssec:appli}
@@
 \end{center}
-  \caption{Other models \rewrite{Caption}
+\caption{
   Computation times and several pieces of data related to the IG and parametrization inferences of fours biological models.
@@
 \end{table}
-}%
-
-\begin{comment}
-\subsection{\warnul{Old version of this section}}
-
-The inference method described in this paper has been implemented as part of
-\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related
-tools.
-Our implementation mainly consists in ASP programs that are solved using Clingo\footnote{Available
-at \url{http://potassco.sourceforge.net}}.
-The IG and parameters inference can be performed using the command
-\texttt{ph2thomas -i model.ph -{}-dot ig.dot}
-where \texttt{model.ph} is the PH model in \textsc{Pint} format and \texttt{ig.dot} is an output of the inferred IG in DOT format.
-The (possibly partial) inferred parametrization will be returned on the standard output.
-The admissible parametrizations enumeration is performed when adding the \texttt{-{}-enumerate}
-parameter to the command.
-
-Applied to the example in \pref{fig:runningPH-2} where cooperations have been defined,
-our method infers the IG and parametrization given in \pref{fig:runningBRN}.
-Regarding the example in \pref{fig:runningPH-1}, the same IG is inferred, as well as for the
-parametrization except for the parameters $K_{a,\{b\},\{c\}}$ and $K_{a,\{c\},\{b\}}$ which are
-undefined (because of the lack of cooperativity between $b$ and $c$).
-In such a case, this partial parametrization allows 36 admissible complete parametrizations, as two
-parameters with 3 potential values could not be inferred.
-If we constrain these latter parameters so that they contain exactly one element, we obtain only 9
-admissible parametrizations.
-
-The current implementation can successfully handle large PH models of BRNs found in the literature
-such as an ERBB receptor-regulated G1/S transition model from \cite{Sahin09} which contains 20
-components and 15 cooperative sorts, and a T-cells receptor model from \cite{Klamt06} which contains 40
-components and 14 cooperative sorts\footnote{Both models are available as examples distributed with \textsc{Pint}.}.
-For each model, IG and parameters inferences are performed together in less than a second
-on a standard desktop computer.
-%\footnote{Using a Dell Inspiron 1720 laptop, with an Intel Core 2 Duo CPU T5550 ($2 \times 1.83\text{GHz}$)
-%and 3.9 Gib memory, on an Ubuntu 11.10 64-bits OS}
-After removing the cooperations from these models (leaving only raw actions), the inferences allow to
-determine 40 parameters out of 195 for the 20 components model, and 77 out of 143 for the 40 components model.
-As we thus have an order of magnitude of respectively $10^{31}$ and $10^{73}$ admissible parametrizations,
-these models would therefore be more efficiently studied as PH than as BRNs.
-We note that the complexity of the method is exponential in the number of regulators of one
-component and linear in the number of components.
-% We note that despite its smaller size in term of components, the ERBB transmission model takes more time to be computed because the biggest cooperative sorts contain more processes (up to 32 processes) than in the T-cells receptor model (up to 8 processes).
-
-A PH model can be built based on information found in the literature about the local influences between components.
-The precision of this knowledge will determine the precision of the modeled activations and inhibitions,
-and some information is likely to help in the representation of cooperations.
-\end{comment}

### 2013-07-23 parts/infer-ig.tex
cleanup

@@
 
-\updated{%
 The Interaction Graph (IG) is an abstract representation of the direct qualitative influences,
@@
 As discussed in \pref{sec:intro}, the IG allows to efficiently characterise global dynamical
-properties for the concrete system, such as the capability for multi-stationarity or oscillation.}
+properties for the concrete system, such as the capability for multi-stationarity or oscillation.
 
-\updated{%
 In a typical biological network modelling process, a prior IG is generally the starting point for
@@
 Hence, deriving the IG directly from the dynamical models lead to more concise IGs, enhancing the
-conclusiveness of static analyses based upon this abstract representation.}
+conclusiveness of static analyses based upon this abstract representation.
 
-\updated{%
 In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
-modelling.}
+modelling.
 %

### 2013-07-23 parts/intro.tex
cleanup

@@
 While the formal checking of dynamical properties is often limited to small networks because of the
-state graph explosion, \updated{an other major} drawback of this framework is the difficulty to
+state graph explosion, an other major drawback of this framework is the difficulty to
 specify Thomas' parameters, especially for large networks.
@@
 This framework can be seen as a special class of formalisms like Petri Nets or Communicating Finite
-State Machines, \updated{where the arity and the effect of synchronisations between components is
-restricted.}
+State Machines, where the arity and the effect of synchronizations between components is
+restricted.
 Thanks to the particular structure of interactions within a PH, very efficient static analysis
@@
 the formal analysis of the qualitative dynamics of BRNs with hundreds or thousands of components,
-\updated{which was impossible with other state-of-the-art approaches \cite{PMR12-MSCS,PAK13-CAV,FPMR13-CS2Bio}}.
+which was impossible with other state-of-the-art approaches \cite{PMR12-MSCS,PAK13-CAV,FPMR13-CS2Bio}.
 
@@
 
-\updated{%
 The aim of this paper is to formally establish links between several layers suited for specifying the
@@
 and Thomas' parameters, which completely specify the functions governing the evolution of
-components.}
-\updatedmax{%
+components.
 Another motivation for the work depicted in this paper is that it constitutes an important step
@@
 we should be able to synthesize families of BRN having some featuring behavioral properties.
-}
 
-\updated{%
 Firstly, we derive rules that, given a PH model, infer the IG corresponding to the dynamics of the
@@
 and the absence of negative cycle to the impossibility for oscillatory behaviours.
-}
 
-\updated{%
 Second, we tackle the characterisation of Thomas' parameters that are compatible with a given PH
@@
 precise dynamical features using a single PH models which gathers all the plausible dynamics.
-}
 
-\updated{%
 Finally, we discuss on an implementation of these inference schemes using Answer Set
 Programming (ASP) \cite{Baral03},  which turns out to be effective for these enumerative searches.
-}%
-\updatedmax{%
 The whole method is also applied to the biological model of ERBB receptor-regulated G1/S transition,
@@
 to refine its dynamics, and study a more precise class of underlying Thomas models.
-}
 %\todo{describe application}
-
-%The outcome of this work is twofold.
-%The first benefit is that such an approach makes it possible to refine the construction of
-%BRNs with a partial and progressively brought knowledge in PH, while being able to export such
-%models in the Thomas' framework.
-%This work thus strengthens the formal link between both modelings.
-%The second feature of our method is that it can be applied on very large BRNs.
 
@@
 
-\updated{%
 The contributions presented in this paper significantly extend and improve the preliminary results
@@
 This allows to apply our methodology to a wider class of models.
-}%
-\updatedmax{%
 Furthermore, some parts have been completely rewritten in order to explain more precisely the efficiency of the ASP implementation,
 and give a new detailed biological application of our work.
-}%
 

### 2013-07-23 parts/asp.tex
TODOs
cleanup
examples in scetion 5 (ASP)

@@
 
-\rewrite{Check and correctly glue this section}
+%\todo{Proofread this whole section}
 
@@
 
-\begin{example}
+\begin{example}\label{ex:asp-component}
 Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
@@
 \begin{example}
-Regarding \pref{ex:infer-param-runningPH-1}, the previous rule will make the following set of atoms belong to every answer set:
+Regarding \pref{ex:infer-param-runningPH-1}, the previous rule together with the facts of \pref{ex:asp-component}
+will give the following answer set:
 \begin{align*}
-  &\{&&\componentlevels(b, 0),
+  &\{&&\component(a, 2),
+  &&\component(b, 1), \\
+  &&&\component(c, 1),
+  &&\componentlevels(a, 0), \\
+  &&&\componentlevels(a, 1),
+  &&\componentlevels(a, 2), \\
+  &&&\componentlevels(b, 0),
   &&\componentlevels(b, 1), \\
   &&&\componentlevels(c, 0),
-  &&\componentlevels(c, 1), \\
-  &&&\componentlevels(a, 0),
-  &&\componentlevels(a, 1), \\
-  &&&\componentlevels(a, 2) &&&&\}
+  &&\componentlevels(c, 1). &&&&\}
 \end{align*}
+%will make the following set of atoms belong to every answer set:
+%\begin{align*}
+%  &\{&&\componentlevels(b, 0),
+%  &&\componentlevels(b, 1), \\
+%  &&&\componentlevels(c, 0),
+%  &&\componentlevels(c, 1), \\
+%  &&&\componentlevels(a, 0),
+%  &&\componentlevels(a, 1), \\
+%  &&&\componentlevels(a, 2),% &&&&\}
+%  &&\component(a, 2), \\
+%  &&&\component(b, 1),
+%  &&\component(c, 1). &&&&\}
+%\end{align*}
 \end{example}
@@
 \begin{example}
-
 The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
-For example, a constraint was written in order to obtain only interval parameter
+For example, a constraint was written in order to obtain only interval parameters
 (thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
@@
 the monotonicity assumption (\pref{pro:param_enum_monotonicity}) removes all parametrizations with: $0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
-In the end, the remaining possible values for the parameters that were not inferred are:
+In the end, the remaining possible values for the two parameters that were not inferred are:
 \begin{align*}

### 2013-07-24 parts/example.tex
improve section Examples

@@
 \pref{ssec:appli} gives a practical and detailed application of our results on a biological model taken from the literature,
-while \pref{ssec:cpu} gives data related to the computation times of our software when applied to several large biological models.
+while \pref{ssec:cpu} gives data related to the results and computation times of our software when applied to several large biological models.
 
@@
 
-This subsection focuses on the study of a model of ERBB receptor-regulated G1/S transition detailed in~\cite{Sahin09}.
+This subsection focuses on the study of the epidermal growth factor (EGF) receptor model detailed in~\cite{Sahin09}.
 This model is represented by an IG containing 20 components and 52 edges.
-A protein named EGF, having no ingoing edge, can be considered as the only input.
-Then, a chain of reactions lead to the activation of protein pRB which is responsible for regulating the cell division,
+A protein named EGF, having no regulator, can be considered as the only input,
+and a chain of reactions leads to the activation of protein pRB which is responsible for regulating the cell division,
 therefore making it essential to prevent cancer development.
@@
 Three models are created from the original IG, with different levels of precision regarding the cooperation between components.
-Model (1) represents a translation of the original IG into Process Hitting, that is,
+The translation from an IG into a Process Hitting is not detailed here as it was previously covered in~\cite{PMR10-TCSB}.
+Model (1) represents a translation of the raw IG into Process Hitting, that is,
 without any knowledge of the Boolean rules (and therefore the cooperations) of the components.
 Model (2) implements some of the rules based on the results of several knockdown experiments.
-Model (3) is the totally refined model with all cooperations implemented given the Boolean functions for all components.
+Model (3) is the totally refined model with all cooperations implemented given the Boolean functions of all components.
 Therefore, those three models can be considered as successive refinements of the original and most general one.
@@
   \begin{tabular}{r|l|l|m{2cm}|m{2.5cm}|m{1.5cm}}
-    \textbf{Model} & \textbf{$|E|$} & \textbf{$|K|$} & \textbf{Inferred parameters} & \textbf{Possible\newline models} & \textbf{Fixed points}
+    \textbf{Model} & $\mathbf{|E|}$ & $\mathbf{|K|}$ & \textbf{Inferred parameters} & \textbf{Possible\newline models} & \textbf{Fixed points}
   \\\hline\hline
@@
 \hfill~%\end{center}
-  \caption{Results of the IG and parameters inference on three models
-  derived the ERBB receptor-regulated G1/S transition model of~\cite{Sahin09}
+  \caption{%
+  Results of the IG and parameters inference on three models
+  derived from the EGF receptor model of~\cite{Sahin09}
   with different precisions in the definition of the cooperations.
   Model (1) contains no cooperations between the components.
-  Several cooperations were included in model (2) under the form of 14 cooperative sorts.
-  Model (3) models all the Boolean functions between components using 22 cooperative sorts.}
+  Several cooperations were included in model (2) under the form of 14 cooperative sorts
+  and all of them were included in Model (3) under the form of 22 cooperative sorts.
+%  Model (3) models all the Boolean functions between components using 22 cooperative sorts.
+  The second column gives the number of edges in the IG inferred with \pref{pps:inference-IG}
+  (the number of nodes is always the number of components in the model, that is, 20).
+  The third column gives the number of parameters in the model (given the IG),
+  the fourth column gives the number of parameters that could be inferred using \pref{pps:param_K},
+  and the fifth column consequently gives the number fo compatible models with the studied PH model,
+  which exponentially depends on the number of parameters that could not be inferred.
+  Finally, the last column gives the number of fixed points in the PH model,
+  computed with another existing PH tool included in \textsc{Pint}.
+  }
   \label{tb:egfr20}
@@
 As all cooperations are fully defined in this model, all the parameters are inferred and only one Thomas model can be derived.
-We note also that this PH model is the only one containing at least one fixed point.
+We note also that this PH model is the only one containing at least one fixed point,
+which indicates the possibility of a steady state once the input signal is completely propagated.
 
@@
 
-The current implementation can successfully handle large PH models of BRNs found in the literature
-such as the ERBB receptor-regulated G1/S transition model from~\cite{Sahin09} presented in the last subsection,
-and a T cell receptor model from~\cite{Klamt06} which contains 40 components and 14 cooperative
+The current implementation can successfully handle large PH models of BRNs found in the literature such as:
+\begin{itemize}
+  \item the EGF receptor model from~\cite{Sahin09} with 20 components presented in the last subsection,
+  \item a T cell receptor model described as an IG in~\cite{Klamt06}, which contains 40 components and 14 cooperative
     sorts\footnote{All models mentioned in this section are available as examples distributed with \textsc{Pint}.}.
+\end{itemize}
 For each model, IG and parameters inferences are performed together in less than a second
@@
 Bigger models related to the aforementioned systems were also tested with our implementation:
-an other T cell receptor model described in~\cite{SaezRodriguez2007} containing 94 components
-and an EGFR/ErbB signaling model described in~\cite{Samaga2009} containing 104 components.
-\todo{Mention how these models were created?}
+\begin{itemize}
+  \item a model of the T cell receptor with 94 components, described in~\cite{SaezRodriguez2007},
+  \item a model of the EGF receptor with 104 components, described in~\cite{Samaga2009}.
+\end{itemize}
+These two models were obtained in a previous work by an automatic translation from the CellNetAnalyzer formalism.
+%\todo{Mention how these models were created?}
 
@@
 
-\begin{table}
+\begin{table}[ht]
 \begin{center}
-  \begin{tabular}{r|c|c|c|c}
-    \textbf{Model} & \textbf{Components} & \textbf{\# CS} & \textbf{$\Delta t$ IG} & \textbf{$\Delta t$ K}
+  \begin{tabular}{r|c|c|c|c|c}
+    \textbf{Model} & $\mathbf{|\Gamma|}$ & $\mathbf{|\PHs \setminus \Gamma|}$ & \textbf{$\Delta t$ IG} & \textbf{$\Delta t$ K} & $\mathbf{|K|}$
 \\\hline\hline
-    ERBB-G1/S \cite{Sahin09} & $20$ & $22$ & <1s & <1s
+    EGF receptor \cite{Sahin09} & $20$ & $22$ & <1s & <1s & 192
 \\\hline
-    T cell receptor \cite{Klamt06} & $40$ & $14$ & <1s & <1s
+    T cell receptor \cite{Klamt06} & $40$ & $14$ & <1s & <1s & 143
 \\\hline
-    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & 10s & ---
+    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & 10s & --- & $2.1\cdot10^{9}$
 \\\hline
-    EGFR/ErbB signaling \cite{Samaga2009} & $104$ & $89$ & 3min 20s & ---
+    EGF receptor \cite{Samaga2009} & $104$ & $89$ & 3min 20s & --- & $4.2\cdot10^{6}$
   \end{tabular}
 \end{center}
-\caption{
-  Computation times and several pieces of data related to the IG and parametrization inferences of fours biological models.
-  The third column gives the number of cooperative sorts used in the models to model joint actions,
-  and the fourth (resp.~fifth) column gives the computation times of the IG inference (resp.~the parametrization inference).
+\caption{%
+  Computation times and several pieces of information related to the IG and parametrization inferences of four biological models.
+  The second column gives the number of components of each model and
+  the third column gives the number of cooperative sorts used to model joint actions.
+  The fourth (resp.~fifth) column gives the computation times of the IG inference (resp.~the parametrization inference).
+  The last column gives the number of parameters in each model.
+  Due to the too high number of parameters, parametrization inference could not be performed on the $94$ and $104$ component models.
   }

### 2013-07-24 parts/infer-ig.tex
minor fixes in section IG Inference

@@
 the level of $b$ makes possible the increase (decrease) of the level of $a$,
-and is directly inspired from the works of~\cite{Richard2010378}.
+which is directly inspired from the works of~\cite{Richard2010378}.
 To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
-Such groups are given by the set $X(a)$ that is the set of connected components in the graph linking two regulators
+Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
 $b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
@@
 To infer influence of a component $b$ on a component $a$,
-we observe of the direction of evolution of $a$ for different active processes of $b$.
+we observe the direction of evolution of $a$ for different active processes of $b$.
 The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
-This reasoning can also be applied in the case where $a = b$.
-
+%This reasoning can also be applied in the case where $a = b$.
+%
 %To infer influences on a component $a$, we compare several configurations of groups of components of $a$.
@@
 %and we can infer a positive influence if raising the level of $b$ changes the evolution of $a$ 
-
+%
 Formally, for a given group $g$ of regulators of a component $a$
 (that is, a minimal set of components regulating $a$ through common cooperative sorts),
-and a configuration $\ctx$ on $g$,
-$\irB_a(\sigma)$ is the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
+and a configuration $\ctx$ on $g$, we note
+$\irB_a(\sigma)$ the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
 %This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
@@
 \begin{align}
+% Arcs a -> b, a ≠ b
 \begin{split}\label{eq:edges-inference}
-%  \forall b\in\PHpredecgene{a} \cup \{ a \}, \forall s \in \{ +, - \}, \\
   \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
@@
 \end{split}\\
-%
-%%% Tentative de réduire le nombre d'auto-arcs superflus
-%
+% Auto-arcs
 \begin{split}\label{eq:edges-inference-auto}
@@
 \end{split}
-%
-%%% Ancienne version de l'inférence
-%
-%\begin{split}\label{eq:edges-inference-pos}
-%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
-%  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
-%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j < a_k
-%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \quad^*}\\
-%%    &\qquad \warnul{a_j \leq \PHget{\sigma}{a} < a_k \vee a_j < \PHget{\sigma}{a} \leq a_k \quad^{**}}\\
-%%    &\qquad \warnul{a_j < a_k \quad^{***}}
-%\end{split}\\
-%\begin{split}\label{eq:edges-inference-neg}
-%  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
-%  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
-%    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-%    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), a_j > a_k
-%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \quad^*}\\
-%%    &\qquad \warnul{a_j > \PHget{\sigma}{a} \geq a_k \vee a_j \geq \PHget{\sigma}{a} > a_k \quad^{**}}\\
-%    %&\qquad \warnul{a_j > a_k \quad^{***}}
-%\end{split}
 \end{align}
@@
 \end{proposition}
-%
-%\todo{\underline{$^*$Strict inference}}\\
-%\todo{\underline{$^{**}$Large inference}}\\
-%\todo{\underline{$^{***}$Non-unitary-style inference}}
 

### 2013-07-24 parts/asp.tex
remove comments + typos

@@
 \section{Answer Set Programming implementation concepts}\label{sec:impl}
-
-%\todo{Proofread this whole section}
 
@@
   &&&\componentlevels(c, 0),
-  &&\componentlevels(c, 1). &&&&\}
+  &&\componentlevels(c, 1). &\}
 \end{align*}
-%will make the following set of atoms belong to every answer set:
-%\begin{align*}
-%  &\{&&\componentlevels(b, 0),
-%  &&\componentlevels(b, 1), \\
-%  &&&\componentlevels(c, 0),
-%  &&\componentlevels(c, 1), \\
-%  &&&\componentlevels(a, 0),
-%  &&\componentlevels(a, 1), \\
-%  &&&\componentlevels(a, 2),% &&&&\}
-%  &&\component(a, 2), \\
-%  &&&\component(b, 1),
-%  &&\component(c, 1). &&&&\}
-%\end{align*}
 \end{example}
@@
 which is a shorthand for “the set of atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
-%\rewrite{Reformulate?}
 
@@
 
-%so that
-%inferred parameters with \pref{pps:param_K} keep their value and
-%parameters that could not be inferred take any possible value amongst the expression levels of the related component: $K_{a, \omega_p} \subset \segm{0}{l_a} \wedge K_{a, \omega_p} \neq \emptyset$.
-
 \begin{example}\label{ex:cardinality}
@@
 \subsection{Constraints}\label{sssec:constraints}
-%Finally, constraints can be used to filter answer sets that model parametrizations which are impossible or do not respect the assumptions of \pref{ssec:admissible-K}.
 Finally, a constraint is a rule with no $head$ part:
@@
 All these candidates respect the other properties of \pref{ssec:admissible-K} and no more candidates are filtered.
-%The other properties do not filter other values for these parameters;
 We thus obtain the results of \pref{ex:enum-param-runningPH-1}.
 \end{example}
-
-\begin{comment}
----------------
-
-In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
-For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
-and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
-In other words, $\paramlabel(a, p)$ states that a parameter $K^p_{a,\omega}$ exists.
-Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 7 parameters, and to state the existence of each of them, the following facts are used:
-\begin{align*}
-  &\paramlabel(a,1).
-  &\paramlabel(b,1).\\
-  &\paramlabel(a,2).
-  &\paramlabel(b,2).\\
-  &\paramlabel(a,3).\\
-  &\paramlabel(a,4).
-  &\paramlabel(c,1).
-\end{align*}
-
-A set (such as sets of resources $\omega$) can be defined in ASP with a series of atoms. \rewrite{Find a more precise sentence.}
-For example, for a given parameter $K^p_{a,\omega}$, expressing $x \in \omega$ is done by the atom: $\paramres(a, p, x)$.
-Then, the sets of resources of \pref{ex:infer-param-runningPH-1} is defined by:
-$$\begin{array}{>{\hspace{.8em}}l|>{\phantom{\raisebox{.6em}{xx}}}l}
-  \hspace{-.8em}\text{Parameter} & \text{Corresponding ASP fact} \\
-\hline
-  K^0_{a,\emptyset} & \\
-  K^1_{a,\{b\}} & \paramres(a,1,b). \\
-  K^2_{a,\{c\}} & \paramres(a,2,c). \\
-  K^3_{a,\{b,c\}} & \paramres(a,3,b). \quad \paramres(a,3,c). \phantom{\raisebox{-.8em}{xx}} \\ \hline
-  K^0_{b,\emptyset} & \\
-  K^1_{b,\{a\}} & \paramres(b,1,a). \phantom{\raisebox{-1em}{xx}} \\ \hline
-  K^0_{c,\emptyset} &
-\end{array}$$
-The absence of an atom $\paramres(a,1,c)$, for example, states that $c$ is not a resource for the parameter $K^1_{a,\{b\}}$ (indeed, $c \notin \{b\}$).
-
-\todo{Continue refactoring.}
-\rewrite{Explain variables}
-
-Rules allow more detailed declarations than facts as they have a body (right-hand part below) containing constraints and allowing to use variables, while facts only have a head (left-hand part).
-For instance, in order to define the set of expression levels of a component, we declare:
-$$\atom{component\_levels}(\var{X}, 0..\var{M}) \la \atom{component}(\var{X}, \var{M}).$$
-where the atom $\atom{component}(X, M)$ stands for the existence of a component $X$ with a maximum level $M$.
-Considering this declaration, any possible answer for the atom $\atom{component\_levels}$ will be found by binding all possible values of its two terms with all existing $\atom{component}$ facts:
-the existence of an answer $\atom{component\_levels}(a, k)$ will depend on the existence of a term $a$, which is bound with $\var{X}$, and an integer $k$, constrained by: $0 \leq k \leq \var{M}$.
-
-Cardinalities are convenient to enumerate all possible parametrizations by creating multiple answer sets.
-A cardinality (denoted hereafter with curly brackets) gives any number of possible answers for some atoms between a lower and upper bounds.
-For example,
-\begin{align*}
-  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
-  & \qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
-\end{align*}
-where $\atom{param}(\var{X}, \var{P}, \var{I})$ stands for: $\var{I} \in K^\var{P}_{\var{X},\omega}$,
-means that any parameter of component $\var{X}$ and label $\var{P}$ must contain at least one level value ($\var{I}$) in the possible expression levels of $\var{X}$.
-Indeed, the lower bound is $1$, forcing at least one element in the parameter, but no upper bound is specified, allowing up to any number of answers.
-The body (right-hand side) of the rule also checks for the existence of a parameter of $\var{X}$ with label $\var{P}$,
-and constrains that the parametrization inference was not conclusive for the considered parameter (“$\nota$” stands for negation by failure: $\nota \atom{L}$ becomes true if $\atom{L}$ is not true).
-Such a constraint gives multiple results as any set of $\atom{param}$ atoms satisfying the cardinality will lead to a new global set of answers.
-In this way, we enumerate all possible parametrizations which respects the results of parameters inference,
-but completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K}.
-
-We rely on integrity constraints to filter only admissible parametrizations.
-An integrity constraint is a rule with no head, that makes an answer set unsatisfiable if its body turns out to be true.
-Hence, if we suppose that:
-\begin{itemize}
-  \item the $\atom{less\_active}(a, p, q)$ atom means that $K^p_{a,\omega}$ stands for a configuration with less activating regulators than $K^q_{a,\omega'}$, % (\ie $A \subset A'$),
-  \item the $\atom{param\_inf}(a, p, q)$ atom means: $K^p_{a,\omega} \leqsegm K^q_{a,\omega'}$,
-\end{itemize}
-then the monotonicity assumption (\pref{pro:param_enum_monotonicity}) is formulated as the following integrity constraint:
-$$\la \atom{less\_active}(\var{X}, \var{P}, \var{Q}), \nota \atom{param\_inf}(\var{X}, \var{P}, \var{Q}).$$
-which removes all parametrization results where parameters $K^\var{P}_{\var{X},\omega}$ and $K^\var{Q}_{\var{X},\omega'}$ exist such that $\var{X}$ is less activated by $\omega$ than $\omega'$ %$A \subset A'$
-but $K^\var{Q}_{\var{X},\omega'} \ltsegm K^\var{P}_{\var{X},\omega}$,
-thus violating the monotonicity assumption.
-Of course, other assumptions can be formulated in the same way.
-\end{comment}
 

### 2013-07-24 parts/example.tex
remove comments + typos

@@
 \section{Examples}\label{sec:examples}
-
-%\todo{Proofread this section}
 
@@
 or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
-
-\begin{comment}
-\todo{Check this paragraph}
-Applied to the example in \pref{fig:runningPH-2} where cooperations have been defined,
-our method infers the IG and parametrization given in \pref{fig:runningBRN}.
-Regarding the example in \pref{fig:runningPH-1}, the same IG is inferred, as well as for the
-parametrization except for the parameters $K_{a,\{b\},\{c\}}$ and $K_{a,\{c\},\{b\}}$ which are
-undefined (because of the lack of cooperativity between $b$ and $c$).
-In such a case, this partial parametrization allows 36 admissible complete parametrizations, as two
-parameters with 3 potential values could not be inferred.
-If we constrain these latter parameters so that they contain exactly one element, we obtain only 9
-admissible parametrizations.
-\end{comment}
 
@@
 These two models were obtained in a previous work by an automatic translation from the CellNetAnalyzer formalism.
-%\todo{Mention how these models were created?}
 

### 2013-07-24 parts/infer-ig.tex
remove comments + typos

@@
 positive and/or negative, between the components of the system.
-As discussed in \pref{sec:intro}, the IG allows to efficiently characterise global dynamical
+As discussed in \pref{sec:intro}, the IG allows to efficiently characterize global dynamical
 properties for the concrete system, such as the capability for multi-stationarity or oscillation.
 
-In a typical biological network modelling process, a prior IG is generally the starting point for
+In a typical biological network modeling process, a prior IG is generally the starting point for
 the formal system specification.
@@
 In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
-modelling.
-%
-%In order to infer a complete BRN, one has to find the Interaction Graph (IG) first, as some
-%constraints on the parametrization rely on it.
-%Inferring the IG is an abstraction step which consists in determining the global influence of
-%components on each of its successors.
-%
-%
+modeling.
 This section first introduces the notion of focal processes within a PH (\pref{ssec:focal})
@@
 The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
-%This reasoning can also be applied in the case where $a = b$.
-%
-%To infer influences on a component $a$, we compare several configurations of groups of components of $a$.
-%Indeed, to study the influence of any component $b$ on $a$, it is sufficient to study the (unique) group of influence
-%$g \in X(a)$ which $b$ belongs to
-%(that is, the unique set of components regulating $a$ through common cooperative sorts containing $b$).
-%This reasoning relies on the observation of the direction of evolution of $a$ for different active processes of $b$.
-%and we can infer a positive influence if raising the level of $b$ changes the evolution of $a$ 
-%
 Formally, for a given group $g$ of regulators of a component $a$
@@
   &\forall g \in X(a), \forall \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c,
-%    \sigma' = \ctx_{\reg{a}}(\sigma),\\
   \irB_a(\sigma) \DEF 
@@
   \end{cases}\\
-%\end{split}\\
-%\begin{split}
   &\text{where: } \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
@@
 \end{example}
-
-\begin{comment}
-\subsection{Structure-based Interaction Inference}\label{ssec:infer-IG-structure}
-
-At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
-($\PHs \setminus \Gamma$) that will not appear in the IG. 
-We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
-by considering direct actions and possible intermediate cooperative sorts.
-The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
-(\pref{eq:regulators}).
-\begin{align}
-\begin{split}
-\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
-(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
-                                   & \quad \quad c^0 = b \wedge c^n = a \\
-                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
-                   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
-\end{split}
-\label{eq:ph_predec}
-\\
-\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
-\label{eq:regulators}
-\end{align}
-
-Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
-refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
-If $g=\{b\}$, we simple note $\ctx_b(\sigma)$.
-This set is composed of the active processes of sorts in $g$, and the focal process (assumed
-unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
-The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
-relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
-$\upsilon$ are defined in $\sigma$.
-When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
-the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
-where $\uplus$ stands for a union of states (which are sets of processes).
-Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
-always terminates.
-\begin{align}
-\forall g\subset \Gamma,
-  \ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
-\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
-\label{eq:ctx-sigma}
-\\
-\upsilon(\sigma) & \DEF
-\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
-  \upsilon'\in\PHdirectpredec{\upsilon} \wedge
-  \upsilon'\in\PHs\setminus\Gamma })
-\label{eq:cooperative-eval}
-\end{align}
-
-This first inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
-We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
-the level of $b$ makes possible the increase (decrease) of the level of $a$.
-To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
-Such groups are given by the set $X(a)$ that is the set of connected components in the graph linking two regulators
-$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
-\begin{equation}
-X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
-        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
-        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
-\label{eq:influence-groups}
-\end{equation}
-
-This reasoning can then be straightforwardly applied when $b \neq a$ (\pref{eq:edges-inference-b}).
-Indeed, to study the influence of any component $b$ on $a$, it is sufficient to study the (unique) group of influence
-$g \in X(a)$ which $b$ belongs to. %, that is, the (unique) set in $X_a$ containing $b$, denoted $X_a(b)$.
-%The reasoning mentioned in the previous paragraph can then be straightforwardly applied when $b \neq a$ (\pref{eq:edges-inference-b}),
-%by considering this 
-Indeed, given a configuration $\sigma\in\prod_{c\in g \cup \{ a \}} L_c$, 
-$\focals(a,\{a_i\},\ctx_b(\sigma))$ gives the bounces that a given process $a_i$ can make in the
-context $\ctx_b(\sigma)$.
-We note $\sigma\{b_i\}$ the configuration $\sigma$ where the process of sort $b$ has been replaced
-by $b_i$.
-If there exists $b_i,b_{i+1}\in L_b$ such that one bounce in 
-$\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_i\}))$
-has a lower (resp. higher) level that one bounce in
-$\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_{i+1}\}))$, then
-$b$ as positive (resp. negative) influence on $a$ with a maximum threshold $t=i+1$.
-
-Finally, we infer that $a$ has a self-influence if its current level can have an impact on its own
-evolution in a given configuration $\sigma$ of any group of influence on $a$ (\pref{eq:edges-inference-a}).
-Given $a_i,a_{i+1}\in L_a$, we pick $a_j\in\focals(a,\{a_i\},\ctx_g(\sigma\{a_i\}))$ and
-$a_k\in\focals(a,\{a_{i+1}\},\ctx_g(\sigma\{a_{i+1}\}))$.
-If $k=j+1$, we can not conclude as there is no difference in the evolution of both levels.
-If $k\neq j+1$ and $k-j\neq 0$, then $a_i$ and $a_{i+1}$ have divergent evolutions: we infer an
-influence of sign of $k-j$ at threshold $i+1$.
-We note that some aspects of this inference are arbitrary and may impact the number of parameters to
-infer in the next section.
-In particular, in some cases, the use of intervals for Thomas' parameters drops the requirement of
-inferring a self-activation.
-
-\pref{pps:inference-edges} details the inference of all existing influences between components occurring
-with a threshold $t$.
-These influences are split into positive and negative ones, and represent possible edges in the final IG.
-We do not consider the cases where a component has no visible influence on another.
-\begin{proposition}[Influences inference]\label{pps:inference-edges}
-We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any
-$a\in\Gamma$ by:
-\begin{align}
-\begin{split}
-\forall b\in\PHpredecgene{a}, b &\neq a, \\
-b\xrightarrow t a \in \hat{E}_s & \Longleftrightarrow
-  \exists g \in X(a), b \in g, 
-  \exists \sigma\in\textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \exists b_i,b_{i+1}\in \PHl_b,\\
-& \qquad\qquad
-        \exists a_{j}\in\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_i\})), \\
-& \qquad\qquad
-        \exists a_{k}\in\focals(a, \{\sigma[a]\}, \ctx_b(\sigma\{b_{i+1}\})), \\
-& \qquad\qquad\qquad
-                        s = \f{sign}(k - j) \wedge t = i+1
-\end{split}
-\label{eq:edges-inference-b}
-\\
-\begin{split}
-a \xrightarrow t a \in \hat{E}_s & \Longleftrightarrow
-\exists g \in X(a), \sigma \in \textstyle\prod_{b\in g \cup \{ a \}} \PHl_b,
-      \exists a_i,a_{i+1}\in \PHl_a, \\
-& \qquad\qquad
-        \exists a_{j}\in\focals(a, \{a_i\}, \ctx_g(\sigma\{a_i\})), \\
-& \qquad\qquad
-        \exists a_{k}\in\focals(a, \{a_{i+1}\},  \ctx_g(\sigma\{a_{i+1}\})), \\
-& \qquad\qquad\quad
-      k \neq j+1
-        \wedge s = \f{sign}(k - j) \wedge t = i+1
-\end{split}
-\label{eq:edges-inference-a}
-\end{align}
-where $s \in \{ +, - \}$,
-$\f{sign}(n) = + \EQDEF n > 0$,
-$\f{sign}(n) = - \EQDEF n < 0$,
-and $\f{sign}(0) \DEF 0$.
-\end{proposition}
-
-We are now able to infer the edges of the final IG by considering positive and negative influences
-(\pref{pps:inference-IG}).
-We infer a positive (resp. negative) edge if there exists a corresponding influence with the same
-sign. If an influence is both positive and negative, we infer an unsigned edge. In the end, the
-threshold of each edge is the minimum threshold for which an influence has been found.
-%As unsigned edges represent ambiguous interactions, no threshold is inferred.
-\begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
-We infer $\IG = (\Gamma,E)$ using \pref{pps:inference-edges} as follows:
-\begin{align*}
-E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
-  \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
-E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
-  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
-E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_- \\
-  & \qquad\qquad\qquad \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\}
-\end{align*}
-\end{proposition}
-
-\begin{example}
-The IG inference from the PH of \pref{fig:runningPH-2} gives
-$\hat{E}_+ = \{b \xrightarrow{1} a, c \xrightarrow{1} a\}$ and 
-$\hat{E}_- = \{a \xrightarrow{2} b\}$, corresponding to the IG of \pref{fig:runningBRN}.
-No self-influence are inferred. %($X(a) = \{ \{b,c\} \}$, $X(b)=\{ \{a\}\}$, and $X(c)=\emptyset$).
-\rewrite{Check this example}
-\end{example}
-
-\todo{Insist on the absence of non-functional edges in the inferred IG (contrary to biologists' IGs)}
-
-\subsection{Dynamics-based Interaction Inference}\label{ssec:infer-IG-dynamics}
-
-This second inference is complementary to the one presented in the previous subsection.
-It is based on an approach considering the global dynamics of a component rather than independent dynamics of each regulator,
-%in that, it is analogous to standard IG inferences from discrete maps \cite{RiCo07}.
-and is inspired from the definition of Interaction Graph built from a Transition Graph as described in~\cite{Richard2010378}.
-Contrary to the previously presented method, we focus on a configuration (or: sub-state) $\sigma$
-on the whole set $\reg{a}$ of predecessors of a component $a$ (rather than on a subset of cooperating regulators).
-
-For a given configuration $\sigma$ of a component $a$ and its regulators,
-$\irB_a(\sigma)$ is the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
-This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
-and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
-If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
-%
-\begin{align}
-\begin{split}\label{eq:possible-actions}
-  \forall \sigma &\in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \sigma' = \ctx_{\reg{a}}(\sigma),\\
-  & \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
-  & \qquad\qquad\qquad\qquad\qquad \PHget{\sigma'}{b} = b_i \wedge \PHget{\sigma'}{a} = a_j \}\\
-\end{split}
-\end{align}
-%
-\begin{align}
-\begin{split}\label{eq:possible-bounces}
-  \forall \sigma &\in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c,
-%    \sigma' = \ctx_{\reg{a}}(\sigma),\\
-  \irB_a(\sigma) \DEF 
-  \begin{cases}
-    \irF_a(\sigma)
-      & \text{ if } \irF_a(\sigma) \neq \emptyset\\
-    \{ \PHget{\sigma}{a} \}
-      & \text{ if } \irF_a(\sigma) = \emptyset
-  \end{cases}
-\end{split}
-\end{align}
-
-The main idea of the inference detailed in \pref{pps:inference-edges-dynamics} is close to the one of \pref{ssec:infer-IG-structure}:
-the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
-there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop) (\pref{eq:edges-inference-richard-pos} \& \eqref{eq:edges-inference-richard-neg}).
-the main difference lies in the fact that this inference focuses on local evolutions (the possibility for $a$ to bounce away from its original state)
-rather than global evolutions (by considering focal processes denoting the overall influence).
-Therefore, an influence is inferred if the possible bounces of $a$ in $\sigma$ depend on the level of $b$.
-%
-\begin{proposition}[Influences inference]\label{pps:inference-edges-dynamics}
-We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
-\begin{align}
-\begin{split}\label{eq:edges-inference-richard-pos}
-  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
-  b\xrightarrow{t+1} a \in \hat{E}_+ \Longleftrightarrow\ &
-    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
-    &\qquad a_j \leq \PHget{\sigma}{a} < a_k \warnul{\vee a_j < \PHget{\sigma}{a} \leq a_k \quad^*}
-\end{split}\\
-\begin{split}\label{eq:edges-inference-richard-neg}
-  \forall b\in\PHpredecgene{a}, \qquad\qquad\quad\\
-  b\xrightarrow{t+1} a \in \hat{E}_- \Longleftrightarrow\ &
-    \exists \sigma \in \textstyle\prod_{c\in \reg{a} \cup \{ a \}} \PHl_c, \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}),\\
-    &\qquad a_j > \PHget{\sigma}{a} \geq a_k \warnul{\vee a_j \geq \PHget{\sigma}{a} > a_k \quad^*}
-\end{split}
-\end{align}
-\end{proposition}
-%
-\todo{\underline{$^*$Keep these large conditions?}}
-
-Finally it is possible to infer the final IG by using the same reasoning as \pref{pps:inference-IG}:
-a positive (negative) edge in inferred if there exist positives (negative) influences only, otherwise an unsigned edge in inferred.
-We point out that the main use of this alternative inference is the fact that the conditions used to infer each influence
-in \pref{eq:edges-inference-richard-pos} \& \eqref{eq:edges-inference-richard-neg}
-allows to apply some results about the links between the structure of an IG and the dynamics of the model.
-These results consist mainly in Thomas' conjectures, that focus on the presence of positive and negative circuits in the IG, and state that:
-\begin{itemize}
-  \item the presence of a positive circuit in the IG is a necessary condition for several fixed points,
-  \item the presence of a negative circuit in the IG is a necessary condition for sustained oscillations.
-\end{itemize}
-The sign of a circuit is defined by the number of negative edges it contains;
-given the nature of unsigned edges, the presence of such an edge in a circuit would count for the presence of both a positive and negative circuit.
-These conjectures have been demonstrated in~\cite{Richard2010378} \towrite{Other references! (multi-stationarity \& boolean models)}.
-
-\todo{Example}
-\end{comment}

### 2013-07-26 main.tex
big application example + computation results on other examples
check TODOs in Introduction
finish proofreading of Examples
TODOs
cleanup
fixes after Morgan's review

@@
 \newrefformat{ex}{Example~\ref{#1}}
+\newrefformat{tb}{Table~\ref{#1}}
 \def\pref{\prettyref}
@@
 %\usepackage{color} % Couleurs du texte
-\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
-\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
-\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
-\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
-\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
+%\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
+%\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
+%\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
+%\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
+%\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
 \newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
-\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
-\def\updated#1{{\color{magenta}#1}}
+%\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
+%\def\updated#1{{\color{magenta}#1}}
+%\newcommand{\updatedmax}[1]{{\color{violet}#1}}
 %\newcommand{\new}[1]{#1}
@@
 % Id est
-\newcommand{\ie}{i.e. }
+\newcommand{\ie}{i.e.~}
 
@@
 formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
+In addition to be able to use the existing results on this formalism,
+obtaining the Thomas representation of a PH model allows the iteratively refine a system.
 
@@
 complementary features.
-First, the inference of the Interaction Graph (IG) from a PH models allows a compact representation of
+First, the inference of the Interaction Graph (IG) from a PH model gives a compact representation of
 the regulations that are effective for the dynamics.
 From this IG, well-known results can be applied to derive global dynamical properties for the
-modelled BRN.
+modeled BRN.
 Second, the inference of all Thomas models that are compatible with a given PH is tackled.
@@
 which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
-\todo{application}
 \end{abstract}

### 2013-07-30 main.tex
abstract

@@
 formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
-In addition to be able to use the existing results on this formalism,
-obtaining the Thomas representation of a PH model allows the iteratively refine a system.
+%In addition to be able to use the existing results on this formalism,
+%obtaining the Thomas representation of a PH model allows to iteratively refine a model.
+Linking this representation to PH would not only allow to use the existing results on this formalism,
+but also to iteratively refine the model of a system.
 

### 2013-08-01 parts/asp.tex
begin fixes after Inoue's feedback

@@
 
-Answer Set Programming (ASP) is a logic programming paradigm \cite{Baral03},
+Answer Set Programming (ASP) is a logic programming paradigm \cite{Baral03, Baral10},
 which has been chosen to address the enumeration of all admissible parametrizations.
@@
 \begin{itemize}
-  \item ASP efficiently tackles the inherent complexity of the models used, thus allowing a fast execution of the formal tools defined in this paper,
-  \item it is convenient to enumerate a large set of possible answers,
-  \item and it allows to easily constrain the answers according to some properties.
+  \item ASP efficiently tackles the inherent complexity of the used models, thus enabling us fast execution of the formal tools defined in this paper,
+  \item ASP can efficiently enumerate a large set of possible answers,
+  \item and it is easy to constrain the answers according to some properties.
 \end{itemize}
@@
 
-\subsection{Simple rules}\label{sssec:simple_rules}
+\subsection{Simple rules and answer sets}\label{sssec:simple_rules}
 ASP is based on a set of rules of of the form:
@@
 In the case of \emph{simple rules} (as opposed to the \emph{cardinality rules} of \pref{sssec:cardinality_rules}), the $head$ is also an atom ($\atom{H}$).
-Such a rule states that if all atoms $\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n$ are true
+Such a rule, whose formal semantics is defined below,
+schematically states that if all atoms $\atom{A}_1, \atom{A}_2, \dots, \atom{A}_n$ are true
 and all atoms $\atom{B}_1, \atom{B}_2, \dots, \atom{B}_m$ are not true (negation by failure), then $\atom{H}$ has to be true.
-Solving an ASP program means finding an \emph{answer set}, that is a minimal set of true atoms that respect all the rules.
-Several answer sets can be solution to the same ASP program and the solver can be directed to enumerate them all.
 
@@
 is composed of the predicate $\predicate{p}$ and $r$ arguments: $x_1, x_2, \dots, x_r$.
-Each argument can be either an constant or a variable, %value -> constant
-a constant being a representation of a piece of data (component name, expression level, …)
-and a variable is used to represent any possible existing constant which respects the rules.
-In this paper, a variable is always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, $\var{Q}$, …)
-while constants are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …).
+Each argument is either a \emph{constant}, which is a representation of a piece of data (component name, expression level, …),
+or a \emph{variable}, which is in fact a shorthand for any possible constants (variables are detailed below).
+In this paper, constants are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …)
+while variables are always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, $\var{Q}$, …).
+We do not use function symbols as arguments in this work.
 
-A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
+\begin{example}\label{ex:asp-atom}
+Consider a PH model such as \pref{ex:infer-param-runningPH-1}:
+in order to state the existence of each component, we use a predicate called $\component$ with two arguments:
+\begin{align*}
+  &\component(x, n).
+\end{align*}
+where $x$ is the name of the component and $l_x = n$ is its maximum level of expression.
+\end{example}
+
+An ASP program is a set of rules as described above.
+Solving an ASP program means finding an \emph{answer set}, which is a minimal set of atoms that respect all the rules.
+In order to formally define this notion of answer set,
+let us define a \emph{definite rule} as a rule with no negation of atoms (noted “$\nota$” above),
+and a \emph{definite program} as a program containing only definite rules.
+Let $X$ be a set of atoms: a definite rule is satisfied by $X$ if $\atom{H} \in X$,
+or if $\exists i \in \segm{1}{n}, A_i \notin X$.
+Given this definition of satisfaction, we define the answer set of a definite program $\Pi$
+as the (unique) minimal set $X$ of atoms that satisfies all the rules in $\Pi$.
+
+In order to consider the general case,
+for all non-definite program $\Pi$ and set $X$ of atoms, we denote $\Pi^X$ the reduct of $\Pi$ w.r.t.~$X$, defined from $\Pi$ by:
+\begin{enumerate}
+  \item deleting all the rules that have a negation of atom $\nota \atom{B}_i$ in the body where $\atom{B}_i \in X$, and
+  \item removing all negations of atoms in the bodies of the remaining rules.
+\end{enumerate}
+Then, a set of atoms $X$ is an answer set of a non-definite program $\Pi$ if it is the answer set of $\Pi^X$, which is a definite program.
+We note that several answer sets can be solution to the same non-definite program, and in practice the solver can be directed to enumerate them all.
+
+%\begin{example}
+%  As an example not related to the current work, consider the following ASP program:
+%  \begin{align*}
+%    a &\la \nota b.\\
+%    b &\la \nota a.
+%  \end{align*}
+%  This program has the two following answer sets: $\{a\}$ and $\{b\}$.
+%\end{example}
+
+\subsection{Facts}
+A rule with no $body$ part is called a \emph{fact}, and its $head$ atom consequently has to belong to all answer sets.
 For instance, the information describing the studied model (the original PH model and the inferred IG and parameters) are expressed in ASP using facts.
@@
 \begin{example}\label{ex:asp-component}
-Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
+In order to define the 3 components of \pref{ex:infer-param-runningPH-1}, we use the following program:
 \begin{align*}
@@
 \end{align*}
-These three atoms are built with the predicate $\component$.
-Furthermore, $a$, $b$, $c$, $1$ and $2$ are constants.
+This program contains only facts using the predicate $\component$ defined in \pref{ex:asp-atom},
+and $a$, $b$, $c$, $1$ and $2$ are constants.
 \end{example}
 
+\subsection{Variables}
 To describe the sets of all expression levels of each component (\ie the set $\segm{0}{l_a}$ for each $a \in \Gamma$),

### 2013-08-02 parts/asp.tex
fixes after Inoue's review

@@
 \section{Answer Set Programming implementation concepts}\label{sec:impl}
-
-\newcommand{\ti}[1]{\texttt{\textit{#1}}}
-\newcommand{\aspil}[1]{\texttt{#1}}
-\newcommand{\asp}[1]{\begin{itemize} \item[] \aspil{#1} \end{itemize}}
 
@@
 %\newcommand{\predicate}{\mathbf}
-\newcommand{\predicate}[1]{#1}
+\newcommand{\predicate}[1]{\mathit{#1}}
 \newcommand{\la}{\leftarrow}
@@
 and a \emph{definite program} as a program containing only definite rules.
-Let $X$ be a set of atoms: a definite rule is satisfied by $X$ if $\atom{H} \in X$,
-or if $\exists i \in \segm{1}{n}, A_i \notin X$.
+Let $S$ be a set of atoms: a definite rule is satisfied by $S$ if $\atom{H} \in S$,
+or if $\exists i \in \segm{1}{n}, A_i \notin S$.
 Given this definition of satisfaction, we define the answer set of a definite program $\Pi$
-as the (unique) minimal set $X$ of atoms that satisfies all the rules in $\Pi$.
+as the (unique) minimal set $S$ of atoms that satisfies all the rules in $\Pi$.
 
 In order to consider the general case,
-for all non-definite program $\Pi$ and set $X$ of atoms, we denote $\Pi^X$ the reduct of $\Pi$ w.r.t.~$X$, defined from $\Pi$ by:
+for all non-definite program $\Pi$ and set $S$ of atoms, we denote $\Pi^S$ the reduct of $\Pi$ w.r.t.~$S$, defined from $\Pi$ by:
 \begin{enumerate}
-  \item deleting all the rules that have a negation of atom $\nota \atom{B}_i$ in the body where $\atom{B}_i \in X$, and
+  \item deleting all the rules that have a negation of atom $\nota \atom{B}_i$ in the body where $\atom{B}_i \in S$, and
   \item removing all negations of atoms in the bodies of the remaining rules.
 \end{enumerate}
-Then, a set of atoms $X$ is an answer set of a non-definite program $\Pi$ if it is the answer set of $\Pi^X$, which is a definite program.
+Then, a set of atoms $S$ is an answer set of a non-definite program $\Pi$ if it is the answer set of $\Pi^S$, which is a definite program.
 We note that several answer sets can be solution to the same non-definite program, and in practice the solver can be directed to enumerate them all.
@@
 Variables here come in handy to enumerate each possible constant $k$ for each component $a$:
-during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible constants it could represent.
+before solving, any rule containing variables is grounded, that is replaced by an equivalent set of rules with constants only.
+%during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible constants it could represent.
 The following rule, for example, contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$) and enumerates the set of possible expression levels of each component in the system:
@@
 \end{align*}
-In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
-The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} \mid \atom{Q}$,
+In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$,
+and for which the condition in the body of the cardinality rule is met.
+The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \{ \atom{P} \mid \atom{Q} \}$,
 which is a shorthand for “the set of atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
@@
 but no upper bound is specified ($\infty$) for the size of each parameter
-(which is already bounded by the number of possible expression levels of the related component).
+(which is in fact already bounded by the number of possible expression levels of the related component).
 %each parameter contains at least $1$ value, as stated by the lower bound, and has no upper bound ($\infty$) but the number of expression levels of its component.
@@
 However, all steps of the inference presented in this paper (\pref{sec:infer-IG} \& \ref{sec:infer-K})
-were implemented in and benefited from this programming paradigm, although in significantly different ways.
+were implemented in and benefited from this programming paradigm.
+%, although in significantly different ways.

### 2013-08-02 parts/discussion.tex
minor change in Conclusion
fixes after Inoue's review

@@
 \section{Conclusion and Discussion}
-
-\todo{Complete/extend}
 
@@
 The PH allows an abstract representation of BRNs dynamics (allowing incomplete knowledge on the
-cooperation between components) that can not be exactly represented in Ren\'e Thomas' formalism by a
+cooperation between components) that cannot be exactly represented in Ren\'e Thomas' formalism by a
 single instance of BRN parametrization.
@@
 The enumeration of such parametrizations is efficiently tackled using Answer Set Programming.
-We illustrate the overall method on several large biological models.
+We illustrated the overall method with several results on large biological models.
 
@@
 in the IG.
-Because of its atomicity, the PH allows to specify a range of cooperations that can not be
+Because of its atomicity, the PH allows to specify a range of cooperations that cannot be
 completely captured by a single instance of BRN multiplexes, then encouraging the inference of a set
@@
 
+\todo{More perspectives}
+
 \paragraph{Acknowledgement}

### 2013-08-02 parts/grn.tex
cleanup
remove comments + typos
fixes after Inoue's review

@@
 The sign of an edge denotes the kind of regulation it models: it can be positive ($+$), negative ($-$) or unsigned ($\uns$),
-the latter meaning that the nature of the regulation can not be reduced to a simple activation or inhibition.
+the latter meaning that the nature of the regulation cannot be reduced to a simple activation or inhibition.
 Regarding the dynamics, an edge $\GRNedge{a}{s}{t}{b}$ states that $a$ influences the evolution of $b$ in a certain way when its expression level is above or equal to the threshold $t$;
@@
 \label{eq:grn-regulators}
-  \GRNreg{b} \DEF \{ b\in\Gamma\mid \exists \GRNedge{a}{s}{t}{b} \in E \}
+  \GRNreg{b} \DEF \{ a\in\Gamma\mid \exists \GRNedge{a}{s}{t}{b} \in E \}
 \end{align}
@@
 
-\updated{
 \paragraph{Interaction Graph of the dynamics}
@@
 (e.g., \cite{Richard2010378,PR11-SASB}).
-}
 
@@
 let $s$ be a state such that $\GRNres{a}{s} = \omega$.
-From \pref{def:dynamics}, if $\GRNget{s}{a} \in K_{a,\omega}$, then $a$ can not evolve in $s$;
+From \pref{def:dynamics}, if $\GRNget{s}{a} \in K_{a,\omega}$, then $a$ cannot evolve in $s$;
 therefore, $a$ has three (local) fixed points ($i$, $i+1$ and $i+2$).
-This behavior can not be reproduced with single value parameters, because it is impossible to distinguish three different sets of resources corresponding to the three involved states.
+This behavior cannot be reproduced with single value parameters, because it is impossible to distinguish three different sets of resources corresponding to the three involved states.
 At most two steady-states can be modeled by using a self-regulation $\GRNedge{a}{s}{t}{a}$, allowing to distinguish two cases ($\GRNget{s}{a} < t$ and $\GRNget{s}{a} \geq t$), which is not enough.

### 2013-08-02 parts/discussion.tex
remove last TODOs

@@
 Several extensions of the presented work are now to be considered.
-First, the inference of BRN multiplexes \cite{BernotMultiplexes} may be of practical interest 
+First, further work on the inference of the IG would lead to better understand and possibly avoid
+the presence of some of the current spurious auto-edges.
+Second, the inference of BRN multiplexes \cite{BernotMultiplexes} may be of practical interest 
 as they allow to implicitly reduce the possible parametrizations by making cooperations appear
@@
 of compatible ones.
-Second, in order to improve the performances in the IG inference, we will consider projection operations on
+Finally, in order to improve the performances in the IG inference, we will consider projection operations on
 the PH structure to undo cooperations between components and reduce the cardinality of
 configurations to explore by making the interactions independent.
-
-\todo{More perspectives}
 

### 2013-08-02 parts/discussion.tex
new perspective in Conclusion

@@
 Several extensions of the presented work are now to be considered.
-First, further work on the inference of the IG would lead to better understand and possibly avoid
+First, the link between successively refined models of a system could be formally studied.
+Indeed, it is convenient to refine a PH model by removing actions or adding cooperations;
+the study of such process would allow to predict behavioral changes and lead to more accurate models.
+Such work would also require further study on the inference of the IG in order to better understand and possibly avoid
 the presence of some of the current spurious auto-edges.

### 2013-08-02 parts/example.tex
reference to CellNetAnalyzer
global proofread; final version 1

@@
 Clingo\footnote{Available at \url{http://potassco.sourceforge.net}}.
-The IG and parameters inference can be performed using the following command:\\
-  \hspace*{1.2em}\texttt{ph2thomas -i model.ph -{}-dot ig.dot}\\
-where \texttt{model.ph} is the PH model in \textsc{Pint} format,
-and \texttt{ig.dot} is an output of the inferred IG in DOT format.
+The IG and parameter inferences can be performed using the following command:\\
+  \hspace*{\parindent}\texttt{ph2thomas -i model.ph -{}-dot ig.dot}\\
+where \texttt{model.ph} is the PH model file in \textsc{Pint} format,
+and \texttt{ig.dot} is an output file to write the inferred IG in DOT format.
 The (possibly partial) inferred parametrization will be returned on the standard output.
@@
   and all of them were included in Model (3) under the form of 22 cooperative sorts.
-%  Model (3) models all the Boolean functions between components using 22 cooperative sorts.
   The second column gives the number of edges in the IG inferred with \pref{pps:inference-IG}
@@
   the fourth column gives the number of parameters that could be inferred using \pref{pps:param_K},
-  and the fifth column consequently gives the number fo compatible models with the studied PH model,
+  and the fifth column consequently gives the number of compatible models with the studied PH model,
   which exponentially depends on the number of parameters that could not be inferred.
   Finally, the last column gives the number of fixed points in the PH model,
-  computed with another existing PH tool included in \textsc{Pint}.
+  computed with another existing PH tool provided with \textsc{Pint}.
   }
@@
 with one additional positive auto-edge on the only input EGF (which is due to its absence of regulators).
-The only parameters that could be inferred are the parameters for the extreme cases
-(all activators present and all inhibitors absent, and the reverse).
+The only parameters that could be inferred are the parameters for the extreme cases of regulation
+(all activators present and all inhibitors absent, and the opposite).
 This first model therefore abstracts a large number of Thomas models as a lot of parameters are left undecided.
@@
 in \pref{ssec:PH} was used to reduce the number of processes in each cooperative sort.
-We note that the added cooperations allowed to infer about half the parameters.
-However, the number of possible Thomas models that can be inferred from this PH is still significant
+As a result, the added cooperations allowed to infer about half the parameters;
+however, the number of possible Thomas models that can be inferred from this PH is still significant
 because of the numerous remaining unknown parameters.
-Furthermore, the inferred IG contains one edge less than the original IG. This is due to the fact that
-one of the Boolean functions could in fact be simplified in a way that this component did not appear anymore.
+Furthermore, we note that the inferred IG contains one edge less than the original IG. This is due to the fact that
+one of the Boolean functions could in fact be simplified in a way that a component did not appear anymore in it.
 No edge have therefore been inferred by our method in this case.
@@
 As all cooperations are fully defined in this model, all the parameters are inferred and only one Thomas model can be derived.
-We note also that this PH model is the only one containing at least one fixed point,
-which indicates the possibility of a steady state once the input signal is completely propagated.
+We note also that this PH model is the only one containing at least one fixed point.
+In fact, the three found steady states include the two states that correspond to a complete propagation of the input signal,
+that is, in the case where EGF is active and in the case where it is not.
+The two other models contain no fixed point because some cooperations are not fully defined,
+leading to oscillations that are a consequence of the nondeterministic behavior.
 
@@
 \begin{itemize}
-  \item the EGF receptor model from~\cite{Sahin09} with 20 components presented in the last subsection,
-  \item a T cell receptor model described as an IG in~\cite{Klamt06}, which contains 40 components and 14 cooperative
-    sorts\footnote{All models mentioned in this section are available as examples distributed with \textsc{Pint}.}.
+  \item the EGF receptor model from~\cite{Sahin09} with 20 components presented in the last
+    subsection\footnote{All models mentioned in this section are available as examples distributed with \textsc{Pint}.},
+  \item a T cell receptor model described as an IG in~\cite{Klamt06}, which contains 40 components and 14 cooperative sorts.
 \end{itemize}
@@
 \end{itemize}
-These two models were obtained in a previous work by an automatic translation from the CellNetAnalyzer formalism.
+These two models were obtained in a previous work by an automatic translation from the CellNetAnalyzer~\cite{klamt2007structural} formalism.
 

### 2013-08-03 parts-asp.tex
all source files in the same folder (TCS submission)

--- 2013-08-02 parts/asp.tex
@@
 \end{align*}
-where $x$ is the name of the component and $l_x = n$ is its maximum level of expression.
+where $x$ is the name of the component and $l_x = n$ is its maximum expression level.
 \end{example}
@@
 and a \emph{definite program} as a program containing only definite rules.
-Let $S$ be a set of atoms: a definite rule is satisfied by $S$ if $\atom{H} \in S$,
+Let $S$ be a set of atoms: a definite rule is \emph{satisfied} by $S$ if $\atom{H} \in S$,
 or if $\exists i \in \segm{1}{n}, A_i \notin S$.
@@
 Then, a set of atoms $S$ is an answer set of a non-definite program $\Pi$ if it is the answer set of $\Pi^S$, which is a definite program.
-We note that several answer sets can be solution to the same non-definite program, and in practice the solver can be directed to enumerate them all.
+We note that several answer sets can be solution to the same non-definite program, and in practice a solver can be directed to enumerate them all.
 
@@
 
-\subsection{Facts}
-A rule with no $body$ part is called a \emph{fact}, and its $head$ atom consequently has to belong to all answer sets.
+Note that it is possible to define a simple rule with no $body$ part.
+Such a rule is called a \emph{fact}, and its $head$ atom consequently has to belong to all answer sets.
 For instance, the information describing the studied model (the original PH model and the inferred IG and parameters) are expressed in ASP using facts.
-In particular, the predicate $\component$ allows to define all components belonging to the inferred IG.
-Thus, an atom $\component(a, m)$ states that $a$ is a component of the IG and that $l_a = m$.
 
@@
 Variables here come in handy to enumerate each possible constant $k$ for each component $a$:
-before solving, any rule containing variables is grounded, that is replaced by an equivalent set of rules with constants only.
+before solving, any rule containing variables is \emph{grounded}, that is, replaced by an equivalent set of rules with constants only.
 %during the solving, any rule containing variables is duplicated in order to replace each variable by all the possible constants it could represent.
@@
 \subsection{Constraints}\label{sssec:constraints}
-Finally, a constraint is a rule with no $head$ part:
+Finally, a \emph{constraint} is a rule with no $head$ part:
 \begin{align*}
@@
 Then, due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
-the monotonicity assumption (\pref{pro:param_enum_monotonicity}) removes all parametrizations with: $0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
+the monotonicity assumption (\pref{pro:param_enum_monotonicity} and expressed in ASP above) removes all parametrizations with:
+$0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
 In the end, the remaining possible values for the two parameters that were not inferred are:

### 2013-08-03 parts-discussion.tex
all source files in the same folder (TCS submission)

--- 2013-08-02 parts/discussion.tex
@@
 Then, based on a prior inference of Ren\'e Thomas' parametrization for BRNs from a PH model, we
-delimit the set of compatible Thomas' parametrizations that are compatible with the PH dynamics,
+delimit the set of admissible Thomas' parametrizations that are compatible with the PH dynamics,
 and give arguments for their correctness.
@@
 The enumeration of such parametrizations is efficiently tackled using Answer Set Programming.
-We illustrated the overall method with several results on large biological models.
+We illustrate the overall method with several results on large biological models.
 
@@
 Indeed, it is convenient to refine a PH model by removing actions or adding cooperations;
-the study of such process would allow to predict behavioral changes and lead to more accurate models.
+the study and formalization of such process would allow to predict behavioral changes and lead to more accurate models.
 Such work would also require further study on the inference of the IG in order to better understand and possibly avoid

### 2013-08-03 parts-grn.tex
all source files in the same folder (TCS submission)

--- 2013-08-02 parts/grn.tex
@@
 \end{definition}
-\todo{Discuss on the widespread use of discrete parameters + binary resources}
+
+%\todo{Discuss on the widespread use of discrete parameters + binary resources}
 An IG and a parametrization make up a complete BRN.
@@
 Let us define
-$f^a(s) = s[a]$ if $s[a]\in K_{a, \GRNres{a}{s}}$, 
-$f^a(s) = s[a]+1$ if $s[a] < K_{a, \GRNres{a}{s}}$, 
+$f^a(s) = s[a]$ if $s[a]\in K_{a, \GRNres{a}{s}}$
 and
-$f^a(s) = s[a]-1$ if $s[a] > K_{a, \GRNres{a}{s}}$.
+$f^a(s) = s[a]+\delta^a(s)$ otherwise.
 A positive (resp. negative) interaction from $b$ to $a$ is inferred if there exists a state $s$ such
-that increasing $b$ in $s$ would increase (resp. decrease) $f^a$:
-$f^a(s\{b_i\}) < \text{ (resp. $>$) } f^a(s\{b_{i+1}\})$, 
+that increasing $b$ in $s$ would increase (resp. decrease) $f^a$; in other words, if
+there exists a state $s$ so that:
+$f^a(s\{b_i\}) < \text{(resp. $>$) } f^a(s\{b_{i+1}\})$, 
 where $i < l_b$ and $s\{b_i\}$ denotes the state $s$ where the component $b$ is assigned to $b_i$.
@@
 The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$ in the IG, \ie the number of components that $a$ regulates.
-Indeed, each discrete expression level in $\segm{0}{l_a}$ stands for an arbitrary set of concentration levels for which $a$ has a certain influence on several other components it regulates.
+Indeed, each discrete expression level in $\segm{0}{l_a}$ stands for an arbitrary set of concentration levels of the protein coded by $a$
+and for which it has a certain influence on several other components that it regulates.
 Allowing more expression levels makes several levels useless as they represent the same influence of $a$ than some other levels;
@@
 it would be more relevant to have $l_a = 1$ and $\GRNedgef{a}{-}{1}{b} \in E$, as $a$ has only one outgoing edge,
-but the value $l_a = 2$ makes the example more interesting, therefore $\GRNedgef{a}{-}{2}{b} \in E$ can be arbitrarily chosen.
+but the value $l_a = 2$ makes the example more interesting; this is why $\GRNedgef{a}{-}{2}{b} \in E$ can be arbitrarily chosen.
 

### 2013-08-03 parts-infer-ig.tex
all source files in the same folder (TCS submission)

--- 2013-07-24 parts/infer-ig.tex
@@
 there is no cycle between cooperative sorts, and that
-sorts being never hit (\ie{} serving as an invariant environment) are components.
+sorts being never hit (\ie serving as an invariant environment) are components.
 
@@
 At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
-($\PHs \setminus \Gamma$) that will not appear in the IG. 
+that will not appear in the IG ($\PHs \setminus \Gamma$).
 We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
@@
 refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
-If $g=\{b\}$, we simple note $\ctx_b(\sigma)$.
+If $g=\{b\}$, we simply note $\ctx_b(\sigma)$.
 This set is composed of the active processes of sorts in $g$, and the focal process (assumed
@@
 We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
+% Arcs a -> b, a ≠ b
 \begin{align}
-% Arcs a -> b, a ≠ b
 \begin{split}\label{eq:edges-inference}
@@
     &\qquad s = \f{sign}(k - a)
-\end{split}\\
-% Auto-arcs
+\end{split}
+\end{align}
+% Auto-arcs depuis les groupes de régulateurs
+\begin{align}
 \begin{split}\label{eq:edges-inference-auto}
@@
     &\qquad s = \f{sign}(a_k - a_j)
-\end{split}\\
+\end{split}
+\end{align}
+% Auto-arcs des composants sans prédécesseurs
+\begin{align}
 \begin{split}\label{eq:edges-inference-noreg}
@@
 \end{align}
-where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
+where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$ \enspace.
 \end{proposition}

### 2013-08-03 parts-infer-param.tex
all source files in the same folder (TCS submission)

--- 2013-07-09 parts/infer-param.tex
@@
 \begin{align*}
-  K_{a,\{a,b\}} &= 1 &
-  K_{a,\{a,c\}} &= 1
+  K_{a,\{a,b\}} &= \segm{1}{1} &
+  K_{a,\{a,c\}} &= \segm{1}{1}
 \end{align*}
@@
 Given the \pref{pps:param_K}, we see that in some cases, the inference of the targeted parameter is impossible.
-This can be due to a lack of cooperation between regulators: when two regulators independently hit a component, their actions can have opposite effects, leading to either an indeterministic evolution or to oscillations.
-Such an indeterminism is not possible in a BRN as in a given configuration of regulators, a component can only have an interval attractor, and eventually reaches a steady-state.
+This can be due to a lack of cooperation between regulators:
+when two regulators independently hit a component, their actions can have opposite effects, leading to two possible evolutions.
+Such an indeterminism is not possible in Thomas modeling as in a given configuration of regulators,
+a component can only have an interval attractor, and can therefore evolve in only one direction.
 In order to avoid such inconclusive cases, one has to ensure that no such behavior is allowed by
 either removing undesired actions or using cooperative sorts to prevent opposite influences between
-regulators.
+concurrent regulators.
 
@@
 When building a BRN, one has to find the parametrization that best describes the desired behavior of the studied system.
-Complexity is inherent to this process as the number of possible parametrizations for a given IG is exponential w.r.t. the number of components.
+Complexity is inherent to this process as the number of possible parametrizations for a given IG is exponential w.r.t.~the number of components.
 However, the method of parameters inference presented in this section gives some information about necessary parameters given a certain dynamics described by a PH.

### 2013-08-03 parts-intro.tex
all source files in the same folder (TCS submission)

--- 2013-07-23 parts/intro.tex
@@
 This framework can be seen as a special class of formalisms like Petri Nets or Communicating Finite
-State Machines, where the arity and the effect of synchronizations between components is
+State Machines, where the arity and the effect of synchronizations between components are
 restricted.
@@
 The obtained IG relates only components that have actual influence in the dynamics.
-This typically results into IGs that contain less interaction than an hypothetical starting prior
+This typically results into IGs that contain less interactions than an hypothetical starting prior
 IG, as the specification of component evolutions may reveal non-functional regulations.
@@
 positive cycle to the impossibility for multi-stationarity (distinct attractors),
-and the absence of negative cycle to the impossibility for oscillatory behaviours.
+and the absence of negative cycle to the impossibility for oscillatory behaviors.
 
-Second, we tackle the characterisation of Thomas' parameters that are compatible with a given PH
+Second, we tackle the inference of Thomas' parameters that are compatible with a given PH
 model, \ie that fully specify the evolution of components while respecting the cooperations
@@
 few cooperations are specified.
-This highlights the fact that the PH is very suited to analyse dynamics of large-scale networks
+This highlights the fact that the PH is very suitable to analyze dynamics of large-scale networks
 where the knowledge on cooperations is often incomplete:
 instead of dealing with a possibly non-tractable number of Thomas models, one can already capture
-precise dynamical features using a single PH models which gathers all the plausible dynamics.
+precise dynamical features using a single PH model which gathers all the plausible dynamics.
 
@@
 Finally, it must be noticed that we are not interested in this paper in the derivation of one PH
-from a BRN (which was previously described in \cite{PMR10-TCSB}) but, on the contrary, to finding
+from a BRN (which was previously described in \cite{PMR10-TCSB}) but, on the contrary, in finding
 out a set of BRNs from one PH.
@@
 The contributions presented in this paper significantly extend and improve the preliminary results
-presented in \cite{FPIMR12-CMSB}.
-In addition of improving the efficiency and accuracy of the IG inference, we have added support for
+introduced in \cite{FPIMR12-CMSB}.
+In addition to the improvement of the efficiency and accuracy of the IG inference, we have added support for
 unsigned regulations, that are regulations being positive or negative depending on a particular

### 2013-08-03 parts-ph.tex
all source files in the same folder (TCS submission)

--- 2013-07-09 parts/ph.tex
@@
 of the realization of an instantaneous cooperation.
+However, it is possible to obtain the exact behavior of the cooperations by using the semantics with prioritized actions defined in~\cite{FPMR13-CS2Bio}.
+Although this new semantics is not further described here, the inferences given in \pref{sec:infer-IG} \& \ref{sec:infer-K} do ignore the temporal shift induced by the cooperative sorts.
 

### 2013-09-09 parts-infer-ig.tex
fix mistake in influences inference equations; version 1.2

@@
     &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+    &\qquad s = \f{sign}(k - j)
+\end{split}
+\end{align}
+% Auto-arcs depuis les groupes de régulateurs
+\begin{align}
+\begin{split}\label{eq:edges-inference-auto}
+  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a),
+  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+    &\qquad \{ a_t, a_{t+1} \} \subset \PHl_a \wedge a_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{a_{t+1}\}), \\
+    &\qquad s = \f{sign}(k - j)
+\end{split}
+\end{align}
+% Auto-arcs des composants sans prédécesseurs
+\begin{align}
+\begin{split}\label{eq:edges-inference-noreg}
+  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
+  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
+    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
+    &\qquad s = \f{sign}(k - j)
+\end{split}
+\end{align}
+\begin{comment}
+% Arcs a -> b, a ≠ b
+\begin{align}
+\begin{split}\label{eq:edges-inference}
+  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
+  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
+  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
     &\qquad s = \f{sign}(k - a)
@@
 \end{align}
+\end{comment}
 where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$ \enspace.

### 2014-08-29 main.tex
add keywords; final version 1.1
all source files in the same folder (TCS submission)
add TODOs after the review

@@
 
-\input{macros/macros}
-\input{macros/macros-ph}
-\input{macros/macros-grn}
+\input{macros}
+\input{macros-ph}
+\input{macros-grn}
 
 % Commandes À FAIRE
-%\usepackage{color} % Couleurs du texte
-%\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
-%\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
-%\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
+\usepackage{color} % Couleurs du texte
+\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
 %\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
-%\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
-\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
-%\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
-%\def\updated#1{{\color{magenta}#1}}
-%\newcommand{\updatedmax}[1]{{\color{violet}#1}}
-%\newcommand{\new}[1]{#1}
 
@@
 The Process Hitting (PH) is a recently introduced framework to model concurrent processes.
+\todo{The reviewer did not understand the following sentence:}
 Its major originality lies in a specific restriction on the causality of actions, which
@@
 
+\todo{Rather name Kaufman as the creator of BRNs.}
 On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
@@
 
+\begin{keyword}
+qualitative modeling \sep
+model abstraction \sep
+interaction graph \sep
+René Thomas' parameters \sep
+answer set programming
+\end{keyword}
+
 \end{frontmatter}
 
-\input{parts/intro}
+\input{parts-intro}
 
 \section{Frameworks}\label{sec:frameworks}
-\input{parts/ph}
-\input{parts/grn}
+\input{parts-ph}
+\input{parts-grn}
 
-\input{parts/infer-ig}
-\input{parts/infer-param}
-\input{parts/asp}
-\input{parts/example}
-\input{parts/discussion}
+\input{parts-infer-ig}
+\input{parts-infer-param}
+\input{parts-asp}
+\input{parts-example}
+\input{parts-discussion}
 

### 2014-08-29 parts-grn.tex
add TODOs after the review

@@
 \subsection{Thomas' modeling}
+
+\todo{Name Kaufman as the creator of BRNs.}
 
@@
 
+\todo{Discuss about the meaning of signs (especially $+$ and $-$) and their link with parameters.
+Especially, these signs introduce redundancy (and possibly contradiction) with the parameters.
+\begin{itemize}
+  \item We can discuss more about the choice we made here: the signs are only informative
+    (it eases the read of the IG), but are used as a basis for the parameters enumeration;
+  \item The reviewer suggests to remove signs, or only present them as an abstraction of the monotonicity assumption, given the levels and parameters of a BRN;
+  \item It is also possible to add the monotonicity assumption in the definition of a BRN to avoid contradictions.
+\end{itemize}}
+
 \begin{definition}[Interaction Graph]
@@
 \end{definition}
+
+\todo{The notation $\GRNreg{b}$ is poorly chosen; should be replaced by $E^{-1}(b)$ for instance.}
 
@@
 \paragraph{Discussion on maximum level restrictions}
+\todo{The reviewer says that the discussion on the number of maximum levels has
+already been fully developed in other articles related to Thomas and that it is not
+necessary to develop this just to state that a PH may accept any number of levels.}
 The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$ in the IG, \ie the number of components that $a$ regulates.
@@
 \paragraph{Discussion on unsigned edges}
+\todo{Discuss the meaning of signs (cf. supra).
+Also, the reviewer proposes to move this discussion before in the text,
+and to define unsigned IGs before signed IGs as the latter is just a consequence of the monotonicity assumption.}
 Unsigned edges allow to model regulations whose trend is more complex than a simple activation or inhibition.

### 2014-08-29 parts-infer-ig.tex
add TODOs after the review

@@
 \section{Interaction Graph Inference from Process Hitting}\label{sec:infer-IG}
+
+\todo{In this while section: add informal explanations and maybe more examples.\\
+Especially:
+\begin{itemize}
+  \item Intuitively detail the motivations
+  \item Give the intuition of the methods (esp. $\focals$)
+  \item Insist on the use/motivation of the definitions
+\end{itemize}
+The reviewer also proposes to use a biological running example (such as $\lambda$-phage).}
 
@@
 properties for the concrete system, such as the capability for multi-stationarity or oscillation.
+
+\todo{The reviewer says about the following paragraph: “The inference of an
+IG from a dynamics is interesting and important in biology. However the
+motivations expressed here are not well argued. In particular I don't find
+this “common”. Some bibliographical references are required to consolidate
+the arguments or the sentence has to be rephrased.”}
 
@@
 \subsection{Focal Processes}\label{ssec:focal}
+
+\todo{The two following subsections (down to “Interaction Inference”) have a lot of definitions but few examples and not enough informal explanations, especially on their use.
+Especially:
+\begin{itemize}
+  \item What kinds of concept do represent $\focals$?
+  \item What is the importance of a strict context?
+  \item Add an outline guiding the reader through these subsections.
+\end{itemize}}
 
@@
 \begin{example}
-The IG inference of the PH of \pref{fig:runningPH-1} gives the
+The IG inference of the PH of \pref{fig:runningPH-2} gives the
 IG in \pref{fig:BRN-inf1}, containing the following edges:
@@
 
-The inference of the refined PH of \pref{fig:runningPH-2} gives the same IG.
+The inference of the PH of \pref{fig:runningPH-1}
+(without refinement with cooperative sort) gives the same IG.
 

### 2014-08-29 parts-intro.tex
add TODOs after the review

@@
 specify Thomas' parameters, especially for large networks.
+\todo{The reviewer proposes to nuance this with several papers:
+\begin{itemize}
+  \item Corblin, F., Tripodi, S., Fanchon, E., Ropers, D., \& Trilling, L. (2009).
+    A declarative constraint-based method for analyzing discrete genetic
+    regulatory networks. Bio Systems, 98(2), 91104.
+  \item Bollman, D., Coln-Reyes, O., \& Orozco, E. (2007). Fixed points in
+    discrete models for regulatory genetic networks. EURASIP journal on
+    bioinformatics \& systems biology, 2007, 97356.
+  \item Delaplace, F., Klaudel, H., Melliti, T., \& Sen, S. (2012). Analysis
+    of modular organisation of interaction networks based on asymptotic
+    dynamics. In D. Gilbert \& M. Heiner (Eds.), CMSB 2012
+  \item Scalable Steady State Analysis of Boolean Biological Regulatory Net-
+    works Ay F, Xu F, Kahveci T (2009) Scalable Steady State Analysis
+    of Boolean Biological Regulatory Networks. PLoS ONE 4(12): e7992.
+\end{itemize}
+This can be in turn be nuanced as some papers only focus on rather specific aspects (like steady-states for instance).}
 
@@
 cooperations (associated influences) between components.
+\todo{Explain the notion of “cooperation” (informally) to be understandable until the definition of cooperative sorts.}
 This allows to model BRNs with a partial knowledge on precise evolution functions for components

### 2014-08-29 parts-ph.tex
add TODOs after the review

@@
 \begin{example}
+\todo{\pref{fig:runningPH-1} lacks readability (arrows hitting $a$ are too small).}
 \pref{fig:runningPH-1} represents a PH $(\PHs,\PHl,\PHa)$ with
@@
 
+\todo{The reviewer did not understand the following paragraph about the time shift.
+He wonders if it is really important to the understanding of the rest of the paper. Three possibilities here:
+\begin{enumerate}
+  \item Add more (informal? formal?) explanations,
+  \item Remove (or substantially reduce) this paragraph,
+  \item Use priorities and prioritized cooperative sorts to remove the temporal shift.
+\end{enumerate}
+The last possibility would be formally the most accurate but also the most time-consuming one.}
+
 While the construction of cooperation in PH allows to encode any boolean functions

### 2014-09-03 response.tex
discusson on the review

\documentclass[11pt]{article}
% vim:set spell spelllang=en:
\usepackage{color}
\usepackage{fullpage}

% ----- parameters ----------------------------------------------------------------
\def\BibTex{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
         T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
%\def\baselinestretch{1.5}
\usepackage{amsmath}
\usepackage{amssymb}

% to import paper citations and refs
% generated with 
% cat ../v5/paper.aux  | sed 's/\\@/% \\@/' > paper.aux
%\input{paper.aux}

% ----------------------------------------------------------------------------------
\newcounter{pointno}
\setcounter{pointno}{0}
\def\thepoint{\arabic{pointno}}
\newcommand{\ptno}[1]{\textcolor{blue}{\refstepcounter{pointno} \thepoint. #1}}
\newcommand{\question}[1]{\bigskip{\it\ptno{#1}}}
\def\answer{}

\def\TODO#1{\textbf{[TODO: #1]}}
\def\DISCUSS#1{\textbf{DISCUSSION(#1): }}

\setlength{\parindent}{0pt}

\def\V{\mathcal V}
\def\O{\mathcal O}

\def\zero{\mathbf 0}
\def\R{\mathbb R}
\def\Rpos{\R_{\geq 0}}
\def\Rspos{\R_{>0}}
\def\Q{\mathbb Q}
\def\Qpos{\Q_{\geq 0}}
\def\Qspos{\Q_{>0}}
\def\Z{\mathbb Z}
\def\Zpos{\Z_{\geq 0}}
\def\Zspos{\Z_{> 0}}

\def\mspan{\mathrm{span}}
\def\lp{\mathrm{lowerpoint}}
\def\orderings{\mathrm{orderings}}

\def\inner#1#2{#1#2}
%\def\inner#1#2{\langle #1,#2 \rangle}

\def\apply{\bullet}

\def\B{\mathcal B}

\begin{document}

\section*{Comments and answers to the reviewer}

We thank the reviewer for th comments and  suggestions.

\TODO{summary of response + changes}

Please find below our detailed answers.

\question{Beside, the definition of RT model, including signs that can be
deduced
from the K parameters must be justified or clarify. For a computer scientist
it is always surprising to have redundant pieces of information without accounting their matching.
With your definition I can write $a \xrightarrow+ b$ and I can
set $K_{b,\emptyset} = 1$ and $K_{b,\{a\}} = 0$.
On the one side I define a as an activator and
on the other side I define a as an inhibitor with K parameters. As signs do
not enter in the interpretation, why they are included in the formalism as
necessary ``comments''?}

\DISCUSS{LP}
We have to clarify this point, while insisting that we want to stick with this
standard definition (due to Thomas). Arguments:
\begin{enumerate}
\item this design allows to decouple the IG from the dynamics, which is
interesting for modelling biological systems as the IG is often a prior
knowledge with possible inconsistencies with the dynamics we want to encode
with.
A further step may then infer back the IG of the actual dynamics (here, an
abstraction of the encoded dynamics) and compare it with the prior IG. Typically, if the inferred IG is included in the prior
one, we have consistency.
\item Additional constraints (as discussed in sect XX) on parameters can also
ensure consistency between the prior IG and the encoded dynamics.
For instance, monotonicity (which imposes an ordering between the value of
parameters), but also lighter constraints (like imposing minimum and maximum
values).
Hence, the K definition is very generic, ranging from totally decoupled
specification to very constrained valuation w.r.t to the prior IG.
\item Several alternative definition exist, with various degree of equivalency.
We picked this definition as it the original and standard one in the Thomas
community (give reference to Thomas' articles).
\end{enumerate}

\TODO{Answer inc. text modification}

\question{
(P.1) Abstract: the following sentence is not understandable. Either more explanations
are needed or simply suppress it.
``Its major originality lies in a specific restriction on the causality of actions, which makes the formal
analysis of very large systems tractable.''
}

\answer
\DISCUSS{LP}
proposition: ``The PH is a particular class of Indeterministic Asynchronous Automata
Network (or safe Petri nets) on which have been designed static analysis for
dynamical properties that are tractable on very large systems.''

\TODO{Modify sentence + answer (quote)}

\question{
The BRNs were introduced by S. Kauffman in 1969. His work also covers
“numerous theoretical works”. Therefore, it would be interesting to mention this
fact. Ren\'e Thomas is not the father of BRN applied to regulatory
circuit dynamics. Beside, the definition of the BRN for S. Kaufman consists
of a Boolean system dynamics with a Boolean function which is a normal
way to define a dynamical system. The R.T.'s formalization is slightly unconventional, namely not based on a stepwise function.}

\answer
\DISCUSS{LP} yes, we have to mention Kauffman of course, while stressing the
differences brought by Thomas (async + parameters).

\question{
For Boolean networks, symbolic methods and parsimonious generation of
states exist to circumvent the generation of the whole state graph and then
scaling up the performances of the analysis (e.g. equilibria computation).
Therefore, claiming that it is a ``major difficulty'' is partly true only.
}

\answer
\DISCUSS{LP} yes, we have to mention those work, while stressing that they have
hard time with transient dynamics on very large networks, partly due to
concurrency, which makes model-checking very hard even with symbolic techniques.

\DISCUSS{LP} how such a discussion is relevant for the paper? Can't we simply
say that PH allows to model indeterministic Boolean and multi-valued networks
and remove (or drastically reduce) this discussion on model-checking?

\question{
(P.3 and 6) The term “cooperation” is often mentioned. It is not self-
explanatory for the reader. In particular in P.3 “associated influences” is
not explanatory at all. If I have missed it I suggest to more clearly emphasize it. This comment does not concern cooperative sort which is clearly
explained, but the notion of cooperation that cooperative sort encompasses.
A definition of “cooperation” is required (formal or not) or more explanations of the notion of cooperation in this context.
}

\DISCUSS{LP} we should avoid this term in introduction, and only use once
formally defined.
In the introduction, maybe we should mention that
PH models can embed additional automata that encode for some (partial) logic
gates. Such additional automata that can be considered as ``internal''
variables, and we may want to mask them in abstract representations of the
systems, such as IG and Thomas' parameters.

\question{
(P. 6) The arrows representing the hitting of sort a fig 1. are too small.
}

\answer
\TODO{improve figure}

\question{
(P. 8)
``This allows the existence of interleaving of actions leading .. semantics with
prioritized actions defined in [13]''.
Again I don't understand the message related to your own notion of cooperation. Is it really important to
mention these elements?}

\answer
\DISCUSS{LP} I would go for removing this discussion.

\question{
(P. 9) Interaction Graph. $\Gamma$ refers to the set of vertices. However
$\Gamma^{−1}(b)$
requires to seek the pre-image in E. I find the choice of $\Gamma$ to annotate
a function giving the regulators confusing. If we admit that E defines a
relation, $Eb$ or $E^{-1} (b)$ are more natural notations for the pre-image or more
simply $b \rightarrow$ for the image and $\rightarrow b$ for the pre image. These notations
are classically used for relation to qualify these sets of elements and do not
require additional notations.
}

\answer
\TODO{Apply the changes + quick answer}

\question{
(P. 8) You mention that Thomas formalism lies on two complementary descriptions of the system IG and K parameters. However There exists some
redundancies for these description because the sign may be deduced from the
parameters representing monotonous properties of the parameters. Moreover, The dynamics is not deduced from signs. IG graph with levels and
parameters appears to suffice for fully describing the RT model. It would
be more parsimonious (for an article in TCS) to define an IG with levels
only, and parameters corresponding to Thomas’ model and then explain
that signs are abstractions of the parameters corresponding to monotony
properties.
}

\answer
\DISCUSS{LP} See 1. The definitions need to be more clear with more textual
explanation for CS people.

\question{
(P. 11) Definition of the sign. The term of monotony is classically associated
to the inference of signs and should be clearly mentioned. Here it is hidden
in your definition. The discussion on the number of maximum levels has
already been fully developed in other articles related to RT formalisms. Is
it necessary here just for mentioning that PH may accept any level?
}

\answer
\DISCUSS{LP} Yes, we should make clear that we are looking for monotonic
influences - but not necessarily at the global scope, which mean that we can
find two contradictory monotonic influences, leading to those ``unsigned'' edges
in the IG.
And yes, we should just mention that PH models can have an arbitrary number of
levels.

\question{
(P. 12) The order of your explanations makes the text confusing. Signs
are components of the definition of an IG. After you argue that it is useless
because parameters suffice to deduce the state graph. You mention that they
can be inferred although they must be included in the definition of an IG.
The formalization is not clear. Unsigned IG could/must be presented before
signed IG which is an abstraction of the monotony of the K parameters.
}

\answer
\DISCUSS{LP} Same as 1 and 9.

\question{
(P. 12) ``However, it is common that the prior IG actually refers to interaction
.. enhancing the conclusiveness of static analyses''.
The inference of an IG from a dynamics is interesting and important in biology. However the
motivations expressed here are not well argued. In particular I don't find
this "common". Some bibliographical references are required to consolidate
the arguments or the sentence has to be rephrased.
}

\answer
\DISCUSS{LP}
Indeed, the motivation is quite weak. Some arguments:
\begin{enumerate}
\item Abstract representation of influences
\item Consistency with prior knowledge: if the modelling started from an IG, we may want to
verify that the dynamical models is consistent w.r.t. to the prior IG
\item + remove non-used influences (IG reduction).
\end{enumerate}

\question{
(P. 13 -14 - 15) 3 pages contain 5 definitions with no example and no
intuition on their use but for $\Gamma$ and focal. More explanations are required:
what kinds of concept do represent focal, what is the importance of a strict
context, where is the outline guiding the reader through the dense area of
definitions located in three pages only? These pages must be rewritten and
extended to understand the core of the method.
}

\answer
\DISCUSS{LP}redo this section

\question{
(P. 18) Figure 5 shows a result of Fig 3. However the detail concerns example
in Fig. 1?  Examples and the implementation part are more clear. However, the
biological example is only used to evaluate performances of the implementation. A concrete (biological) modeling example used as a guideline for
illustrating the different notions throughout the article would certainly help
the reader.
}

\answer
\DISCUSS{LP} use Lambda phage as a running example (or dedicate a subsection to
apply the whole pipeline to it).

\end{document}

### 2014-09-03 main.tex
intro

@@
 \documentclass[preprint]{elsarticle}
+% vi:spell spelllang=en
 
@@
 
+\def\modMF#1{\textcolor{teal}{#1}}
+\def\modLP#1{\textcolor{magenta}{#1}}
+\def\modMM#1{\textcolor{blue}{#1}}
+\def\modOR#1{\textcolor{olive}{#1}}
+
 % Id est
@@
 
-\title{Constructing Biological Regulatory Networks from Process Hitting models}
+\title{\modLP{Identification of} Biological Regulatory Networks from Process Hitting models}
 %\\\rewriteil{Or possibly: “Complementarity between PH, IG and Thomas modeling”}
@@
 \ead{Maxime.Folschette@irccyn.ec-nantes.fr}
-\author[eth]{Lo\"ic Paulev\'e}
+\author[lri]{Lo\"ic Paulev\'e}
 \author[nii]{Katsumi Inoue}
@@
 2-1-2, Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan.}
-\address[eth]{BISON group, Automatic Control Laboratory, ETH Zürich\\
-Physikstrasse 3, 8092 Zurich, Switzerland.}
+\address[lri]{CNRS, Laboratoire de Recherche en Informatique UMR CNRS 8623\\
+Universit\'e Paris-Sud, 91405 Orsay Cedex, France}
 
 \begin{abstract}
-\parindent 0.5cm
-The Process Hitting (PH) is a recently introduced framework to model concurrent processes.
-\todo{The reviewer did not understand the following sentence:}
-Its major originality lies in a specific restriction on the causality of actions, which
-makes the formal analysis of very large systems tractable.
-PH is suitable to model Biological Regulatory Networks (BRNs) with complete or partial
-knowledge of cooperations between regulators by defining the most permissive dynamics
-with respect to these constraints.
+\modMF{Maxime}
+\modLP{Lo\"ic}
+\modMM{Morgan}
+\modOR{Olivier}
 
-\todo{Rather name Kaufman as the creator of BRNs.}
-On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
-formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
+%\parindent 0.5cm
+The Process Hitting (PH) \modLP{is a particular class of Indeterministic Asynchronous Automata
+Network (or safe Petri nets) on which have been recently designed several static analyses for
+dynamical properties that are tractable on very large systems.}
+
+\modLP{In this paper, we address the formal identification of qualitative
+models of Biological Regulatory Networks (BRNs) from PH models.}
+First, the inference of the Interaction Graph (IG) from a PH model
+\modLP{summarizes the signed influences between the components} that are effective for the dynamics.
+Second, we provide the inference of all Ren\'e Thomas models of BRNs 
+that are compatible with a given PH.
+As the PH allows the specification of indeterministic interactions between
+components, our inference emphasizes the ability of PH to deal with large BRNs with incomplete knowledge
+on interactions, where Thomas' approach fails because of the combinatorics of parameters.
+
+The inference of corresponding Thomas models is provided using Answer Set Programming,
+which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
+
+%\todo{Rather name Kaufman as the creator of BRNs.}
+%On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
+%formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
 %In addition to be able to use the existing results on this formalism,
 %obtaining the Thomas representation of a PH model allows to iteratively refine a model.
-Linking this representation to PH would not only allow to use the existing results on this formalism,
-but also to iteratively refine the model of a system.
+%Linking this representation to PH would not only allow to use the existing results on this formalism,
+%but also to iteratively refine the model of a system.
 
-This paper establishes formal links between these two formalisms, permitting to appreciate their
-complementary features.
-First, the inference of the Interaction Graph (IG) from a PH model gives a compact representation of
-the regulations that are effective for the dynamics.
-From this IG, well-known results can be applied to derive global dynamical properties for the
-modeled BRN.
-Second, the inference of all Thomas models that are compatible with a given PH is tackled.
-As the PH allows partial specification of cooperation between components, this inference
-emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
-cooperations, where Thomas' approach fails because of the combinatorics of parameters.
-
-The inference of corresponding Thomas' models is provided using Answer Set Programming,
-which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
+%This paper establishes formal links between these two formalisms, permitting to appreciate their
+%complementary features.
 \end{abstract}

### 2014-09-03 main.tex
abstract

@@
 
-The inference of corresponding Thomas models is provided using Answer Set Programming,
-which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
+The inference of corresponding Thomas models is \modLP{implemented} using Answer Set Programming,
+which allows an efficient enumeration of (possibly numerous) compatible parametrizations.
 

### 2014-09-12 parts-infer-ig.tex
update

@@
 
-\subsection{Focal Processes}\label{ssec:focal}
+\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}
 
-\todo{The two following subsections (down to “Interaction Inference”) have a lot of definitions but few examples and not enough informal explanations, especially on their use.
-Especially:
-\begin{itemize}
-  \item What kinds of concept do represent $\focals$?
-  \item What is the importance of a strict context?
-  \item Add an outline guiding the reader through these subsections.
-\end{itemize}}
-
-Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
-processes} w.r.t. a given context (a set of processes that are potentially present).
-When such a context applies, we expect to (always) reach one focal process in a bounded number of
-actions.
-
-For $S_a\subseteq L_a$ and a context (set of processes) $\ctx$, let us define as $\PHa(S_a,\ctx)$
-the set of actions on the sort $a$ having their hitter in $\ctx$ and target in $S_a$
-(\pref{eq:PHa-ctx});
-and the digraph $(V, E)$ where arcs are the bounces within the sort $a$ triggered by actions
-in $\PHa(S_a,\ctx)$ (\pref{eq:bounce-graph}).
-$\focals(a,S_a,\ctx)$ denotes the set of focal processes of sort $a$ in the scope of
-$\PHa(S_a,\ctx)$ (\pref{def:focals}).
-\begin{align}
-\PHa(S_a,\ctx) & \DEF \{ \PHfrappe{b_i}{a_j}{a_k}\in\PHa \mid b_i\in\ctx \wedge a_j\in S_a \}
-\label{eq:PHa-ctx}
-\\
-\begin{split}
-E  & \DEF \{(a_j,a_k)\in (S_a \times \PHl_a) \mid 
-			\exists\PHfrappe{b_i}{a_j}{a_k}\in \PHa(S_a,\ctx) \}
-\\
-V & \DEF S_a \cup \{ a_k\in L_a\mid \exists (a_j,a_k)\in E\}
-\end{split}
-\label{eq:bounce-graph}
-\end{align}
-
-\begin{definition}[$\focals(a,S_a,\ctx)$]\label{def:focals}
-The set of processes that are focal for processes in $S_a$ in the scope of $\PHa(S_a,\ctx)$
-are given by:
-%$\focals(a,S_a,\ctx)$ is the set of focal processes of sort $a$ in the context $\ctx$:
-\[
-\focals(a,S_a,\ctx) \DEF
-\begin{cases}
-\{ a_i \in V \mid \nexists (a_i,a_j)\in E\} & \text{if the digraph $(V,E)$ is acyclic},\\
-\emptyset & \text{otherwise.}\\
-\end{cases}
-\]
-\end{definition}
-
-We note $\PHl(\ctx)$ the set of states $s\in L$ such that $\forall a\in\PHsort(\ctx), \PHget{s}{a}\in\ctx$,
-where $\PHsort(\ctx)$ is the set of sorts with processes in $\ctx$.
-We say a sequence of actions $h^1,\dots,h^n$ is \emph{bounce-wise} if and only if
-$\forall m\in\segm{1}{n-1}, \PHbounce(h^m)=\PHtarget(h^{m+1})$.
-From \pref{def:focals}, it derives that:
-\begin{enumerate}
-\item if $\focals(a,S_a,\ctx)=\emptyset$, there exists a 
-state $s\in \PHl(\ctx\cup S_a)$ such that $\forall n\in\mathbb N$ there
-exists a bounce-wise sequence of actions $h^1,\dots,h^{n+1}$ in $\PHa(S_a,\ctx)$ 
-with $\PHtarget(h^1)\in s$.
-\item if $\focals(a,S_a,\ctx)\neq\emptyset$, for all
-state $s\in \PHl(\ctx\cup S_a)$,
-for any bounce-wise sequence of actions $h^1,\dots,h^n$ in $\PHa(S_a,\ctx)$ where $\PHtarget(h^1)\in
-s$,
-either
- $\PHbounce(h^n) \in \focals(a,S_a,\ctx)$,
-or
-$\exists h^{n+1}\in \PHa(a,\ctx)$ such that $\PHbounce(h^n) = \PHtarget(h^{n+1})$.
-Moreover $n\leq|\PHa(S_a,\ctx)|$ (i.e. no cycle of actions possible).
-\end{enumerate}
-
-It is worth noticing that those bounce-wise sequences of actions may not be successively playable in
-a state $s\in L(\ctx\cup S_a)$.
-Indeed, nothing impose that the hitters of actions are present in $s$.
-In the general case, the playability of those bounce-wise sequences, referred to as \emph{focals
-reachability} may be hard to prove.
-However, in the scope of this paper, the particular contexts used with $\focals$ ensure this property.
-Notably, the rest of this section uses only \emph{strict} contexts (\pref{def:strict-ctx}) which
-allow at most one hitter per sort in the bounce-wise sequences (and thus are present in $s$).
-
-\begin{definition}[Strict context for $S_a$]\label{def:strict-ctx}
-A context (set of processes) $\ctx$ is strict for $S_a\subseteq L_a$ if and only if
-$\{b_i,b_j\} \subset \ctx \wedge b\neq a \Rightarrow i=j$.
-\end{definition}
-
-In other words, assuming focals reachability, if $\focals(a,S_a,\ctx)$ is empty, there exists a
-sequence of actions that may be played an unbound number of times (cycle);
-if it is non-empty, it is ensured that any state in $\PHl(\ctx\cup S_a)$ converges, in a bounded
-number of steps, either to a process in $S_a$ that is not hit by processes in $\ctx$, or to a process in
-$L_a\setminus S_a$.
-
-\begin{example}
-In the PH of \pref{fig:runningPH-1}, we obtain:
-\begin{align*}
-\focals(a,L_a,\{b_0,c_0\}) &= \{ a_0 \}
-&
-\focals(a,L_a,\{b_1,c_1\}) &= \{ a_2 \}
-\\
-\focals(a,L_a,\{b_1,c_0\}) &= \emptyset
-&
-\focals(a,\{a_1\},\{b_1,c_0\}) &= \{ a_0, a_2 \}
-\end{align*}
-\end{example}
-
-\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}
+\todo{rework text with new paper structure (\pref{sec:tr2global})}
 
@@
 
-Any sort that does not act as a component should then be treated as a cooperative sort.
-As explained in \pref{ssec:PH}, the role of a cooperative sort $\upsilon$ is to compute the current
-state of set of cooperating processes.
-Hence, for each sub-state $\sigma$ formed by the sorts hitting $\upsilon$, $\upsilon$ should
-converge to a focal process.
-This is expressed by \pref{pro:wf-cooperative-sort}, where
-the set of sorts having an action on a given sort $a$ is given by 
-$\PHdirectpredec{a}$ (\pref{eq:ph_direct_predec})
-and $\PHproc(\sigma)$ is the set of processes that compose the sub-state $\sigma$.
-
-\begin{equation}
-\forall a \in \PHs, \PHdirectpredec{a} \DEF \{b \in \PHs \mid \exists \PHfrappe{b_i}{a_j}{a_k}\in\PHa \}
-\label{eq:ph_direct_predec}
-\end{equation}
-
-\begin{property}[Well-formed cooperative sort]\label{pro:wf-cooperative-sort}
-A sort $\upsilon\in\PHs$ is a well-formed cooperative sort if and only if
-each configuration $\sigma$ of its predecessors leads $\upsilon$ to a unique focal process,
-denoted by $\upsilon(\sigma)$:
-\[
-\forall \sigma \in {\textstyle\prod_{
-a\in\PHdirectpredec{\upsilon} \wedge a\neq \upsilon}}
-\PHl_{a},
-\focals(\upsilon,\PHl_\upsilon,\PHproc(\sigma)\cup \PHl_\upsilon) = \{ \upsilon(\sigma) \}\]
-\end{property}
-
-Such a property allows a large variety of definitions of a cooperative sort, but
-for the sake of simplicity, does not allow the existence of multiple focal processes.
-While this may be easily extended to (the condition becomes 
-$\focals(\upsilon,\PHl_\upsilon, \PHproc(\sigma)\cup \PHl_\upsilon)\neq\emptyset$), it makes some
-hereafter equations a bit more complex to read as they should handle a set of focal processes instead
-of a unique focal process.
+\todo{maybe we should also support a user-supplied $\Gamma$ and give the
+constraints (included in the prev. def?)}
 
@@
 \end{comment}
-where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$ \enspace.
+where $\f{sign}(x) = 1$ if $x>0$, $\f{sign}(x) = -1$ if $x < 0$, and
+$\f{sign}(x) = 0$ if $x=0$.
 \end{proposition}

### 2014-09-12 parts-ph.tex
update

--- 2013-07-09 parts/ph.tex
@@
 
-We recall here the definition and semantics of the Process Hitting (PH), and its usage to model
+\todo{(LP) rework text, make presentation more CS oriented}
+
+We give the the definition and semantics of the Process Hitting (PH), and its usage to model
 cooperation between concurrent components.
@@
 \caption{\label{fig:runningPH-1}
+\todo{make more visible}
 A Process Hitting (PH) example.
@@
 
-While the construction of cooperation in PH allows to encode any boolean functions
-between cooperating processes \cite{PMR10-TCSB}, it is worth noticing they introduce a temporal
-shift in their application.
-This allows the existence of interleaving of actions leading to a cooperative sort representing a
-past sub-state of the presence of the cooperative processes.
-The resulting behavior is then an over-approximation
-of the realization of an instantaneous cooperation.
-
 \begin{example}

### 2014-09-12 parts-tr2global.tex
update

--- 2013-07-24 parts/infer-ig.tex
@@
 % vim:spell spelllang=en:
-\section{Interaction Graph Inference from Process Hitting}\label{sec:infer-IG}
+\section{From Local Transitions to Global Functions}\label{sec:tr2global}
 
-The Interaction Graph (IG) is an abstract representation of the direct qualitative influences,
-positive and/or negative, between the components of the system.
-As discussed in \pref{sec:intro}, the IG allows to efficiently characterize global dynamical
-properties for the concrete system, such as the capability for multi-stationarity or oscillation.
-
-In a typical biological network modeling process, a prior IG is generally the starting point for
-the formal system specification.
-However, it is common that the prior IG actually refers to interactions that reveal to be 
-non-effective with respect to the dynamics.
-Hence, deriving the IG directly from the dynamical models lead to more concise IGs, enhancing the
-conclusiveness of static analyses based upon this abstract representation.
-
-In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
-modeling.
-This section first introduces the notion of focal processes within a PH (\pref{ssec:focal})
-which is used to characterize well-formed PH for IG inference (\pref{ssec:wf}).
-%and as well used by the parametrization inference presented in \pref{sec:infer-K}.
-The rules for inferring the interactions between components from a PH are
-described in \pref{ssec:infer-IG}.
-We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
-performed.
-
-\subsection{Focal Processes}\label{ssec:focal}
-
+\todo{(LP) motivation, outline}
 Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
@@
 actions.
+
+\subsection{Focal Processes}\label{ssec:focal}
+
+\todo{(LP) more explanations, check strict context importance, more examples?}
 
@@
 
-\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}
+\subsection{Restrictions for Determinism} % or Well-formed Cooperative Sorts
 
-The inference of an IG from a PH assumes that the PH defines two types of sorts:
-the sorts corresponding to BRN components, and the cooperative sorts.
-This leads to the characterization of a \emph{well-formed} PH for IG inference.
-
-The identification of sorts modeling components relies on the observation that their processes
-represent (ordered) qualitative levels.
-Hence an action on such a sort cannot make it bounce to a process at a distance more than one.
-The set of sorts satisfying such a condition is referred to as $\Gamma$
-(\pref{eq:PH-components}).
-Therefore, in the rest of this paper, $\Gamma$ denotes the set of components of the BRN to infer.
-
-\begin{equation}
-\Gamma \DEF \{a \in \PHs \mid \nexists \PHfrappe{b_i}{a_j}{a_k} \in \PHa, |j - k| > 1\} \\
-\label{eq:PH-components}
-\end{equation}
+\todo{(LP) redo text}
 
@@
 of a unique focal process.
-
-Finally, \pref{pro:wf-ph} sums up the conditions for a Process Hitting to be suitable for IG
-inference.
-In addition of having either component sorts or well-formed cooperative sorts, we also require that
-there is no cycle between cooperative sorts, and that
-sorts being never hit (\ie{} serving as an invariant environment) are components.
-
-\begin{property}[Well-formed Process Hitting for IG inference]\label{pro:wf-ph}
-A PH is well-formed for IG inference if and only if the following conditions are verified:
-\begin{itemize}
-\item 
-each sort $a\in\PHs$ either belongs to $\Gamma$, or is a well-formed cooperative sort;
-\item 
-there is no cycle between cooperative sorts
-(the digraph $(\Sigma,\{(a,b)\in(\Sigma\times\Sigma)\mid \exists \PHfrappe{a_i}{b_j}{b_k}\in\PHa
-\wedge a\neq b\wedge \{a,b\}\cap\Gamma=\emptyset \})$ is
-acyclic);
-\item 
-sorts having no action hitting them belong to $\Gamma$
-($\{ a \in \Sigma\mid \nexists \PHfrappe{b_i}{a_j}{a_k}\in\PHa\} \subset \Gamma$).
-\end{itemize}
-\end{property}
-
-\begin{example}
-In the PH of \pref{fig:runningPH-2}, $bc$ is a well-formed cooperative sort as defined in \pref{pro:wf-cooperative-sort}, because:
-\begin{align*}
-\focals(bc, \PHl_{bc}, \{b_0, c_0\} \cup \PHl_{bc}) = \{bc_{00}\} && \focals(bc, \PHl_{bc}, \{b_0, c_1\} \cup \PHl_{bc}) = \{bc_{01}\} \\
-\focals(bc, \PHl_{bc}, \{b_1, c_0\} \cup \PHl_{bc}) = \{bc_{10}\} && \focals(bc, \PHl_{bc}, \{b_1, c_1\} \cup \PHl_{bc}) = \{bc_{11}\}
-\end{align*}
-Hence, both \pref{fig:runningPH-1} and \pref{fig:runningPH-2} are well-formed PH for IG inference
-with $\Gamma = \{a,b,c\}$.
-\end{example}
-
-\subsection{Interaction Inference}\label{ssec:infer-IG}
-
-At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
-($\PHs \setminus \Gamma$) that will not appear in the IG. 
-We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
-by considering direct actions and possible intermediate cooperative sorts.
-The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
-(\pref{eq:regulators}).
-\begin{align}
-\begin{split}
-\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
-(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
-                                   & \quad \quad c^0 = b \wedge c^n = a \\
-                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
-                   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
-\end{split}
-\label{eq:ph_predec}
-\\
-\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
-\label{eq:regulators}
-\end{align}
-
-Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
-refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
-If $g=\{b\}$, we simple note $\ctx_b(\sigma)$.
-This set is composed of the active processes of sorts in $g$, and the focal process (assumed
-unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
-The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
-relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
-$\upsilon$ are defined in $\sigma$.
-When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
-the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
-where $\uplus$ stands for a union of states (which are sets of processes).
-Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
-always terminates.
-\begin{align}
-\forall g\subset \Gamma,
-  \ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
-\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
-\label{eq:ctx-sigma}
-\\
-\upsilon(\sigma) & \DEF
-\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
-  \upsilon'\in\PHdirectpredec{\upsilon} \wedge
-  \upsilon'\in\PHs\setminus\Gamma })
-\label{eq:cooperative-eval}
-\end{align}
-
-This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
-We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
-the level of $b$ makes possible the increase (decrease) of the level of $a$,
-which is directly inspired from the works of~\cite{Richard2010378}.
-To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
-Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
-$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
-\begin{equation}
-X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
-        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
-        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
-\label{eq:influence-groups}
-\end{equation}
-
-To infer influence of a component $b$ on a component $a$,
-we observe the direction of evolution of $a$ for different active processes of $b$.
-The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
-Formally, for a given group $g$ of regulators of a component $a$
-(that is, a minimal set of components regulating $a$ through common cooperative sorts),
-and a configuration $\ctx$ on $g$, we note
-$\irB_a(\sigma)$ the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
-%This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
-%and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
-If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
-%
-\begin{align}
-\begin{split}\label{eq:possible-bounces}
-  &\forall g \in X(a), \forall \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c,
-  \irB_a(\sigma) \DEF 
-  \begin{cases}
-    \irF_a(\sigma)
-      & \text{ if } \irF_a(\sigma) \neq \emptyset\\
-    \{ \PHget{\sigma}{a} \}
-      & \text{ if } \irF_a(\sigma) = \emptyset
-  \end{cases}\\
-  &\text{where: } \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
-  & \qquad\qquad\qquad\qquad\qquad\qquad \PHget{\sigma}{a} = a_j \wedge \PHget{(\ctx_{g}(\sigma))}{b} = b_i \}\\
-\end{split}
-\end{align}
-
-\pref{pps:inference-edges} details the inference of all existing influences between components occurring
-with a threshold $t$.
-The main idea of this inference is that
-the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
-there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
-(\pref{eq:edges-inference}).
-Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
-Furthermore, studying the influences of the groups of regulators of $a$
-allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
-Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
-We ignore the cases where a component has no visible influence on another.
-%
-\begin{proposition}[Influences inference]\label{pps:inference-edges}
-We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
-\begin{align}
-% Arcs a -> b, a ≠ b
-\begin{split}\label{eq:edges-inference}
-  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
-  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(k - a)
-\end{split}\\
-% Auto-arcs
-\begin{split}\label{eq:edges-inference-auto}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
-\end{split}\\
-\begin{split}\label{eq:edges-inference-noreg}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
-    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
-\end{split}
-\end{align}
-where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
-\end{proposition}
-
-We are now able to infer the edges of the final IG by considering positive and negative influences
-(\pref{pps:inference-IG}).
-We infer a positive (resp. negative) edge if there only exist corresponding influences with the same sign.
-If an influence is both positive and negative, we infer an unsigned edge.
-In the end, the threshold of each edge is the minimum threshold for which an influence has been found.
-%
-\begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
-We infer $\IG = (\Gamma,E)$ using \pref{pps:inference-edges} as follows:
-\begin{align*}
-E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
-  \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
-E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
-  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
-E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_- \\
-  & \qquad\qquad\qquad \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\}
-\end{align*}
-\end{proposition}
-
-\begin{example}
-The IG inference of the PH of \pref{fig:runningPH-1} gives the
-IG in \pref{fig:BRN-inf1}, containing the following edges:
-\begin{align*}
-  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{b}{+}{1}{b}, \GRNedgef{c}{+}{1}{c}\}\\
-  E_- &= \{\GRNedgef{a}{-}{2}{b}\} \qquad\qquad\qquad\qquad\qquad
-  E_\uns = \emptyset
-\end{align*}
-This IG is close to the one in \pref{fig:runningBRN} but not equivalent,
-as each component has an additional auto-action.
-The auto-actions on $b$ and $c$ are the consequence of a global stability
-in some configurations: $c$ never evolves, and neither does $b$ when $a_2$ is not active.
-The auto-action on $a$ is mainly caused by its multi-valued nature.
-
-The inference of the refined PH of \pref{fig:runningPH-2} gives the same IG.
-
-\begin{figure}[t]
-\centering
-\scalebox{1.2}{
-\begin{tikzpicture}[grn]
-  \path[use as bounding box] (-1.3,-0.75) rectangle (3.5,1.5);
-  \node[inner sep=0] (a) at (2,0) {a};
-  \node[inner sep=0] (b) at (0,0) {b};
-  \node[inner sep=0] (c) at (2,1.2) {c};
-  \path[->]
-    (b) edge[bend right] node[elabel, below=-2pt] {$+1$} (a)
-    (c) edge node[elabel, right=-2pt] {$+1$} (a)
-    (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b)
-    (b) edge[in=-15+180, out=15+180, loop] node[elabel, left=-2pt] {+1} (b)
-    (c) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (c)
-    (a) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (a);
-\end{tikzpicture}
-}
-\caption{\label{fig:BRN-inf1}
-  Result of the IG inference performed on the PH of \pref{fig:runningPH-2}.
-}
-\end{figure}
-
-\end{example}
-
-\begin{example}
-If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
-then two unsigned edges towards $b$ are inferred instead of the previous signed edges:
-\begin{align*}
-  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{c}{+}{1}{c}\}\\
-  E_- &= \emptyset \qquad\qquad\qquad\qquad
-  E_\uns = \{\GRNedgef{a}{\uns}{2}{b}, \GRNedgef{b}{\uns}{1}{b}\}
-\end{align*}
-This is due to the fact that the actions $\PHfrappe{a_2}{b_1}{b_0}$ and $\PHfrappe{a_2}{b_0}{b_1}$
-introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
-\end{example}

### 2014-09-15 parts-ph.tex
PH section: rel w/ Petri net + update cooperation paragraph

@@
 \label{ssec:PH}
-
-\todo{(LP) rework text, make presentation more CS oriented}
 
@@
 
+\medskip
+
+\modLP{PH can be seens as a particular sub-class of safe Petri nets \cite{BC92},
+i.e. Petri nets with at most one token per place, having
+groups of multually exclusive places (processes) acting as the sorts, and where each
+transition has at most one read-arc (connected to the \emph{hitter}), and
+exactly one incoming arc (from the \emph{target}) and one outgoing arc (to the
+\emph{bounce}).
+The relationship between PH and Petri nets is detailed in \cite{TR-PHvsPN-12}.}
+
 \begin{example}
-\todo{\pref{fig:runningPH-1} lacks readability (arrows hitting $a$ are too small).}
 \pref{fig:runningPH-1} represents a PH $(\PHs,\PHl,\PHa)$ with
@@
 
-\paragraph{Modeling cooperation}
-As described in \cite{PMR10-TCSB}, the cooperation between several processes to make another process bounce can be
-expressed in PH by building a \emph{cooperative sort}.
+\paragraph{Modeling conjunctions}
+\modLP{PH restricts the causality of transitions within sorts to the presence of
+at most one process in another sort (the hitter).
+One may want to model transitions that should be taken only when several
+processes are present, i.e., modeling conjunctions -- or cooperations -- between the presence of
+several processes.
+As described in \cite{PMR10-TCSB}, such cooperations can be encoded in PH with
+the use of some intermediate sorts, so-called \emph{cooperative sorts}, that
+merge information on the state of cooperating sorts.}
 \pref{fig:PH-cooperativity} shows an example of cooperation between processes $b_1$ and $c_1$ to
@@
 special treatment regarding the semantics of related actions.
+\modLP{%
+It is also worth noticing that the use of such intermediate sorts may trigger
+spurious transitions due to potential incoherences between the state of the
+cooperative sort and the actual state of cooperating sorts.
+However, such phenomena can be fixed by the use of priorities between actions \cite{FPMR13-CS2Bio}.}
 

### 2014-09-15 parts-infer-param.tex
add TODOs after the review
update
new 'Well-formed PH for K inference' property (to be completed)

@@
 \section{Parametrization inference}\label{sec:infer-K}
+
+\todo{In this while section: add informal explanations and maybe more examples (see previous section).}
 
@@
 
+\todo{makes it standalone: repeat the constraints for IG inference that are
+actually necessary + check that we can do if if the IG is provided (then what
+are the constraints on this IG)}
+
 \begin{property}[Well-formed PH for parameter inference]\label{pro:wf-ph-K}
@@
 verifies the following property:
+\todo{Define $c(\sigma)$ for $\sigma \in \textstyle\prod_{d \in \PHpredec{c}} \PHl_d$
+(\ie the unique focal point)}
 \begin{align*}
   \begin{split}
-  \forall a \in \Gamma&, \forall b\in \GRNreg{a},
-          \forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
-  & \quad \forall c \in \PHs, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a}\setminus \Gamma \wedge b\in\PHdirectpredec{c})), \\
-  & \qquad
-                          \PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
-                                  \PHfrappe{b_j}{c_k}{c_l}\in\PHa
+%   \forall a \in \Gamma&, \forall b\in \GRNreg{a},
+%           \forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
+%   & \quad \forall c \in \PHs, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a}\setminus \Gamma \wedge b\in\PHdirectpredec{c})), \\
+%   & \qquad
+%                           \PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
+%                                   \PHfrappe{b_j}{c_k}{c_l}\in\PHa
+      &
+    \forall a \in \Gamma,
+    \forall g \in X(a),
+    \forall \sigma, \sigma' \in \textstyle\prod_{d \in g} \PHl_d,
+    \forall b \in g,
+    b \in \GRNreg{c},
+    \forall c \in \PHdirectpredec{a} \setminus \Gamma,
+      \\&\qquad
+    b \in \reg{c},
+    \forall (i,j \in \levels{b}{a} \vee i,j \in \ulevels{b}{a}),
+%    \forall b_i \in \PHl_b,
+    b_i \in \sigma,
+%    b_i \in \levels{b}{a},
+%    \forall b_j \in \levels{b}{a},
+      \\&\qquad\qquad
+    \sigma' = \sigma\{b_j\},
+    %c_{\sigma} \in \PHl_c \wedge c_{\sigma'} \in \PHl_c,  %facultatif
+    \PHfrappe{c(\sigma)}{a_i}{a_j} \in \PHa \Leftrightarrow
+    \PHfrappe{c(\sigma')}{a_i}{a_j} \in \PHa
+  \end{split}\\
+  \begin{split}
+    & \forall a \in \Gamma, \forall b \in \GRNreg{a},
+      \forall (i,j \in \levels{b}{a} \vee i,j \in \ulevels{b}{a}), \\
+    & \qquad \forall c \in \PHs, b \neq a,
+      \PHfrappe{b_i}{a_k}{a_l} \in \PHa \Leftrightarrow
+      \PHfrappe{b_j}{a_k}{a_l} \in \PHa
   \end{split}

### 2014-09-17 parts-tr2global.tex
intro of sect. 3 + section 3.1

@@
 
-\todo{(LP) motivation, outline}
-Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
-processes} w.r.t. a given context (a set of processes that are potentially present).
-When such a context applies, we expect to (always) reach one focal process in a bounded number of
-actions.
+Process Hitting (PH) and Thomas modelling rely on two different paradigms for the specification of
+network dynamics:
+Thomas parametrization defines a global function
+which associates to each component and each resource configuration a level toward which the component
+evolve in the given configuration.
+Such a function is always deterministic (whereas the associated asynchronous dynamics can be
+indeterministic due to concurrent evolution of components).
+In contrast, 
+PH models specify a list of local transitions between the levels (processes) of the components (sorts)
+that are conditioned with the presence of other processes.
+In addition, PH allows the use of intermediate sorts, such as cooperative sorts, that do not refer
+to the components of the network.
+
+Hence, identifying Thomas models from PH requires a step that lump a set of transitions that can be
+applied in a given context into a single process that correspond to the level toward which the
+component evolve.
+This step is detailed in \pref{ssec:focal} in which is defined the $\focals$ function associating 
+the farthest reachable processes, called \emph{focal processes}, of a given sort for a given context.
+It is possible that several different focal processes are identified, indicating an indeterministic
+evolution of the sort;
+and it is also possible possible that no focal process exist, when in presence of terminal cycles of
+transitions.
+In those cases, there is no possible Thomas specification for the PH.
+In that way, necessary conditions for Thomas model identification from PH are described in
+\pref{ssec:dfocals}.
 
@@
 
-\todo{(LP) more explanations, check strict context importance, more examples?}
+Given a sort $a\in\Sigma$ and a context delimited by 
+a sub-set of processes of sort $a$ and a sub-state $\sigma$
+of the sorts having an action on $a$,
+$\focals(a,S_a,\sigma)$ (\pref{def:focals}) is the set of processes of sort $a$ towards which
+$a$ will eventually converge in the scope of the context.
+The context delimits the set of states $s\in L$ where $s[a]\in S_a$, and for all 
+sort $b\in\PHpred a, b\neq a$,
+$s[b]=\sigma[b]$.
+$\focals(a,S_a,\sigma)$ is obtained from the digraph where $a_i\in S_a$ is connected to $a_j\in L_a$ only if there
+exists an action $\PHfrappe{b_k}{a_i}{a_j}$ where either $b_k = a_i$ or $b_k$ is in $\sigma$.
+$\focals(a,S_a,\sigma)$ is empty if there is any cyclic terminal connected components;
+otherwise, it is exactly the leafs of the digraph.
 
-For $S_a\subseteq L_a$ and a context (set of processes) $\ctx$, let us define as $\PHa(S_a,\ctx)$
-the set of actions on the sort $a$ having their hitter in $\ctx$ and target in $S_a$
-(\pref{eq:PHa-ctx});
-and the digraph $(V, E)$ where arcs are the bounces within the sort $a$ triggered by actions
-in $\PHa(S_a,\ctx)$ (\pref{eq:bounce-graph}).
-$\focals(a,S_a,\ctx)$ denotes the set of focal processes of sort $a$ in the scope of
-$\PHa(S_a,\ctx)$ (\pref{def:focals}).
-\begin{align}
-\PHa(S_a,\ctx) & \DEF \{ \PHfrappe{b_i}{a_j}{a_k}\in\PHa \mid b_i\in\ctx \wedge a_j\in S_a \}
-\label{eq:PHa-ctx}
-\\
-\begin{split}
-E  & \DEF \{(a_j,a_k)\in (S_a \times \PHl_a) \mid 
-			\exists\PHfrappe{b_i}{a_j}{a_k}\in \PHa(S_a,\ctx) \}
-\\
-V & \DEF S_a \cup \{ a_k\in L_a\mid \exists (a_j,a_k)\in E\}
-\end{split}
-\label{eq:bounce-graph}
-\end{align}
-
-\begin{definition}[$\focals(a,S_a,\ctx)$]\label{def:focals}
-The set of processes that are focal for processes in $S_a$ in the scope of $\PHa(S_a,\ctx)$
-are given by:
-%$\focals(a,S_a,\ctx)$ is the set of focal processes of sort $a$ in the context $\ctx$:
+\begin{definition}[$\focals(a,S_a,\sigma)$]\label{def:focals}
+Given a sort $a\in\Sigma$, a sub-set of its processes $S_a\subset L_a$ 
+and a sub-state
+$\sigma \in {\prod_{b\in\PHdirectpredec{a}, b\neq a} \PHl_{b}}$,
 \[
-\focals(a,S_a,\ctx) \DEF
+\focals(a,S_a,\sigma) \DEF
 \begin{cases}
-\{ a_i \in V \mid \nexists (a_i,a_j)\in E\} & \text{if the digraph $(V,E)$ is acyclic},\\
-\emptyset & \text{otherwise.}\\
+\emptyset & \text{if }\f{tscc}(V,E)\neq\emptyset\\
+\{ a_i \in V \mid \nexists (a_i,a_j)\in E\} & \text{otherwise,}
 \end{cases}
 \]
+where
+\begin{align}
+E  & \DEF \{(a_i,a_j)\in (S_a \times \PHl_a) \mid 
+			\exists\PHfrappe{b_k}{a_i}{a_j}\in \PHa:
+				(b_k = a_i \vee \sigma[b]=k) \}
+\\
+V & \DEF S_a \cup \{ a_j\in L_a\mid \exists (a_i,a_j)\in E\}
+\end{align}
+and $\f{tscc}(V,E)$ are the non-elementary terminal strongly connected components of the digraph
+$(V,E)$:
+\[
+\f{tscc}(V,E) = \{
+W\in\f{scc}(V,E) \mid \card{W}\geq 2\wedge \forall a_i\in W, (a_i,a_j)\in E\Rightarrow a_j\in W\}
+\]
+with $\f{scc}(V,E)$ the strongly connected components of the digraph $(V,E)$.
 \end{definition}
 
-We note $\PHl(\ctx)$ the set of states $s\in L$ such that $\forall a\in\PHsort(\ctx), \PHget{s}{a}\in\ctx$,
-where $\PHsort(\ctx)$ is the set of sorts with processes in $\ctx$.
-We say a sequence of actions $h^1,\dots,h^n$ is \emph{bounce-wise} if and only if
+From \pref{def:focals} can be derived 
+\pref{pps:no-focals} -- if $\focals(a,S_a,\sigma)$ is empty, there exists a limit cycle in the evolution of
+$a$ in the given context;
+and \pref{pps:has-focals} -- if $\focals(a,S_a,\sigma)$ is non empty, all the evolutions of $a$ in the scope
+of the given context eventually terminate, and the resulting processes are in $\focals(a,S_a,\sigma)$.
+In other words, if $\focals(a,S_a,\sigma)$ is empty, there exists a
+sequence of actions that may be played an unbound number of times (cycle) in the given context;
+if it is non-empty, it is ensured that any state in the given context converges 
+in a bounded number of steps to a process in $\focals(a,S_a,\sigma)$.
+
+\begin{proposition}
+\label{pps:no-focals}
+$\focals(a,S_a,\sigma)=\emptyset$ if and only if
+there exists a 
+state $s\in L$
+where
+$s[a]\in S_a$ and
+$\forall b\in \PHpred a, b\neq a, s[b] = \sigma[b]$,
+such that
+$\forall n\in \mathbb N$
+there 
+exists a sequence of actions $h^1,\dots,h^{n+1}$ in $\PHa$
+sequentially playable in $s$ with
+$\PHtarget(h^1)\in S_a$ and
 $\forall m\in\segm{1}{n-1}, \PHbounce(h^m)=\PHtarget(h^{m+1})$.
-From \pref{def:focals}, it derives that:
-\begin{enumerate}
-\item if $\focals(a,S_a,\ctx)=\emptyset$, there exists a 
-state $s\in \PHl(\ctx\cup S_a)$ such that $\forall n\in\mathbb N$ there
-exists a bounce-wise sequence of actions $h^1,\dots,h^{n+1}$ in $\PHa(S_a,\ctx)$ 
-with $\PHtarget(h^1)\in s$.
-\item if $\focals(a,S_a,\ctx)\neq\emptyset$, for all
-state $s\in \PHl(\ctx\cup S_a)$,
-for any bounce-wise sequence of actions $h^1,\dots,h^n$ in $\PHa(S_a,\ctx)$ where $\PHtarget(h^1)\in
-s$,
+\end{proposition}
+
+\begin{proposition}
+\label{pps:has-focals}
+If $\focals(a,S_a,\sigma)\neq\emptyset$, for all
+state $s\in L$
+where
+$s[a]\in S_a$ and
+$\forall b\in \PHpred a, b\neq a, s[b] = \sigma[b]$,
 either
- $\PHbounce(h^n) \in \focals(a,S_a,\ctx)$,
+$\nexists h\in\PHa$ with $\PHtarget(h)\in S_a$  playable in $s$ and
+$s[a]\in\focals(a,S_a,\sigma)$;
 or
-$\exists h^{n+1}\in \PHa(a,\ctx)$ such that $\PHbounce(h^n) = \PHtarget(h^{n+1})$.
-Moreover $n\leq|\PHa(S_a,\ctx)|$ (i.e. no cycle of actions possible).
-\end{enumerate}
-
-It is worth noticing that those bounce-wise sequences of actions may not be successively playable in
-a state $s\in L(\ctx\cup S_a)$.
-Indeed, nothing impose that the hitters of actions are present in $s$.
-In the general case, the playability of those bounce-wise sequences, referred to as \emph{focals
-reachability} may be hard to prove.
-However, in the scope of this paper, the particular contexts used with $\focals$ ensure this property.
-Notably, the rest of this section uses only \emph{strict} contexts (\pref{def:strict-ctx}) which
-allow at most one hitter per sort in the bounce-wise sequences (and thus are present in $s$).
-
-\begin{definition}[Strict context for $S_a$]\label{def:strict-ctx}
-A context (set of processes) $\ctx$ is strict for $S_a\subseteq L_a$ if and only if
-$\{b_i,b_j\} \subset \ctx \wedge b\neq a \Rightarrow i=j$.
-\end{definition}
-
-In other words, assuming focals reachability, if $\focals(a,S_a,\ctx)$ is empty, there exists a
-sequence of actions that may be played an unbound number of times (cycle);
-if it is non-empty, it is ensured that any state in $\PHl(\ctx\cup S_a)$ converges, in a bounded
-number of steps, either to a process in $S_a$ that is not hit by processes in $\ctx$, or to a process in
-$L_a\setminus S_a$.
+there exists a sequence of actions $h^1,\dots,h^n$ in $\PHa$ sequentially playable in $s$ with
+$\PHtarget(h^1)\in S_a$ and
+$\forall m\in\segm{1}{n-1}, \PHbounce(h^m)=\PHtarget(h^{m+1})$
+where
+ $\PHbounce(h^n) \in \focals(a,S_a,\sigma)$
+ and
+ where 
+$\nexists h^{n+1}\in \PHa$ with $\PHtarget(h^{n+1})=\PHbounce(h^n)$ 
+that is playable in $s\play h^1\play \cdots \play h^n$.
+\end{proposition}
 
@@
 \begin{align*}
-\focals(a,L_a,\{b_0,c_0\}) &= \{ a_0 \}
+\focals(a,L_a,\state{b_0,c_0}) &= \{ a_0 \}
 &
-\focals(a,L_a,\{b_1,c_1\}) &= \{ a_2 \}
+\focals(a,L_a,\state{b_1,c_1}) &= \{ a_2 \}
 \\
-\focals(a,L_a,\{b_1,c_0\}) &= \emptyset
+\focals(a,L_a,\state{b_1,c_0}) &= \emptyset
 &
-\focals(a,\{a_1\},\{b_1,c_0\}) &= \{ a_0, a_2 \}
+\focals(a,\{a_1\},\state{b_1,c_0}) &= \{ a_0, a_2 \}
 \end{align*}
@@
 \subsection{Restrictions for Determinism} % or Well-formed Cooperative Sorts
+\label{ssec:dfocals}
 

### 2014-09-17 parts-tr2global.tex
sect 3.2

@@
 
-\todo{(LP) redo text}
+As discussed at the beginning of this section, and as formally described in the above sub-section, 
+in the scope of a given context,
+a PH sort may converge to one among several different processes (indeterminism) or may never converge
+(infinite loop).
 
-Any sort that does not act as a component should then be treated as a cooperative sort.
-As explained in \pref{ssec:PH}, the role of a cooperative sort $\upsilon$ is to compute the current
-state of set of cooperating processes.
-Hence, for each sub-state $\sigma$ formed by the sorts hitting $\upsilon$, $\upsilon$ should
-converge to a focal process.
-This is expressed by \pref{pro:wf-cooperative-sort}, where
-the set of sorts having an action on a given sort $a$ is given by 
-$\PHdirectpredec{a}$ (\pref{eq:ph_direct_predec})
-and $\PHproc(\sigma)$ is the set of processes that compose the sub-state $\sigma$.
+\todo{this statement is too strong if we consider intervals, but correct if we are not}
+It results that deriving Thomas specification for dynamics of a PH model
+requires that every sort converges toward a unique focal process, i.e., that every sort has a
+deterministic behaviour w.r.t. any context, as stated by \pref{pro:deterministic-sort}.
 
-\begin{equation}
-\forall a \in \PHs, \PHdirectpredec{a} \DEF \{b \in \PHs \mid \exists \PHfrappe{b_i}{a_j}{a_k}\in\PHa \}
-\label{eq:ph_direct_predec}
-\end{equation}
-
-\begin{property}[Well-formed cooperative sort]\label{pro:wf-cooperative-sort}
-A sort $\upsilon\in\PHs$ is a well-formed cooperative sort if and only if
-each configuration $\sigma$ of its predecessors leads $\upsilon$ to a unique focal process,
-denoted by $\upsilon(\sigma)$:
+\begin{property}[Deterministic sort]\label{pro:deterministic-sort}
+A sort $a\in\PHs$ is a deterministic sort if and only if
+each configuration $\sigma$ of its predecessors leads $a$ to a unique focal process,
+denoted $a(\sigma)$:
 \[
 \forall \sigma \in {\textstyle\prod_{
-a\in\PHdirectpredec{\upsilon} \wedge a\neq \upsilon}}
-\PHl_{a},
-\focals(\upsilon,\PHl_\upsilon,\PHproc(\sigma)\cup \PHl_\upsilon) = \{ \upsilon(\sigma) \}\]
+b\in\PHdirectpredec{\upsilon} \wedge b\neq a}}
+\PHl_{b},
+\focals(a,\PHl_a,\sigma) = \{ a(\sigma) \}\]
 \end{property}
 
-Such a property allows a large variety of definitions of a cooperative sort, but
-for the sake of simplicity, does not allow the existence of multiple focal processes.
-While this may be easily extended to (the condition becomes 
-$\focals(\upsilon,\PHl_\upsilon, \PHproc(\sigma)\cup \PHl_\upsilon)\neq\emptyset$), it makes some
-hereafter equations a bit more complex to read as they should handle a set of focal processes instead
-of a unique focal process.
+\todo{examples}

### 2014-09-17 parts-tr2global.tex
typo

@@
 which associates to each component and each resource configuration a level toward which the component
-evolve in the given configuration.
+eventually evolves in the given configuration.
 Such a function is always deterministic (whereas the associated asynchronous dynamics can be

### 2014-09-17 parts-tr2global.tex
update

@@
 
-\todo{this statement is too strong if we consider intervals, but correct if we are not}
-It results that deriving Thomas specification for dynamics of a PH model
-requires that every sort converges toward a unique focal process, i.e., that every sort has a
-deterministic behaviour w.r.t. any context, as stated by \pref{pro:deterministic-sort}.
+From \pref{def:focals}, we call a \emph{deterministic sort} \pref{pro:deterministic-sort} a sort $a$
+that has a single focal process for any possible configuration of its predecessors in the scope of
+$L_a$.
 
@@
 
-\todo{examples}
+\todo{examples?}
+
+For the inference of the interaction graph and Thomas parameters from a PH model, the following
+sections assume that \emph{all the cooperative sorts are deterministic}.
+\todo{add more explanation depending on the text of the next sections}

### 2014-09-17 parts-tr2global.tex
minor

@@
 
-From \pref{def:focals}, we call a \emph{deterministic sort} \pref{pro:deterministic-sort} a sort $a$
-that has a single focal process for any possible configuration of its predecessors in the scope of
+We call a \emph{deterministic sort} \pref{pro:deterministic-sort} a sort $a$
+that has a single focal process for each possible configuration of its predecessors in the scope of
 $L_a$.

### 2014-09-18 parts-tr2global.tex
minor changes in sect 3

@@
 In those cases, there is no possible Thomas specification for the PH.
-In that way, necessary conditions for Thomas model identification from PH are described in
-\pref{ssec:dfocals}.
+The \pref{ssec:dfocals} discusses on constraints for obtaining deterministic
+(local) behaviours in PH.
 
@@
 From \pref{def:focals} can be derived 
-\pref{pps:no-focals} -- if $\focals(a,S_a,\sigma)$ is empty, there exists a limit cycle in the evolution of
+\pref{pro:no-focals} -- if $\focals(a,S_a,\sigma)$ is empty, there exists a limit cycle in the evolution of
 $a$ in the given context;
-and \pref{pps:has-focals} -- if $\focals(a,S_a,\sigma)$ is non empty, all the evolutions of $a$ in the scope
+and \pref{pro:has-focals} -- if $\focals(a,S_a,\sigma)$ is non empty, all the evolutions of $a$ in the scope
 of the given context eventually terminate, and the resulting processes are in $\focals(a,S_a,\sigma)$.
@@
 
-\begin{proposition}
-\label{pps:no-focals}
+\begin{property}
+\label{pro:no-focals}
 $\focals(a,S_a,\sigma)=\emptyset$ if and only if
@@
 $\forall m\in\segm{1}{n-1}, \PHbounce(h^m)=\PHtarget(h^{m+1})$.
-\end{proposition}
+\end{property}
 
-\begin{proposition}
-\label{pps:has-focals}
+\begin{property}
+\label{pro:has-focals}
 If $\focals(a,S_a,\sigma)\neq\emptyset$, for all
@@
 that is playable in $s\play h^1\play \cdots \play h^n$.
-\end{proposition}
+\end{property}
 
@@
 
-We call a \emph{deterministic sort} \pref{pro:deterministic-sort} a sort $a$
+We call a \emph{deterministic sort} \pref{def:deterministic-sort} a sort $a$
 that has a single focal process for each possible configuration of its predecessors in the scope of
@@
 
-\begin{property}[Deterministic sort]\label{pro:deterministic-sort}
+\begin{definition}[Deterministic sort]\label{def:deterministic-sort}
 A sort $a\in\PHs$ is a deterministic sort if and only if
@@
 \focals(a,\PHl_a,\sigma) = \{ a(\sigma) \}\]
-\end{property}
+\end{definition}
 

### 2014-09-23 parts-phage-lambda.tex
start new running example with phage lambda

\begin{example}
\begin{align*}
  \PHs &= \{ cI, cro, cII, N \} \\
  \PHl_{cI} &= \{ cI_0, cI_1, cI_2 \} &
  \PHl_{cro} &= \{ cro_0, cro_1, cro_2, cro_3 \} \\
  \PHl_{cII} &= \{ cII_0, cII_1 \} &
  \PHl_{N} &= \{ N_0, N_1 \} \\
\end{align*}
  % TODO : figure PH du modèle
\end{example}

\begin{example}
  The IG inference of \pref{pps:inference-IG} applied to the PH of \todo{ref}
  produces the IG given in \pref{fig:phage-lambda-ig}.
  This IG contains all edges included in the original model except the
  self-influence on $cI$ : $\GRNedgef{cI}{+}{2}{cI} \notin E$;
  this is due to the fact that an equivalent dynamics exists without this edge.
  \begin{figure}[t]
  \centering
  \scalebox{1.2}{
  \begin{tikzpicture}[grn]
    %\path[use as bounding box] (-1.3,-0.75) rectangle (3.5,1.5);
    \node[inner sep=0] (cI) at (0,2.5) {cI};
    \node[inner sep=0] (cro) at (2.5,2.5) {cro};
    \node[inner sep=0] (cII) at (0,0) {cII};
    \node[inner sep=0] (N) at (2.5,0) {N};
    \path[->]
      (cI) edge[bend right] node[elabel, above=-4pt] {$-2$} (cro)
           edge[bend right] node[elabel, left=-2pt] {$-2$} (cII)
           edge[pos=.85] node[elabel, above=-2pt] {$-1$} (N)
      (cro) edge[bend right] node[elabel, above=-4pt] {$-1$} (cI)
            edge[loop right] node[elabel, below=-2pt] {$-3$} (cro)
            edge[pos=.65] node[elabel, below=-2pt] {$-3$} (cII)
            edge node[elabel, right=-4pt] {$-2$} (N)
      (cII) edge[bend right] node[elabel, left=-3pt] {$+1$} (cI)
      (N) edge node[elabel, above=-4pt] {$+1$} (cII);
  \end{tikzpicture}
  }
  \caption{\label{fig:phage-lambda-ig}
    Result of the IG inference performed on the PH of \todo{ref}.
  }
  \end{figure}
\end{example}

\begin{example}
  The parameters inference of \pref{pps:param_K} applied to the PH of \todo{ref}
  is conclusive for all parameters.
  The complete parametrization produced is available in \pref{tb:phage-lambda-k}.
  \begin{table}[t]
  \begin{align*}
    K_{cI, \emptyset} &= \{2\} &
    K_{cI, \{cro\}} &= \{0\} \\
    K_{cI, \{cII\}} &= \{2\} &
    K_{cI, \{cII;cro\}} &= \{2\} \\
    K_{N, \emptyset} &= \{1\} &
    K_{N, \{cro\}} &= \{0\} \\
    K_{N, \{cI\}} &= \{0\} &
    K_{N, \{cI;cro\}} &= \{0\} \\
    K_{cII, \emptyset} &= \{0\} &
    K_{cII, \{cro\}} &= \{0\} \\
    K_{cII, \{cI\}} &= \{0\} &
    K_{cII, \{cI;cro\}} &= \{0\} \\
    K_{cII, \{N\}} &= \{1\} &
    K_{cII, \{N;cro\}} &= \{0\} \\
    K_{cII, \{N;cI\}} &= \{0\} &
    K_{cII, \{N;cI;cro\}} &= \{0\} \\
    K_{cro, \emptyset} &= \{3\} &
    K_{cro, \{cro\}} &= \{2\} \\
    K_{cro, \{cI\}} &= \{0\} &
    K_{cro, \{cI;cro\}} &= \{2\}
  \end{align*}
  \caption{\label{tb:phage-lambda-k}
    Result of the parameters inference performed on the PH of \todo{ref}.
  }
  \end{table}
\end{example}

### 2014-09-23 parts-phage-lambda.tex
parts of the PH of phage lambda

@@
 \end{align*}
-  % TODO : figure PH du modèle
+
+\begin{figure}[p]
+\centering
+\scalebox{1.3}{
+\begin{tikzpicture}
+  %\path[use as bounding box] (-4,-1.9) rectangle (4.5,3.9);
+
+  % Sortes
+  \TSort{(0,3.5)}{cI}{3}{l}
+  \TSort{(0,-1.5)}{cro}{4}{l}
+  \TSort{(7,1.5)}{N}{2}{r}
+
+  % Sorte coopérative
+  \TSetTick{cIcro}{0}{cI~$<$~1\\cro~$<$~2}
+  \TSetTick{cIcro}{1}{cI~$<$~1\\cro~$\geq$~2}
+  \TSetTick{cIcro}{2}{cI~$\geq$~1\\cro~$<$~2}
+  \TSetTick{cIcro}{3}{cI~$\geq$~1\\cro~$\geq$~2}
+  \TSort{(4,0.5)}{cIcro}{4}{r}
+
+  % Actions de MàJ
+  \THit{cI_1}{}{cIcro_0}{.west}{cIcro_2}
+  \THit{cI_1}{}{cIcro_1}{.160}{cIcro_3}
+  \THit{cI_2}{}{cIcro_0}{.west}{cIcro_2}
+  \THit{cI_2}{}{cIcro_1}{.160}{cIcro_3}
+  \path[bounce,bend left]
+    \TBounce{cIcro_0}{}{cIcro_2}{.240}
+    \TBounce{cIcro_1}{}{cIcro_3}{.south west} ;
+  \THit{cI_0}{}{cIcro_2}{.160}{cIcro_0}
+  \THit{cI_0}{}{cIcro_3}{.150}{cIcro_1}
+  \path[bounce,bend right]
+    \TBounce{cIcro_2}{}{cIcro_0}{.north west}
+    \TBounce{cIcro_3}{}{cIcro_1}{.120} ;
+  \THit{cro_2}{}{cIcro_0}{.210}{cIcro_1}
+  \THit{cro_2}{}{cIcro_2}{.200}{cIcro_3}
+  \THit{cro_3}{}{cIcro_0}{.210}{cIcro_1}
+  \THit{cro_3}{}{cIcro_2}{.200}{cIcro_3}
+  \path[bounce,bend left]
+    \TBounce{cIcro_0}{}{cIcro_1}{.240}
+    \TBounce{cIcro_2}{}{cIcro_3}{.south} ;
+  \THit{cro_0}{}{cIcro_3}{.190}{cIcro_2}
+  \THit{cro_0}{}{cIcro_1}{.200}{cIcro_0}
+  \THit{cro_1}{}{cIcro_3}{.190}{cIcro_2}
+  \THit{cro_1}{}{cIcro_1}{.200}{cIcro_0}
+  \path[bounce,bend right]
+    \TBounce{cIcro_1}{}{cIcro_0}{.north}
+    \TBounce{cIcro_3}{}{cIcro_2}{.120} ;
+
+  % Actions sortantes
+  \THit{cIcro_3}{}{N_0}{.west}{N_1}
+  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.south} ;
+  \THit{cIcro_2}{}{N_1}{.west}{N_0}
+  \path[bounce,bend right] \TBounce{N_1}{}{N_0}{.north} ;
+\end{tikzpicture}
+}
+\caption{\label{fig:phage-lambda-cIcro-N}%
+  A portion of the PH model used in this section modeling the phage lambda immunity response.
+  This part of the whole model only depicts the interactions towards $N$ from components
+  $cI$ and $cro$.
+  The processes of the cooperative sort $cIcro$ are labeled by the configuration they represent.
+}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\scalebox{1.3}{
+\begin{tikzpicture}
+  %\path[use as bounding box] (-4,-1.9) rectangle (4.5,3.9);
+
+  % Sortes
+  \TSort{(0,1)}{cI}{3}{l}
+  \TSort{(4,0)}{cro}{4}{r}
+
+  % Actions sortantes
+  \THit{cro_3}{out=north east,in=south east,selfhit,min distance=50pt}{cro_3}{.south east}{cro_2}
+    \path[bounce, bend left] \TBounce{cro_3}{}{cro_2}{.north east} ;
+
+  \THit{cI_2}{}{cro_3}{.210}{cro_2}
+    \path[bounce,bend right=40] \TBounce{cro_3}{}{cro_2}{.west} ;
+  \THit{cI_2}{}{cro_2}{.210}{cro_1}
+    \path[bounce,bend right=40] \TBounce{cro_2}{}{cro_1}{.west} ;
+  \THit{cI_2}{}{cro_1}{.210}{cro_0}
+    \path[bounce,bend right=40] \TBounce{cro_1}{}{cro_0}{.west} ;
+
+  \THit{cI_1}{}{cro_0}{.110}{cro_1}
+    \path[bounce,bend left] \TBounce{cro_0}{}{cro_1}{.south} ;
+  \THit{cI_1}{}{cro_1}{.110}{cro_2}
+    \path[bounce,bend left] \TBounce{cro_1}{}{cro_2}{.south} ;
+  \THit{cI_1}{}{cro_2}{.110}{cro_3}
+    \path[bounce,bend left] \TBounce{cro_2}{}{cro_3}{.south} ;
+
+  \THit{cI_0}{}{cro_0}{.140}{cro_1}
+    \path[bounce,bend left] \TBounce{cro_0}{}{cro_1}{.240} ;
+  \THit{cI_0}{}{cro_1}{.140}{cro_2}
+    \path[bounce,bend left] \TBounce{cro_1}{}{cro_2}{.240} ;
+  \THit{cI_0}{}{cro_2}{.140}{cro_3}
+    \path[bounce,bend left] \TBounce{cro_2}{}{cro_3}{.240} ;
+\end{tikzpicture}
+}
+\caption{\label{fig:phage-lambda-cIcro-cro}%
+  A portion of the PH model used in this section modeling the phage lambda immunity response.
+  This part of the whole model only depicts the interactions towards $cro$ from components
+  $cI$ and $cro$.
+}
+\end{figure}
 \end{example}
@@
   }
-  \caption{\label{fig:phage-lambda-ig}
+  \caption{\label{fig:phage-lambda-ig}%
     Result of the IG inference performed on the PH of \todo{ref}.
@@
   \end{align*}
-  \caption{\label{tb:phage-lambda-k}
+  \caption{\label{tb:phage-lambda-k}%
     Result of the parameters inference performed on the PH of \todo{ref}.

### 2014-09-23 parts-infer-ig.tex
new defs for IG inference

@@
 
+\def\fp{\Phi}
+
+\modLP{
+\begin{equation}
+\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid \sigma[a] = i \wedge
+					\focals(a, \{a_i\}, \ctx_{\PHpredecgene{a}}(\sigma)) =
+					\{a_i\} \}
+\label{eq:fp}
+\end{equation}
+}
+
 To infer influence of a component $b$ on a component $a$,
@@
 %
+
+\modLP{
+$\forall g\in X(a), \forall \sigma\in L(g\cup\{a\}), i=\sigma[a]$:
 \begin{align}
-\begin{split}\label{eq:possible-bounces}
-  &\forall g \in X(a), \forall \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c,
-  \irB_a(\sigma) \DEF 
-  \begin{cases}
-    \irF_a(\sigma)
-      & \text{ if } \irF_a(\sigma) \neq \emptyset\\
-    \{ \PHget{\sigma}{a} \}
-      & \text{ if } \irF_a(\sigma) = \emptyset
-  \end{cases}\\
-  &\text{where: } \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
-  & \qquad\qquad\qquad\qquad\qquad\qquad \PHget{\sigma}{a} = a_j \wedge \PHget{(\ctx_{g}(\sigma))}{b} = b_i \}\\
-\end{split}
-\end{align}
+a_i \in \irB_a^g(\sigma) &\EQDEF \exists\sigma'\in\fp(a_i): \sigma\subset\sigma'
+\label{eq:B-no-bounce}
+\\
+\forall a_j\in L_a, j\neq i,
+a_j \in \irB_a^g(\sigma) &\EQDEF \focals(a, \{a_j\}, \ctx_{g}(\sigma)) = \{a_j\}
+\label{eq:B-bounce}
+\end{align}}
 
@@
 We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
-% Arcs a -> b, a ≠ b
-\begin{align}
-\begin{split}\label{eq:edges-inference}
-  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
-  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(k - j)
-\end{split}
-\end{align}
-% Auto-arcs depuis les groupes de régulateurs
-\begin{align}
-\begin{split}\label{eq:edges-inference-auto}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a),
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ a_t, a_{t+1} \} \subset \PHl_a \wedge a_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{a_{t+1}\}), \\
-    &\qquad s = \f{sign}(k - j)
-\end{split}
-\end{align}
-% Auto-arcs des composants sans prédécesseurs
-\begin{align}
-\begin{split}\label{eq:edges-inference-noreg}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
-    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
-    &\qquad s = \f{sign}(k - j)
-\end{split}
-\end{align}
-\begin{comment}
-% Arcs a -> b, a ≠ b
-\begin{align}
-\begin{split}\label{eq:edges-inference}
-  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
-  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(k - a)
-\end{split}
-\end{align}
-% Auto-arcs depuis les groupes de régulateurs
-\begin{align}
-\begin{split}\label{eq:edges-inference-auto}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
-  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
-\end{split}
-\end{align}
-% Auto-arcs des composants sans prédécesseurs
-\begin{align}
-\begin{split}\label{eq:edges-inference-noreg}
-  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
-  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
-    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
-    &\qquad s = \f{sign}(a_k - a_j)
-\end{split}
-\end{align}
-\end{comment}
-where $\f{sign}(x) = 1$ if $x>0$, $\f{sign}(x) = -1$ if $x < 0$, and
-$\f{sign}(x) = 0$ if $x=0$.
+\modLP{
+\begin{align*}
+  \forall b\in\PHpredecgene{a}\cup\{a\}, \forall s \in \{ +, - \}, \\
+  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), 
+  	b\neq a \Rightarrow b \in g, \\
+	\qquad & \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
+    \qquad & \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
+    \qquad & \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+	\qquad & k\geq t+1 \wedge j \leq t \wedge s=+ \\
+	\qquad & \vee k<t+1 \wedge j>t \wedge s=-
+\end{align*}
+}
 \end{proposition}

### 2014-09-24 parts-infer-ig.tex
IG inference++
oops
small fixe
improve inference eq
simplify eq 13
tr2global++

@@
 
-In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
-modeling.
-This section first introduces the notion of focal processes within a PH (\pref{ssec:focal})
-which is used to characterize well-formed PH for IG inference (\pref{ssec:wf}).
-%and as well used by the parametrization inference presented in \pref{sec:infer-K}.
-The rules for inferring the interactions between components from a PH are
-described in \pref{ssec:infer-IG}.
-We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
+We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
+the level of $b$ makes possible the increase (decrease) of the level of $a$,
+which is directly inspired from the works of~\cite{Richard2010378}.
+
+We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ and a parition of sorts
+$\Gamma \cup \Delta=\PHs$ satisfying \pref{pro:wf-ph} on which the IG inference is to be
 performed.
+\todo{outline}
 
-\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}
+\subsection{Inference of Influences}\label{ssec:infer-IG}
 
@@
 
-The inference of an IG from a PH assumes that the PH defines two types of sorts:
-the sorts corresponding to BRN components, and the cooperative sorts.
-This leads to the characterization of a \emph{well-formed} PH for IG inference.
-
-The identification of sorts modeling components relies on the observation that their processes
-represent (ordered) qualitative levels.
-Hence an action on such a sort cannot make it bounce to a process at a distance more than one.
-The set of sorts satisfying such a condition is referred to as $\Gamma$
-(\pref{eq:PH-components}).
-Therefore, in the rest of this paper, $\Gamma$ denotes the set of components of the BRN to infer.
-
-\begin{equation}
-\Gamma \DEF \{a \in \PHs \mid \nexists \PHfrappe{b_i}{a_j}{a_k} \in \PHa, |j - k| > 1\} \\
-\label{eq:PH-components}
-\end{equation}
-
-\todo{maybe we should also support a user-supplied $\Gamma$ and give the
-constraints (included in the prev. def?)}
-
-Finally, \pref{pro:wf-ph} sums up the conditions for a Process Hitting to be suitable for IG
-inference.
-In addition of having either component sorts or well-formed cooperative sorts, we also require that
-there is no cycle between cooperative sorts, and that
-sorts being never hit (\ie serving as an invariant environment) are components.
-
-\begin{property}[Well-formed Process Hitting for IG inference]\label{pro:wf-ph}
-A PH is well-formed for IG inference if and only if the following conditions are verified:
-\begin{itemize}
-\item 
-each sort $a\in\PHs$ either belongs to $\Gamma$, or is a well-formed cooperative sort;
-\item 
-there is no cycle between cooperative sorts
-(the digraph $(\Sigma,\{(a,b)\in(\Sigma\times\Sigma)\mid \exists \PHfrappe{a_i}{b_j}{b_k}\in\PHa
-\wedge a\neq b\wedge \{a,b\}\cap\Gamma=\emptyset \})$ is
-acyclic);
-\item 
-sorts having no action hitting them belong to $\Gamma$
-($\{ a \in \Sigma\mid \nexists \PHfrappe{b_i}{a_j}{a_k}\in\PHa\} \subset \Gamma$).
-\end{itemize}
-\end{property}
-
-\begin{example}
-In the PH of \pref{fig:runningPH-2}, $bc$ is a well-formed cooperative sort as defined in \pref{pro:wf-cooperative-sort}, because:
-\begin{align*}
-\focals(bc, \PHl_{bc}, \{b_0, c_0\} \cup \PHl_{bc}) = \{bc_{00}\} && \focals(bc, \PHl_{bc}, \{b_0, c_1\} \cup \PHl_{bc}) = \{bc_{01}\} \\
-\focals(bc, \PHl_{bc}, \{b_1, c_0\} \cup \PHl_{bc}) = \{bc_{10}\} && \focals(bc, \PHl_{bc}, \{b_1, c_1\} \cup \PHl_{bc}) = \{bc_{11}\}
-\end{align*}
-Hence, both \pref{fig:runningPH-1} and \pref{fig:runningPH-2} are well-formed PH for IG inference
-with $\Gamma = \{a,b,c\}$.
-\end{example}
-
-\subsection{Interaction Inference}\label{ssec:infer-IG}
-
 At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
 that will not appear in the IG ($\PHs \setminus \Gamma$).
-We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
-by considering direct actions and possible intermediate cooperative sorts.
-The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
-(\pref{eq:regulators}).
-\begin{align}
-\begin{split}
-\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
-(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
-                                   & \quad \quad c^0 = b \wedge c^n = a \\
-                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
-                   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
-\end{split}
-\label{eq:ph_predec}
-\\
-\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
-\label{eq:regulators}
-\end{align}
-
-Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
-refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
-If $g=\{b\}$, we simply note $\ctx_b(\sigma)$.
-This set is composed of the active processes of sorts in $g$, and the focal process (assumed
-unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
-The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
-relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
-$\upsilon$ are defined in $\sigma$.
-When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
-the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
-where $\uplus$ stands for a union of states (which are sets of processes).
-Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
-always terminates.
-\begin{align}
-\forall g\subset \Gamma,
-  \ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
-\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
-\label{eq:ctx-sigma}
-\\
-\upsilon(\sigma) & \DEF
-\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
-  \upsilon'\in\PHdirectpredec{\upsilon} \wedge
-  \upsilon'\in\PHs\setminus\Gamma })
-\label{eq:cooperative-eval}
-\end{align}
-
-This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
-We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
-the level of $b$ makes possible the increase (decrease) of the level of $a$,
-which is directly inspired from the works of~\cite{Richard2010378}.
-To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
-Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
-$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
-\begin{equation}
-X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
-        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
-        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
-\label{eq:influence-groups}
-\end{equation}
 
@@
 \forall a_j\in L_a, j\neq i,
-a_j \in \irB_a^g(\sigma) &\EQDEF \focals(a, \{a_j\}, \ctx_{g}(\sigma)) = \{a_j\}
+a_j \in \irB_a^g(\sigma) &\EQDEF
+\exists b_k\in\ctx_{g}(\sigma): \PHfrappe{b_k}{a_i}{a_j}\in\PHa
 \label{eq:B-bounce}
@@
 \begin{proposition}[Influences inference]\label{pps:inference-edges}
-We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
+We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp.
+$\hat{E}_-$) by
 \modLP{
-\begin{align*}
-  \forall b\in\PHpredecgene{a}\cup\{a\}, \forall s \in \{ +, - \}, \\
-  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), 
-  	b\neq a \Rightarrow b \in g, \\
-	\qquad & \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
-    \qquad & \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
-    \qquad & \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
+\begin{align}
+\begin{split}
+\forall a\in\Gamma, \forall b\in\PHpredecgene{a}, b\neq a, & \forall s \in \{ +, - \}, \forall t < l_b\\
+  b \xrightarrow{t+1} a \in \hat{E}_s \EQDEF & \exists g \in X(a), 
+  	b \in g, \exists \sigma \in L(g\cup\{a\}), \\
+    \qquad & \exists a_j \in \irB_a^g(\sigma\{b_t\}), \exists a_k \in \irB_a^g(\sigma\{b_{t+1}\}), \\
+	\qquad & k > j \wedge s=+  \vee j < t \wedge s=-
+\end{split}
+\\
+\begin{split}
+ \forall a\in\Gamma: a\in\PHpredecgene{a}, & \forall s \in \{ +, - \}, \forall t < l_a\\
+  a \xrightarrow{t+1} a \in \hat{E}_s \EQDEF & \exists g \in X(a), 
+  	 \exists \sigma \in L(g\cup\{a\}), \\
+    \qquad & \exists a_j \in \irB_a^g(\sigma\{a_t\}), \exists a_k \in \irB_a^g(\sigma\{a_{t+1}\}), \\
 	\qquad & k\geq t+1 \wedge j \leq t \wedge s=+ \\
 	\qquad & \vee k<t+1 \wedge j>t \wedge s=-
-\end{align*}
+\end{split}
+\\
+\forall a\in\Gamma: \PHpredecgene{a} = \emptyset, &
+	\forall t\in\{0,\cdots,l_a-1\},
+	a \xrightarrow{t+1} a \in \hat{E}_+
+\end{align}
 }
@@
 
-\begin{example}
+\subsection{Examples}
+
 The IG inference of the PH of \pref{fig:runningPH-2} gives the
@@
 
-\end{example}
-
-\begin{example}
 If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
@@
 introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
-\end{example}

### 2014-09-24 parts-tr2global.tex
tr2global++

@@
 (local) behaviours in PH.
+\todo{introduce \pref{ssec:split-sorts}}
 
@@
 \[
-\forall \sigma \in {\textstyle\prod_{
-b\in\PHdirectpredec{\upsilon} \wedge b\neq a}}
-\PHl_{b},
+\forall \sigma \in L(\PHdirectpredec{\upsilon}),
 \focals(a,\PHl_a,\sigma) = \{ a(\sigma) \}\]
@@
 \todo{add more explanation depending on the text of the next sections}
+
+\subsection{Separating components from cooperative sorts}
+\label{ssec:split-sorts}
+
+The identification of a BRN from a PH assumes that the PH defines two types of sorts:
+the sorts corresponding to BRN components -- noted $\Gamma$ --, and the cooperative
+\modLP{(intermediate)} sorts -- noted $\Delta$ --
+\modLP{which should not appear in the BRN}.
+\modLP{In this subsection, we first give a criteria for identifying the sorts of a PH
+as either a component or cooperative sort.
+Then, we characterize the \emph{well-formed} PH for IG inference.}
+
+\modLP{The delimitation of sorts modeling components relies on the observation that their processes
+represent (ordered) qualitative levels.
+Hence an action on such a sort cannot make it bounce to a process at a distance more than one.
+This set of sorts in denoted by $\hat\Gamma$ (\pref{eq:PH-components}), whereas the
+set of cooperative sorts is denoted by $\hat\CSorts$ (\pref{eq:PH-coop}).}
+\begin{align}
+\hat\Gamma  &\DEF \{a \in \PHs \mid \nexists \PHfrappe{b_i}{a_j}{a_k} \in \PHa, |j - k| > 1\}
+\label{eq:PH-components}\\
+\hat\CSorts &\DEF \PHs \setminus \hat\Gamma
+\label{eq:PH-coop}
+\end{align}
+
+\modLP{Given a PH and a partition of its sorts in components $\Gamma$ and cooperative
+sorts $\Delta$,
+\pref{pro:wf-ph} establishes conditions for BRN identification:
+in addition of having $\Gamma$ compatible with $\hat\Gamma$ and cooperative
+sorts being deterministic (\pref{def:deterministic-sort})},
+we also require that there is no cycle between cooperative sorts, and that
+sorts being never hit (\ie serving as an invariant environment) are components.
+
+\begin{property}[Well-formed Process Hitting for BRN identification]\label{pro:wf-ph}
+A PH is well-formed for BRN identification with
+$\PHs = \Gamma\cup\CSorts$ only if the following conditions hold:
+\begin{enumerate}
+\item 
+\modLP{$\Gamma\cap\CSorts=\emptyset$, $\Gamma\subseteq\hat\Gamma$ and $\forall \upsilon\in\CSorts$, 
+$\upsilon$ is a deterministic sort (\pref{def:deterministic-sort})};
+\item 
+there is no cycle between cooperative sorts
+(the digraph $(\Sigma,\{(a,b)\in(\Sigma\times\Sigma)\mid \exists \PHfrappe{a_i}{b_j}{b_k}\in\PHa
+\wedge a\neq b\wedge \{a,b\}\cap\Gamma=\emptyset \})$ is
+acyclic);
+\item 
+sorts having no action hitting them belong to $\Gamma$
+($\{ a \in \Sigma\mid \nexists \PHfrappe{b_i}{a_j}{a_k}\in\PHa\} \subset \Gamma$).
+\end{enumerate}
+\end{property}
+
+\begin{example}
+In the PH of \pref{fig:runningPH-2}, $bc$ is a deterministic sort as defined in
+\pref{def:deterministic-sort}:
+\begin{align*}
+\focals(bc, \PHl_{bc}, \state{b_0, c_0}) = \{bc_{00}\} &&
+\focals(bc, \PHl_{bc}, \state{b_0, c_1}) = \{bc_{01}\} \\
+\focals(bc, \PHl_{bc}, \state{b_1, c_0}) = \{bc_{10}\} && 
+\focals(bc, \PHl_{bc}, \state{b_1, c_1}) = \{bc_{11}\}
+\end{align*}
+Hence, both \pref{fig:runningPH-1} and \pref{fig:runningPH-2} are well-formed PH
+for BRN identification with $\Gamma = \{a,b,c\}$ and $\CSorts=\{bc\}$.
+\end{example}
+
+\modLP{Assuming a PH and a partition $\Gamma$ and $\CSorts$ of its sort that satisfy
+\pref{pro:wf-ph}, }
+\pref{def:ph-pred} characterizes the set of predecessors of a sort $a$, that is, the sorts influencing $a$
+by considering direct actions and possible intermediate cooperative sorts.
+The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
+(\pref{eq:regulators}).
+\modLP{%
+\begin{definition}[$\PHpredec{a}$ and $\PHpredecgene{a}$]
+\label{def:ph-pred}
+\label{def:ph-reg}
+Given $a\in\PHs$, $\PHpredec{a}\subseteq\PHs$ is the smallest set satisfying the following
+conditions:
+\begin{align}
+\PHdirectpredec{a} &\subseteq \PHpredec{a}
+\\
+\upsilon\in\PHpredec{a}\cap\CSorts \Rightarrow \PHpredec{\upsilon}&\subseteq\PHpredec{a}
+\end{align}
+The components that are predecessors of $a$ are referred to as $\PHpredecgene{a}$:
+\begin{equation}
+\PHpredecgene{a} \DEF \PHpredec{a} \cap \Gamma
+\label{eq:regulators}
+\end{equation}
+\end{definition}}
+
+\todo{update text}
+This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
+To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
+Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
+$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on
+$a$ (\pref{def:influence-groups}).
+\modLP{%
+\begin{definition}[Partition $X(a)$ of $\PHpredecgene{a}$]
+\label{def:influence-groups}
+Given $a\in\PHs$,
+$X(a)$ is the finest partition of $\PHpredecgene{a}$ such that
+for any $b,c\in\PHpredecgene{a}$, $b\neq c$, if there exists a cooperative sort
+$\upsilon\in\PHdirectpredec{a}\cap\CSorts$ such that 
+$\{b,c\}\subseteq\PHpredecgene{\upsilon}$ then
+there exists $g\in X(a)$ such that $\{b,c\}\subseteq g$.
+\end{definition}
+}
+
+\todo{update text and improve definition}
+Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
+refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
+If $g=\{b\}$, we simply note $\ctx_b(\sigma)$.
+This set is composed of the active processes of sorts in $g$, and the focal process (assumed
+unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
+The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
+relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
+$\upsilon$ are defined in $\sigma$.
+When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
+the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
+where $\uplus$ stands for a union of states (which are sets of processes).
+Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
+always terminates.
+\modLP{%
+\begin{definition}[Configuration extension $\ctx_a^g(\sigma)$]
+\label{def:ctx-sigma}
+Given $a\in\PHs$, a group $g\in X(a)$ and a configuration 
+$\sigma\in L(g)$
+\todo{$L(g\cup \{a\})$ ?},
+$\ctx_a^g(\sigma)$ is the configuration $\sigma$ with the corresponding
+processes of the cooperative sorts in $\PHdirectpredec{a}$:
+\begin{align}
+\ctx_a^g(\sigma) & \DEF \sigma \uplus \state{\upsilon(\sigma) \mid\upsilon \in\PHdirectpredec{a} \cap \CSorts} 
+\label{eq:ctx-sigma}
+\\
+\upsilon(\sigma') & \DEF
+\begin{cases}
+\focals(\upsilon,L_\upsilon,\sigma') & \text{if
+$\PHdirectpredec{\upsilon}\subseteq\f{dom}(\sigma')$}
+\\
+\upsilon(\sigma' \uplus \state{\upsilon'(\sigma') \mid 
+  \upsilon'\in\PHdirectpredec{\upsilon} \cap \CSorts })
+  & \text{otherwise.}
+\end{cases}
+\label{eq:cooperative-eval}
+\end{align}
+\end{definition}
+}

### 2014-09-25 main.tex
start new running example with phage lambda
Modifications of abstract and intro
Modifications of abstract and intro according to reviewers' comments

@@
 %\parindent 0.5cm
-The Process Hitting (PH) \modLP{is a particular class of Indeterministic Asynchronous Automata
-Network (or safe Petri nets) on which have been recently designed several static analyses for
-dynamical properties that are tractable on very large systems.}
+\modOR{Qualitative formalisms offer a well-established alternative to the more traditionally used differential equation models of Biological Regulatory Networks (BRNs).
+These formalisms, introduced by Kauffman, Glas and Thomas, led to numerous theoretical works and practical tools to understand emerging behaviors.} \modMM{The analysis of dynamics of very large models is however a rather hard problem, which led us to previously introduce a new framework:} the Process Hitting (PH), \modMM{which} \modLP{is a particular class of Indeterministic Asynchronous Automata
+Network (or safe Petri nets).} \modMM{Its major advantage lies in the efficiency of several static analyses recently designed to assess dynamical properties, making it possible to tackle very large models.}
 
@@
 
+\input{parts-tr2global}
 \input{parts-infer-ig}
@@
 \input{parts-asp}
+\input{parts-phage-lambda}
 \input{parts-example}

### 2014-09-25 parts-grn.tex
Modifications of abstract and intro
Modifications of abstract and intro according to reviewers' comments

@@
 
-\todo{Name Kaufman as the creator of BRNs.}
+%\todo{Name Kaufman as the creator of BRNs.}
 
-We concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
+\modMM{
+The principle of qualitative modeling was introduced as synchronous Boolean networks by Stuart Kauffman on the one hand \cite{kauffman1969metabolic}, asynchronous Ren\'e Thomas’ networks on the other \cite{Thomas73}.
+Both models have been considered of interest and led to numerous works.
+In the following, we choose to focus on  Thomas' modelling, with its asynchronous and multi-level semantics.}
+
+\modMM{In this paragraph, we} concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
 \cite{Richard06,BernotSemBRN}.

### 2014-09-25 parts-intro.tex
Modifications of abstract and intro
Modifications of abstract and intro according to reviewers' comments

--- 2013-08-03 parts-intro.tex
@@
 With the purpose of analyzing these systems, they are often modeled as graphs which make it possible to determine the possible evolutions of all the interacting components of the system.
-Indeed, besides continuous models of physicists, often designed through systems of ordinary
-differential equations, a discrete modeling approach was initiated by René Thomas in 1973
-\cite{Thomas73}.
+\modOR{Besides continuous models of physicists, generally designed through systems of ordinary differential equations, modeling regulatory networks by means of Boolean networks has become popular in the wake of Kauffman work \cite{kauffman1969metabolic}.
+We based our work on a logical formalism developed by Ren\'e Thomas from 1973 \cite{Thomas73} that generalizes upon Boolean networks in the sense that it allows variables to have more than two values and transitions between states to occur asynchronously.}
 
@@
 specify Thomas' parameters, especially for large networks.
+\modOR{Other works have also been carried out with the aim of dealing with large systems (e.g.
+\cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generaly focus on some restricted properties,} \modOR{as finding out steady states.}
+%\todo{The reviewer proposes to nuance this with several papers:
+%\begin{itemize}
+%  \item Corblin, F., Tripodi, S., Fanchon, E., Ropers, D., \& Trilling, L. (2009).
+%    A declarative constraint-based method for analyzing discrete genetic
+%    regulatory networks. Bio Systems, 98(2), 91104.
+%  \item Bollman, D., Coln-Reyes, O., \& Orozco, E. (2007). Fixed points in
+%    discrete models for regulatory genetic networks. EURASIP journal on
+%    bioinformatics \& systems biology, 2007, 97356.
+%  \item Delaplace, F., Klaudel, H., Melliti, T., \& Sen, S. (2012). Analysis
+%    of modular organisation of interaction networks based on asymptotic
+%    dynamics. In D. Gilbert \& M. Heiner (Eds.), CMSB 2012
+%  \item Scalable Steady State Analysis of Boolean Biological Regulatory Net-
+%    works Ay F, Xu F, Kahveci T (2009) Scalable Steady State Analysis
+%    of Boolean Biological Regulatory Networks. PLoS ONE 4(12): e7992.
+%\end{itemize}
+%This can be in turn be nuanced as some papers only focus on rather specific aspects (like steady-states for instance).}
 
-In order to address the formal analysis of dynamical properties within very large BRNs, we recently
+In order to address the formal analysis of additional dynamical properties \modMM{(e.g., state reachabilities)} within very large BRNs, we recently
 introduced in \cite{PMR10-TCSB} a new formalism, named the \emph{Process Hitting} (PH), to model
@@
 cooperations (associated influences) between components.
+\modOR{The concept of cooperation refers to the} \modMM{way two (or more) components jointly influence a third one.
+In other words, it captures the logical functions stating how} \modOR{various elements coalesce and act together upon an other element among the network.}
+%\todo{Explain the notion of “cooperation” (informally) to be understandable until the definition of cooperative sorts.}
+
+The modeling and analysis of the qualitative dynamics of large interaction net- works face two important challenges.
+For modeling, it is very rare that all logical functions governing the dynamics are known, even if the interaction graph is con- sidered as prior knowledge.
+One technique would then be to enumerate all logical networks that are compatible with the prior knowledge on the interaction graph and partial logical functions.
+But, with large networks, such an enumeration suffers from high combinatorics.
+For analysis, because the size of the state transition graph grows exponentially with the number of components, proving dynamical properties becomes intractable for networks with hundreds or thousands of elements.
+In order to illustrate the challenge of modeling interactions with partial knowledge,
+
 This allows to model BRNs with a partial knowledge on precise evolution functions for components

### 2014-09-25 parts-tr2global.tex
section 3

@@
 
-Hence, identifying Thomas models from PH requires a step that lump a set of transitions that can be
-applied in a given context into a single process that correspond to the level toward which the
+Hence, identifying Thomas models from PH requires to lump a set of transitions that can be
+applied in a given configuration into a single process that correspond to the level toward which the
 component evolve.
 This step is detailed in \pref{ssec:focal} in which is defined the $\focals$ function associating 
-the farthest reachable processes, called \emph{focal processes}, of a given sort for a given context.
+the farthest reachable processes, called \emph{focal processes}, of a given sort
+for a given configuration.
 It is possible that several different focal processes are identified, indicating an indeterministic
@@
 (local) behaviours in PH.
-\todo{introduce \pref{ssec:split-sorts}}
+
+Finally, \pref{ssec:split-sorts} addresses the separation of PH sorts into
+components and intermediate cooperate sorts, the latter having to be masked in Thomas models.
+It notably defines the predecessors, (group of) regulators of a
+sort, and the extension of a configuration to incorporate the state of the
+cooperative sorts.
 
@@
 
-Given a sort $a\in\Sigma$ and a context delimited by 
+Given a sort $a\in\Sigma$ and a configuration delimited by 
 a sub-set of processes of sort $a$ and a sub-state $\sigma$
@@
 $\focals(a,S_a,\sigma)$ (\pref{def:focals}) is the set of processes of sort $a$ towards which
-$a$ will eventually converge in the scope of the context.
-The context delimits the set of states $s\in L$ where $s[a]\in S_a$, and for all 
+$a$ will eventually converge in the scope of the configuration.
+The configuration delimits the set of states $s\in L$ where $s[a]\in S_a$, and for all 
 sort $b\in\PHpred a, b\neq a$,
@@
 \pref{pro:no-focals} -- if $\focals(a,S_a,\sigma)$ is empty, there exists a limit cycle in the evolution of
-$a$ in the given context;
+$a$ in the given configuration;
 and \pref{pro:has-focals} -- if $\focals(a,S_a,\sigma)$ is non empty, all the evolutions of $a$ in the scope
-of the given context eventually terminate, and the resulting processes are in $\focals(a,S_a,\sigma)$.
+of the given configuration eventually terminate, and the resulting processes are in $\focals(a,S_a,\sigma)$.
 In other words, if $\focals(a,S_a,\sigma)$ is empty, there exists a
-sequence of actions that may be played an unbound number of times (cycle) in the given context;
-if it is non-empty, it is ensured that any state in the given context converges 
+sequence of actions that may be played an unbound number of times (cycle) in the
+given configuration;
+if it is non-empty, it is ensured that any state in the configuration converges 
 in a bounded number of steps to a process in $\focals(a,S_a,\sigma)$.
@@
 As discussed at the beginning of this section, and as formally described in the above sub-section, 
-in the scope of a given context,
+in the scope of a given configuration,
 a PH sort may converge to one among several different processes (indeterminism) or may never converge
@@
 \end{definition}
-
-\todo{examples?}
 
@@
 
-\modLP{Assuming a PH and a partition $\Gamma$ and $\CSorts$ of its sort that satisfy
+\modLP{Assuming a PH and a split $\Gamma$ and $\CSorts$ of its sort that satisfy
 \pref{pro:wf-ph}, }
-\pref{def:ph-pred} characterizes the set of predecessors of a sort $a$, that is, the sorts influencing $a$
-by considering direct actions and possible intermediate cooperative sorts.
+\pref{def:ph-pred} characterizes the set of predecessors of a sort $a$ as the sorts influencing $a$
+through direct actions or intermediate cooperative sorts.
 The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
@@
 
-\todo{update text}
-This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
-To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
-Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
-$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on
-$a$ (\pref{def:influence-groups}).
 \modLP{%
+As described in the introduction of this section, the identification of Thomas
+models from PH relies on the inference of focal processes in the
+different configurations of the regulators of each component.
+In order to reduce the scope of the configurations that need to be enumerated, 
+we introduce the notion of \emph{group of regulators} of sort $a$ as a set of
+components that have a joint influence on $a$.
+More precisely, $b$ and $c$ are in the same group of regulators of $a$ if there exists
+an intermediate cooperative sort $\upsilon$ that hits $a$ such that $b$ and $c$
+are regulators of $\upsilon$.
+Those groups form a partition of $\PHpredecgene{a}$.
+We denote $X(a)$ the finest partition in groups of
+regulators (\pref{def:influence-groups}), i.e. where each group is minimal.
 \begin{definition}[Partition $X(a)$ of $\PHpredecgene{a}$]
@@
 
-\todo{update text and improve definition}
-Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
-refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
-If $g=\{b\}$, we simply note $\ctx_b(\sigma)$.
-This set is composed of the active processes of sorts in $g$, and the focal process (assumed
-unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
-The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
-relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
-$\upsilon$ are defined in $\sigma$.
-When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
-the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
-where $\uplus$ stands for a union of states (which are sets of processes).
-Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
-always terminates.
 \modLP{%
+If the cooperative sorts are all deterministic (\pref{def:deterministic-sort}),
+it is sufficient to specify the configuration $\sigma$ of a group $g$ of regulators of a sort $a$
+to obtain the configuration of all the sorts hitting $a$ directly.
+Indeed, because of the absence of cycles between cooperative sorts
+(\pref{pro:wf-ph}), one can recursively
+evaluate the focal process of each cooperative sort that hit $a$ with respect to
+the configuration $\sigma$ of the regulators $g$.
+Such an extended configuration is denoted by $\ctx_a^g(\sigma)$, formalized in
+\pref{def:ctx-sigma}
 \begin{definition}[Configuration extension $\ctx_a^g(\sigma)$]
@@
 Given $a\in\PHs$, a group $g\in X(a)$ and a configuration 
-$\sigma\in L(g)$
-\todo{$L(g\cup \{a\})$ ?},
-$\ctx_a^g(\sigma)$ is the configuration $\sigma$ with the corresponding
+$\sigma\in L(g)$, $\ctx_a^g(\sigma)$ is the configuration $\sigma$ with the corresponding
 processes of the cooperative sorts in $\PHdirectpredec{a}$:

### 2014-09-25 parts-infer-ig.tex
section 3
section 4.1

@@
 \section{Interaction Graph Inference from Process Hitting}\label{sec:infer-IG}
-
-\todo{In this while section: add informal explanations and maybe more examples.\\
-Especially:
-\begin{itemize}
-  \item Intuitively detail the motivations
-  \item Give the intuition of the methods (esp. $\focals$)
-  \item Insist on the use/motivation of the definitions
-\end{itemize}
-The reviewer also proposes to use a biological running example (such as $\lambda$-phage).}
 
@@
 
-We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
-the level of $b$ makes possible the increase (decrease) of the level of $a$,
-which is directly inspired from the works of~\cite{Richard2010378}.
-
-We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ and a parition of sorts
+We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ and a split of sorts
 $\Gamma \cup \Delta=\PHs$ satisfying \pref{pro:wf-ph} on which the IG inference is to be
 performed.
-\todo{outline}
+The inference of the IG is described in \pref{ssec:infer-IG}, and is illustrated
+on small examples in \pref{ssec:infer-ig-examples}.
 
@@
 
-\todo{rework text with new paper structure (\pref{sec:tr2global})}
-
-At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
-that will not appear in the IG ($\PHs \setminus \Gamma$).
-
 \def\fp{\Phi}
 
+We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
+the level of $b$ makes possible the increase (decrease) of the level of $a$,
+following the standard IG inference from Boolean and discrete networks \cite{Richard2010378}.
+
 \modLP{
+Given a process $a_i$ and a configuration $\sigma$ of a group $g$ of regulators of $a$,
+we denote by
+$\irB_a^g(\sigma)$ (\pref{def:B-bounce}) the set of processes of sort $a$ towards which $a_i$ can bounce
+in the configuration $\sigma$.
+$\irB_a^g(\sigma)$ matches with the focal processes of $a$ in the scope of $a_i$
+(\pref{eq:B-bounce})
+except when no bounce can occur in $\sigma$:
+in such a case, we need to ensure that there exists a configuration of the other
+groups of regulators of $a$ where $a_i$ is never hit (\pref{eq:B-no-bounce}).
+The set of configuration of regulators of $a$ where $a_i$ is never hit is given
+by $\fp(a_i)$ (\pref{eq:fp}).
+\begin{definition}[$\irB_a^g(\sigma)$]
+\label{def:B-bounce}
+Given $a\in\PHs$, $g\in X(a)$ and $\sigma\in L(\PHpredecgene{a}\cup\{a\})$
+such that $a_i\in\sigma$,
+$\irB_a^g(\sigma)\subseteq L_a$ is defined as follows:
+\begin{align}
+\forall a_j\in L_a, j\neq i,
+a_j \in \irB_a^g(\sigma) &\EQDEF
+a_j \in \focals(a, \{a_i\}, \ctx_a^g(\sigma))
+%\exists b_k\in\ctx_{g}(\sigma): \PHfrappe{b_k}{a_i}{a_j}\in\PHa
+\label{eq:B-bounce}
+\\
+a_i \in \irB_a^g(\sigma) &\EQDEF \exists\sigma'\in\fp(a_i): \sigma\subset\sigma'
+\label{eq:B-no-bounce}
+\end{align}
+where $\fp(a_i) \subseteq  L(\PHpredecgene{a}\cup\{a\}$ is defined as follows:
 \begin{equation}
-\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid \sigma[a] = i \wedge
-					\focals(a, \{a_i\}, \ctx_{\PHpredecgene{a}}(\sigma)) =
+\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid a_i\in\sigma \wedge
+					\focals(a, \{a_i\}, \ctx_a^{\PHpredecgene{a}}(\sigma)) =
 					\{a_i\} \}
@@
 \end{equation}
-}
-
-To infer influence of a component $b$ on a component $a$,
-we observe the direction of evolution of $a$ for different active processes of $b$.
-The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
-Formally, for a given group $g$ of regulators of a component $a$
-(that is, a minimal set of components regulating $a$ through common cooperative sorts),
-and a configuration $\ctx$ on $g$, we note
-$\irB_a(\sigma)$ the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
-%This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
-%and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
-If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
-%
-
-\modLP{
-$\forall g\in X(a), \forall \sigma\in L(g\cup\{a\}), i=\sigma[a]$:
-\begin{align}
-a_i \in \irB_a^g(\sigma) &\EQDEF \exists\sigma'\in\fp(a_i): \sigma\subset\sigma'
-\label{eq:B-no-bounce}
-\\
-\forall a_j\in L_a, j\neq i,
-a_j \in \irB_a^g(\sigma) &\EQDEF
-\exists b_k\in\ctx_{g}(\sigma): \PHfrappe{b_k}{a_i}{a_j}\in\PHa
-\label{eq:B-bounce}
-\end{align}}
+\end{definition}}
 
@@
 with a threshold $t$.
-The main idea of this inference is that
-the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
-there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
-(\pref{eq:edges-inference}).
+\modLP{The inference relies on finding configurations $\sigma$ of group of regulators of $a$
+such that the increase of one of the regulator $b\in g$ from $b_t$ to $b_{t+1}$
+makes possible the increase (decrease) of $a$ (\pref{eq:edges-inference} when
+$b\neq a$ and \pref{eq:edges-inference-auto} when $b=a$)}.
 Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
-Furthermore, studying the influences of the groups of regulators of $a$
-allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
-Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
-We ignore the cases where a component has no visible influence on another.
+When $a$ has no regulators, it derives that there is a positive self-influence
+of $a$ for each level above $1$ (\pref{eq:edges-inference-noreg}).
 %
@@
 \end{split}
+\label{eq:edges-inference}
 \\
@@
 \end{split}
+\label{eq:edges-inference-auto}
 \\
@@
 	a \xrightarrow{t+1} a \in \hat{E}_+
+\label{eq:edges-inference-noreg}
 \end{align}
@@
 \begin{align*}
-E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
+E_+ &= \{ \GRNedge{a}{+}{t}{b} \in \hat E_+ \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
   \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
-E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
+E_- &= \{ \GRNedge{a}{-}{t}{b} \in \hat E_- \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
   \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
@@
 \subsection{Examples}
+\label{ssec:infer-ig-examples}
 

### 2014-09-25 parts-infer-ig.tex
section 4

@@
 
-\todo{The reviewer says about the following paragraph: “The inference of an
-IG from a dynamics is interesting and important in biology. However the
-motivations expressed here are not well argued. In particular I don't find
-this “common”. Some bibliographical references are required to consolidate
-the arguments or the sentence has to be rephrased.”}
-
-In a typical biological network modeling process, a prior IG is generally the starting point for
-the formal system specification.
-However, it is common that the prior IG actually refers to interactions that reveal to be 
-non-effective with respect to the dynamics.
-Hence, deriving the IG directly from the dynamical models lead to more concise IGs, enhancing the
-conclusiveness of static analyses based upon this abstract representation.
+\modLP{The inference of the IG also allows to check for the
+consistency of the model with respect to prior knowledge on the influences.
+This is the case for Thomas models, where a prior IG is required to specify the 
+parameters of the dynamics.
+Following \todo{ref Adrien, autre}, the IG of a dynamical model is consistent
+with a prior IG if it is a sub-graph of the prior IG.
+On the one hand, having influences in the model not references in the prior IG
+indicates either mistakes in the encoding of the dynamics, or potential
+influences that might need to be verified in the biological system.
+One the other hand, some influences specified in the prior IG may not have been
+necessary to encode the desired dynamics, indicating some non-minimality of the
+prior IG \todo{more refs}.}
 

### 2014-09-26 parts-infer-ig.tex
fix section 4.1

@@
 following the standard IG inference from Boolean and discrete networks \cite{Richard2010378}.
-
-\modLP{
-Given a process $a_i$ and a configuration $\sigma$ of a group $g$ of regulators of $a$,
-we denote by
-$\irB_a^g(\sigma)$ (\pref{def:B-bounce}) the set of processes of sort $a$ towards which $a_i$ can bounce
-in the configuration $\sigma$.
-$\irB_a^g(\sigma)$ matches with the focal processes of $a$ in the scope of $a_i$
-(\pref{eq:B-bounce})
-except when no bounce can occur in $\sigma$:
-in such a case, we need to ensure that there exists a configuration of the other
-groups of regulators of $a$ where $a_i$ is never hit (\pref{eq:B-no-bounce}).
-The set of configuration of regulators of $a$ where $a_i$ is never hit is given
-by $\fp(a_i)$ (\pref{eq:fp}).
-\begin{definition}[$\irB_a^g(\sigma)$]
-\label{def:B-bounce}
-Given $a\in\PHs$, $g\in X(a)$ and $\sigma\in L(\PHpredecgene{a}\cup\{a\})$
-such that $a_i\in\sigma$,
-$\irB_a^g(\sigma)\subseteq L_a$ is defined as follows:
-\begin{align}
-\forall a_j\in L_a, j\neq i,
-a_j \in \irB_a^g(\sigma) &\EQDEF
-a_j \in \focals(a, \{a_i\}, \ctx_a^g(\sigma))
-%\exists b_k\in\ctx_{g}(\sigma): \PHfrappe{b_k}{a_i}{a_j}\in\PHa
-\label{eq:B-bounce}
-\\
-a_i \in \irB_a^g(\sigma) &\EQDEF \exists\sigma'\in\fp(a_i): \sigma\subset\sigma'
-\label{eq:B-no-bounce}
-\end{align}
-where $\fp(a_i) \subseteq  L(\PHpredecgene{a}\cup\{a\}$ is defined as follows:
-\begin{equation}
-\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid a_i\in\sigma \wedge
-					\focals(a, \{a_i\}, \ctx_a^{\PHpredecgene{a}}(\sigma)) =
-					\{a_i\} \}
-\label{eq:fp}
-\end{equation}
-\end{definition}}
+This can be seen as looking for changes of derivatives in the dynamics of $a$
+when the configuration changes.
 
@@
 with a threshold $t$.
-\modLP{The inference relies on finding configurations $\sigma$ of group of regulators of $a$
+\modLP{The inference relies on finding configurations $\sigma$ of groups $g$ of regulators of $a$
 such that the increase of one of the regulator $b\in g$ from $b_t$ to $b_{t+1}$
-makes possible the increase (decrease) of $a$ (\pref{eq:edges-inference} when
-$b\neq a$ and \pref{eq:edges-inference-auto} when $b=a$)}.
-Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
-When $a$ has no regulators, it derives that there is a positive self-influence
+changes the sign of the evolution of $a$.
+If $b\neq a$ (\pref{eq:edges-inference}), this is achieved by finding a configuration $\sigma$ of $g$ and $a$ such
+that there exists a bounce $a_j$ from $a_i$ in $\sigma\{b_t\}$ which is different
+from a bounce $a_k$ from $a_i$ in $\sigma\{b_{t+1}\}$.
+If so, $b$ regulates $a$ at threshold $t+1$ with the sign of $k-j$.
+The case where  $b = a$ (\pref{eq:edges-inference-auto}) is similar: we look for
+$a_t$ such that there exists a bounce in opposite direction in $\sigma\{a_t\}$
+than in $\sigma\{a_{t+1}\}$.
+If there is no bounce in one of these two configurations, we also need to ensure
+that the configuration corresponds to a local fixed point between all the
+regulators of $a$ (instead of in scope of $g$).
+The set of configuration where no bounce occur on $a_i$ is given given by
+$\fp(a_i)$ (\pref{eq:fp}).
+In \pref{eq:edges-inference} and \pref{eq:edges-inference-auto} we use
+$\irB_a^g(\sigma)$ as a shortcut for $\focals$, which returns the set of
+possible bounces of $a$ in the configuration $\sigma$.}
+Finally, when $a$ has no regulators, it derives that there is a positive self-influence
 of $a$ for each level above $1$ (\pref{eq:edges-inference-noreg}).
-%
 \begin{proposition}[Influences inference]\label{pps:inference-edges}
@@
     \qquad & \exists a_j \in \irB_a^g(\sigma\{b_t\}), \exists a_k \in \irB_a^g(\sigma\{b_{t+1}\}), \\
-	\qquad & k > j \wedge s=+  \vee j < t \wedge s=-
+	\qquad & k > j \wedge s=+  \vee k < j \wedge s=-
 \end{split}
@@
     \qquad & \exists a_j \in \irB_a^g(\sigma\{a_t\}), \exists a_k \in \irB_a^g(\sigma\{a_{t+1}\}), \\
-	\qquad & k\geq t+1 \wedge j \leq t \wedge s=+ \\
+	\qquad & 
+	(j=t \Rightarrow \exists \sigma'\in\fp(a_t): \sigma\{a_t\}\subseteq \sigma') \\
+	\qquad & \wedge 
+	(k=t+1 \Rightarrow \exists \sigma'\in\fp(a_{t+1}): \sigma\{a_{t+1}\}\subseteq \sigma') \\
+	\qquad & \wedge k\geq t+1 \wedge j \leq t \wedge s=+ \\
 	\qquad & \vee k<t+1 \wedge j>t \wedge s=-
@@
 \end{align}
+where $\irB_a^g(\sigma) \DEF \focals(a, \{a_i\}, \ctx_a^g(\sigma))$ with
+$a_i\in\sigma$,
+and where $\fp(a_i) \subseteq  L(\PHpredecgene{a}\cup\{a\})$ is defined as follows:
+\begin{equation}
+\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid a_i\in\sigma \wedge
+					\focals(a, \{a_i\}, \ctx_a^{\PHpredecgene{a}}(\sigma)) =
+					\{a_i\} \}
+\label{eq:fp}
+\end{equation}
 }
@@
 In the end, the threshold of each edge is the minimum threshold for which an influence has been found.
+\modLP{Note that as regulations matches with sign changes of the evolution of the
+regulated components,
+if two positive (negative) regulations from $b$ to $a$ have been inferred,
+necessarily, a third regulation of opposite sign has been inferred in between.
+The only exception is when $a$ is never hit ($\PHpredecgene{a} = \emptyset$), in
+such a case, we arbitrarily pick the minimum threshold (1) for the
+self-activation.}
 %
@@
 \begin{align*}
+E_- &= \{ \GRNedge{a}{-}{t}{b} \in \hat E_- \mid \nexists a \xrightarrow{t'} b
+\in \hat{E}_+ \} \\
 E_+ &= \{ \GRNedge{a}{+}{t}{b} \in \hat E_+ \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
   \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
-E_- &= \{ \GRNedge{a}{-}{t}{b} \in \hat E_- \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
-  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
 E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_- \\

### 2014-09-26 parts-grn.tex
start updating sect 5 notations

@@
 
-\paragraph{Discussion on interval parameters}
-Whereas most of the literature use single value parameters (e.g., $K_{a,\omega} = i$), we introduce
-in this paper intervals as parameter values (e.g., $K_{a,\omega} = \segm{i}{j}$) for specifying
-the discrete functions for components.
-These interval parameters are mainly introduced for convenience, and it is worth noticing that they
-do not add extra expressivity for modeling multivalued networks in the general case.
-However, interval parameters are more expressive than Thomas' parameters with single value.
-First, we remark that if  $K_{a,\omega} = \segm{i}{j}$ with $i < j$, then, necessarily there exists
-a positive edge from $a$ to $a$ in the interaction graph of the dynamics.
-Then, let us consider the case where $K_{a,\omega} = \segm{i}{i+2}$ (interval of three values), and
-let $s$ be a state such that $\GRNres{a}{s} = \omega$.
-From \pref{def:dynamics}, if $\GRNget{s}{a} \in K_{a,\omega}$, then $a$ cannot evolve in $s$;
-therefore, $a$ has three (local) fixed points ($i$, $i+1$ and $i+2$).
-This behavior cannot be reproduced with single value parameters, because it is impossible to distinguish three different sets of resources corresponding to the three involved states.
-At most two steady-states can be modeled by using a self-regulation $\GRNedge{a}{s}{t}{a}$, allowing to distinguish two cases ($\GRNget{s}{a} < t$ and $\GRNget{s}{a} \geq t$), which is not enough.
+\todo{(LP) parameters equivalence:
+if $\GRNedge{a}{+}{t}{a}$,
+$K_{a,R\setminus \{a\}} = t
+\equiv
+K_{a,R\setminus \{a\}} = u$
+for $t \leq u \leq l_a$;
+$K_{a,R\cup \{a\}} = t-1
+\equiv
+K_{a,R\cup \{a\}} = u$
+for $0 \leq u < t$.
+If $\GRNedge{a}{-}{t}{a}$,
+$K_{a,R\cup \{a\}} = t
+\equiv
+K_{a,R\cup \{a\}} = u$
+for $t \leq u \leq l_a$;
+$K_{a,R\setminus \{a\}} = t-1
+\equiv
+K_{a,R\cup \{a\}} = u$
+for $0 \setminus u < t$.
+- any other equivalence?}

### 2014-09-26 parts-infer-param.tex
TODO on parameters equivalence
forgot to highlight my changes
tr2global++
start updating sect 5 notations

@@
 
-\todo{In this while section: add informal explanations and maybe more examples (see previous section).}
-
+\todo{update text: IG can be given by the user + details}
 Given the IG inferred from a PH as presented in the previous section, one can find the discrete parameters
@@
 It relies on an exhaustive enumeration of all predecessors of each component in order to find attractor processes
-and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations (\pref{ssec:infer-K}).
+and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations (\pref{ssec:infer-K})
+\todo{not clear}.
 The last step consists of the enumeration of all compatible complete parametrizations (\pref{ssec:admissible-K}) given this
 set of inferred parameters, the PH dynamics and some biological constraints on parameters.
+\todo{not clear}
 
@@
 These results are equivalent to those presented in \cite{PMR10-TCSB}, with notation adapted to be shared with the previous section.
-In addition, we introduce the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
+In addition, we characterize the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
 which implies that %the inferred IG does not contain any unsigned interactions, and thus can be seen as the
@@
 A PH is well-formed for parameter inference if and only if
-it is well-formed for IG inference, and
-the IG $(\Gamma, E)$ inferred by \pref{pps:inference-IG}
-verifies the following property:
+it is well-formed for BRN identification (\pref{pro:wf-ph}) and
+that the associated IG $(\Gamma, E)$ verifies the following property:
+\modLP{
+\begin{align*}
+\forall b\rightarrow a\in E&, b\neq a,
+	\forall \sigma\in L(\PHpredecgene{a}),\\
+    \forall (i,j & \in \levels{b}{a} \vee i,j \in \ulevels{b}{a}),
+	\forall a_k \in L_a,
+	\\
+&	\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_i\})))
+	=\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_j\})))
+\end{align*}
+\begin{align*}
+\forall a\in\Gamma, &\forall b\in\PHpredecgene{a},
+b\neq a \wedge b\rightarrow a\notin E,\\
+	\forall \sigma&\in L(\PHpredecgene{a}), 
+    \forall i,j \in L_b,
+	\forall a_k \in L_a,
+	\\
+&	\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_i\})))
+	=\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_j\})))
+\end{align*}
+}
 \todo{Define $c(\sigma)$ for $\sigma \in \textstyle\prod_{d \in \PHpredec{c}} \PHl_d$
 (\ie the unique focal point)}
+\modMF{
 \begin{align*}
@@
 \end{align*}
+}
 \end{property}
@@
   \label{eq:param_context}
-  \forall b\in\Gamma,~
-  C_{a,\omega}^b & \DEF \begin{cases}
-    \levels{b}{a} & \text{if $b \in \omega$,}\\
-    \ulevels{b}{a} & \text{if $b \notin \omega$,}\\
-    L_b            & \text{otherwise;}\\
+  \forall b\rightarrow a\in E,~
+  \sigma_{a,\omega}^b & \DEF \begin{cases}
+    \modLP{\min}(\levels{b}{a}) & \text{if $b \in \omega$,}\\
+    \modLP{\min}(\ulevels{b}{a}) & \text{if $b \notin \omega$}
   \end{cases}
   \\
-  \label{eq:param_context_coop}
-  \forall \upsilon \in \PHpredec{a}\setminus\Gamma,~
-  C_{a,\omega}^\upsilon & \DEF \{
-  \upsilon(\sigma) \mid \sigma \in \textstyle\prod_{c\in\PHdirectpredec{\upsilon}}C_{a,\omega}^c \}
+  \modLP{
+  \forall b\in\PHpredecgene a, b\rightarrow a\notin E,~
+  \sigma_{a,\omega}^b} & \modLP{\DEF b_0}
   \\
-  C_{a,\omega} & \DEF \textstyle\bigcup_{b\in\PHpredec{a}} C^b_{a,\omega}
+  \sigma_{a,\omega} & \DEF
+  \modLP{\ctx_a^{\PHpredecgene{a}}(
+  	\state{\sigma^b_{a,\omega} \mid  b\in\PHpredecgene{a}})}
   \label{eq:K-ctx}
+  \\
+  \modLP{
+  \forall a\in\Gamma,~
+  S_a(\ctx)} &\modLP{\DEF
+	\begin{cases}
+	\{a_i\} & \text{if }a_i\in \ctx\\
+	L_a & \text{otherwise.}
+	\end{cases}}
 \end{align}
@@
 \pref{eq:param_context_coop}.
-Hence $K_{a,\omega} = \focals(a,C^a_{a,\omega},C_{a,\omega})$ if this latter is a non-empty interval
-(\pref{pps:param_K}).
+Hence \modLP{$K_{a,\omega} = \focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})$ if this latter is
+a singleton, denoting a deterministic behaviour (\pref{pps:param_K})}.
 
@@
 $a$.
-If $\focals(a,C^a_{a,\omega},C_{a,\omega})=\segm{a_i}{a_j}$ is a non-empty interval, then $K_{a,\omega} = \segm{i}{j}$.
+If $\focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})= \{ a_i \}$, then \modLP{$K_{a,\omega} = i$}.
 \end{proposition}
@@
 
+\todo{update notations}
 \begin{property}[Parameter validity]\label{pro:K-valid}

### 2014-10-01 parts-infer-param.tex
typo
layout
section 5++

@@
 
-\todo{update text: IG can be given by the user + details}
-Given the IG inferred from a PH as presented in the previous section, one can find the discrete parameters
-that model the behavior of the studied PH using the method presented in the following.
-It relies on an exhaustive enumeration of all predecessors of each component in order to find attractor processes
-and returns a possibly incomplete parametrization, given the exhaustiveness of the cooperations (\pref{ssec:infer-K})
-\todo{not clear}.
-The last step consists of the enumeration of all compatible complete parametrizations (\pref{ssec:admissible-K}) given this
-set of inferred parameters, the PH dynamics and some biological constraints on parameters.
-\todo{not clear}
+\modLP{
+Given a PH and an IG (either arbitrary, or inferred following the previous section), this section
+addresses the identification of the parametrization for the corresponding Thomas model.}
+
+\modLP{
+As described in \pref{ssec:infer-K}, this identifcation relies on the computation of the focal
+processes for each configuration of the parameters.
+When there is a unique focal process of a component for a given configuration of its regulators,
+the focal process matches with the value of the associated Thomas parameter.}
+
+\modLP{
+However, in the general case, as described in \pref{sec:tr2global} and even when the PH is
+well-formed for parameter inference (characterized in the next subsection by \pref{pro:wf-ph-K}),
+there maybe several focal processes (non-deterministic behaviour) or none (cyclic behaviour).  Such
+a case occurs notably when the PH encodes the union of several Boolean or discrete networks, as
+described in \cite{PMR10-TCSB,PCFMR14-chapterLMBS}.}
+
+\modLP{
+Hence, we propose in \pref{ssec:admissible-K} the notion of \emph{compatible} parametrization with respect
+ot the PH dynamics: a Thomas model is compatible with a PH if Thomas dynamics is included in the PH
+dynamics, i.e., all the transitions in the Thomas model are possible transitions in the PH.
+This relaxed notion of parameterization compatibility allows to enumerate all parametrizations
+compatible with a given PH, i.e. all Thomas model whose dynamics is included in the PH dynamics.}
 
@@
 
-This subsection presents some results related to the inference of independent discrete parameters from a given PH.
-These results are equivalent to those presented in \cite{PMR10-TCSB}, with notation adapted to be shared with the previous section.
+This subsection addresses the inference of independent discrete parameters from a given PH.
+The inference is equivalent to the one in \cite{PMR10-TCSB}.
 In addition, we characterize the well-formed PH for parameter inference property (\pref{pro:wf-ph-K}),
@@
 %and that 
-any process in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior
-regarding $a$.
-
-\todo{makes it standalone: repeat the constraints for IG inference that are
-actually necessary + check that we can do if if the IG is provided (then what
-are the constraints on this IG)}
+any process in $\levels{b}{a}$ (resp. $\ulevels{b}{a}$) share the same behavior regarding $a$.
 
@@
 it is well-formed for BRN identification (\pref{pro:wf-ph}) and
-that the associated IG $(\Gamma, E)$ verifies the following property:
+that the associated IG $(\Gamma, E)$ verifies that:
 \modLP{
-\begin{align*}
-\forall b\rightarrow a\in E&, b\neq a,
+\begin{align}
+\begin{split}
+\forall b\rightarrow a&\in E, b\neq a,
 	\forall \sigma\in L(\PHpredecgene{a}),\\
@@
 	=\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_j\})))
-\end{align*}
-\begin{align*}
+\end{split}
+\\
+\begin{split}
 \forall a\in\Gamma, &\forall b\in\PHpredecgene{a},
@@
 	\forall \sigma&\in L(\PHpredecgene{a}), 
-    \forall i,j \in L_b,
+    \forall b_i,b_j \in L_b,
 	\forall a_k \in L_a,
@@
 	=\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_j\})))
-\end{align*}
-}
-\todo{Define $c(\sigma)$ for $\sigma \in \textstyle\prod_{d \in \PHpredec{c}} \PHl_d$
-(\ie the unique focal point)}
-\modMF{
-\begin{align*}
-  \begin{split}
-%   \forall a \in \Gamma&, \forall b\in \GRNreg{a},
-%           \forall (i,j\in\levels{b}{a} \vee i,j\in\ulevels{b}{a}), \\
-%   & \quad \forall c \in \PHs, ( (b\neq a\wedge c=a) \vee (c\in\PHpredec{a}\setminus \Gamma \wedge b\in\PHdirectpredec{c})), \\
-%   & \qquad
-%                           \PHfrappe{b_i}{c_k}{c_l}\in\PHa \Leftrightarrow
-%                                   \PHfrappe{b_j}{c_k}{c_l}\in\PHa
-      &
-    \forall a \in \Gamma,
-    \forall g \in X(a),
-    \forall \sigma, \sigma' \in \textstyle\prod_{d \in g} \PHl_d,
-    \forall b \in g,
-    b \in \GRNreg{c},
-    \forall c \in \PHdirectpredec{a} \setminus \Gamma,
-      \\&\qquad
-    b \in \reg{c},
-    \forall (i,j \in \levels{b}{a} \vee i,j \in \ulevels{b}{a}),
-%    \forall b_i \in \PHl_b,
-    b_i \in \sigma,
-%    b_i \in \levels{b}{a},
-%    \forall b_j \in \levels{b}{a},
-      \\&\qquad\qquad
-    \sigma' = \sigma\{b_j\},
-    %c_{\sigma} \in \PHl_c \wedge c_{\sigma'} \in \PHl_c,  %facultatif
-    \PHfrappe{c(\sigma)}{a_i}{a_j} \in \PHa \Leftrightarrow
-    \PHfrappe{c(\sigma')}{a_i}{a_j} \in \PHa
-  \end{split}\\
-  \begin{split}
-    & \forall a \in \Gamma, \forall b \in \GRNreg{a},
-      \forall (i,j \in \levels{b}{a} \vee i,j \in \ulevels{b}{a}), \\
-    & \qquad \forall c \in \PHs, b \neq a,
-      \PHfrappe{b_i}{a_k}{a_l} \in \PHa \Leftrightarrow
-      \PHfrappe{b_j}{a_k}{a_l} \in \PHa
 \end{split}
-\end{align*}
+\end{align}
 }
@@
 
-Let $K_{a,\omega}$ be the parameter we want to infer for a given component $a \in \Gamma$
-and $\omega \subset \GRNreg{a}$ a set of its regulators.
-This inference, as for the IG inference, relies on the search of focal processes of the component for the given configuration of its regulators.
+Let $K_{a,\omega}$ be a Thomas parameter for a given component $a \in \Gamma$ 
+with $\omega \subset \GRNreg{a}$ a set of its regulators.
+As described in \pref{ssec:thomas}, $K_{a,\omega}$ specifies to which values $a$ eventually evolves
+in the configuration matching with $\omega$.
 
-For each sort $b \in \GRNreg{a}$, we define a context $C^b_{a,\omega}$ in \pref{eq:param_context} that contains all processes representing the influence of the resources in the configuration modeled by $\omega$.
-The context of a cooperative sort $\upsilon$ that regulates $a$ is given in
-\pref{eq:param_context_coop} as the set of focal processes matching the current configuration.
-$C_{a,\omega}$ refers to the union of all these contexts (\pref{eq:K-ctx}).
+\modLP{The configuration of the PH corresponding to $\omega$ is given as follows.
+For each component $b \in \GRNreg{a}$, we define $\sigma^b_{a,\omega}$ (\pref{eq:param_context}) as 
+the process of $b$ with the level in 
+$\levels{b}{a}$ if $b\in\omega$, or in $\ulevels{b}{a}$ if $b\not\omega$.
+Because of \pref{pro:wf-ph-K}, if several possible levels exist, this process can be chosen
+arbitrarily.
+If $b$ does not regulate $a$ in the IG, all processes of sort $b$ should have the same action on
+$a$, so the process $b_0$ is arbitrarily selected (\pref{eq:param_context_free}).
+The configuration $\sigma_{a,\omega}$ corresponding to $\omega$ (\pref{eq:K-ctx})
+is then obtained by extending the configuration of the
+regulators of $a$ to the (deterministic) cooperative sorts (\pref{def:ctx-sigma} in
+\pref{ssec:split-sorts}).
+Finally, we denote by $S_a(\sigma)$ (\pref{eq:K-domain}) the set of processes of sort $a$ that are
+compatible with a configuration $\sigma$: if $a$ is specified in $\ctx$, then $S_a(\sigma) = \{a_i\}$
+where $a_i\in\sigma$; otherwise $S_a(\sigma)$ is the all the processes of sort $a$, $L_a$.}
 \begin{align}
@@
   \sigma_{a,\omega}^b} & \modLP{\DEF b_0}
+  \label{eq:param_context_free}
   \\
@@
   \forall a\in\Gamma,~
-  S_a(\ctx)} &\modLP{\DEF
+  S_a(\sigma)} &\modLP{\DEF
 	\begin{cases}
@@
 	\end{cases}}
+\label{eq:K-domain}
 \end{align}
 
-The parameter $K_{a,\omega}$ specifies to which values $a$ eventually evolves as long as the context
-$C_{a,\omega}$ holds, which is precisely the definition of the $\focals$ function
-(\pref{def:focals} in \pref{ssec:focal}),
-where the focals reachability property can be derived from \pref{pro:wf-ph-K} and
-\pref{eq:param_context_coop}.
-Hence \modLP{$K_{a,\omega} = \focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})$ if this latter is
+Therefore, we obtain that
+\modLP{$K_{a,\omega} = \focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})$ if this latter is
 a singleton, denoting a deterministic behaviour (\pref{pps:param_K})}.
@@
 \label{pps:param_K}
-Let $(\PHs, \PHl, \PHh)$ be a Process Hitting well-formed for parameter inference, and $\IG = (\Gamma, E)$ the inferred IG.
+Let $(\PHs, \PHl, \PHh)$ be a Process Hitting \modLP{with $\PHs=\Gamma\cup\CSorts$} 
+well-formed for parameter inference, and $\IG = (\Gamma, E)$ the inferred IG.
 Let $A$ (resp. $B$) $\subseteq \Gamma$ be the set of regulators that activate (resp. inhibit) a sort
@@
 \begin{align*}
-K_{a, \emptyset} &= \segm{0}{0} &
-K_{b, \emptyset} &= \segm{0}{0} \\
-K_{a, \{a\}} &= \segm{0}{0} &
-K_{b, \{a\}} &= \segm{0}{0} \\
-K_{a, \{c\}} &= \segm{1}{1} &
-K_{b, \{b\}} &= \segm{1}{1} \\
-K_{a, \{b\}} &= \segm{1}{1} &
-K_{b, \{a,b\}} &= \segm{0}{0} \\
-K_{a, \{b,c\}} &= \segm{1}{1} &
-K_{c, \emptyset} &= \segm{0}{0} \\
-K_{a, \{a,b,c\}} &= \segm{2}{2} &
-K_{c, \{c\}} &= \segm{1}{1}
+K_{a, \emptyset} &= 0&
+K_{b, \emptyset} &= 0 \\
+K_{a, \{a\}} &= 0 &
+K_{b, \{a\}} &= 0 \\
+K_{a, \{c\}} &= 1 &
+K_{b, \{b\}} &= 1 \\
+K_{a, \{b\}} &= 1 &
+K_{b, \{a,b\}} &= 0 \\
+K_{a, \{b,c\}} &= 1 &
+K_{c, \emptyset} &= 0 \\
+K_{a, \{a,b,c\}} &= 2 &
+K_{c, \{c\}} &= 1
 %K_{a, \{a,b\}} = ?
@@
 \begin{align*}
-  K_{a,\{a,b\}} &= \segm{1}{1} &
-  K_{a,\{a,c\}} &= \segm{1}{1}
+  K_{a,\{a,b\}} &= 1 &
+  K_{a,\{a,c\}} &= 1
 \end{align*}

### 2014-10-24 parts-grn.tex
section 5++
Tentative de remodelage de 2.2

@@
 \subsection{Thomas' modeling}
+\label{ssec:thomas}
 
@@
 Both models have been considered of interest and led to numerous works.
-In the following, we choose to focus on  Thomas' modelling, with its asynchronous and multi-level semantics.}
+In the following, we choose to focus on  Thomas' modelling, with its asynchronous, unitary and multi-level semantics.}
 
@@
 \emph{Interaction Graph} (IG) models the structure of the system by defining the components'
-mutual influences and the conditions of these influences.
-The \emph{parametrization} then specifies the levels towards which tends a component when a given
-configuration of its regulators applies.
+mutual influences and the conditions of these influences. \modMM{
+Then t}he \emph{parametrization} \modMM{
+removes the ambiguity when a component is targeted by (at least) two different influences.
+In other words, it} specifies the levels towards which a component tends when a given configuration of its regulators applies.
 
@@
 when its expression level is strictly below this threshold, another kind of influence is expressed, which usually consists of the opposite influence.
+
+\modMM{The IG is not sufficient to unambiguously define the dynamics of the system.
+In particular, a component may be regulated by both a positive and negative interaction derived from two different components.
+To refine the dynamics, a parametrization had been added to the model \cite{snoussi1989qualitative}.
+This parametrization allows to define focal points, i.e.
+it clarifies the level towards which a component evolves depending on the expression of its regulators.
+Some of the information included in the IG and the parametrization may appear as slighltly redundant.
+Especially, if the single parametrization were enriched with additional constraints (e.g., monotonicity requiring an ordering on the value of parameters), then it may directly capture the signs.
+However, distinguishing the IG and parametrization allows to decouple the graph (which is generally one of the prior knowledge from the biologists) from the constraints on its dynamics.
+It opened the way to numerous studies (e.g., necessary condition on sustained oscillations) solely based on the structure of the IG, e.g., the analysis of positive or negative circuits.}
+
 %For a regulation to take place (activation or inhibition), the expression level of its head component has to be higher than its threshold;

### 2014-10-24 parts-example.tex
intro
complete phage lambda example

--- 2013-08-02 parts/example.tex
@@
 
-This section details the implementation and usage of the piece of software developed for this work.
-\pref{ssec:appli} gives a practical and detailed application of our results on a biological model taken from the literature,
-while \pref{ssec:cpu} gives data related to the results and computation times of our software when applied to several large biological models.
+This section aims at giving several applications of our work in order to
+understand its results and range of applications.
+First, \pref{ssec:phage-lambda} gives a detailed application of our IG inference method
+by studying parts of a model of the phage lambda immunity control.
+% the functioning of our method
+% and the implementation of the software developed for this work.
+% First, \pref{ssec:phage-lambda} aims at giving a detailed example of the application of our method
+% on parts of a model of the phage lambda immunity control.
+%We then explain the use of the implementation and usage of the software developed for this work:
+Then, \pref{ssec:appli} gives a practical application of our results on a biological model
+of epithelial growth factor receptor taken from the literature.
+Finally, \pref{ssec:cpu} gives data related to the results and computation times of our software
+when applied to several large biological models.
 
@@
 The inference methods described in this paper have been implemented as part of
-\textsc{Pint}\footnote{Available at \url{http://process.hitting.free.fr}}, which gathers PH related tools.
+\textsc{Pint}\footnote{Available at \url{http://loicpauleve.name/pint}}, which gathers PH related tools.
 Our implementation mainly consists in ASP programs that are solved using
@@
 or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
+\todo{Remove \texttt{full-enumerate}?}
+
+\input{parts-phage-lambda}
 

### 2014-10-24 parts-infer-ig.tex
complete phage lambda example

--- 2014-09-25 parts-infer-ig.tex
@@
 
-We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ and a split of sorts
-$\Gamma \cup \Delta=\PHs$ satisfying \pref{pro:wf-ph} on which the IG inference is to be
+In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
+modeling.
+This section first introduces the notion of focal processes within a PH (\pref{ssec:focal})
+which is used to characterize well-formed PH for IG inference (\pref{ssec:wf}).
+%and as well used by the parametrization inference presented in \pref{sec:infer-K}.
+The rules for inferring the interactions between components from a PH are
+described in \pref{ssec:infer-IG}.
+We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
 performed.
-The inference of the IG is described in \pref{ssec:infer-IG}, and is illustrated
-on small examples in \pref{ssec:infer-ig-examples}.
 
-\subsection{Inference of Influences}\label{ssec:infer-IG}
+\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}
+
+\todo{rework text with new paper structure (\pref{sec:tr2global})}
+
+The inference of an IG from a PH assumes that the PH defines two types of sorts:
+the sorts corresponding to BRN components, and the cooperative sorts.
+This leads to the characterization of a \emph{well-formed} PH for IG inference.
+
+The identification of sorts modeling components relies on the observation that their processes
+represent (ordered) qualitative levels.
+Hence an action on such a sort cannot make it bounce to a process at a distance more than one.
+The set of sorts satisfying such a condition is referred to as $\Gamma$
+(\pref{eq:PH-components}).
+Therefore, in the rest of this paper, $\Gamma$ denotes the set of components of the BRN to infer.
+
+\begin{equation}
+\Gamma \DEF \{a \in \PHs \mid \nexists \PHfrappe{b_i}{a_j}{a_k} \in \PHa, |j - k| > 1\} \\
+\label{eq:PH-components}
+\end{equation}
+
+\todo{maybe we should also support a user-supplied $\Gamma$ and give the
+constraints (included in the prev. def?)}
+
+Finally, \pref{pro:wf-ph} sums up the conditions for a Process Hitting to be suitable for IG
+inference.
+In addition of having either component sorts or well-formed cooperative sorts, we also require that
+there is no cycle between cooperative sorts, and that
+sorts being never hit (\ie serving as an invariant environment) are components.
+
+\begin{property}[Well-formed Process Hitting for IG inference]\label{pro:wf-ph}
+A PH is well-formed for IG inference if and only if the following conditions are verified:
+\begin{itemize}
+\item 
+each sort $a\in\PHs$ either belongs to $\Gamma$, or is a well-formed cooperative sort;
+\item 
+there is no cycle between cooperative sorts
+(the digraph $(\Sigma,\{(a,b)\in(\Sigma\times\Sigma)\mid \exists \PHfrappe{a_i}{b_j}{b_k}\in\PHa
+\wedge a\neq b\wedge \{a,b\}\cap\Gamma=\emptyset \})$ is
+acyclic);
+\item 
+sorts having no action hitting them belong to $\Gamma$
+($\{ a \in \Sigma\mid \nexists \PHfrappe{b_i}{a_j}{a_k}\in\PHa\} \subset \Gamma$).
+\end{itemize}
+\end{property}
+
+\begin{example}
+In the PH of \pref{fig:runningPH-2}, $bc$ is a well-formed cooperative sort as defined in \pref{pro:wf-cooperative-sort}, because:
+\begin{align*}
+\focals(bc, \PHl_{bc}, \{b_0, c_0\} \cup \PHl_{bc}) = \{bc_{00}\} && \focals(bc, \PHl_{bc}, \{b_0, c_1\} \cup \PHl_{bc}) = \{bc_{01}\} \\
+\focals(bc, \PHl_{bc}, \{b_1, c_0\} \cup \PHl_{bc}) = \{bc_{10}\} && \focals(bc, \PHl_{bc}, \{b_1, c_1\} \cup \PHl_{bc}) = \{bc_{11}\}
+\end{align*}
+Hence, both \pref{fig:runningPH-1} and \pref{fig:runningPH-2} are well-formed PH for IG inference
+with $\Gamma = \{a,b,c\}$.
+\end{example}
+
+\subsection{Interaction Inference}\label{ssec:infer-IG}
+
+At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
+that will not appear in the IG ($\PHs \setminus \Gamma$).
+We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
+by considering direct actions and possible intermediate cooperative sorts.
+The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
+(\pref{eq:regulators}).
+\begin{align}
+\begin{split}
+\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
+(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
+                                   & \quad \quad c^0 = b \wedge c^n = a \\
+                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
+                   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
+\end{split}
+\label{eq:ph_predec}
+\\
+\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
+\label{eq:regulators}
+\end{align}
+
+Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
+refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
+If $g=\{b\}$, we simply note $\ctx_b(\sigma)$.
+This set is composed of the active processes of sorts in $g$, and the focal process (assumed
+unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
+The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
+relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
+$\upsilon$ are defined in $\sigma$.
+When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
+the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
+where $\uplus$ stands for a union of states (which are sets of processes).
+Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
+always terminates.
+\begin{align}
+\forall g\subset \Gamma,
+  \ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
+\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
+\label{eq:ctx-sigma}
+\\
+\upsilon(\sigma) & \DEF
+\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
+  \upsilon'\in\PHdirectpredec{\upsilon} \wedge
+  \upsilon'\in\PHs\setminus\Gamma })
+\label{eq:cooperative-eval}
+\end{align}
+
+This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
+We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
+the level of $b$ makes possible the increase (decrease) of the level of $a$,
+which is directly inspired from the works of~\cite{Richard2010378}.
+To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
+Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
+$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
+\begin{equation}
+X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
+        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
+        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
+\label{eq:influence-groups}
+\end{equation}
 
@@
 
-We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
-the level of $b$ makes possible the increase (decrease) of the level of $a$,
-following the standard IG inference from Boolean and discrete networks \cite{Richard2010378}.
-
 \modLP{
-Given a process $a_i$ and a configuration $\sigma$ of a group $g$ of regulators of $a$,
-we denote by
-$\irB_a^g(\sigma)$ (\pref{def:B-bounce}) the set of processes of sort $a$ towards which $a_i$ can bounce
-in the configuration $\sigma$.
-$\irB_a^g(\sigma)$ matches with the focal processes of $a$ in the scope of $a_i$
-(\pref{eq:B-bounce})
-except when no bounce can occur in $\sigma$:
-in such a case, we need to ensure that there exists a configuration of the other
-groups of regulators of $a$ where $a_i$ is never hit (\pref{eq:B-no-bounce}).
-The set of configuration of regulators of $a$ where $a_i$ is never hit is given
-by $\fp(a_i)$ (\pref{eq:fp}).
-\begin{definition}[$\irB_a^g(\sigma)$]
-\label{def:B-bounce}
-Given $a\in\PHs$, $g\in X(a)$ and $\sigma\in L(\PHpredecgene{a}\cup\{a\})$
-such that $a_i\in\sigma$,
-$\irB_a^g(\sigma)\subseteq L_a$ is defined as follows:
-\begin{align}
-\forall a_j\in L_a, j\neq i,
-a_j \in \irB_a^g(\sigma) &\EQDEF
-a_j \in \focals(a, \{a_i\}, \ctx_a^g(\sigma))
-%\exists b_k\in\ctx_{g}(\sigma): \PHfrappe{b_k}{a_i}{a_j}\in\PHa
-\label{eq:B-bounce}
-\\
-a_i \in \irB_a^g(\sigma) &\EQDEF \exists\sigma'\in\fp(a_i): \sigma\subset\sigma'
-\label{eq:B-no-bounce}
-\end{align}
-where $\fp(a_i) \subseteq  L(\PHpredecgene{a}\cup\{a\}$ is defined as follows:
 \begin{equation}
-\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid a_i\in\sigma \wedge
-					\focals(a, \{a_i\}, \ctx_a^{\PHpredecgene{a}}(\sigma)) =
+\fp(a_i) \DEF \{\sigma\in L(\PHpredecgene{a}\cup\{a\}) \mid \sigma[a] = i \wedge
+					\focals(a, \{a_i\}, \ctx_{\PHpredecgene{a}}(\sigma)) =
 					\{a_i\} \}
@@
 \end{equation}
-\end{definition}}
+}
+
+To infer influence of a component $b$ on a component $a$,
+we observe the direction of evolution of $a$ for different active processes of $b$.
+The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
+Formally, for a given group $g$ of regulators of a component $a$
+(that is, a minimal set of components regulating $a$ through common cooperative sorts),
+and a configuration $\ctx$ on $g$, we note
+$\irB_a(\sigma)$ the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
+%This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
+%and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
+If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
+%
+
+\modLP{
+$\forall g\in X(a), \forall \sigma\in L(g\cup\{a\}), i=\sigma[a]$:
+\begin{align}
+a_i \in \irB_a^g(\sigma) &\EQDEF \exists\sigma'\in\fp(a_i): \sigma\subset\sigma'
+\label{eq:B-no-bounce}
+\\
+\forall a_j\in L_a, j\neq i,
+a_j \in \irB_a^g(\sigma) &\EQDEF
+\exists b_k\in\ctx_{g}(\sigma): \PHfrappe{b_k}{a_i}{a_j}\in\PHa
+\label{eq:B-bounce}
+\end{align}}
 
@@
 with a threshold $t$.
-\modLP{The inference relies on finding configurations $\sigma$ of group of regulators of $a$
-such that the increase of one of the regulator $b\in g$ from $b_t$ to $b_{t+1}$
-makes possible the increase (decrease) of $a$ (\pref{eq:edges-inference} when
-$b\neq a$ and \pref{eq:edges-inference-auto} when $b=a$)}.
+The main idea of this inference is that
+the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
+there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
+(\pref{eq:edges-inference}).
 Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
-When $a$ has no regulators, it derives that there is a positive self-influence
-of $a$ for each level above $1$ (\pref{eq:edges-inference-noreg}).
+Furthermore, studying the influences of the groups of regulators of $a$
+allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
+Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
+We ignore the cases where a component has no visible influence on another.
 %
@@
 \begin{align}
-\begin{split}
+\begin{split}\label{eq:inference-edges-ba}
 \forall a\in\Gamma, \forall b\in\PHpredecgene{a}, b\neq a, & \forall s \in \{ +, - \}, \forall t < l_b\\
@@
     \qquad & \exists a_j \in \irB_a^g(\sigma\{b_t\}), \exists a_k \in \irB_a^g(\sigma\{b_{t+1}\}), \\
-	\qquad & k > j \wedge s=+  \vee j < t \wedge s=-
+	\qquad & k > j \wedge s=+  \vee \modMF{k < j} \wedge s=-
 \end{split}
-\label{eq:edges-inference}
 \\
-\begin{split}
+\begin{split}\label{eq:inference-edges-aa}
  \forall a\in\Gamma: a\in\PHpredecgene{a}, & \forall s \in \{ +, - \}, \forall t < l_a\\
@@
       \exists \sigma \in L(g\cup\{a\}), \\
-    \qquad & \exists a_j \in \irB_a^g(\sigma\{a_t\}), \exists a_k \in \irB_a^g(\sigma\{a_{t+1}\}), \\
+    \qquad & \exists a_j \in \irB_a^{g \modMF{\cup \{a\}}}(\sigma\{a_t\}),
+      \exists a_k \in \irB_a^{g \modMF{\cup \{a\}}}(\sigma\{a_{t+1}\}), \\
 	\qquad & k\geq t+1 \wedge j \leq t \wedge s=+ \\
@@
 \end{split}
-\label{eq:edges-inference-auto}
 \\
@@
 	a \xrightarrow{t+1} a \in \hat{E}_+
-\label{eq:edges-inference-noreg}
 \end{align}
@@
 \begin{align*}
-E_+ &= \{ \GRNedge{a}{+}{t}{b} \in \hat E_+ \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
+E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
   \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
-E_- &= \{ \GRNedge{a}{-}{t}{b} \in \hat E_- \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
+E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
   \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
@@
 
-\subsection{Examples}
-\label{ssec:infer-ig-examples}
-
+\begin{example}
 The IG inference of the PH of \pref{fig:runningPH-2} gives the
@@
 
+\end{example}
+
+\begin{example}
 If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
@@
 introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
+\end{example}

### 2014-10-24 parts-phage-lambda.tex
complete phage lambda example

@@
-\begin{example}
+\modMF{
+\subsection{Detailed example on the bacteriophage lambda immunity control model}
+}
+\label{ssec:phage-lambda}
+
+\newcommand{\cI}{\f{cI}}
+\newcommand{\cro}{\f{cro}}
+\newcommand{\cII}{\f{cII}}
+\newcommand{\N}{\f{N}}
+
+\newcommand{\tcIcrocII}{cI\mbox{-}cro\mbox{-}cII}
+\newcommand{\tcIcroN}{cI\mbox{-}cro\mbox{-}N}
+\newcommand{\tcIcro}{cI\mbox{-}cro}
+\newcommand{\cIcrocII}{\f{\tcIcrocII}}
+\newcommand{\cIcroN}{\f{\tcIcroN}}
+\newcommand{\cIcro}{\f{\tcIcro}}
+
+In order to illustrate the IG inference developed as \pref{sec:infer-IG},
+this first subsection focuses on a model of the lambda phage immunity control,
+which has been widely studied mainly for its particular switch mechanism,
+allowing this virus to “chose” between lysis and lysogenization.
+These two possible responses are characterized by different dynamics that lead to two
+separate attractors.
+
+We consider here the model with four genes proposed in \cite{thieffry_dynamical_1995},
+which will serve as example to detail the inference method proposed in the previous sections.
+This model can be straightforwardly represented under the form of a Process Hitting
+with 4 components %($\Gamma = \{ \cI, \cro, \cII, \N \}$)
+and 3 cooperative sorts: %($\PHs \setminus \Gamma = \{ \cIcrocII, \cIcroN, \cIcro \}$).
+$$\PHs = \{ \underbrace{\cI, \cro, \cII, \N}_{\Gamma}, \cIcrocII, \cIcroN, \cIcro \}$$
+The four components have respectively 3, 4 2 and 2 expression levels,
+which we denote:
 \begin{align*}
-  \PHs &= \{ cI, cro, cII, N \} \\
-  \PHl_{cI} &= \{ cI_0, cI_1, cI_2 \} &
-  \PHl_{cro} &= \{ cro_0, cro_1, cro_2, cro_3 \} \\
-  \PHl_{cII} &= \{ cII_0, cII_1 \} &
-  \PHl_{N} &= \{ N_0, N_1 \} \\
+  \PHl_{\cI} &= \{ \cI_0, \cI_1, \cI_2 \} &
+  \PHl_{\cro} &= \{ \cro_0, \cro_1, \cro_2, \cro_3 \} \\
+  \PHl_{\cII} &= \{ \cII_0, \cII_1 \} &
+  \PHl_{\N} &= \{ \N_0, \N_1 \} \\
 \end{align*}
+
+The full Process Hitting model is not represented here due to a lack of space,
+but excerpts are given in \pref{fig:phage-lambda-cIcro-N} and \pref{fig:phage-lambda-cIcro-cro}.
+
+\subsubsection{Regulation of $\cI$ on $\N$}
+
+We first study the influence of $\cI$ on $\N$.
+The relevant part of the Process Hitting model for this problem is depicted in
+\pref{fig:phage-lambda-cIcro-N}.
+It involves a third component $\cro$ and the cooperative sort $\cIcro$;
+indeed, by studying the full model, we find:
+\[\PHpredec{\N} = \{ \cI, \cro, \cIcro \} \qquad
+  \PHpredecgene{\N} = \{ \cI, \cro \} \qquad
+  X(\N) = \{ \{ \cI, \cro \} \}\]
+Thus, we only have to focus on the group of regulators $\{ \cI, \cro \}$
+in order to analyze the regulation $\cI \rightarrow \N$.
+The idea behind the IG inference developed in \pref{sec:infer-IG}
+is to make the level of the considered regulator vary (here, $\cI$)
+while keeping constant all the other regulators in the same group (here, only $\cro$)
+and to observe the results on the evolution of the observed component (here, $\N$).
+In other words, if increasing the sole level of $\cI$ makes $\N$ tend to decrease,
+then a negative local influence of $\cI$ on $\N$ is inferred.
 
@@
   % Sortes
+  \TSetSortLabel{cI}{$\cI$}
   \TSort{(0,3.5)}{cI}{3}{l}
+  \TSetSortLabel{cro}{$\cro$}
   \TSort{(0,-1.5)}{cro}{4}{l}
-  \TSort{(7,1.5)}{N}{2}{r}
+  \TSetSortLabel{N}{$\N$}
+  \TSort{(7.5,1.5)}{N}{2}{r}
 
   % Sorte coopérative
-  \TSetTick{cIcro}{0}{cI~$<$~1\\cro~$<$~2}
-  \TSetTick{cIcro}{1}{cI~$<$~1\\cro~$\geq$~2}
-  \TSetTick{cIcro}{2}{cI~$\geq$~1\\cro~$<$~2}
-  \TSetTick{cIcro}{3}{cI~$\geq$~1\\cro~$\geq$~2}
+  \newcommand{\ttable}[3]{\scriptsize%
+    $#1\left\{\begin{array}{@{}l}\cI #2 1\\\cro #3 2\end{array}\right.$}
+  \TSetTick{cIcro}{0}{\ttable{FF}{<}{<}}
+  \TSetTick{cIcro}{1}{\ttable{FT}{<}{\leq}}
+  \TSetTick{cIcro}{2}{\ttable{TF}{\leq}{<}}
+  \TSetTick{cIcro}{3}{\ttable{TT}{\leq}{\leq}}
+  \TSetSortLabel{cIcro}{$\cIcro$}
   \TSort{(4,0.5)}{cIcro}{4}{r}
@@
   % Actions sortantes
-  \THit{cIcro_3}{}{N_0}{.west}{N_1}
-  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.south} ;
-  \THit{cIcro_2}{}{N_1}{.west}{N_0}
+  \THit{cIcro_3}{}{N_1}{.west}{N_0}
   \path[bounce,bend right] \TBounce{N_1}{}{N_0}{.north} ;
+  \THit{cIcro_2}{}{N_0}{.north west}{N_1}
+  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.260} ;
+  \THit{cIcro_1}{}{N_0}{.west}{N_1}
+  \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.230} ;
+  \THit{cIcro_0}{}{N_0}{.south west}{N_1}
+  \path[bounce,bend left=40] \TBounce{N_0}{}{N_1}{.200} ;
 \end{tikzpicture}
@@
   A portion of the PH model used in this section modeling the phage lambda immunity response.
-  This part of the whole model only depicts the interactions towards $N$ from components
-  $cI$ and $cro$.
-  The processes of the cooperative sort $cIcro$ are labeled by the configuration they represent.
+  This part of the whole model only depicts the interactions towards $\N$ from components
+  $\cI$ and $\cro$.
+  The processes of the cooperative sort $\cIcro$ are labeled by the configuration they represent,
+  given the values of $\cI$ and $\cro$ regarding their respective thresholds.
 }
 \end{figure}
+
+We first note that in the case where $\cro_3$ is present, we have from \pref{eq:ctx-sigma}:
+\begin{align*}
+  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \cI_0, \cro_3, \cIcro_{FT} \} \\
+  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \cI_1, \cro_3, \cIcro_{TT} \} \\
+  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \cI_2, \cro_3, \cIcro_{TT} \}
+\end{align*}
+We can then deduce the following directions of evolution:
+\begin{align*}
+  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} &&
+    \text{as } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) &&
+    \text{(from \pref{eq:B-no-bounce})} \\
+  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} && &&
+    \text{(from \pref{eq:B-bounce})} \\
+  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \} && &&
+    \text{(from \pref{eq:B-bounce})}
+\end{align*}
+Therefore, \pref{eq:inference-edges-ba} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
+The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
+do not add more information on the influences towards $\N$.
+Thus, given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cI}{-}{1}{\N} \in E$.
+
+We note that an equivalent work on $\cro$ also gives the second influence on $\N$:
+$\GRNedgef{\cro}{-}{2}{\N} \in E$.
+
+\subsubsection{Self-regulation of $\cro$ due to $\cI$}
+
+We now wish to study the self-regulation of $\cro$.
+For this, we will show that studying only the actions self-hitting $\cro$ is not enough;
+indeed, it is required to also study the interactions with $\cI$.
+We have the following results from the whole model:
+\[\PHpredec{\cro} = \{ \cI, \cro \} \qquad
+  \PHpredecgene{\cro} = \{ \cI, \cro \} \qquad
+  X(\cro) = \{ \{ \cI \}, \{ \cro \} \}\]
+The interactions of $\cro$ and $\cI$ on $\cro$ have been represented in
+\pref{fig:phage-lambda-cIcro-cro}.
 
@@
   A portion of the PH model used in this section modeling the phage lambda immunity response.
-  This part of the whole model only depicts the interactions towards $cro$ from components
-  $cI$ and $cro$.
+  This part of the whole model only depicts the interactions towards $\cro$ from components
+  $\cI$ and $\cro$.
 }
 \end{figure}
-\end{example}
 
-\begin{example}
-  The IG inference of \pref{pps:inference-IG} applied to the PH of \todo{ref}
-  produces the IG given in \pref{fig:phage-lambda-ig}.
+Let us first focus on the group of regulators $\{ \cro \}$.
+Of course, $\forall \cro_i \in \PHl_{\cro}, \ctx_{\{ \cro \}}(\PHetat{\cro_i}) = \{ \cro_i \}$.
+However, we note that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
+therefore $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$, and:
+%, there exist no sub-state $\sigma \in \Phi(\cro_1)$
+\begin{align*}
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_3}) &= \{ \cro_2 \} && &&
+    \text{(from \pref{eq:B-bounce})} \\
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_2}) &= \emptyset &&
+    \text{because } \Phi(\cro_2) = \emptyset &&
+    \text{(from \pref{eq:B-no-bounce})} \\
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_1}) &= \emptyset &&
+    \text{because } \Phi(\cro_1) = \emptyset &&
+    \text{(from \pref{eq:B-no-bounce})} \\
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_0}) &= \{ \cro_0 \} &&
+    \text{because } \PHetat{\cI_2, \cro_0} \in \Phi(\cro_0) &&
+    \text{(from \pref{eq:B-no-bounce})}
+\end{align*}
+
+We thus cannot infer any local influence by observing the group of regulators $\{ \cro \}$.
+Let us now focus instead on the group of regulators $\{ \cI \}$.
+One again, we trivially have:
+$\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
+  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}$.
+Furthermore:
+\begin{align*}
+%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
+%   \text{(from \pref{eq:B-bounce})} \\
+%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
+%   \text{(from \pref{eq:B-bounce})} \\
+  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_2}) &= \{ \cro_3 \} &&
+    \text{(from \pref{eq:B-bounce})} \\
+  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3}) &= \{ \cro_2 \} &&
+    \text{(from \pref{eq:B-bounce})}
+\end{align*}
+Therefore, \pref{eq:inference-edges-aa} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
+% The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
+% do not add more information on the influences towards $\N$.
+Finally, as no more information is added by observing the other levels of $\cro$ and $\cI$,
+and given \pref{pps:inference-IG}, it comes: $\GRNedgef{\cro}{-}{3}{\cro} \in E$.
+
+\subsubsection{Conclusion on the phage lambda immunity control}
+
+The IG inference of \pref{pps:inference-IG} applied to the whole Process Hitting modeling
+the phage lambda immunity response produces the IG given in \pref{fig:phage-lambda-ig}.
 This IG contains all edges included in the original model except the
-  self-influence on $cI$ : $\GRNedgef{cI}{+}{2}{cI} \notin E$;
+self-influence on $\cI$ (namely, $\GRNedgef{\cI}{+}{2}{\cI} \notin E$);
 this is due to the fact that an equivalent dynamics exists without this edge.
+
 \begin{figure}[t]
@@
   %\path[use as bounding box] (-1.3,-0.75) rectangle (3.5,1.5);
-    \node[inner sep=0] (cI) at (0,2.5) {cI};
-    \node[inner sep=0] (cro) at (2.5,2.5) {cro};
-    \node[inner sep=0] (cII) at (0,0) {cII};
-    \node[inner sep=0] (N) at (2.5,0) {N};
+  \node[inner sep=0] (cI) at (0,2.5) {$\cI$};
+  \node[inner sep=0] (cro) at (2.5,2.5) {$\cro$};
+  \node[inner sep=0] (cII) at (0,0) {$\cII$};
+  \node[inner sep=0] (N) at (2.5,0) {$\N$};
   \path[->]
@@
 \caption{\label{fig:phage-lambda-ig}%
-    Result of the IG inference performed on the PH of \todo{ref}.
+  Result of the IG inference performed on the Process Hitting model
+  of the phage lambda immunity response
+  taken from \cite{thieffry_dynamical_1995}.
 }
 \end{figure}
-\end{example}
 
-\begin{example}
-  The parameters inference of \pref{pps:param_K} applied to the PH of \todo{ref}
-  is conclusive for all parameters.
+Finally, we note that the parameters inference of \pref{pps:param_K} applied to the same model,
+and using the output of the previous IG inference is conclusive for all parameters.
 The complete parametrization produced is available in \pref{tb:phage-lambda-k}.
+
 \begin{table}[t]
 \begin{align*}
-    K_{cI, \emptyset} &= \{2\} &
-    K_{cI, \{cro\}} &= \{0\} \\
-    K_{cI, \{cII\}} &= \{2\} &
-    K_{cI, \{cII;cro\}} &= \{2\} \\
-    K_{N, \emptyset} &= \{1\} &
-    K_{N, \{cro\}} &= \{0\} \\
-    K_{N, \{cI\}} &= \{0\} &
-    K_{N, \{cI;cro\}} &= \{0\} \\
-    K_{cII, \emptyset} &= \{0\} &
-    K_{cII, \{cro\}} &= \{0\} \\
-    K_{cII, \{cI\}} &= \{0\} &
-    K_{cII, \{cI;cro\}} &= \{0\} \\
-    K_{cII, \{N\}} &= \{1\} &
-    K_{cII, \{N;cro\}} &= \{0\} \\
-    K_{cII, \{N;cI\}} &= \{0\} &
-    K_{cII, \{N;cI;cro\}} &= \{0\} \\
-    K_{cro, \emptyset} &= \{3\} &
-    K_{cro, \{cro\}} &= \{2\} \\
-    K_{cro, \{cI\}} &= \{0\} &
-    K_{cro, \{cI;cro\}} &= \{2\}
+  K_{\cI, \emptyset} &= \{2\} &
+  K_{\cI, \{\cro\}} &= \{0\} \\
+  K_{\cI, \{\cII\}} &= \{2\} &
+  K_{\cI, \{\cII;cro\}} &= \{2\} \\
+  K_{\N, \emptyset} &= \{1\} &
+  K_{\N, \{\cro\}} &= \{0\} \\
+  K_{\N, \{\cI\}} &= \{0\} &
+  K_{\N, \{\cI;cro\}} &= \{0\} \\
+  K_{\cII, \emptyset} &= \{0\} &
+  K_{\cII, \{\cro\}} &= \{0\} \\
+  K_{\cII, \{\cI\}} &= \{0\} &
+  K_{\cII, \{\cI;cro\}} &= \{0\} \\
+  K_{\cII, \{\N\}} &= \{1\} &
+  K_{\cII, \{\N;cro\}} &= \{0\} \\
+  K_{\cII, \{\N;cI\}} &= \{0\} &
+  K_{\cII, \{\N;cI;cro\}} &= \{0\} \\
+  K_{\cro, \emptyset} &= \{3\} &
+  K_{\cro, \{\cro\}} &= \{2\} \\
+  K_{\cro, \{\cI\}} &= \{0\} &
+  K_{\cro, \{\cI;cro\}} &= \{2\}
 \end{align*}
 \caption{\label{tb:phage-lambda-k}%
-    Result of the parameters inference performed on the PH of \todo{ref}.
+  Result of the parameters inference performed on the Process Hitting model
+  of the phage lambda immunity response
+  taken from \cite{thieffry_dynamical_1995}.
+  \todo{Remove the curly brackets $\{\}$?}
 }
 \end{table}
-\end{example}
+
+We note two main differences between the parametrization obtained by our inference
+and the one proposed in \cite{thieffry_dynamical_1995}:
+\begin{itemize}
+  \item The absence of the self-regulation of $\cI$ automatically reduces the size
+    of the parametrization, but this does not modify the dynamics, as mentioned previously;
+  \item Our inference returns $K_{\cro, \{\cI;cro\}} = 2$ instead of originally $0$;
+    once again, this does not change the dynamics, and $1$ would also have been
+    an acceptable value.
+\end{itemize}

### 2014-10-24 response.tex
update

@@
 \DISCUSS{LP}
-We have to clarify this point, while insisting that we want to stick with this
-standard definition (due to Thomas). Arguments:
-\begin{enumerate}
-\item this design allows to decouple the IG from the dynamics, which is
-interesting for modelling biological systems as the IG is often a prior
-knowledge with possible inconsistencies with the dynamics we want to encode
-with.
-A further step may then infer back the IG of the actual dynamics (here, an
-abstraction of the encoded dynamics) and compare it with the prior IG. Typically, if the inferred IG is included in the prior
-one, we have consistency.
-\item Additional constraints (as discussed in sect XX) on parameters can also
-ensure consistency between the prior IG and the encoded dynamics.
-For instance, monotonicity (which imposes an ordering between the value of
-parameters), but also lighter constraints (like imposing minimum and maximum
-values).
-Hence, the K definition is very generic, ranging from totally decoupled
-specification to very constrained valuation w.r.t to the prior IG.
-\item Several alternative definition exist, with various degree of equivalency.
-We picked this definition as it the original and standard one in the Thomas
-community (give reference to Thomas' articles).
-\end{enumerate}
+Thomas' formalism with discrete parameters indeed includes some redundant piece of information in the sense that signs could be captured by the parameters with additional constraints, e.g.
+monotonous behaviors (which imposes an ordering between the value of parameters).
+Speaking from a computer science oriented point of view, the information encapsulated into a Thomas' network could be reduced to the set of nodes and the corresponding logical parameters.
+The single parametrization is very generic, ranging from totally decoupled specification to very constrained valuation w.r.t to the interaction graph.
+This would however be a disruption with the Thomas framework as it has been considered for more than 25 years.
 
-\TODO{Answer inc. text modification}
+Moreover signs and edges have been an inspiration for numerous works with regard to the analysis of networks.
+As showed for example in THOMAS, R.
+\& D'ARI, R., "Biological feedback", CRC Press, Boca Raton , Florida, USA, (1990), 316 pp), the analysis of some dynamical properties (e.g.
+sustained oscillations) can derive from the study of positive or negative circuits.
+
+In addition, having both the interaction graph and the parametrization in the model allows to decouple the IG from the dynamics, which is interesting for modeling biological systems as the IG is often a prior knowledge with possible inconsistencies with the dynamics we want to encode with.
+A further step may then infer back the IG of the actual dynamics (here, an abstraction of the encoded dynamics) and compare it with the prior IG.
+Typically, if the inferred IG is included in the prior one, we have consistency.
+
+Finally, several alternative definitions exist, with various degree of equivalency.
+As a consequence, we chose here to stick to the original formalism and picked this definition as it the original and standard one in the Thomas' community.
+
+Nevertheless, to remove ambiguity and comply with the reviewer's suggestions, we redesigned section 2.2, emphasizing such a choice.
+
+%\TODO{Answer inc. text modification}
 
@@
 systems, such as IG and Thomas' parameters.
+\DISCUSS{LP} the general the notion of cooperation should be defined in the
+introduction.
 

### 2014-10-25 parts-infer-param.tex
some typos in K inference

@@
 	\\
-&	\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_i\})))
-	=\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_j\})))
+&	\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_i\}))
+	=\focals(a,\{a_k\},\ctx_a^{\PHpredecgene{a}}(\sigma\{b_j\}))
 \end{split}
@@
 the process of $b$ with the level in 
-$\levels{b}{a}$ if $b\in\omega$, or in $\ulevels{b}{a}$ if $b\not\omega$.
+$\levels{b}{a}$ if $b\in\omega$, or in $\ulevels{b}{a}$ if $b\notin\omega$.
 Because of \pref{pro:wf-ph-K}, if several possible levels exist, this process can be chosen
@@
 compatible with a configuration $\sigma$: if $a$ is specified in $\ctx$, then $S_a(\sigma) = \{a_i\}$
-where $a_i\in\sigma$; otherwise $S_a(\sigma)$ is the all the processes of sort $a$, $L_a$.}
+where $a_i\in\sigma$; otherwise $S_a(\sigma)$ contains all the processes of sort $a$, $L_a$.}
 \begin{align}

### 2014-10-25 parts-grn.tex
todo--

@@
 
-\todo{Discuss about the meaning of signs (especially $+$ and $-$) and their link with parameters.
-Especially, these signs introduce redundancy (and possibly contradiction) with the parameters.
-\begin{itemize}
-  \item We can discuss more about the choice we made here: the signs are only informative
-    (it eases the read of the IG), but are used as a basis for the parameters enumeration;
-  \item The reviewer suggests to remove signs, or only present them as an abstraction of the monotonicity assumption, given the levels and parameters of a BRN;
-  \item It is also possible to add the monotonicity assumption in the definition of a BRN to avoid contradictions.
-\end{itemize}}
-
 \begin{definition}[Interaction Graph]
@@
 \end{definition}
-
-\todo{The notation $\GRNreg{b}$ is poorly chosen; should be replaced by $E^{-1}(b)$ for instance.}
 
@@
 \end{definition}
-The notations “$<$” and “$>$” comparing an integer and an interval were defined at the end of \pref{sec:intro}.
 
@@
 
-\paragraph{Discussion on maximum level restrictions}
-\todo{The reviewer says that the discussion on the number of maximum levels has
-already been fully developed in other articles related to Thomas and that it is not
-necessary to develop this just to state that a PH may accept any number of levels.}
-The maximum level $l_a$ of a component $a$ is usually chosen equal to the number of outgoing edges of $a$ in the IG, \ie the number of components that $a$ regulates.
-Indeed, each discrete expression level in $\segm{0}{l_a}$ stands for an arbitrary set of concentration levels of the protein coded by $a$
-and for which it has a certain influence on several other components that it regulates.
-Allowing more expression levels makes several levels useless as they represent the same influence of $a$ than some other levels;
-on the contrary, reducing the maximum level requires that several thresholds on outgoing edges of $a$ share the same value, which is biologically unlikely.
-However, these considerations are usually taken into account when one has to directly build a model in this framework, but are not mandatory.
-In the scope of this paper, we allow any number of levels for components, without considering the number of outgoing edges,
-as the number of processes in a PH sort is not constrained in any way.
-For instance, in the example of \pref{fig:runningBRN},
-it would be more relevant to have $l_a = 1$ and $\GRNedgef{a}{-}{1}{b} \in E$, as $a$ has only one outgoing edge,
-but the value $l_a = 2$ makes the example more interesting; this is why $\GRNedgef{a}{-}{2}{b} \in E$ can be arbitrarily chosen.
-
-\paragraph{Discussion on unsigned edges}
-\todo{Discuss the meaning of signs (cf. supra).
-Also, the reviewer proposes to move this discussion before in the text,
-and to define unsigned IGs before signed IGs as the latter is just a consequence of the monotonicity assumption.}
-Unsigned edges allow to model regulations whose trend is more complex than a simple activation or inhibition.
-It has to be noted that using unsigned edges in addition to the regular positive and negative ones does not add expressivity to the framework.
-Indeed, signs have no impact on the parametrization (\pref{def:param}) or the dynamics (\pref{def:dynamics}).
-They do have an impact later in this paper, when admissible parametrizations are enumerated (\pref{ssec:admissible-K}).
-However, signs in an IG are useful to understand the trend of a regulation.
-PH actions are very atomistic and the inference of the underlying BRN proposed in \pref{sec:infer-IG}
-allows to find the global trends of regulations between components,
-which may turn out to be difficult to perform by hand.
-As signs have no direct influence on the semantics, unsigned edges can be convenient to model an IG with partial knowledge on some regulations.
-
 \todo{(LP) parameters equivalence:

### 2014-10-25 parts-phage-lambda.tex

@@
 \modMF{
-\subsection{Detailed example on the bacteriophage lambda immunity control model}
+\subsection{Detailed example: the bacteriophage lambda immunity control}
 }
@@
   \TSetTick{cIcro}{0}{\ttable{FF}{<}{<}}
-  \TSetTick{cIcro}{1}{\ttable{FT}{<}{\leq}}
-  \TSetTick{cIcro}{2}{\ttable{TF}{\leq}{<}}
-  \TSetTick{cIcro}{3}{\ttable{TT}{\leq}{\leq}}
+  \TSetTick{cIcro}{1}{\ttable{FT}{<}{\geq}}
+  \TSetTick{cIcro}{2}{\ttable{TF}{\geq}{<}}
+  \TSetTick{cIcro}{3}{\ttable{TT}{\geq}{\geq}}
   \TSetSortLabel{cIcro}{$\cIcro$}
@@
 \end{align*}
-We can then deduce the following directions of evolution:
+We can then deduce the following directions of evolution from \pref{eq:irB}:
 \begin{align*}
-  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} &&
-    \text{as } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) &&
-    \text{(from \pref{eq:B-no-bounce})} \\
-  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} && &&
-    \text{(from \pref{eq:B-bounce})} \\
-  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \} && &&
-    \text{(from \pref{eq:B-bounce})}
+  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} \\
+%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
+%     \text{(from \pref{eq:B-no-bounce})} \\
+  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} \\
+%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
+%     \text{(from \pref{eq:B-bounce})} \\
+  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \}
+%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
+%     \text{(from \pref{eq:B-bounce})}
 \end{align*}
-Therefore, \pref{eq:inference-edges-ba} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
+          %      eq:inference-edges-ba
+Therefore, \pref{eq:edges-inference} gives: $\cI \xrightarrow{1} \N \in \hat{E}_-$.
 The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
@@
 Let us first focus on the group of regulators $\{ \cro \}$.
-Of course, $\forall \cro_i \in \PHl_{\cro}, \ctx_{\{ \cro \}}(\PHetat{\cro_i}) = \{ \cro_i \}$.
-However, we note that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
-therefore $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$, and:
-%, there exist no sub-state $\sigma \in \Phi(\cro_1)$
+Of course, $\forall \cro_i \in \PHl_{\cro}, \ctx_{\{ \cro \}}(\PHetat{\cro_i}) = \{ \cro_i \}$,
+and from \pref{eq:irB} and \eqref{eq:fp}:
 \begin{align*}
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_3}) &= \{ \cro_2 \} && &&
-    \text{(from \pref{eq:B-bounce})} \\
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_2}) &= \emptyset &&
-    \text{because } \Phi(\cro_2) = \emptyset &&
-    \text{(from \pref{eq:B-no-bounce})} \\
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_1}) &= \emptyset &&
-    \text{because } \Phi(\cro_1) = \emptyset &&
-    \text{(from \pref{eq:B-no-bounce})} \\
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_0}) &= \{ \cro_0 \} &&
-    \text{because } \PHetat{\cI_2, \cro_0} \in \Phi(\cro_0) &&
-    \text{(from \pref{eq:B-no-bounce})}
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_3}) &= \{ \cro_2 \} && \\
+%     \text{(from \pref{eq:B-bounce})} \\
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_2}) &= \{ \cro_1 \} && \text{and} &
+    \Phi&(\cro_2) = \emptyset \\
+%     \text{(from \pref{eq:B-no-bounce})} \\
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_1}) &= \{ \cro_1 \} && \text{and} &
+    \Phi&(\cro_1) = \emptyset \\
+%     \text{(from \pref{eq:B-no-bounce})} \\
+  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_0}) &= \{ \cro_0 \} && \text{and} &
+    \PHetat{\cI_2, \cro_0} \in \Phi&(\cro_0)
+%     \text{(from \pref{eq:B-no-bounce})}
 \end{align*}
+We thus cannot infer any local influence from \pref{eq:edges-inference-auto}
+by observing the group of regulators $\{ \cro \}$,
+because $\Phi(\cro_1) = \Phi(\cro_2) = \emptyset$.
+This is due to the fact that all processes of $\cI$ can hit $\cro_1$ and $\cro_2$;
+these processes can therefore not be considered stable in any configuration.
 
-We thus cannot infer any local influence by observing the group of regulators $\{ \cro \}$.
 Let us now focus instead on the group of regulators $\{ \cI \}$.
@@
   \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}$.
-Furthermore:
+Furthermore, from \pref{eq:irB}, we especially find:
 \begin{align*}
@@
 %   \text{(from \pref{eq:B-bounce})} \\
-  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_2}) &= \{ \cro_3 \} &&
-    \text{(from \pref{eq:B-bounce})} \\
-  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3}) &= \{ \cro_2 \} &&
-    \text{(from \pref{eq:B-bounce})}
+  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_2}) &= \{ \cro_3 \} \\
+%     \text{(from \pref{eq:B-bounce})} \\
+  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3}) &= \{ \cro_2 \}
+%     \text{(from \pref{eq:B-bounce})}
 \end{align*}
-Therefore, \pref{eq:inference-edges-aa} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
+Therefore, \pref{eq:edges-inference} gives: $\cro \xrightarrow{3} \cro \in \hat{E}_-$.
 % The configurations using the other processes of $\cro$ ($\cro_0$, $\cro_1$ and $\cro_2$)
@@
   K_{\cI, \emptyset} &= \{2\} &
-  K_{\cI, \{\cro\}} &= \{0\} \\
+  K_{\N, \emptyset} &= \{1\} \\
+  K_{\cI, \{\cro\}} &= \{0\} &
+  K_{\N, \{\cro\}} &= \{0\} \\
   K_{\cI, \{\cII\}} &= \{2\} &
-  K_{\cI, \{\cII;cro\}} &= \{2\} \\
-  K_{\N, \emptyset} &= \{1\} &
-  K_{\N, \{\cro\}} &= \{0\} \\
-  K_{\N, \{\cI\}} &= \{0\} &
+  K_{\N, \{\cI\}} &= \{0\} \\
+  K_{\cI, \{\cII;cro\}} &= \{2\} &
   K_{\N, \{\cI;cro\}} &= \{0\} \\
+  \\
   K_{\cII, \emptyset} &= \{0\} &
-  K_{\cII, \{\cro\}} &= \{0\} \\
+  K_{\cro, \emptyset} &= \{3\} \\
+  K_{\cII, \{\cro\}} &= \{0\} &
+  K_{\cro, \{\cro\}} &= \{2\} \\
   K_{\cII, \{\cI\}} &= \{0\} &
-  K_{\cII, \{\cI;cro\}} &= \{0\} \\
-  K_{\cII, \{\N\}} &= \{1\} &
+  K_{\cro, \{\cI\}} &= \{0\} \\
+  K_{\cII, \{\cI;cro\}} &= \{0\} &
+  K_{\cro, \{\cI;cro\}} &= \{2\} \\
+  K_{\cII, \{\N\}} &= \{1\} \\
   K_{\cII, \{\N;cro\}} &= \{0\} \\
-  K_{\cII, \{\N;cI\}} &= \{0\} &
-  K_{\cII, \{\N;cI;cro\}} &= \{0\} \\
-  K_{\cro, \emptyset} &= \{3\} &
-  K_{\cro, \{\cro\}} &= \{2\} \\
-  K_{\cro, \{\cI\}} &= \{0\} &
-  K_{\cro, \{\cI;cro\}} &= \{2\}
+  K_{\cII, \{\N;cI\}} &= \{0\} \\
+  K_{\cII, \{\N;cI;cro\}} &= \{0\}
 \end{align*}

### 2014-10-30 parts-phage-lambda.tex
phage lambda++

@@
 
+\newcommand{\tcrocII}{cro\mbox{-}cII}
+\newcommand{\crocII}{\f{\tcrocII}}
+
 In order to illustrate the IG inference developed as \pref{sec:infer-IG},
@@
 with 4 components %($\Gamma = \{ \cI, \cro, \cII, \N \}$)
-and 3 cooperative sorts: %($\PHs \setminus \Gamma = \{ \cIcrocII, \cIcroN, \cIcro \}$).
-$$\PHs = \{ \underbrace{\cI, \cro, \cII, \N}_{\Gamma}, \cIcrocII, \cIcroN, \cIcro \}$$
+and 4 cooperative sorts: %($\PHs \setminus \Gamma = \{ \cIcrocII, \cIcroN, \cIcro \}$).
+\[\PHs = \{ \underbrace{\cI, \cro, \cII, \N}_{\Gamma},
+  \underbrace{\cIcrocII, \cIcroN, \cIcro, \crocII}_{\Delta} \}\]
 The four components have respectively 3, 4 2 and 2 expression levels,

### 2014-10-30 parts-phage-lambda.tex
phage lambda++++

@@
 \subsection{Detailed example: the bacteriophage lambda immunity control}
+\label{ssec:phage-lambda}
 }
-\label{ssec:phage-lambda}
 
@@
 but excerpts are given in \pref{fig:phage-lambda-cIcro-N} and \pref{fig:phage-lambda-cIcro-cro}.
+The whole set of actions is given in appendix.
 
@@
 \begin{align*}
-  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \cI_0, \cro_3, \cIcro_{FT} \} \\
-  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \cI_1, \cro_3, \cIcro_{TT} \} \\
-  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \cI_2, \cro_3, \cIcro_{TT} \}
+  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0})
+    &= \{ \cI_0, \cro_3, \N_0, \cIcro_{FT} \} \\
+  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0})
+    &= \{ \cI_1, \cro_3, \N_0, \cIcro_{TT} \} \\
+  \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0})
+    &= \{ \cI_2, \cro_3, \N_0, \cIcro_{TT} \}
 \end{align*}
@@
 \begin{align*}
-  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} \\
-%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
+  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0}) &= \{ \N_1 \} \\
+%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
 %     \text{(from \pref{eq:B-no-bounce})} \\
-  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} \\
-%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
+  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0}) &= \{ \N_0 \} \\
+%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
 %     \text{(from \pref{eq:B-bounce})} \\
-  \irB_\N^{\{ \cI, \cro \}}(\PHetat{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \}
-%     \text{because } \PHetat{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
+  \irB_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0}) &= \{ \N_0 \}
+%     \text{because } \state{\cI_0, \cro_3, \N_0} \in \Phi(\N_0) \\
 %     \text{(from \pref{eq:B-bounce})}
@@
 Let us first focus on the group of regulators $\{ \cro \}$.
-Of course, $\forall \cro_i \in \PHl_{\cro}, \ctx_{\{ \cro \}}(\PHetat{\cro_i}) = \{ \cro_i \}$,
+Of course:
+\[\forall \cro_i \in \PHl_{\cro}, \ctx_{\cro}^{\{ \cro \}}(\state{\cro_i}) = \{ \cro_i \}\]
 and from \pref{eq:irB} and \eqref{eq:fp}:
 \begin{align*}
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_3}) &= \{ \cro_2 \} && \\
+  \irB_{\cro}^{\{ \cro \}}(\state{\cro_3}) &= \{ \cro_2 \} && \\
 %     \text{(from \pref{eq:B-bounce})} \\
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_2}) &= \{ \cro_1 \} && \text{and} &
+  \irB_{\cro}^{\{ \cro \}}(\state{\cro_2}) &= \{ \cro_1 \} && \text{and} &
     \Phi&(\cro_2) = \emptyset \\
 %     \text{(from \pref{eq:B-no-bounce})} \\
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_1}) &= \{ \cro_1 \} && \text{and} &
+  \irB_{\cro}^{\{ \cro \}}(\state{\cro_1}) &= \{ \cro_1 \} && \text{and} &
     \Phi&(\cro_1) = \emptyset \\
 %     \text{(from \pref{eq:B-no-bounce})} \\
-  \irB_{\cro}^{\{ \cro \}}(\PHetat{\cro_0}) &= \{ \cro_0 \} && \text{and} &
-    \PHetat{\cI_2, \cro_0} \in \Phi&(\cro_0)
+  \irB_{\cro}^{\{ \cro \}}(\state{\cro_0}) &= \{ \cro_0 \} && \text{and} &
+    \state{\cI_2, \cro_0} \in \Phi&(\cro_0)
 %     \text{(from \pref{eq:B-no-bounce})}
@@
 One again, we trivially have:
-$\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
-  \ctx_{\{ \cI, \cro \}}(\PHetat{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}$.
+\[\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
+  \ctx_{\cro}^{\{ \cI \}}(\state{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}\]
 Furthermore, from \pref{eq:irB}, we especially find:
 \begin{align*}
-%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
+%   \irB_{\cro}^{\{ \cI, \cro \}}(\state{\cI_0, \cro_0}) &= \{ \cro_1 \} &&
 %   \text{(from \pref{eq:B-bounce})} \\
-%   \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
+%   \irB_{\cro}^{\{ \cI, \cro \}}(\state{\cI_0, \cro_1}) &= \{ \cro_2 \} &&
 %   \text{(from \pref{eq:B-bounce})} \\
-  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_2}) &= \{ \cro_3 \} \\
+  \irB_{\cro}^{\{ \cI \}}(\state{\cI_0, \cro_2}) &= \{ \cro_3 \} \\
 %     \text{(from \pref{eq:B-bounce})} \\
-  \irB_{\cro}^{\{ \cI, \cro \}}(\PHetat{\cI_0, \cro_3}) &= \{ \cro_2 \}
+  \irB_{\cro}^{\{ \cI \}}(\state{\cI_0, \cro_3}) &= \{ \cro_2 \}
 %     \text{(from \pref{eq:B-bounce})}
@@
 \begin{align*}
-  K_{\cI, \emptyset} &= \{2\} &
-  K_{\N, \emptyset} &= \{1\} \\
-  K_{\cI, \{\cro\}} &= \{0\} &
-  K_{\N, \{\cro\}} &= \{0\} \\
-  K_{\cI, \{\cII\}} &= \{2\} &
-  K_{\N, \{\cI\}} &= \{0\} \\
-  K_{\cI, \{\cII;cro\}} &= \{2\} &
-  K_{\N, \{\cI;cro\}} &= \{0\} \\
+  K_{\cI, \emptyset} &= 2 &
+  K_{\N, \emptyset} &= 1\\
+  K_{\cI, \{\cro\}} &= 0&
+  K_{\N, \{\cro\}} &= 0\\
+  K_{\cI, \{\cII\}} &= 2&
+  K_{\N, \{\cI\}} &= 0\\
+  K_{\cI, \{\cII;cro\}} &= 2&
+  K_{\N, \{\cI;cro\}} &= 0\\
   \\
-  K_{\cII, \emptyset} &= \{0\} &
-  K_{\cro, \emptyset} &= \{3\} \\
-  K_{\cII, \{\cro\}} &= \{0\} &
-  K_{\cro, \{\cro\}} &= \{2\} \\
-  K_{\cII, \{\cI\}} &= \{0\} &
-  K_{\cro, \{\cI\}} &= \{0\} \\
-  K_{\cII, \{\cI;cro\}} &= \{0\} &
-  K_{\cro, \{\cI;cro\}} &= \{2\} \\
-  K_{\cII, \{\N\}} &= \{1\} \\
-  K_{\cII, \{\N;cro\}} &= \{0\} \\
-  K_{\cII, \{\N;cI\}} &= \{0\} \\
-  K_{\cII, \{\N;cI;cro\}} &= \{0\}
+  K_{\cII, \emptyset} &= 0&
+  K_{\cro, \emptyset} &= 3\\
+  K_{\cII, \{\cro\}} &= 0&
+  K_{\cro, \{\cro\}} &= 2\\
+  K_{\cII, \{\cI\}} &= 0&
+  K_{\cro, \{\cI\}} &= 0\\
+  K_{\cII, \{\cI;cro\}} &= 0&
+  K_{\cro, \{\cI;cro\}} &= 2\\
+  K_{\cII, \{\N\}} &= 1\\
+  K_{\cII, \{\N;cro\}} &= 0\\
+  K_{\cII, \{\N;cI\}} &= 0\\
+  K_{\cII, \{\N;cI;cro\}} &= 0
 \end{align*}
@@
   taken from \cite{thieffry_dynamical_1995}.
-  \todo{Remove the curly brackets $\{\}$?}
 }

### 2014-10-30 parts-grn.tex
Minor typo fix

@@
 \modMM{In this paragraph, we} concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
-\cite{Richard06,BernotSemBRN}.
+\cite{thomas1990biological, Richard06,BernotSemBRN}.
 In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t.~the presented
@@
 However, distinguishing the IG and parametrization allows to decouple the graph (which is generally one of the prior knowledge from the biologists) from the constraints on its dynamics.
-It opened the way to numerous studies (e.g., necessary condition on sustained oscillations) solely based on the structure of the IG, e.g., the analysis of positive or negative circuits.}
+It opened the way to numerous studies (e.g., necessary condition on sustained oscillations) solely based on the structure of the IG (e.g., the analysis of positive or negative circuits).}
 

### 2014-10-30 parts-asp.tex
switch from interval parameters to integer parameters (minus the running example)

@@
   \item $\paramlabel(a, p)$ states that $p$ is a valid label for a set of resources of component $a$ (and therefore $K_{a,\omega_p}$ is a valid parameter);
-  \item $\param(a, p, i)$ states that: $i \in K_{a, \omega_p}$;
+  \item $\param(a, p, i)$ states that: $K_{a, \omega_p} = i$;
   \item $\inferedparam(a, p)$ states that the parameter inference of $K_{a, \omega_p}$ was conclusive (\pref{pps:param_K}).
@@
 Indeed, this rule applies to any possible parameter $\var{P}$ of any component $\var{A}$ ($\paramlabel$) whose value is still unknown ($\nota \inferedparam$),
-and states that any expression level $\var{I}$ of this component ($\componentlevels$) can belong to the value of the parameter ($\param$).
-Furthermore, the lower bound is $1$, which forces each enumerated parameter to contain at least one value,
-but no upper bound is specified ($\infty$) for the size of each parameter
-(which is in fact already bounded by the number of possible expression levels of the related component).
+and states that any expression level $\var{I}$ of this component ($\componentlevels$)
+is a candidate value for the parameter ($\param$).
+Furthermore, both the lower and upper bounds are $1$,
+which force each enumerated parameter to have exactly one value.
+%but no upper bound is specified ($\infty$) for the size of each parameter
+%(which is in fact already bounded by the number of possible expression levels of the related component).
 %each parameter contains at least $1$ value, as stated by the lower bound, and has no upper bound ($\infty$) but the number of expression levels of its component.
@@
 so that if $K_{a, \omega_p}$ could not be inferred by \pref{pps:param_K}, then
-$K_{a, \omega_p} \subset \segm{0}{l_a} \wedge K_{a, \omega_p} \neq \emptyset$
-(thus completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K} or the fact that parameters have to be intervals).
+$K_{a, \omega_p} \in \segm{0}{l_a}$
+(thus completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K}).
 
 \begin{example}\label{ex:cardinality}
+\todo{Refresh this example with integer parameters}
 In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred by \pref{pps:param_K}.
@@
 which thus allows to invalidate answer sets containing some unwanted combinations of atoms.
-In the scope of parameters enumeration, for example, constraints are especially useful to filter parametrizations
-that contain non-interval parameters, or that do not respect the assumptions of \pref{ssec:admissible-K}.
+In the scope of parameters enumeration, for example, constraints are especially useful
+to filter out parametrizations that do not respect the assumptions of \pref{ssec:admissible-K}.
 Indeed, suppose that:
@@
   \item $\lessactive(a, p, q)$ states that $\omega_p$ is a set of resources of $a$ with (loosely) less activators and more inhibitors than $\omega_q$;
-  \item $\paraminf(a, p, q)$ states that: $K_{a,\omega_p} \leqsegm K_{a,\omega_q}$.
+  \item $\paraminf(a, p, q)$ states that: $K_{a,\omega_p} \leq K_{a,\omega_q}$.
 \end{itemize}
@@
 such that $\var{A}$ is less activated by the set of resources $\omega_\var{P}$ than it is by $\omega_\var{Q}$,
-but $K_{\var{A},\omega_\var{Q}} \ltsegm K_{\var{A},\omega_\var{P}}$,
+but $K_{\var{A},\omega_\var{Q}} < K_{\var{A},\omega_\var{P}}$,
 thus violating the monotonicity assumption.
@@
 \begin{example}
+\todo{Refresh this example with integer parameters}
 The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.

### 2014-10-30 parts-grn.tex
switch from interval parameters to integer parameters (minus the running example)

--- 2014-10-25 parts-grn.tex
@@
 For a given component $a \in \Gamma$, and $\omega \subset \GRNreg{a}$ a set of regulators of $a$,
-the \emph{parameter} $K_{a,\omega} = \segm{i}{j}$ is a non-empty interval with with $0 \leq i \leq j \leq l_a$.
+the \emph{parameter} \modMF{$K_{a,\omega} \in \segm{0}{l_a}$ is a non-negative integer}.
 The complete map $K$ of parameters on $\IG$ is called a \emph{parametrization} on $\IG$.
@@
 An IG and a parametrization make up a complete BRN.
-The purpose of intervals as parameters is discussed at the end of this section.
-Regarding the dynamics, an interval parameter $K_{a,\omega}$ is the set of values towards which $a$ will tend
-in the states where its resources are exactly the regulators in $\omega$.
+%The purpose of intervals as parameters is discussed at the end of this section.
+%Regarding the dynamics, an interval parameter $K_{a,\omega}$ is the set of values towards which $a$ will tend
+%in the states where its resources are exactly the regulators in $\omega$.
+\modMF{Regarding the dynamics, a parameter $K_{a,\omega}$ is the value towards which $a$ will tend
+in the states where its resources are exactly the regulators in $\omega$.}
 %as defined in the asynchronous dynamics of \pref{def:dynamics}.
@@
 Indeed, from a given state $s$, a transition to another state $s'$ is possible provided that
-exactly one component $a$ evolves of one expression level towards the closest value of the interval $K_{a,\GRNres{a}{s}}$,
+exactly one component $a$ evolves of one expression level \modMF{towards} $K_{a,\GRNres{a}{s}}$,
 as stated by the definition of the transition relation $\GRNtrans{s}{s'}$ (\pref{def:dynamics}).
-However, $a$ cannot evolve if its expression level already belongs to the interval $K_{a,\GRNres{a}{s}}$.
+However, $a$ cannot evolve if its expression level already \modMF{equals}
+the parameter $K_{a,\GRNres{a}{s}}$.
 
@@
   \forall s, s' \in \GRNstates, \GRNtrans{s}{s'} &\Longleftrightarrow \exists a \in \Gamma,
-  \GRNget{s}{a} \notin K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + \delta^a(s) \\
+  \GRNget{s}{a} \neq K_{a, \GRNres{a}{s}} \wedge \GRNget{s'}{a} = \GRNget{s}{a} + \delta^a(s) \\
     & \qquad\quad \wedge \forall b \in \Gamma, b \neq a \Rightarrow \GRNget{s}{b} = \GRNget{s'}{b}
@@
 Due to the semantics of interval parameters and as $K_{b,\emptyset} = \segm{0}{1}$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
+\todo{Remove interval parameter}
+\todo{Change example to preserve the dynamics?}
 \end{example}
@@
 Let us define
-$f^a(s) = s[a]$ if $s[a]\in K_{a, \GRNres{a}{s}}$
+$f^a(s) = s[a]$ if $s[a] = K_{a, \GRNres{a}{s}}$
 and

### 2014-10-30 parts-infer-ig.tex
adapt notations of the phage lambda example
switch from interval parameters to integer parameters (minus the running example)

--- 2014-09-26 parts-infer-ig.tex
@@
     \exists \sigma \in L(g\cup\{a\}), \\
-    \qquad & \exists a_j \in \irB_a^g(\sigma\{a_t\}), \exists a_k \in \irB_a^g(\sigma\{a_{t+1}\}), \\
+  \qquad & \exists a_j \in \irB_a^{g \modMF{\cup \{a\}}}(\sigma\{a_t\}),
+    \exists a_k \in \irB_a^{g \modMF{\cup \{a\}}}(\sigma\{a_{t+1}\}), \\
   \qquad & 
@@
 \end{align}
-where $\irB_a^g(\sigma) \DEF \focals(a, \{a_i\}, \ctx_a^g(\sigma))$ with
-$a_i\in\sigma$,
-and where $\fp(a_i) \subseteq  L(\PHpredecgene{a}\cup\{a\})$ is defined as follows:
+\modMF{where $\irB_a^g(\sigma) \subset \PHl_a$ is defined as follows:}
+\begin{align}
+\bar\irB_a^g(\sigma) &\DEF \{ a_j \mid \PHfrappe{b_k}{a_i}{a_j}\in\PHa, b\in g \wedge b_k\in\ctx_a^g(\sigma)
+\wedge a_i\in\sigma \}
+\\
+\irB_a^g(\sigma) &\DEF
+\begin{cases}
+\bar\irB_a^g(\sigma) & \text{if }\bar\irB_a^g(\sigma)\neq\emptyset \\
+\{\PHget{\sigma}{a}\} & \text{otherwise}
+\label{eq:irB}
+\end{cases}
+\end{align}
+and $\fp(a_i) \subseteq  L(\PHpredecgene{a}\cup\{a\})$ is defined as follows:
 \begin{equation}
@@
 introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
+\todo{Refresh this example for the new inference}

### 2014-10-30 parts-infer-param.tex
switch from interval parameters to integer parameters (minus the running example)

@@
 This relaxed notion of parameterization compatibility allows to enumerate all parametrizations
-compatible with a given PH, i.e. all Thomas model whose dynamics is included in the PH dynamics.}
+compatible with a given PH, \ie all Thomas model whose dynamics is included in the PH dynamics.}
 
@@
 
-\todo{update notations}
 \begin{property}[Parameter validity]\label{pro:K-valid}
-A parameter $K_{a,\omega}$ is valid w.r.t. the PH if and only if the following equation is verified:
+A parameter $K_{a,\omega}$ is valid w.r.t.~the PH if and only if the following equation is verified:
 \begin{align*}
-  \forall a_i\in C^a_{a,\omega}, a_i \notin K_{a,\omega} \Longrightarrow
+  \forall a_i \in C^a_{a,\omega}, a_i \neq K_{a,\omega} \Longrightarrow
     (& \exists \PHfrappe{c_k}{a_i}{a_j}\in\PHa, c_k \in C^c_{a,\omega} \\
@@
 \begin{align*}
-  \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Longrightarrow \exists \omega \subset \GRNreg{b}, 0 \in K_{b,\omega} \wedge \exists \omega' \subset \GRNreg{b}, l_b \in K_{b,\omega'}
+  \forall b \in \Gamma, \GRNreg{b} \neq \emptyset \Longrightarrow
+    \exists \omega \subset \GRNreg{b}, K_{b,\omega} = 0 \wedge
+    \exists \omega' \subset \GRNreg{b}, K_{b,\omega'} = l_b
 \end{align*}
@@
   \forall A^- \subset \{ a \in \Gamma \mid \GRNedge{a}{-}{t}{b} \in E_- \},\\
-  K_{b,\omega \cup A^-} & \leqsegm K_{b,\omega \cup A^+}
+  K_{b,\omega \cup A^-} & \leq K_{b,\omega \cup A^+}
 \end{align*}
@@
 to find the behavior of the model refined with a cooperative sort described in \pref{fig:runningPH-2}.
+\todo{Refresh this example with integer parameters}
 \end{example}

### 2014-10-30 parts-grn.tex
begin switching from interval parameters to integer parameters in the running example

@@
 \begin{align*}
-  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}\} &
-  E_\uns &= \emptyset \\
-  E_- &= \{\GRNedgef{a}{-}{2}{b}\}
+  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a},
+    \GRNedgef{b}{+}{1}{b}, \GRNedgef{c}{+}{1}{c}\} \\
+  E_- &= \{\GRNedgef{a}{-}{2}{b}\} &&
+    E_\uns = \emptyset \\
 \end{align*}
@@
   \GRNreg{a} &= \{ b, c \} &
-  \GRNreg{b} &= \{ a \} \\
-  \GRNreg{c} &= \emptyset
+  \GRNreg{b} &= \{ a, b \} \\
+  \GRNreg{c} &= \{ c \}
 \end{align*}
@@
 \begin{tikzpicture}[grn]
-  \path[use as bounding box] (-0.3,-0.75) rectangle (2.5,1.5);
+  \path[use as bounding box] (-0.5,-0.75) rectangle (3.2,1.7);
   \node[inner sep=0] (a) at (2,0) {a};
@@
     (c) edge node[elabel, right=-2pt] {$+1$} (a)
-    (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b);
+    (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b)
+    (b) edge[loop, min distance=30pt] node[elabel, above=-5pt] {$+1$} (b)
+    (c) edge[loop right, min distance=30pt] node[elabel, above=0pt, xshift=-10] {$+1$} (c);
 \end{tikzpicture}
@@
 \centering
+\modMF{
 \begin{align*}
-  K_{a,\emptyset} &= \segm{0}{0} & K_{b,\emptyset} &= \segm{0}{1} \\
-  K_{a,\{b\}} &= \segm{1}{1} & K_{b,\{a\}} &= \segm{0}{0} \\
-  K_{a,\{c\}} &= \segm{1}{1} &&\\
-  K_{a,\{b,c\}} &= \segm{2}{2} & K_{c,\emptyset} &= \segm{0}{1}
+  K_{a,\emptyset} &= 0 & K_{b,\emptyset} &= 0 \\
+  K_{a,\{b\}} &= 1 & K_{b,\{a\}} &= 0 \\
+  K_{a,\{c\}} &= 1 & K_{b,\{b\}} &= 1 \\
+  K_{a,\{b,c\}} &= 2  & K_{b,\{a,b\}} &= 0 \\
+  \\
+  K_{c,\emptyset} &= 0 & K_{c,\{c\}} &= 1
 \end{align*}
+}
 \end{minipage}
@@
 This sequence of states ends in a steady state: no evolution is possible in $\GRNstate{a_1, b_0, c_1}$.
-Due to the semantics of interval parameters and as $K_{b,\emptyset} = \segm{0}{1}$, no auto-regulation on $b$ is needed to prevent its evolution when $a$ is not at level $2$.
-\todo{Remove interval parameter}
-\todo{Change example to preserve the dynamics?}
 \end{example}

### 2014-10-30 parts-infer-ig.tex
begin switching from interval parameters to integer parameters in the running example

@@
 
+\modMF{
 The IG inference of the PH of \pref{fig:runningPH-2} gives the
-IG in \pref{fig:BRN-inf1}, containing the following edges:
+IG in \pref{fig:runningBRN}(left), that is, allows to find back all the edges
+and their signs and thresholds.
+This is mainly due to the fact that the cooperation between $b$ and $c$ is well-defined,
+using a (deterministic) cooperative sort.
+}
+
+\modMF{
+Indeed, the inference of the PH of \pref{fig:runningPH-1},
+that is, without refinement with a cooperative sort, gives the following IG:
 \begin{align*}
-  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{b}{+}{1}{b}, \GRNedgef{c}{+}{1}{c}\}\\
+  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{c}{+}{1}{c}\}\\
   E_- &= \{\GRNedgef{a}{-}{2}{b}\} \qquad\qquad\qquad\qquad\qquad
@@
 \end{align*}
-This IG is close to the one in \pref{fig:runningBRN} but not equivalent,
-as each component has an additional auto-action.
-The auto-actions on $b$ and $c$ are the consequence of a global stability
-in some configurations: $c$ never evolves, and neither does $b$ when $a_2$ is not active.
-The auto-action on $a$ is mainly caused by its multi-valued nature.
+This IG is close to the one in \pref{fig:runningBRN} but misses the self-influence
+$\GRNedgef{b}{+}{1}{b}$.
+\todo{Detail why?}
+}
 
-The inference of the PH of \pref{fig:runningPH-1}
-(without refinement with cooperative sort) gives the same IG.
-
-\begin{figure}[t]
-\centering
-\scalebox{1.2}{
-\begin{tikzpicture}[grn]
-  \path[use as bounding box] (-1.3,-0.75) rectangle (3.5,1.5);
-  \node[inner sep=0] (a) at (2,0) {a};
-  \node[inner sep=0] (b) at (0,0) {b};
-  \node[inner sep=0] (c) at (2,1.2) {c};
-  \path[->]
-    (b) edge[bend right] node[elabel, below=-2pt] {$+1$} (a)
-    (c) edge node[elabel, right=-2pt] {$+1$} (a)
-    (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b)
-    (b) edge[in=-15+180, out=15+180, loop] node[elabel, left=-2pt] {+1} (b)
-    (c) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (c)
-    (a) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (a);
-\end{tikzpicture}
-}
-\caption{\label{fig:BRN-inf1}
-  Result of the IG inference performed on the PH of \pref{fig:runningPH-2}.
-}
-\end{figure}
-
+\modMF{
 If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
@@
 introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
-\todo{Refresh this example for the new inference}
+}

### 2014-10-30 parts-infer-param.tex
begin switching from interval parameters to integer parameters in the running example

@@
 \label{ex:infer-param-runningPH-1}
+When applied to the refined PH in \pref{fig:runningPH-2},
+and with the IG of \pref{fig:runningBRN}(left)
+(which is the result of the IG inference on this PH model,
+as explained in \pref{ssec:infer-ig-examples}),
+\label{pps:param_K} infers the parametrization of \pref{fig:runningBRN}(right).
+In other words, the inference of parameters allows to find back all parameters
+of the original model when the cooperations are fully defined.
+\end{example}
 
-Applied to the PH in \pref{fig:runningPH-1}, we infer the following parameters:
-\begin{align*}
-K_{a, \emptyset} &= 0&
-K_{b, \emptyset} &= 0 \\
-K_{a, \{a\}} &= 0 &
-K_{b, \{a\}} &= 0 \\
-K_{a, \{c\}} &= 1 &
-K_{b, \{b\}} &= 1 \\
-K_{a, \{b\}} &= 1 &
-K_{b, \{a,b\}} &= 0 \\
-K_{a, \{b,c\}} &= 1 &
-K_{c, \emptyset} &= 0 \\
-K_{a, \{a,b,c\}} &= 2 &
-K_{c, \{c\}} &= 1
-%K_{a, \{a,b\}} = ?
-%K_{a, \{a,c\}} = ?
-\end{align*}
 $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ cannot be inferred,
 which is a direct consequence of the lack of cooperation between $b$ and $c$ on $a$.
-\end{example}
 
 \begin{example}
-Regarding the refined PH of \pref{fig:runningPH-2}, all parameters can be inferred.
-We obtain the same value for the inferred parameters of \pref{ex:infer-param-runningPH-1},
-together with the following results:
+Regarding the non-refined PH of \pref{fig:runningPH-1},
+with the IG produced by the IG inference described in \pref{ssec:infer-ig-examples}
+(that is, the same as \pref{fig:runningBRN}(left)
+without the self-influence $\GRNedgef{b}{+}{1}{b}$)
+the following parameters are inferred:
 \begin{align*}
-  K_{a,\{a,b\}} &= 1 &
-  K_{a,\{a,c\}} &= 1
+  K_{a,\emptyset} &= 0 &
+  K_{b,\{a\}} &= 0 \\
+%  K_{a,\{b\}} &= 0 \\
+%  K_{a,\{c\}} &= 0 &
+  K_{a,\{b,c\}} &= 2 \\ \\
+%  K_{b,\emptyset} &= 0
+  K_{c,\emptyset} &= 0 &
+  K_{c,\{c\}} &= 1
 \end{align*}
+The parameters $K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$
+could not be inferred,
+mainly due to the lack of precision in the cooperation between $b$ and $c$.
+Note that the parameters $K_{b,\{b\}}$ and $K_{b,\{a,b\}}$ do not exist for this IG,
+because of the absence of any self-influence $b \rightarrow b$.
 \end{example}

### 2014-10-30 parts-tr2global.tex
section 4.1
fix Def 11

@@
 sections assume that \emph{all the cooperative sorts are deterministic}.
-\todo{add more explanation depending on the text of the next sections}
 
@@
 \label{def:ctx-sigma}
-Given $a\in\PHs$, a group $g\in X(a)$ and a configuration 
+Given $a\in\PHs$, a sub-set of regulators $g\subset \PHpredecgene{a}$ and a configuration 
 $\sigma\in L(g)$, $\ctx_a^g(\sigma)$ is the configuration $\sigma$ with the corresponding
@@
   \upsilon'\in\PHdirectpredec{\upsilon} \cap \CSorts })
-  & \text{otherwise.}
+  & \text{otherwise,}
 \end{cases}
@@
 \end{align}
+where $\f{dom}(\sigma')$ is the set of sorts defined in $\sigma'$.
 \end{definition}
 }
+
+\begin{example}
+\todo{}
+\end{example}

### 2014-10-30 parts-tr2global.tex
fix Def 11

@@
 \label{def:ctx-sigma}
-Given $a\in\PHs$, a sub-set of regulators $g\subset \PHpredecgene{a}$ and a configuration 
+Given $a\in\PHs$, a sub-set of regulators $g\subseteq \PHpredecgene{a}$ and a configuration 
 $\sigma\in L(g)$, $\ctx_a^g(\sigma)$ is the configuration $\sigma$ with the corresponding

### 2014-10-30 parts-grn.tex
cleanup bib, make Fig 1 more visible, reference to K equivalence
K-equivalence

@@
 
+\begin{remark}[Parameters equivalence]
+\label{rem:K-equiv}
+\modLP{%
+In the scope of the asychronous dynamics (\pref{def:dynamics}), one can remark
+that some parameters values can be equivalent, i.e. the resulting dynamics is
+the same.
+It is notably the case with auto-regulations:
+if $\GRNedge{a}{s}{t}{a}$,
+then for any $\omega\subseteq \GRNreg{a}$,
+two values $v$ and $v'$ of the parameter 
+$K_{a,\omega\setminus \{a\}}$ are equivalent
+if $v\geq t$ and $v'\geq t$.
+Similary,
+two values $v$ and $v'$ of the parameter
+$K_{a,\omega\cup\{a\}}$ are equivalent 
+if $v<t$ and $v'<t$.}
+\end{remark}
+
 \paragraph{Interaction Graph of the dynamics}
@@
 (e.g., \cite{Richard2010378,PR11-SASB}).
-
-\todo{(LP) parameters equivalence:
-if $\GRNedge{a}{+}{t}{a}$,
-$K_{a,R\setminus \{a\}} = t
-\equiv
-K_{a,R\setminus \{a\}} = u$
-for $t \leq u \leq l_a$;
-$K_{a,R\cup \{a\}} = t-1
-\equiv
-K_{a,R\cup \{a\}} = u$
-for $0 \leq u < t$.
-If $\GRNedge{a}{-}{t}{a}$,
-$K_{a,R\cup \{a\}} = t
-\equiv
-K_{a,R\cup \{a\}} = u$
-for $t \leq u \leq l_a$;
-$K_{a,R\setminus \{a\}} = t-1
-\equiv
-K_{a,R\cup \{a\}} = u$
-for $0 \setminus u < t$.
-- any other equivalence?}

### 2014-10-30 parts-infer-ig.tex
ref IG

@@
 parameters of the dynamics.
-Following \todo{ref Adrien, autre}, the IG of a dynamical model is consistent
+As detailed in \cite{Richard06}, the IG of a dynamical model is consistent
 with a prior IG if it is a sub-graph of the prior IG.
@@
 necessary to encode the desired dynamics, indicating some non-minimality of the
-prior IG \todo{more refs}.}
+prior IG.}
 

### 2014-10-31 parts-infer-ig.tex
unsigned -> non-monotonous

@@
 We infer a positive (resp. negative) edge if there only exist corresponding influences with the same sign.
-If an influence is both positive and negative, we infer an unsigned edge.
+If an influence is both positive and negative, we infer an
+\modLP{non-monotonous} edge.
 In the end, the threshold of each edge is the minimum threshold for which an influence has been found.
@@
 If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
-then two unsigned edges towards $b$ are inferred instead of the previous signed edges:
+then two \modLP{non-monotonous} edges towards $b$ are inferred instead of the previous signed edges:
 \begin{align*}

### 2014-10-31 parts-intro.tex
todo--
unsigned -> non-monotonous

@@
 In addition to the improvement of the efficiency and accuracy of the IG inference, we have added support for
-unsigned regulations, that are regulations being positive or negative depending on a particular
+\modLP{non-monotonous} regulations, that are regulations being positive or negative depending on a particular
 context.
@@
 $\segm{i}{j}$ is the set of integers $\{ i, i+1, \dots, j \}$.
-Given an integer $k$, we note: $k < \segm{i}{j} \EQDEF k < i$ and $k > \segm{i}{j} \EQDEF k > j$.
-We also define:
-\begin{align*}
-  \segm{i_1}{j_1} \leqsegm \segm{i_2}{j_2} &\EQDEF (i_1\leq i_2\wedge j_1\leq j_2) \enspace\text{, and:}\\
-  \segm{i_1}{j_1} \ltsegm \segm{i_2}{j_2} &\EQDEF (i_1<i_2 \wedge j_1\leq j_2) \vee (i_1\leq i_2\wedge j_1 < j_2) \enspace.
-\end{align*}

### 2014-11-03 parts-asp.tex
end switching from interval parameters to integer parameters

@@
 \begin{example}\label{ex:asp-atom}
-Consider a PH model such as \pref{ex:infer-param-runningPH-1}:
+Consider a PH model such as depicted in \pref{fig:runningPH-1}:
 in order to state the existence of each component, we use a predicate called $\component$ with two arguments:
@@
 \begin{example}\label{ex:asp-component}
-In order to define the 3 components of \pref{ex:infer-param-runningPH-1}, we use the following program:
+In order to define the 3 components of \pref{fig:runningPH-1}, we use the following program:
 \begin{align*}
@@
 \begin{example}
-Regarding \pref{ex:infer-param-runningPH-1}, the previous rule together with the facts of \pref{ex:asp-component}
+Regarding \pref{fig:runningPH-1}, the previous rule together with the facts of \pref{ex:asp-component}
 will give the following answer set:
@@
 \begin{example}\label{ex:cardinality}
-\todo{Refresh this example with integer parameters}
-In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred by \pref{pps:param_K}.
-The previous cardinality rule allows to produce 49 parametrizations, in which these two parameters can take all possible values:
+\modMF{%
+In the scope of \pref{ex:infer-param-runningPH-1},
+$K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$
+could not be inferred by \pref{pps:param_K}.
+The previous cardinality rule allows to produce 18 parametrizations,
+in which these three parameters can take all possible values:
 %The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
@@
 \begin{align*}
-  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2}, \{ 0, 2 \} \}^2
+  (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
+    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \} \times \{ 0, 1 \}
+%   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2}, \{ 0, 2 \} \}^2
 \end{align*}
 and all the other parameters keep their inferred values.
+Note that $0$ belongs to the set of candidate values for parameter $K_{b,\emptyset}$
+because the extreme values assumption (\pref{pro:param_enum_extreme})
+has not been taken into account yet.
 %These parametrizations take the form of sets of $\param$ atoms.
-Note that $\{ 0, 2 \}$ belongs to the set of candidate parametrizations
-as no rule specifying that a parameter has to be an interval has been defined yet.
+% Note that $\{ 0, 2 \}$ belongs to the set of candidate parametrizations
+% as no rule specifying that a parameter has to be an interval has been defined yet.
+}%
+\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
 \end{example}
@@
 \begin{example}
-\todo{Refresh this example with integer parameters}
+\modMF{%
 The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
-For example, a constraint was written in order to obtain only interval parameters
-(thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
-Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
+% For example, a constraint was written in order to obtain only interval parameters
+% (thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
+% Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
+% \begin{align*}
+%   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
+% \end{align*}
+% }%
+% 
+% \modMF{%
+For example,
+% , due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
+the extreme values assumption (\pref{pro:param_enum_extreme}) removes all parametrizations with
+$K_{b,\emptyset} = 0$,
+thus reducing the set of candidate values to 9 possible values:
 \begin{align*}
-  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
+  (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
+    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \} \times \{ 1 \}
 \end{align*}
-
-Then, due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
-the monotonicity assumption (\pref{pro:param_enum_monotonicity} and expressed in ASP above) removes all parametrizations with:
-$0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
-In the end, the remaining possible values for the two parameters that were not inferred are:
-\begin{align*}
-  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{1}{1} , \segm{2}{2} , \segm{1}{2} \}^2
-\end{align*}
-All these candidates respect the other properties of \pref{ssec:admissible-K} and no more candidates are filtered.
-We thus obtain the results of \pref{ex:enum-param-runningPH-1}.
+Furthermore, parameter $K_{b,\emptyset} = 1$ doest not respect \pref{pro:K-valid},
+because no action in the model of \pref{fig:runningPH-1} allows to bounce from $b_0$ to $b_1$.
+Therefore, this parameter is removed from all returned parametrizations.
+In the end, 9 partial parametrizations are therefore produced.
+}%
+\todo{Find an example to apply the monotonicity assumption?}
+\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
 \end{example}

### 2014-11-03 parts-example.tex
end switching from interval parameters to integer parameters

@@
 
-The inference methods described in this paper have been implemented as part of
+The inference \modMF{and enumeration} methods described in this paper have been implemented as part of
 \textsc{Pint}\footnote{Available at \url{http://loicpauleve.name/pint}}, which gathers PH related tools.
@@
 The (possibly partial) inferred parametrization will be returned on the standard output.
-The admissible parametrizations enumeration is performed when adding the \texttt{-{}-full-enumerate} parameter to the command,
-or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
-\todo{Remove \texttt{full-enumerate}?}
+\modMF{%
+Instead of the \texttt{-{}-dot ig.dot} option, it is possible to specify
+an alternative IG to perform the parameters inference on
+(instead of using the inferred IG) with the \texttt{-{}-ig model.ig} option,
+where \texttt{model.ig} is an IG model file in \textsc{Pint} format.
+The admissible parametrizations enumeration is performed by adding the \texttt{-{}-enumerate}
+option to the command.
+%or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
+}%
 
@@
 
-\subsection{Biological application}\label{ssec:appli}
+\subsection{Biological application: the epithelial growth factor receptor}\label{ssec:appli}
 

### 2014-11-03 parts-grn.tex

@@
 \cite{thomas1990biological, Richard06,BernotSemBRN}.
-In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t.~the presented
-inference),
-we propose the notion of unsigned edge modeling an interaction whose nature (activation or inhibition) is undetermined,
-and we extend the classical parametrization formalism by setting parameters to intervals of values instead of single values.
-We briefly discuss these additions.
+%In order to enlarge the class of Thomas' models compatible with PH dynamics (w.r.t.~the presented
+%inference),
+%we propose the notion of unsigned edge modeling an interaction whose nature (activation or inhibition) is undetermined,
+%and we extend the classical parametrization formalism by setting parameters to intervals of values instead of single values.
+%We briefly discuss these additions.
 
@@
 The maximum expression level of a component $a$ is denoted $l_a$.
-The sign of an edge denotes the kind of regulation it models: it can be positive ($+$), negative ($-$) or unsigned ($\uns$),
-the latter meaning that the nature of the regulation cannot be reduced to a simple activation or inhibition.
+The sign of an edge denotes the kind of regulation it models: it can be positive
+($+$), negative ($-$) or \modLP{non-monotonous} ($\uns$),
+the latter meaning that \modLP{the regulation have different sign depending on the context \cite{Sene13-TCS}}.
 Regarding the dynamics, an edge $\GRNedge{a}{s}{t}{b}$ states that $a$ influences the evolution of $b$ in a certain way when its expression level is above or equal to the threshold $t$;
@@
 %otherwise, the opposite influence is expressed.
-The uniqueness of each regulation is forced, in order to make the following sections simpler.
-The purpose of unsigned edges is discussed at the end of the current section.
+%The uniqueness of each regulation is forced, in order to make the following sections simpler.
+%The purpose of unsigned edges is discussed at the end of the current section.
 

### 2014-11-03 parts-infer-param.tex
fix Def 11

@@
 
+\modMF{
 \begin{example}
-\label{ex:infer-param-runningPH-1}
+\label{ex:infer-param-runningPH-2}
 When applied to the refined PH in \pref{fig:runningPH-2},
@@
 as explained in \pref{ssec:infer-ig-examples}),
-\label{pps:param_K} infers the parametrization of \pref{fig:runningBRN}(right).
+\pref{pps:param_K} infers the parametrization of \pref{fig:runningBRN}(right).
 In other words, the inference of parameters allows to find back all parameters
@@
 \end{example}
+}
 
-$K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ cannot be inferred,
-which is a direct consequence of the lack of cooperation between $b$ and $c$ on $a$.
-
+\modMF{
 \begin{example}
+\label{ex:infer-param-runningPH-1}
 Regarding the non-refined PH of \pref{fig:runningPH-1},
@@
 \end{example}
+}
 
@@
 one transition in the PH, proving the correctness of our inference.
-We remark that any parameter inferred by \pref{pps:param_K} satisfies this property.
+\modMF{%
+Any parameter that does not satisfy \pref{pro:K-valid} is therefore excluded from the enumeration.
+}%
+We remark that all parameters inferred by \pref{pps:param_K} satisfy this property.
 
@@
 \begin{example}\label{ex:enum-param-runningPH-1}
-The parametrization inferred in \pref{ex:infer-param-runningPH-1} was partial because $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred.
-It is however possible to enumerate all complete and admissible parametrizations
+The parametrization inferred in \pref{ex:infer-param-runningPH-1} was partial
+because \modMF{$K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$} could not be inferred.
+It is however possible to enumerate all \modMF{partial} and admissible parametrizations
 compatible with both the inferred parameters, and the properties of this subsection.
-This enumeration gives 9 different parametrizations which correspond to the 3 possible values
-for each of the two parameters that could not be inferred:
+This enumeration gives 9 different \modMF{partial} parametrizations
+which correspond to the 3 possible values
+\modMF{%
+for both $K_{a,\{b\}}$ and $K_{a,\{c\}}$:
 \begin{align*}
-  K_{a,\{a,b\}} &\in \{ \segm{1}{1}, \segm{1}{2}, \segm{2}{2} \} \\
-  K_{a,\{a,c\}} &\in \{ \segm{1}{1}, \segm{1}{2}, \segm{2}{2} \}
+  K_{a,\{b\}} &\in \{ 0, 1, 2 \} \\
+  K_{a,\{c\}} &\in \{ 0, 1, 2 \}
 \end{align*}
-We note that for all solutions, $0 \notin K_{a,\{a,b\}} \wedge 0 \notin K_{a,\{a,c\}}$.
-This is due to the monotonicity assumption (\pref{pro:param_enum_monotonicity}) which especially states that:
-\begin{align*}
-  K_{a,\{b\}} \leqsegm K_{a,\{a,b\}} \wedge
-  K_{a,\{c\}} \leqsegm K_{a,\{a,c\}}
-\end{align*}
+$K_{b,\emptyset}$ is excluded
+from all return parametrizations because no value is compatible with the previous properties.
+}%
+% We note that for all solutions, $0 \notin K_{a,\{a,b\}} \wedge 0 \notin K_{a,\{a,c\}}$.
+% This is due to the monotonicity assumption (\pref{pro:param_enum_monotonicity}) which especially states that:
+% \begin{align*}
+%   K_{a,\{b\}} \leqsegm K_{a,\{a,b\}} \wedge
+%   K_{a,\{c\}} \leqsegm K_{a,\{a,c\}}
+% \end{align*}
+\todo{Find an example to apply the assumptions (properties 6--8)?}
 
-Finally, we note that $\segm{1}{1}$ belongs to the possible values for both parameters.
+Finally, we note that \modMF{$1$} belongs to the possible values for both parameters.
 Therefore this enumeration allows, from the model in \pref{fig:runningPH-1},
-to find the behavior of the model refined with a cooperative sort described in \pref{fig:runningPH-2}.
-\todo{Refresh this example with integer parameters}
+\modMF{%
+to find the behavior of component $a$ in the model refined with a cooperative sort described in
+\pref{fig:runningPH-2}.
+}%
+% to find the behavior of the model refined with a cooperative sort described in \pref{fig:runningPH-2}.
+\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
 \end{example}

### 2014-11-03 parts-ph.tex
intro of sect. 3 + section 3.1
sect 3.2
tr2global++

@@
 If $a_i\in \PHl_a$, we define the notation $a_i\in s \EQDEF \PHget{s}{a}=a_i$.
+\modLP{%
+Given a set of sorts $g\subset\PHs$,
+$L(g)$ denotes the set of all the sub-states of the sorts in $g$:
+$L(g) \DEF \prod_{b\in g} L_b$.}
+
+\modLP{
+The set of sorts having an action on a given sort $a$ is noted $\PHdirectpredec{a}$
+(\pref{eq:ph_direct_predec}).
+}
+\begin{equation}
+\forall a \in \PHs, \PHdirectpredec{a} \DEF \{b \in \PHs \mid \exists \PHfrappe{b_i}{a_j}{a_k}\in\PHa \}
+\label{eq:ph_direct_predec}
+\end{equation}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[tb]
 \centering
-\scalebox{1.3}{
+\scalebox{1.5}{
 \begin{tikzpicture}
@@
 
-\path[bounce, bend left=60]
+\path[bounce, bend left=30]
 \TBounce{a_1}{very thick}{a_2}{.south}
@@
 ;
-\path[bounce, bend right=60]
+\path[bounce, bend right=40]
 \TBounce{a_2}{}{a_1}{.north}
@@
 
-\path[bounce, bend right=60]
+\path[bounce, bend right=30]
 \TBounce{a_1}{very thick}{a_2}{.south east}
@@
 ;
-\path[bounce, bend left=60]
+\path[bounce, bend left=40]
 \TBounce{a_2}{}{a_1}{.north}
-\TBounce{a_1}{}{a_0}{.north east}
+\TBounce{a_1}{}{a_0}{.north}
 ;
@@
 \caption{\label{fig:runningPH-1}
-\todo{make more visible}
 A Process Hitting (PH) example.

### 2014-11-03 parts-phage-lambda.tex

@@
     of the parametrization, but this does not modify the dynamics, as mentioned previously;
-  \item Our inference returns $K_{\cro, \{\cI;cro\}} = 2$ instead of originally $0$;
-    once again, this does not change the dynamics, and $1$ would also have been
-    an acceptable value.
+  \item Our inference returns $K_{\cro, \{\cI;cro\}} = 2$, \modLP{which is a value
+  equivalent to the original $0$ value according to \pref{rem:K-equiv}:
+  	the resulting dynamics is the same, and $1$ would also have been an
+	acceptable value}.
 \end{itemize}

### 2014-11-03 parts-phage-lambda.tex
examples++

--- 2014-10-30 parts-phage-lambda.tex
@@
   \TSetSortLabel{cI}{$\cI$}
-  \TSort{(0,3.5)}{cI}{3}{l}
+  \TSort{(1,3.5)}{cI}{3}{l}
   \TSetSortLabel{cro}{$\cro$}
-  \TSort{(0,-1.5)}{cro}{4}{l}
+  \TSort{(1,-1.5)}{cro}{4}{l}
   \TSetSortLabel{N}{$\N$}
@@
   % Actions de MàJ
-  \THit{cI_1}{}{cIcro_0}{.west}{cIcro_2}
-  \THit{cI_1}{}{cIcro_1}{.160}{cIcro_3}
-  \THit{cI_2}{}{cIcro_0}{.west}{cIcro_2}
-  \THit{cI_2}{}{cIcro_1}{.160}{cIcro_3}
-  \path[bounce,bend left]
-    \TBounce{cIcro_0}{}{cIcro_2}{.240}
-    \TBounce{cIcro_1}{}{cIcro_3}{.south west} ;
-  \THit{cI_0}{}{cIcro_2}{.160}{cIcro_0}
-  \THit{cI_0}{}{cIcro_3}{.150}{cIcro_1}
-  \path[bounce,bend right]
-    \TBounce{cIcro_2}{}{cIcro_0}{.north west}
-    \TBounce{cIcro_3}{}{cIcro_1}{.120} ;
-  \THit{cro_2}{}{cIcro_0}{.210}{cIcro_1}
-  \THit{cro_2}{}{cIcro_2}{.200}{cIcro_3}
-  \THit{cro_3}{}{cIcro_0}{.210}{cIcro_1}
-  \THit{cro_3}{}{cIcro_2}{.200}{cIcro_3}
-  \path[bounce,bend left]
-    \TBounce{cIcro_0}{}{cIcro_1}{.240}
-    \TBounce{cIcro_2}{}{cIcro_3}{.south} ;
-  \THit{cro_0}{}{cIcro_3}{.190}{cIcro_2}
-  \THit{cro_0}{}{cIcro_1}{.200}{cIcro_0}
-  \THit{cro_1}{}{cIcro_3}{.190}{cIcro_2}
-  \THit{cro_1}{}{cIcro_1}{.200}{cIcro_0}
-  \path[bounce,bend right]
-    \TBounce{cIcro_1}{}{cIcro_0}{.north}
-    \TBounce{cIcro_3}{}{cIcro_2}{.120} ;
+%   \THit{cI_1}{}{cIcro_0}{.west}{cIcro_2}
+%   \THit{cI_1}{}{cIcro_1}{.160}{cIcro_3}
+%   \THit{cI_2}{}{cIcro_0}{.west}{cIcro_2}
+%   \THit{cI_2}{}{cIcro_1}{.160}{cIcro_3}
+%   \path[bounce,bend left]
+%     \TBounce{cIcro_0}{}{cIcro_2}{.240}
+%     \TBounce{cIcro_1}{}{cIcro_3}{.south west} ;
+%   \THit{cI_0}{}{cIcro_2}{.160}{cIcro_0}
+%   \THit{cI_0}{}{cIcro_3}{.150}{cIcro_1}
+%   \path[bounce,bend right]
+%     \TBounce{cIcro_2}{}{cIcro_0}{.north west}
+%     \TBounce{cIcro_3}{}{cIcro_1}{.120} ;
+%   \THit{cro_2}{}{cIcro_0}{.210}{cIcro_1}
+%   \THit{cro_2}{}{cIcro_2}{.200}{cIcro_3}
+%   \THit{cro_3}{}{cIcro_0}{.210}{cIcro_1}
+%   \THit{cro_3}{}{cIcro_2}{.200}{cIcro_3}
+%   \path[bounce,bend left]
+%     \TBounce{cIcro_0}{}{cIcro_1}{.240}
+%     \TBounce{cIcro_2}{}{cIcro_3}{.south} ;
+%   \THit{cro_0}{}{cIcro_3}{.190}{cIcro_2}
+%   \THit{cro_0}{}{cIcro_1}{.200}{cIcro_0}
+%   \THit{cro_1}{}{cIcro_3}{.190}{cIcro_2}
+%   \THit{cro_1}{}{cIcro_1}{.200}{cIcro_0}
+%   \path[bounce,bend right]
+%     \TBounce{cIcro_1}{}{cIcro_0}{.north}
+%     \TBounce{cIcro_3}{}{cIcro_2}{.120} ;
+
+  % MàJ abstraites de la sorte coopératiove
+  \path (1.8, 4.5) edge[coopupdate] (3.2, 3);
+  \path (1.8, 0) edge[coopupdate] (3.2, 1);
 
   % Actions sortantes
-  \THit{cIcro_3}{}{N_1}{.west}{N_0}
-  \path[bounce,bend right] \TBounce{N_1}{}{N_0}{.north} ;
-  \THit{cIcro_2}{}{N_0}{.north west}{N_1}
+  \THit{cIcro_3}{bend right=5}{N_1}{.150}{N_0}
+  \path[bounce,bend right=80,in=220] \TBounce{N_1}{}{N_0}{.130} ;
+  \THit{cIcro_2}{bend right=4}{N_0}{.150}{N_1}
   \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.260} ;
-  \THit{cIcro_1}{}{N_0}{.west}{N_1}
+  \THit{cIcro_1}{bend right=20}{N_0}{.west}{N_1}
   \path[bounce,bend left] \TBounce{N_0}{}{N_1}{.230} ;
-  \THit{cIcro_0}{}{N_0}{.south west}{N_1}
+  \THit{cIcro_0}{bend left=2}{N_0}{.south west}{N_1}
   \path[bounce,bend left=40] \TBounce{N_0}{}{N_1}{.200} ;
@@
   $\cI$ and $\cro$.
-  The processes of the cooperative sort $\cIcro$ are labeled by the configuration they represent,
-  given the values of $\cI$ and $\cro$ regarding their respective thresholds.
+  All the actions updating the cooperative $\cIcro$ are simply depicted by two
+  zigzag arrows.
+  %, which abstract the fact that this cooperative sort is deterministic.
+  Furthermore, as this cooperative sort $\cIcro$ is deterministic,
+  its processes are labeled by the configuration they represent,
+  given by two Boolean variables indicating whether
+  the values of $\cI$ and $\cro$ lie above their respective thresholds ($T$) or not ($F$).
+  For example, the configuration $\state{\cI_1, \cro_0}$ is represented by the process $TF$.
 }
@@
   \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_0, \cro_3, \N_0})
-    &= \{ \cI_0, \cro_3, \N_0, \cIcro_{FT} \} \\
+    &= \state{\cI_0, \cro_3, \N_0, \cIcro_{FT}} \\
   \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_1, \cro_3, \N_0})
-    &= \{ \cI_1, \cro_3, \N_0, \cIcro_{TT} \} \\
+    &= \state{\cI_1, \cro_3, \N_0, \cIcro_{TT}} \\
   \ctx_\N^{\{ \cI, \cro \}}(\state{\cI_2, \cro_3, \N_0})
-    &= \{ \cI_2, \cro_3, \N_0, \cIcro_{TT} \}
+    &= \state{\cI_2, \cro_3, \N_0, \cIcro_{TT}}
 \end{align*}
@@
 Of course:
-\[\forall \cro_i \in \PHl_{\cro}, \ctx_{\cro}^{\{ \cro \}}(\state{\cro_i}) = \{ \cro_i \}\]
+\[\forall \cro_i \in \PHl_{\cro}, \ctx_{\cro}^{\{ \cro \}}(\state{\cro_i}) = \state{\cro_i}\]
 and from \pref{eq:irB} and \eqref{eq:fp}:
@@
 \[\forall \cI_i \in \PHl_{\cI}, \forall \cro_j \in \PHl_{\cro},
-  \ctx_{\cro}^{\{ \cI \}}(\state{\cI_i, \cro_j}) = \{ \cI_i, \cro_j \}\]
+  \ctx_{\cro}^{\{ \cI \}}(\state{\cI_i, \cro_j}) = \state{\cI_i, \cro_j}\]
 Furthermore, from \pref{eq:irB}, we especially find:
@@
   K_{\N, \{\cI\}} &= 0\\
-  K_{\cI, \{\cII;cro\}} &= 2&
-  K_{\N, \{\cI;cro\}} &= 0\\
+  K_{\cI, \{\cII;\cro\}} &= 2&
+  K_{\N, \{\cI;\cro\}} &= 0\\
   \\
@@
   K_{\cro, \{\cI\}} &= 0\\
-  K_{\cII, \{\cI;cro\}} &= 0&
-  K_{\cro, \{\cI;cro\}} &= 2\\
+  K_{\cII, \{\cI;\cro\}} &= 0&
+  K_{\cro, \{\cI;\cro\}} &= 2\\
   K_{\cII, \{\N\}} &= 1\\
-  K_{\cII, \{\N;cro\}} &= 0\\
+  K_{\cII, \{\N;\cro\}} &= 0\\
   K_{\cII, \{\N;cI\}} &= 0\\
-  K_{\cII, \{\N;cI;cro\}} &= 0
+  K_{\cII, \{\N;cI;\cro\}} &= 0
 \end{align*}
@@
     of the parametrization, but this does not modify the dynamics, as mentioned previously;
-  \item Our inference returns $K_{\cro, \{\cI;cro\}} = 2$ instead of originally $0$;
-    once again, this does not change the dynamics, and $1$ would also have been
-    an acceptable value.
+  \item Our inference returns $K_{\cro, \{\cI;\cro\}} = 2$, \modLP{which is a value
+    equivalent to the original $0$ value according to \pref{rem:K-equiv}:
+    the resulting dynamics is the same, and $1$ would also have been an
+    acceptable value}.
 \end{itemize}

### 2014-11-03 parts-grn.tex

@@
 To refine the dynamics, a parametrization had been added to the model \cite{snoussi1989qualitative}.
-This parametrization allows to define focal points, i.e.
-it clarifies the level towards which a component evolves depending on the expression of its regulators.
+This parametrization allows to define focal points, \ie it clarifies the level towards which a component evolves depending on the expression of its regulators.
 Some of the information included in the IG and the parametrization may appear as slighltly redundant.
@@
 In the scope of the asychronous dynamics (\pref{def:dynamics}), one can remark
-that some parameters values can be equivalent, i.e. the resulting dynamics is
+that some parameters values can be equivalent, \ie the resulting dynamics is
 the same.

### 2014-11-03 parts-infer-ig.tex
fix auto-arc inference

@@
 possible bounces of $a$ in the configuration $\sigma$.}
-Finally, when $a$ has no regulators, it derives that there is a positive self-influence
-of $a$ for each level above $1$ (\pref{eq:edges-inference-noreg}).
 \begin{proposition}[Influences inference]\label{pps:inference-edges}
@@
 \begin{split}
- \forall a\in\Gamma: a\in\PHpredecgene{a}, & \forall s \in \{ +, - \}, \forall t < l_a\\
+ \forall a\in\Gamma, & \forall s \in \{ +, - \}, \forall t < l_a\\
   a \xrightarrow{t+1} a \in \hat{E}_s \EQDEF & \exists g \in X(a), 
@@
 \label{eq:edges-inference-auto}
-\\
-\forall a\in\Gamma: \PHpredecgene{a} = \emptyset, &
-	\forall t\in\{0,\cdots,l_a-1\},
-	a \xrightarrow{t+1} a \in \hat{E}_+
-\label{eq:edges-inference-noreg}
 \end{align}

### 2014-11-06 parts-asp.tex
fix running example

@@
 \begin{align*}
-  & 1\ \{\ \param(\var{A}, \var{P}, \var{I}) \mid \componentlevels(\var{A}, \var{I})\ \}\ \infty\ \la \\
+  & 1\ \{\ \param(\var{A}, \var{P}, \var{I}) \mid \componentlevels(\var{A}, \var{I})\ \}\ \modMF{1}\ \la \\
   & \qquad\qquad\qquad \paramlabel(\var{A}, \var{P}), \nota \inferedparam(\var{A}, \var{P}).
@@
 Furthermore, both the lower and upper bounds are $1$,
-which force each enumerated parameter to have exactly one value.
+which forces each enumerated parameter to have exactly one value.
 %but no upper bound is specified ($\infty$) for the size of each parameter
@@
 In the scope of \pref{ex:infer-param-runningPH-1},
-$K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$
+$K_{a,\{b\}}$ and $K_{a,\{c\}}$
 could not be inferred by \pref{pps:param_K}.
-The previous cardinality rule allows to produce 18 parametrizations,
+The previous cardinality rule allows to produce 9 parametrizations,
 in which these three parameters can take all possible values:
-%The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
 %(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
@@
   (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
-    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \} \times \{ 0, 1 \}
-%   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2}, \{ 0, 2 \} \}^2
+    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \}
 \end{align*}
 and all the other parameters keep their inferred values.
-Note that $0$ belongs to the set of candidate values for parameter $K_{b,\emptyset}$
-because the extreme values assumption (\pref{pro:param_enum_extreme})
-has not been taken into account yet.
 %These parametrizations take the form of sets of $\param$ atoms.
@@
 }%
-\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
+\todo{Find an example to apply the assumptions (properties 6--8)?}
 \end{example}
@@
 
-\begin{example}
-\modMF{%
-The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
-% For example, a constraint was written in order to obtain only interval parameters
-% (thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
-% Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
-% \begin{align*}
-%   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
-% \end{align*}
+% \begin{example}
+% \modMF{%
+% The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
+% % For example, a constraint was written in order to obtain only interval parameters
+% % (thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
+% % Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
+% % \begin{align*}
+% %   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
+% % \end{align*}
+% % }%
+% % 
+% % \modMF{%
 % }%
-% 
-% \modMF{%
-For example,
-% , due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
-the extreme values assumption (\pref{pro:param_enum_extreme}) removes all parametrizations with
-$K_{b,\emptyset} = 0$,
-thus reducing the set of candidate values to 9 possible values:
-\begin{align*}
-  (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
-    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \} \times \{ 1 \}
-\end{align*}
-Furthermore, parameter $K_{b,\emptyset} = 1$ doest not respect \pref{pro:K-valid},
-because no action in the model of \pref{fig:runningPH-1} allows to bounce from $b_0$ to $b_1$.
-Therefore, this parameter is removed from all returned parametrizations.
-In the end, 9 partial parametrizations are therefore produced.
-}%
-\todo{Find an example to apply the monotonicity assumption?}
-\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
-\end{example}
+% \todo{Find an example to apply the monotonicity assumption?}
+% \end{example}
 

### 2014-11-06 parts-infer-ig.tex
fix running example

@@
 \modMF{
-The IG inference of the PH of \pref{fig:runningPH-2} gives the
-IG in \pref{fig:runningBRN}(left), that is, allows to find back all the edges
-and their signs and thresholds.
-This is mainly due to the fact that the cooperation between $b$ and $c$ is well-defined,
-using a (deterministic) cooperative sort.
+The IG inference on the PH of \pref{fig:runningPH-2} returns the
+IG in \pref{fig:runningBRN}(left), that is, allows to find back all the expected edges
+alongside with their signs and thresholds.
+Furthermore, the IG inference on the PH of \pref{fig:runningPH-1}, which is not refined,
+gives the same result, despite the absence of the cooperative sort.
+This is due to the fact that the inference of \pref{pps:inference-edges}
+is able to analyze the influences from different regulators independently
+if they do not share a cooperative sort.
 }
@@
 \modMF{
-Indeed, the inference of the PH of \pref{fig:runningPH-1},
-that is, without refinement with a cooperative sort, gives the following IG:
-\begin{align*}
-  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{c}{+}{1}{c}\}\\
-  E_- &= \{\GRNedgef{a}{-}{2}{b}\} \qquad\qquad\qquad\qquad\qquad
-  E_\uns = \emptyset
-\end{align*}
-This IG is close to the one in \pref{fig:runningBRN} but misses the self-influence
-$\GRNedgef{b}{+}{1}{b}$.
-\todo{Detail why?}
-}
-
-\modMF{
-If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
+If an action $\PHfrappe{a_2}{b_0}{b_1}$ is added to the PH of \pref{fig:runningPH-2},
 then two \modLP{non-monotonous} edges towards $b$ are inferred instead of the previous signed edges:

### 2014-11-06 parts-infer-param.tex
fix running example

@@
 \label{ex:infer-param-runningPH-2}
-When applied to the refined PH in \pref{fig:runningPH-2},
-and with the IG of \pref{fig:runningBRN}(left)
-(which is the result of the IG inference on this PH model,
-as explained in \pref{ssec:infer-ig-examples}),
+When applied to the refined PH of \pref{fig:runningPH-2},
+and using the IG of \pref{fig:runningBRN}(left)
+which is the result of the IG inference on this PH model,
+as explained in \pref{ssec:infer-ig-examples},
 \pref{pps:param_K} infers the parametrization of \pref{fig:runningBRN}(right).
@@
 Regarding the non-refined PH of \pref{fig:runningPH-1},
-with the IG produced by the IG inference described in \pref{ssec:infer-ig-examples}
-(that is, the same as \pref{fig:runningBRN}(left)
-without the self-influence $\GRNedgef{b}{+}{1}{b}$)
-the following parameters are inferred:
-\begin{align*}
-  K_{a,\emptyset} &= 0 &
-  K_{b,\{a\}} &= 0 \\
-%  K_{a,\{b\}} &= 0 \\
-%  K_{a,\{c\}} &= 0 &
-  K_{a,\{b,c\}} &= 2 \\ \\
-%  K_{b,\emptyset} &= 0
-  K_{c,\emptyset} &= 0 &
-  K_{c,\{c\}} &= 1
-\end{align*}
-The parameters $K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$
-could not be inferred,
-mainly due to the lack of precision in the cooperation between $b$ and $c$.
-Note that the parameters $K_{b,\{b\}}$ and $K_{b,\{a,b\}}$ do not exist for this IG,
-because of the absence of any self-influence $b \rightarrow b$.
+and using the same IG of \pref{fig:runningBRN}(left),
+the parameters of \pref{fig:runningBRN}(right) are inferred,
+except the parameters $K_{a,\{b\}}$ and $K_{a,\{c\}}$
+for which \pref{pps:param_K} is not conclusive.
+The obtained parametrization is therefore partial.
+This is due to the lack of precision in the cooperation between $b$ and $c$,
+caused by the absence of the cooperative sort.
 \end{example}
@@
 The parametrization inferred in \pref{ex:infer-param-runningPH-1} was partial
-because \modMF{$K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$} could not be inferred.
-It is however possible to enumerate all \modMF{partial} and admissible parametrizations
+because \modMF{$K_{a,\{b\}}$ and $K_{a,\{c\}}$} could not be inferred.
+It is however possible to enumerate all admissible parametrizations
 compatible with both the inferred parameters, and the properties of this subsection.
-This enumeration gives 9 different \modMF{partial} parametrizations
+This enumeration gives 9 different parametrizations
 which correspond to the 3 possible values
@@
 for both $K_{a,\{b\}}$ and $K_{a,\{c\}}$:
-\begin{align*}
-  K_{a,\{b\}} &\in \{ 0, 1, 2 \} \\
-  K_{a,\{c\}} &\in \{ 0, 1, 2 \}
-\end{align*}
-$K_{b,\emptyset}$ is excluded
-from all return parametrizations because no value is compatible with the previous properties.
+%\[(K_{a,\{b\}}; K_{a,\{c\}}) \in \{ 0, 1, 2 \} \times \{ 0, 1, 2 \}\]
+\[K_{a,\{b\}} \in \{ 0, 1, 2 \} \qquad \text{and} \qquad K_{a,\{c\}} \in \{ 0, 1, 2 \}\]
 }%
@@
 Therefore this enumeration allows, from the model in \pref{fig:runningPH-1},
-\modMF{%
-to find the behavior of component $a$ in the model refined with a cooperative sort described in
+to find the behavior of the model refined with a cooperative sort described in
 \pref{fig:runningPH-2}.
-}%
-% to find the behavior of the model refined with a cooperative sort described in \pref{fig:runningPH-2}.
-\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
 \end{example}

### 2014-11-06 models/inference.sh
script for the benchmarks

@@ -1,0 +1,12 @@
+#!/dev/bash
+
+echo " --- ${1}${3}.ph --> ${1}${2}.{dot, lp, out, debug} ---"
+
+echo " --- ${1}${3}.ph --> ${1}${2}.{dot, lp, out, debug} ---" > $1$2.debug
+
+gnome-terminal --command="tail -f $1$2.debug"
+
+ph2thomas -i $1$3.ph --debug --dot $1$2.dot --debug-asp $1$2.lp > $1$2.out 2>> $1$2.debug
+
+#diff <(sort ${1}${3}.dot) <(sort ${1}${2}.dot)
+

### 2014-11-06 parts-asp.tex
add modMF

@@
 is a candidate value for the parameter ($\param$).
+\modMF{
 Furthermore, both the lower and upper bounds are $1$,
 which forces each enumerated parameter to have exactly one value.
+}
 %but no upper bound is specified ($\infty$) for the size of each parameter
@@
 so that if $K_{a, \omega_p}$ could not be inferred by \pref{pps:param_K}, then
+\modMF{
 $K_{a, \omega_p} \in \segm{0}{l_a}$
+}
 (thus completely disregarding the notion of admissible parametrizations given in \pref{ssec:admissible-K}).

### 2014-11-07 parts-example.tex
examples++
update computation times
phrase obsolète

@@
 
+\modMF{%
 This section aims at giving several applications of our work in order to
@@
 by studying parts of a model of the phage lambda immunity control.
+}%
 % the functioning of our method
@@
 \\\hline
-    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & 10s & --- & $2.1\cdot10^{9}$
+    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & 100s & <1s & 578
 \\\hline
-    EGF receptor \cite{Samaga2009} & $104$ & $89$ & 3min 20s & --- & $4.2\cdot10^{6}$
+    EGF receptor \cite{Samaga2009} & $104$ & $89$ & 200s & 2.5s & 27496
   \end{tabular}
@@
   The last column gives the number of parameters in each model.
-  Due to the too high number of parameters, parametrization inference could not be performed on the $94$ and $104$ component models.
   }

### 2014-11-07 parts-example.tex
corrections

@@
 The composition of all models and the results of the inferences are summarized in \pref{tb:computation}.
-We note that due to a very high number of parameters, no parameters inference could be performed on the $94$ and $104$ component models.
-These models would therefore be more efficiently studied as PH than as BRNs.
-Finally, we note that the complexity of the method is exponential in the number of regulators of one
-component and linear in the number of components.
+%We note that due to a very high number of parameters, no parameters inference could be performed on the $94$ and $104$ component models.
+%These models would therefore be more efficiently studied as PH than as BRNs.
+%Finally, we note that the complexity of the method is exponential in the number of regulators of one
+%component and linear in the number of components.
 

### 2014-11-07 parts-infer-ig.tex
corrections

@@
 such that the increase of one of the regulator $b\in g$ from $b_t$ to $b_{t+1}$
-changes the sign of the evolution of $a$.
+changes the direction of the evolution of $a$.
 If $b\neq a$ (\pref{eq:edges-inference}), this is achieved by finding a configuration $\sigma$ of $g$ and $a$ such
@@
     (k=t+1 \Rightarrow \exists \sigma'\in\fp(a_{t+1}): \sigma\{a_{t+1}\}\subseteq \sigma') \\
-  \qquad & \wedge k\geq t+1 \wedge j \leq t \wedge s=+ \\
-    \qquad & \vee k<t+1 \wedge j>t \wedge s=-
+  \qquad & \wedge (k\geq t+1 \wedge j \leq t \wedge s=+ \\
+    \qquad & \vee k<t+1 \wedge j>t \wedge s=-)
 \end{split}
@@
 We infer a positive (resp. negative) edge if there only exist corresponding influences with the same sign.
-If an influence is both positive and negative, we infer an
+If an influence is both positive and negative, we infer a
 \modLP{non-monotonous} edge.

### 2014-11-07 parts-infer-param.tex
corrections

@@
 Finally, we denote by $S_a(\sigma)$ (\pref{eq:K-domain}) the set of processes of sort $a$ that are
-compatible with a configuration $\sigma$: if $a$ is specified in $\ctx$, then $S_a(\sigma) = \{a_i\}$
+compatible with a configuration $\sigma$: if $a$ is specified in $\sigma$, then $S_a(\sigma) = \{a_i\}$
 where $a_i\in\sigma$; otherwise $S_a(\sigma)$ contains all the processes of sort $a$, $L_a$.}
@@
 	\begin{cases}
-	\{a_i\} & \text{if }a_i\in \ctx\\
+	\{a_i\} & \text{if }a_i\in \sigma\\
 	L_a & \text{otherwise.}

### 2014-11-07 parts-intro.tex
corrections

@@
 In other words, it captures the logical functions stating how} \modOR{various elements coalesce and act together upon an other element among the network.}
-%\todo{Explain the notion of “cooperation” (informally) to be understandable until the definition of cooperative sorts.}
-
-The modeling and analysis of the qualitative dynamics of large interaction net- works face two important challenges.
-For modeling, it is very rare that all logical functions governing the dynamics are known, even if the interaction graph is con- sidered as prior knowledge.
-One technique would then be to enumerate all logical networks that are compatible with the prior knowledge on the interaction graph and partial logical functions.
-But, with large networks, such an enumeration suffers from high combinatorics.
-For analysis, because the size of the state transition graph grows exponentially with the number of components, proving dynamical properties becomes intractable for networks with hundreds or thousands of elements.
-In order to illustrate the challenge of modeling interactions with partial knowledge,
-
-This allows to model BRNs with a partial knowledge on precise evolution functions for components
-by capturing the largest (the most general) dynamics.
+\modLP{In particular, PH allows to specify partial (non-deterministic) cooperations that superpose the dynamics of 
+candidate fully-specified (deterministic) cooperations.
+This permit to model BRNs with a partial knowledge on precise evolution functions for components
+by capturing the largest (the most general) dynamics, without a combinatorics enumeration of
+compatible Thomas models.}
 
@@
 few cooperations are specified.
-This highlights the fact that the PH is very suitable to analyze dynamics of large-scale networks
+This highlights that the PH is suitable to analyze dynamics of large-scale networks
 where the knowledge on cooperations is often incomplete:

### 2014-11-07 parts-ph.tex
corrections

@@
 
-\modLP{PH can be seens as a particular sub-class of safe Petri nets \cite{BC92},
+\modLP{PH is a particular sub-class of safe Petri nets \cite{BC92},
 i.e. Petri nets with at most one token per place, having
@@
 Hence, $bc_{01}$ represents the sub-state $\state{b_0,c_1}$, and so on.
-Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the status of the sorts $b$
+Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the state of the sorts $b$
 and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).

### 2014-11-07 parts-tr2global.tex
fix auto-arc inference
corrections

@@
 
-We call a \emph{deterministic sort} \pref{def:deterministic-sort} a sort $a$
+We call a \emph{deterministic sort} (\pref{def:deterministic-sort}) a sort $a$
 that has a single focal process for each possible configuration of its predecessors in the scope of
@@
 there exists $g\in X(a)$ such that $\{b,c\}\subseteq g$.
+If $\PHpredecgene{a} = \emptyset$, $X(a)\DEF \{\emptyset\}$.
 \end{definition}
@@
 \begin{example}
-\todo{}
+In the PH of \pref{fig:runningPH-2},
+$\ctx_a^{\{b,c\}}(\state{b_0,c_1}) = \state{b_0,c_1,bc_{01}}$
+and
+$\ctx_a^{\{b,c\}}(\state{b_1,c_1}) = \state{b_0,c_1,bc_{11}}$.
 \end{example}

### 2014-11-10 response.tex
update response to the reviewer

@@
 % vim:set spell spelllang=en:
+
+%\usepackage{babel}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
 \usepackage{color}
@@
 
-\section*{Comments and answers to the reviewer}
+\section*{Summary of the main changes}
 
-We thank the reviewer for th comments and  suggestions.
+\TODO{Detail more/less these changes?}
 
-\TODO{summary of response + changes}
+\begin{itemize}
+  \item The title has been changed from “Constructing...” to “Identification of...”.
+  \item The abstract and the introduction have been improved to better contextualize our work.
+  \item New explanations have been added at the beginning of every section
+    to detail their objectives and methods.
+  \item The explanations about cooperations in Process Hitting have been partially rewritten.
+  \item In the section defining Thomas modeling (subsection 2.2):
+    \begin{itemize}
+      \item “Unsigned edges” are now called “non-monotonous edges”,
+        and explanations have been added to explain their meaning;
+      \item We now only consider single-value parameters
+        (as opposed to previous interval parameters);
+      \item A property has been added to characterize equivalent parametrizations
+        (see Remark 1).
+    \end{itemize}
+  \item A new section has been created (section 3)
+    in order to better separate general definitions (esp. the notions of focal processes,
+    deterministic sorts, components/cooperative sorts separation,
+    regulators and configuration extension)
+    from the IG inference propositions of section 4.
+  \item The IG inference method has been improved in order to get more accurate results
+    (see Proposition 1).
+  \item Property 4 (Well-formed PH for parameter inference) has been improved
+    to handle more general forms of PH models,
+    namely, with cooperative sorts containing more than 4 processes.
+  \item Equations of section 5 (19 to 22) have been adapted to single-value parameters.
+  \item All examples in sections 2 to 5 have been actualized to the new formalism
+    and propositions.
+  \item A detailed example has been added in section 7 (see subsection 7.1)
+    where we detail the application of our method to some parts of a model of
+    the bacteriophage lambda immunity control.
+  \item Actualized results are given in subsection 7.3,
+    related to several improvements brought to our method.
+\end{itemize}
+
+\section*{Summary of our answers to the reviewer}
+
+We thank the reviewer for these comments and  suggestions.
+The reviewer's main concerns about our work were the readability of our paper and the consistency problems posed by our definition of Thomas' modeling.
+
+The readability has been improved by rewriting large parts of the section previously numbered 3, by:
+\begin{itemize}
+  \item Splitting it into two sections, in order to handle apart the general definitions and the propositions linked to the IG inference;
+  \item Adding some paragraphs of explanations at the beginning of sections and before the most complex definitions;
+  \item Adding running examples in these sections;
+  \item Dealing with a detailed example of a model of the bacteriophage lambda immunity control (subsection 7.1) to illustrate our method.
+\end{itemize}
+
+Regarding the definition of Thomas' modeling, we completely understand the reviewer's concerns and comments.
+These definitions are indeed redundant and some pieces of informations can be contradictory (namely, the parametrization vs.
+the sings on the edges that describe the kind of influences between the components).
+However, we chose to keep the current definition of Thomas' models. Our motivations are the following:
+\begin{itemize}
+  \item We wish to stay close to the traditional definition of this formalism, in order not to destabilize the reader;
+  \item Introducing new definitions such as an unsigned IG, or new constrains on the IG such as a monotonicity property may add unnecessary complexity;
+  \item The signs in the IG are interesting in themselves and lead to several works on IG-based static analyses.
+\end{itemize}
 
 Please find below our detailed answers.
+
+\section*{Comments and answers to the reviewer}
 
@@
 
-\DISCUSS{LP}
+% \DISCUSS{LP}
 Thomas' formalism with discrete parameters indeed includes some redundant piece of information in the sense that signs could be captured by the parameters with additional constraints, e.g.
 monotonous behaviors (which imposes an ordering between the value of parameters).
-Speaking from a computer science oriented point of view, the information encapsulated into a Thomas' network could be reduced to the set of nodes and the corresponding logical parameters.
-The single parametrization is very generic, ranging from totally decoupled specification to very constrained valuation w.r.t to the interaction graph.
+Speaking from a computer science oriented point of view, the information encapsulated into a Thomas' network could be reduced to a parametrization (which contains the name of each component, and where each maximum level can be inferred from the maximum value found in the related parameters) by simply adding information about the thresholds.
+Furthermore, the parametrization is very generic, ranging from totally decoupled specification to very constrained valuation w.r.t.
+the interaction graph.
 This would however be a disruption with the Thomas framework as it has been considered for more than 25 years.
@@
 As showed for example in THOMAS, R.
-\& D'ARI, R., "Biological feedback", CRC Press, Boca Raton , Florida, USA, (1990), 316 pp), the analysis of some dynamical properties (e.g.
+\& D'ARI, R., “Biological feedback” (CRC Press, Boca Raton , Florida, USA, 1990, 316 pp), the analysis of some dynamical properties (e.g.
 sustained oscillations) can derive from the study of positive or negative circuits.
@@
 
-Nevertheless, to remove ambiguity and comply with the reviewer's suggestions, we redesigned section 2.2, emphasizing such a choice.
+Nevertheless, to remove ambiguity and comply with the reviewer's suggestions, we redesigned section 2.2, emphasizing such a choice, especially with this excerpt:
+“distinguishing the IG and parametrization allows to decouple the graph (which is generally one of the prior knowledge from the biologists) from the constraints on its dynamics.
+It opened the way to numerous studies (e.g., necessary condition on sustained oscillations) solely based on the structure of the IG (e.g., the analysis of positive or negative circuits)”.
+
+% \DISCUSS{MF} I just changed a little bit the first paragraph.
 
@@
 \answer
-\DISCUSS{LP}
-proposition: ``The PH is a particular class of Indeterministic Asynchronous Automata
-Network (or safe Petri nets) on which have been designed static analysis for
-dynamical properties that are tractable on very large systems.''
+% \DISCUSS{LP}
+% proposition: ``The PH is a particular class of Indeterministic Asynchronous Automata
+% Network (or safe Petri nets) on which have been designed static analysis for
+% dynamical properties that are tractable on very large systems.''
 
-\TODO{Modify sentence + answer (quote)}
+This comment has been taken into account and the sentence has been replaced by:
+“The analysis of dynamics of very large models is however a rather hard problem, which led us to previously introduce a new framework: the Process Hitting (PH), which is a particular class of Indeterministic Asynchronous Automata Network (or safe Petri nets).
+Its major advantage lies in the efficiency of several static analyses recently designed to assess dynamical properties, making it possible to tackle very large models.”
+
+%\TODO{Modify sentence + answer (quote)}
 
@@
 \answer
-\DISCUSS{LP} yes, we have to mention Kauffman of course, while stressing the
-differences brought by Thomas (async + parameters).
+% \DISCUSS{LP} yes, we have to mention Kauffman of course, while stressing the
+% differences brought by Thomas (async + parameters).
+
+This comment has been taken into account and we added a reference to Kauffman in the abstract and in the introduction.
+However, we do not insist more on this as we only use BRNs as later extended and defined by Thomas in our work.
 
@@
 \answer
-\DISCUSS{LP} yes, we have to mention those work, while stressing that they have
-hard time with transient dynamics on very large networks, partly due to
-concurrency, which makes model-checking very hard even with symbolic techniques.
+We added some references to such methods, but did not emphasis on this as it is not the main subject of our work.
+The main interesting feature of Process Hitting in our work is that it is able to model multy-valued and indeterministic systems.
 
-\DISCUSS{LP} how such a discussion is relevant for the paper? Can't we simply
-say that PH allows to model indeterministic Boolean and multi-valued networks
-and remove (or drastically reduce) this discussion on model-checking?
+% \DISCUSS{LP} yes, we have to mention those work, while stressing that they have
+% hard time with transient dynamics on very large networks, partly due to
+% concurrency, which makes model-checking very hard even with symbolic techniques.
+% 
+% \DISCUSS{LP} how such a discussion is relevant for the paper? Can't we simply
+% say that PH allows to model indeterministic Boolean and multi-valued networks
+% and remove (or drastically reduce) this discussion on model-checking?
 
@@
 
-\DISCUSS{LP} we should avoid this term in introduction, and only use once
-formally defined.
-In the introduction, maybe we should mention that
-PH models can embed additional automata that encode for some (partial) logic
-gates. Such additional automata that can be considered as ``internal''
-variables, and we may want to mask them in abstract representations of the
-systems, such as IG and Thomas' parameters.
-\DISCUSS{LP} the general the notion of cooperation should be defined in the
-introduction.
+We took this comment into account and added more explanations in the introduction of paper to better explain what we meant by “cooperations” or “associated influences”, and by changing a little bit the terminology for better understanding.
+We especially added the following sentences just after the first occurrence of this notion of cooperation:
+“The concept of cooperation refers to the way two (or more) components jointly influence a third one.
+In other words, it captures the logical functions stating how various elements coalesce and act together upon an other element among the network.”
+
+% \DISCUSS{LP} we should avoid this term in introduction, and only use once
+% formally defined.
+% In the introduction, maybe we should mention that
+% PH models can embed additional automata that encode for some (partial) logic
+% gates. Such additional automata that can be considered as ``internal''
+% variables, and we may want to mask them in abstract representations of the
+% systems, such as IG and Thomas' parameters.
+% \DISCUSS{LP} the general the notion of cooperation should be defined in the
+% introduction.
 
@@
 \answer
-\TODO{improve figure}
+% \TODO{improve figure}
+We took this comment into account and made the figure more readable.
 
@@
 \answer
-\DISCUSS{LP} I would go for removing this discussion.
+% \DISCUSS{LP} I would go for removing this discussion.
+We took this comment into account and removed this part of the discussion as it was more confusing than explanatory regarding our work.
 
@@
 (P. 9) Interaction Graph. $\Gamma$ refers to the set of vertices. However
-$\Gamma^{−1}(b)$
+$\Gamma^{-1}(b)$
 requires to seek the pre-image in E. I find the choice of $\Gamma$ to annotate
@@
 \answer
-\TODO{Apply the changes + quick answer}
+% \TODO{Apply the changes + quick answer}
+This notation was intended to match the corresponding one for the Process Hitting models, but it was indeed a poor choice. We took this comment into account and changed this notation to $E^{-1}(b)$.
 
@@
 \answer
-\DISCUSS{LP} See 1. The definitions need to be more clear with more textual
-explanation for CS people.
+Please refer to our answer to question 1.
+There are indeed several less redundant ways to represent a BRN but we decided to keep the definition proposed by Thomas.
+Even if the parametrization is present, the signs on the edges still have an informative value.
+Furthermore, we do consider the signs in the case of the parameters enumeration.
+% \DISCUSS{LP} See 1. The definitions need to be more clear with more textual
+% explanation for CS people.
 
@@
 \answer
+We took the second comment into account and removed the discussion about maximum levels which was indeed unnecessary.
+
 \DISCUSS{LP} Yes, we should make clear that we are looking for monotonic
@@
 levels.
+
+\DISCUSS{MF} In fact I don't really understand this comment --- I fist thought it was related to questions 1 and 9 but in fact it is not.
+There is no inference on page 11; was the reviewer talking about the dynamics?
+In the end, has the discussion about monotonic influences been added to the paper?
 
@@
 \answer
-\DISCUSS{LP} Same as 1 and 9.
+Please refer to our answers to questions 1 and 9.
+A definition of unsigned IG could indeed be included, but we do not think that it would make the matter clearer because our method is not related to unsigned IGs.
+% \DISCUSS{LP} Same as 1 and 9.
 
@@
 \answer
-\DISCUSS{LP}
-Indeed, the motivation is quite weak. Some arguments:
-\begin{enumerate}
-\item Abstract representation of influences
-\item Consistency with prior knowledge: if the modelling started from an IG, we may want to
-verify that the dynamical models is consistent w.r.t. to the prior IG
-\item + remove non-used influences (IG reduction).
-\end{enumerate}
+We took this comment into account and added the following explanations (2nd paragraph of section 4):
+“The inference of the IG also allows to check for the consistency of the model
+with respect to prior knowledge on the influences. This is the case for Thomas
+models, where a prior IG is required to specify the parameters of the dynamics.
+As detailed in A.
+Richard, J.-P.
+Comet, G.
+Bernot, “Formal Methods for Modeling Biological Regulatory Networks” (Modern Formal Methods and Applications, 2006, pp.
+83--122), the IG of a dynamical model is consistent with a prior IG
+if it is a sub-graph of the prior IG. On the one hand, having influences in the
+model not referenced in the prior IG indicates either mistakes in the encoding
+of the dynamics, or potential influences that might need to be verified in the
+biological system. One the other hand, some influences specified in the prior IG
+may not have been necessary to encode the desired dynamics, indicating some
+non-minimality of the prior IG.”
+
+% \DISCUSS{LP}
+% Indeed, the motivation is quite weak. Some arguments:
+% \begin{enumerate}
+% \item Abstract representation of influences
+% \item Consistency with prior knowledge: if the modelling started from an IG, we may want to
+% verify that the dynamical models is consistent w.r.t. to the prior IG
+% \item + remove non-used influences (IG reduction).
+% \end{enumerate}
 
@@
 \answer
-\DISCUSS{LP}redo this section
+We took this comment into account and made major changes to this section, by splitting it in two and adding explanations and examples for most definitions.
+% \DISCUSS{LP}redo this section
 
@@
 \answer
-\DISCUSS{LP} use Lambda phage as a running example (or dedicate a subsection to
-apply the whole pipeline to it).
+We took these comments into account:
+\begin{itemize}
+  \item We have actualized the running examples, especially in order to fix this inconsistency between the explanations and the figure.
+  \item A subsection has been added (subsection 7.1) in order to detail an example of our method on the bacteriophage lambda immunity control, as suggested.
+\end{itemize}
+
+% \DISCUSS{LP} use Lambda phage as a running example (or dedicate a subsection to
+% apply the whole pipeline to it).
 

### 2014-11-10 response.tex
response++

@@
 
-\TODO{Detail more/less these changes?}
+We thank the reviewer for these helpful comments and suggestions.
+The reviewer's main concerns about our work were the readability of our paper and the consistency problems posed by our definition of Thomas' modeling.
 
+We made the following main changes to the article:
 \begin{itemize}
@@
   \item A new section has been created (section 3)
-    in order to better separate general definitions (esp. the notions of focal processes,
-    deterministic sorts, components/cooperative sorts separation,
-    regulators and configuration extension)
-    from the IG inference propositions of section 4.
-  \item The IG inference method has been improved in order to get more accurate results
-    (see Proposition 1).
-  \item Property 4 (Well-formed PH for parameter inference) has been improved
-    to handle more general forms of PH models,
-    namely, with cooperative sorts containing more than 4 processes.
-  \item Equations of section 5 (19 to 22) have been adapted to single-value parameters.
-  \item All examples in sections 2 to 5 have been actualized to the new formalism
-    and propositions.
+    in order to better separate general definitions 
+    from the Interaction Graph (IG) inference propositions of section 4.
+  %\item The IG inference method has been improved in order to get more accurate results
+  %  (see Proposition 1).
+  %\item Property 4 (Well-formed PH for parameter inference) has been improved
+  %  to handle more general forms of PH models,
+  %  namely, with cooperative sorts containing more than 4 processes.
+  %\item Equations of section 5 (19 to 22) have been adapted to single-value parameters.
+  \item Some definitions and propositions have been updated to make them more simple and readable
+  %All examples in sections 2 to 5 have been actualized to the new formalism
+  %  and propositions.
   \item A detailed example has been added in section 7 (see subsection 7.1)
@@
     the bacteriophage lambda immunity control.
-  \item Actualized results are given in subsection 7.3,
-    related to several improvements brought to our method.
-\end{itemize}
-
-\section*{Summary of our answers to the reviewer}
-
-We thank the reviewer for these comments and  suggestions.
-The reviewer's main concerns about our work were the readability of our paper and the consistency problems posed by our definition of Thomas' modeling.
-
-The readability has been improved by rewriting large parts of the section previously numbered 3, by:
-\begin{itemize}
-  \item Splitting it into two sections, in order to handle apart the general definitions and the propositions linked to the IG inference;
-  \item Adding some paragraphs of explanations at the beginning of sections and before the most complex definitions;
-  \item Adding running examples in these sections;
-  \item Dealing with a detailed example of a model of the bacteriophage lambda immunity control (subsection 7.1) to illustrate our method.
-\end{itemize}
-
-Regarding the definition of Thomas' modeling, we completely understand the reviewer's concerns and comments.
-These definitions are indeed redundant and some pieces of informations can be contradictory (namely, the parametrization vs.
-the sings on the edges that describe the kind of influences between the components).
-However, we chose to keep the current definition of Thomas' models. Our motivations are the following:
-\begin{itemize}
-  \item We wish to stay close to the traditional definition of this formalism, in order not to destabilize the reader;
-  \item Introducing new definitions such as an unsigned IG, or new constrains on the IG such as a monotonicity property may add unnecessary complexity;
-  \item The signs in the IG are interesting in themselves and lead to several works on IG-based static analyses.
+  %\item Actualized results are given in subsection 7.3,
+  %  related to several improvements brought to our method.
 \end{itemize}
@@
 
+Having both the (signed) interaction graph (IG) and the parametrization in the model allows to decouple the IG from the dynamics, which is interesting for modeling biological systems as the IG is often a prior knowledge with possible inconsistencies with the dynamics we want to encode with.
+A further step may then infer back the IG of the actual dynamics (here, an abstraction of the encoded dynamics) and compare it with the prior IG.
+Typically, if the inferred IG is included in the prior one, we have consistency.
+Conversely, the signs of the prior IG can be used to constraint the possible parametrizations to
+ensure consistency with the signed interactions (e.g. by imposing an ordering between parameters
+values to ensure monotonicity, see section 5.2).
+
 % \DISCUSS{LP}
-Thomas' formalism with discrete parameters indeed includes some redundant piece of information in the sense that signs could be captured by the parameters with additional constraints, e.g.
+%Thomas' formalism with discrete parameters indeed includes some redundant piece of information in the sense that signs could be captured by the parameters with additional constraints, e.g.
 monotonous behaviors (which imposes an ordering between the value of parameters).
@@
 
-Moreover signs and edges have been an inspiration for numerous works with regard to the analysis of networks.
+%Moreover signs and edges have been an inspiration for numerous works with regard to the analysis of networks.
 As showed for example in THOMAS, R.
@@
 
-In addition, having both the interaction graph and the parametrization in the model allows to decouple the IG from the dynamics, which is interesting for modeling biological systems as the IG is often a prior knowledge with possible inconsistencies with the dynamics we want to encode with.
-A further step may then infer back the IG of the actual dynamics (here, an abstraction of the encoded dynamics) and compare it with the prior IG.
-Typically, if the inferred IG is included in the prior one, we have consistency.
-
-Finally, several alternative definitions exist, with various degree of equivalency.
+Several alternative definitions exist, with various degree of equivalency.
 As a consequence, we chose here to stick to the original formalism and picked this definition as it the original and standard one in the Thomas' community.
 
-Nevertheless, to remove ambiguity and comply with the reviewer's suggestions, we redesigned section 2.2, emphasizing such a choice, especially with this excerpt:
-“distinguishing the IG and parametrization allows to decouple the graph (which is generally one of the prior knowledge from the biologists) from the constraints on its dynamics.
+We added such a discussion in the introduction of section 2.2.
+%“distinguishing the IG and parametrization allows to decouple the graph (which is generally one of the prior knowledge from the biologists) from the constraints on its dynamics.
 It opened the way to numerous studies (e.g., necessary condition on sustained oscillations) solely based on the structure of the IG (e.g., the analysis of positive or negative circuits)”.
@@
 % \TODO{Apply the changes + quick answer}
-This notation was intended to match the corresponding one for the Process Hitting models, but it was indeed a poor choice. We took this comment into account and changed this notation to $E^{-1}(b)$.
+%This notation was intended to match the corresponding one for the Process Hitting models, but it was indeed a poor choice. 
+We took this comment into account and changed this notation to $E^{-1}(b)$.
 
@@
 \answer
-Please refer to our answer to question 1.
-There are indeed several less redundant ways to represent a BRN but we decided to keep the definition proposed by Thomas.
+Please refer to our answer to question 1.% There are indeed several less redundant ways to represent a BRN but we decided to keep the definition proposed by Thomas.
 Even if the parametrization is present, the signs on the edges still have an informative value.
@@
 \answer
-We took the second comment into account and removed the discussion about maximum levels which was indeed unnecessary.
-
-\DISCUSS{LP} Yes, we should make clear that we are looking for monotonic
-influences - but not necessarily at the global scope, which mean that we can
-find two contradictory monotonic influences, leading to those ``unsigned'' edges
-in the IG.
-And yes, we should just mention that PH models can have an arbitrary number of
-levels.
-
-\DISCUSS{MF} In fact I don't really understand this comment --- I fist thought it was related to questions 1 and 9 but in fact it is not.
-There is no inference on page 11; was the reviewer talking about the dynamics?
-In the end, has the discussion about monotonic influences been added to the paper?
+We added mentions to the notion of monotony in the definition of the signs of influences from the
+dynamics.\\
+As suggested, we removed the discussion about maximum levels which was indeed unnecessary.
 
@@
 \answer
-We took this comment into account and made major changes to this section, by splitting it in two and adding explanations and examples for most definitions.
-% \DISCUSS{LP}redo this section
+We took this comment into account and revised the structure of this section.
+We created a new section dedicated to the main definitions used by the IG and parameters inference
+section.
+Most of these definitions are related to the step of inferring global influences between components
+from local transitions specifications.
+We added explanations and examples to make such definitions more understandable.
 
@@
 \answer
-We took these comments into account:
-\begin{itemize}
-  \item We have actualized the running examples, especially in order to fix this inconsistency between the explanations and the figure.
-  \item A subsection has been added (subsection 7.1) in order to detail an example of our method on the bacteriophage lambda immunity control, as suggested.
-\end{itemize}
-
-% \DISCUSS{LP} use Lambda phage as a running example (or dedicate a subsection to
-% apply the whole pipeline to it).
+As suggested, we added a new subsection (7.1) dedicated to a didactic application of our method to
+a concrete biological modeling example, on the bacteriophage lambda immunity control.
+We have also actualized the running examples, especially in order to fix this inconsistency between the explanations and the figure.
 

### 2014-11-10 parts-grn.tex
response++

@@
 where $i < l_b$ and $s\{b_i\}$ denotes the state $s$ where the component $b$ is assigned to $b_i$.
-Then, such an IG can be used to derive global properties on the dynamics
-(e.g., \cite{Richard2010378,PR11-SASB}).
+\modLP{If all the interactions from $b$ to $a$ are positive (resp. negative), there is a positive
+(resp. negative) edge from $b$ to $a$ in the IG, denoting a monotonous influence of $b$ on $a$.
+In the case when there exists two differently signed interactions, the edge in the IG is said
+non-monotonous.
+This abstract representation of the dynamical system can then} be used to derive global properties on
+the dynamics (e.g., \cite{Richard2010378,PR11-SASB}).

### 2014-11-10 response.tex
response++

@@
 Typically, if the inferred IG is included in the prior one, we have consistency.
-Conversely, the signs of the prior IG can be used to constraint the possible parametrizations to
+Conversely, the signs of the prior IG can be used to constrain the possible parametrizations to
 ensure consistency with the signed interactions (e.g. by imposing an ordering between parameters
-values to ensure monotonicity, see section 5.2).
+values to ensure monotonicity, see section 5.2) -- in such a case, in your example, we would impose
+that $K_{b,\{a\}} \geq K_{b,\emptyset}$.
 

### 2014-11-10 parts-infer-param.tex
finish running example + minor fixes

@@
 \begin{property}[Extreme values assumption]
+\label{pro:param_enum_extreme}
 Let $\IG = (\Gamma, E)$ be an IG. A parametrization $K$ on $\IG$ satisfies the \emph{extreme values assumption} if and only if:
-\label{pro:param_enum_extreme}
 \begin{align*}
@@
 \[K_{a,\{b\}} \in \{ 0, 1, 2 \} \qquad \text{and} \qquad K_{a,\{c\}} \in \{ 0, 1, 2 \}\]
-}%
-% We note that for all solutions, $0 \notin K_{a,\{a,b\}} \wedge 0 \notin K_{a,\{a,c\}}$.
-% This is due to the monotonicity assumption (\pref{pro:param_enum_monotonicity}) which especially states that:
-% \begin{align*}
-%   K_{a,\{b\}} \leqsegm K_{a,\{a,b\}} \wedge
-%   K_{a,\{c\}} \leqsegm K_{a,\{a,c\}}
-% \end{align*}
-\todo{Find an example to apply the assumptions (properties 6--8)?}
+We note that all these solutions satisfy Properties
+\ref{pro:K-valid} to \ref{pro:param_enum_monotonicity}.
+All the parametrizations obtained from these combinations are thus admissible.
+}
 
-Finally, we note that \modMF{$1$} belongs to the possible values for both parameters.
+Finally, we note that the value \modMF{$1$} belongs to the possible values for both parameters.
 Therefore this enumeration allows, from the model in \pref{fig:runningPH-1},

### 2014-11-10 parts-asp.tex
move TODO
finish running example + minor fixes
minor corrections

@@
 or a \emph{variable}, which is in fact a shorthand for any possible constants (variables are detailed below).
-In this paper, constants are either numerical or consist of a single lowercase letter (e.g.~$a$, $b$, $c$, $1$, $2$, …)
-while variables are always denoted by a single capital letter (e.g.~$\var{A}$, $\var{P}$, $\var{Q}$, …).
+In this paper, constants are either numerical or consist of a single lowercase letter (\eg $a$, $b$, $c$, $1$, $2$, …)
+while variables are always denoted by a single capital letter (\eg $\var{A}$, $\var{P}$, $\var{Q}$, …).
 We do not use function symbols as arguments in this work.
@@
 \begin{align*}
-  (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
+  (K_{a,\{b\}}; K_{a,\{c\}}) \in
     \{ 0, 1, 2 \} \times \{ 0, 1, 2 \}
@@
 }%
-\todo{Find an example to apply the assumptions (properties 6--8)?}
 \end{example}
@@
 
-% \begin{example}
-% \modMF{%
-% The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
-% % For example, a constraint was written in order to obtain only interval parameters
-% % (thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
-% % Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
-% % \begin{align*}
-% %   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
-% % \end{align*}
-% % }%
-% % 
-% % \modMF{%
-% }%
-% \todo{Find an example to apply the monotonicity assumption?}
-% \end{example}
+\begin{example}
+\modMF{%
+All the candidate values enumerated in \pref{ex:cardinality}
+already respect Properties \ref{pro:K-valid} to \ref{pro:param_enum_monotonicity}.
+Therefore, the constraints encoding these properties do not filter any solution.
+}
+\end{example}
 

### 2014-11-10 parts-grn.tex
minor corrections

@@
 Some of the information included in the IG and the parametrization may appear as slighltly redundant.
-Especially, if the single parametrization were enriched with additional constraints (e.g., monotonicity requiring an ordering on the value of parameters), then it may directly capture the signs.
+Especially, if the single parametrization were enriched with additional constraints (\eg monotonicity requiring an ordering on the value of parameters), then it may directly capture the signs.
 However, distinguishing the IG and parametrization allows to decouple the graph (which is generally one of the prior knowledge from the biologists) from the constraints on its dynamics.
-It opened the way to numerous studies (e.g., necessary condition on sustained oscillations) solely based on the structure of the IG (e.g., the analysis of positive or negative circuits).}
+It opened the way to numerous studies (\eg necessary condition on sustained oscillations) solely based on the structure of the IG (\eg the analysis of positive or negative circuits).}
 
@@
 This abstract representation of the dynamical system can then} be used to derive global properties on
-the dynamics (e.g., \cite{Richard2010378,PR11-SASB}).
+the dynamics (\eg \cite{Richard2010378,PR11-SASB}).

### 2014-11-10 parts-infer-ig.tex
minor corrections

@@
 with a prior IG if it is a sub-graph of the prior IG.
-On the one hand, having influences in the model not references in the prior IG
+On the one hand, having influences in the model not referenced in the prior IG
 indicates either mistakes in the encoding of the dynamics, or potential

### 2014-11-10 parts-infer-param.tex
minor corrections

@@
 ot the PH dynamics: a Thomas model is compatible with a PH if Thomas dynamics is included in the PH
-dynamics, i.e., all the transitions in the Thomas model are possible transitions in the PH.
+dynamics, \ie all the transitions in the Thomas model are possible transitions in the PH.
 This relaxed notion of parameterization compatibility allows to enumerate all parametrizations

### 2014-11-10 parts-intro.tex
minor corrections

@@
 Thomas' modeling has motivated numerous works around the link between the Interaction Graph (IG)
-(summarizing the global influences between components) and the possible dynamics (e.g.,
+(summarizing the global influences between components) and the possible dynamics (\eg
 \cite{RiCo07,RRT08}),
-model reduction (e.g., \cite{Naldi09}), formal checking of dynamics (e.g., \cite{Richard06,Naldi07}), 
-and the incorporation of time (e.g., \cite{Siebert06,Ahmad08}) and probability
-(e.g., \cite{Twardziok10-CMSB}) dimensions, to name but a few.
+model reduction (\eg \cite{Naldi09}), formal checking of dynamics (\eg \cite{Richard06,Naldi07}), 
+and the incorporation of time (\eg \cite{Siebert06,Ahmad08}) and probability
+(\eg \cite{Twardziok10-CMSB}) dimensions, to name but a few.
 
@@
 specify Thomas' parameters, especially for large networks.
-\modOR{Other works have also been carried out with the aim of dealing with large systems (e.g.
-\cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generaly focus on some restricted properties,} \modOR{as finding out steady states.}
+\modOR{Other works have also been carried out with the aim of dealing with large systems (\eg \cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generaly focus on some restricted properties,} \modOR{as finding out steady states.}
 %\todo{The reviewer proposes to nuance this with several papers:
@@
 
-In order to address the formal analysis of additional dynamical properties \modMM{(e.g., state reachabilities)} within very large BRNs, we recently
+In order to address the formal analysis of additional dynamical properties \modMM{(\eg state reachabilities)} within very large BRNs, we recently
 introduced in \cite{PMR10-TCSB} a new formalism, named the \emph{Process Hitting} (PH), to model
@@
 Indeed, when one wants to modify some behaviors
-(e.g.~so as to avoid the reachability of some states or to create stable states)
+(\eg so as to avoid the reachability of some states or to create stable states)
 we intend to be able to get upstream the results in accordance at the regulatory level.
@@
 The most known are the Thomas' conjectures (that have been proved since,
-e.g.~\cite{RRT08,RiCo07,Richard2010378}
+\eg \cite{RRT08,RiCo07,Richard2010378}
 for Boolean and discrete networks),

### 2014-11-10 parts-ph.tex
minor corrections

@@
 \modLP{PH is a particular sub-class of safe Petri nets \cite{BC92},
-i.e. Petri nets with at most one token per place, having
+\ie Petri nets with at most one token per place, having
 groups of multually exclusive places (processes) acting as the sorts, and where each
@@
 One may want to model transitions that should be taken only when several
-processes are present, i.e., modeling conjunctions -- or cooperations -- between the presence of
+processes are present, \ie modeling conjunctions -- or cooperations -- between the presence of
 several processes.
@@
 Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the state of the sorts $b$
-and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
+and $c$ (\eg $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
 Then, it is process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of

### 2014-11-10 parts-infer-ig.tex
ensure some notations are introduced

@@
 If $b\neq a$ (\pref{eq:edges-inference}), this is achieved by finding a configuration $\sigma$ of $g$ and $a$ such
-that there exists a bounce $a_j$ from $a_i$ in $\sigma\{b_t\}$ which is different
+that there exists a bounce $a_j$ from $a_i$ in $\sigma\{b_t\}$ (\ie the state $\sigma$ where the
+process of sort $b$ has been replaced with $b_t$) which is different
 from a bounce $a_k$ from $a_i$ in $\sigma\{b_{t+1}\}$.

### 2014-11-10 parts-tr2global.tex
minor corrections
ensure some notations are introduced

@@
 We denote $X(a)$ the finest partition in groups of
-regulators (\pref{def:influence-groups}), i.e. where each group is minimal.
+regulators (\pref{def:influence-groups}), \ie where each group is minimal.
 \begin{definition}[Partition $X(a)$ of $\PHpredecgene{a}$]
@@
 \end{align}
-where $\f{dom}(\sigma')$ is the set of sorts defined in $\sigma'$.
+where $\f{dom}(\sigma')$ is the set of sorts defined in $\sigma'$
+and $\uplus$ denotes the union of two (sub)states.
 \end{definition}

### 2014-11-10 parts-infer-ig.tex
remove obsolete sentence

@@
 The set of configuration where no bounce occur on $a_i$ is given given by
-$\fp(a_i)$ (\pref{eq:fp}).
-In \pref{eq:edges-inference} and \pref{eq:edges-inference-auto} we use
-$\irB_a^g(\sigma)$ as a shortcut for $\focals$, which returns the set of
-possible bounces of $a$ in the configuration $\sigma$.}
+$\fp(a_i)$ (\pref{eq:fp}).}
 \begin{proposition}[Influences inference]\label{pps:inference-edges}

### 2014-11-11 parts-asp.tex
corrections mineures

@@
+% vi:spell spelllang=en:
 \section{Answer Set Programming implementation concepts}\label{sec:impl}
@@
 which has been chosen to address the enumeration of all admissible parametrizations.
-The motivations are following:
-\begin{itemize}
-  \item ASP efficiently tackles the inherent complexity of the used models, thus enabling us fast execution of the formal tools defined in this paper,
-  \item ASP can efficiently enumerate a large set of possible answers,
-  \item and it is easy to constrain the answers according to some properties.
-\end{itemize}
-We now synthesize some key points to better make the reader understand our ASP implementation with the enumeration example.
+%The motivations are following:
+ASP can efficiently enumerate a large set of possible answers,
+  and it is easy to constrain the answers according to some properties.
+Applied to the models and constraints defined in this papers,
+ASP appears to be efficient for tackling the inherent complexity of the enumeration of parametrizations.
+
+Hereafter, we synthesize some key points of our ASP implementation with the enumeration example.
 
 \subsection{Simple rules and answer sets}\label{sssec:simple_rules}
-ASP is based on a set of rules of of the form:
+ASP is based on a set of rules of the form:
 \begin{align*}

### 2014-11-11 parts-example.tex
minor corrections
corrections mineures

@@
+% vi:spell spelllang=en:
 \section{Examples}\label{sec:examples}
@@
 In order to reproduce such requirements, the Boolean functions of their successors,
-that is CDK4, CDK6, prB, p21, p27, IGF1R, MYC, CycD1 and CycE1,
+that are CDK4, CDK6, prB, p21, p27, IGF1R, MYC, CycD1 and CycE1,
 were modeled as cooperative sorts, if needed.
@@
 \\\hline
-    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & 100s & <1s & 578
+    T cell receptor \cite{SaezRodriguez2007} & $94$ & $39$ & \modMF{100s} & \modMF{<1s} & \modMF{578}
 \\\hline
-    EGF receptor \cite{Samaga2009} & $104$ & $89$ & 200s & 2.5s & 27496
+    EGF receptor \cite{Samaga2009} & $104$ & $89$ & \modMF{200s} & \modMF{2.5s} & \modMF{27496}
   \end{tabular}

### 2014-11-11 parts-infer-ig.tex
corrections mineures

@@
 regulators of $a$ (instead of in scope of $g$).
-The set of configuration where no bounce occur on $a_i$ is given given by
+The set of configurations where no bounce occur on $a_i$ is given by
 $\fp(a_i)$ (\pref{eq:fp}).}
@@
 The IG inference on the PH of \pref{fig:runningPH-2} returns the
-IG in \pref{fig:runningBRN}(left), that is, allows to find back all the expected edges
+IG in \pref{fig:runningBRN}(left): it finds back all the expected edges
 alongside with their signs and thresholds.

### 2014-11-11 parts-infer-param.tex
corrections mineures

@@
+% vi:spell spelllang=en:
 \section{Parametrization inference}\label{sec:infer-K}
@@
 \modLP{
-As described in \pref{ssec:infer-K}, this identifcation relies on the computation of the focal
+As described in \pref{ssec:infer-K}, this identification relies on the computation of the focal
 processes for each configuration of the parameters.
@@
 well-formed for parameter inference (characterized in the next subsection by \pref{pro:wf-ph-K}),
-there maybe several focal processes (non-deterministic behaviour) or none (cyclic behaviour).  Such
+there may be several focal processes (non-deterministic behaviour) or none (cyclic behaviour).  Such
 a case occurs notably when the PH encodes the union of several Boolean or discrete networks, as
@@
 \label{pps:param_K}
-Let $(\PHs, \PHl, \PHh)$ be a Process Hitting \modLP{with $\PHs=\Gamma\cup\CSorts$} 
-well-formed for parameter inference, and $\IG = (\Gamma, E)$ the inferred IG.
-Let $A$ (resp. $B$) $\subseteq \Gamma$ be the set of regulators that activate (resp. inhibit) a sort
-$a$.
-If $\focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})= \{ a_i \}$, then \modLP{$K_{a,\omega} = i$}.
+Let $(\PHs, \PHl, \PHh)$ be a Process Hitting \modLP{with $\PHs=\Gamma\cup\CSorts$} and an
+associated IG $\IG = (\Gamma, E)$ well-formed for parameter inference.
+For all $a\in\Gamma$, for any $\omega\subseteq \GRNreg{a}$,
+if $\focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})= \{ a_i \}$, then \modLP{$K_{a,\omega} = i$}.
 \end{proposition}
@@
 \pref{pps:param_K} infers the parametrization of \pref{fig:runningBRN}(right).
-In other words, the inference of parameters allows to find back all parameters
+The inference of parameters finds back all parameters
 of the original model when the cooperations are fully defined.

### 2014-11-11 parts-intro.tex
corrections mineures

@@
 specify Thomas' parameters, especially for large networks.
-\modOR{Other works have also been carried out with the aim of dealing with large systems (\eg \cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generaly focus on some restricted properties,} \modOR{as finding out steady states.}
+\modOR{Other works have also been carried out with the aim of dealing with large systems (\eg \cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generally focus on some restricted properties,} \modOR{as finding out steady states.}
 %\todo{The reviewer proposes to nuance this with several papers:

### 2014-11-11 parts-ph.tex
corrections mineures

@@
+% vi:spell spelllang=en:
 \subsection{The Process Hitting framework}
@@
 
-We give the the definition and semantics of the Process Hitting (PH), and its usage to model
+We give the definition and semantics of the Process Hitting (PH), and its usage to model
 cooperation between concurrent components.
@@
 \ie Petri nets with at most one token per place, having
-groups of multually exclusive places (processes) acting as the sorts, and where each
+groups of mutually exclusive places (processes) acting as the sorts, and where each
 transition has at most one read-arc (connected to the \emph{hitter}), and
@@
 In this BRN, $b$ and $c$ activate $a$, while $a$ inhibits $b$.
-The inhibition of $b$ by $a$ is only effective when $a$ is at level $2$ (noted: $a_2$);
+The inhibition of $b$ by $a$ is only effective when $a$ is at level $2$ (noted $a_2$);
 in the other cases, $b$ cannot evolve in any direction.

### 2014-11-11 parts-tr2global.tex
corrections mineures

@@
 evolution of the sort;
-and it is also possible possible that no focal process exist, when in presence of terminal cycles of
+and it is also possible that no focal process exist, when in presence of terminal cycles of
 transitions.
@@
 components and intermediate cooperate sorts, the latter having to be masked in Thomas models.
-It notably defines the predecessors, (group of) regulators of a
+It notably defines the predecessors, the (groups of) regulators of a
 sort, and the extension of a configuration to incorporate the state of the
@@
 \begin{align}
-\ctx_a^g(\sigma) & \DEF \sigma \uplus \state{\upsilon(\sigma) \mid\upsilon \in\PHdirectpredec{a} \cap \CSorts} 
+\ctx_a^g(\sigma) & \DEF \sigma \uplus \state{\upsilon(\sigma) \mid\upsilon \in\PHdirectpredec{a} \cap \CSorts
+\wedge \PHpredecgene{\upsilon}\subseteq g} 
 \label{eq:ctx-sigma}

### 2014-11-11 parts-asp.tex
typo

@@
   and it is easy to constrain the answers according to some properties.
-Applied to the models and constraints defined in this papers,
+Applied to the models and constraints defined in this paper,
 ASP appears to be efficient for tackling the inherent complexity of the enumeration of parametrizations.

### 2014-11-12 response.tex
response++
last minor corrections

@@
   \item The abstract and the introduction have been improved to better contextualize our work.
-  \item New explanations have been added at the beginning of every section
+  \item Additional explanations have been added at the beginning of every section
     to detail their objectives and methods.
@@
 This comment has been taken into account and the sentence has been replaced by:
-“The analysis of dynamics of very large models is however a rather hard problem, which led us to previously introduce a new framework: the Process Hitting (PH), which is a particular class of Indeterministic Asynchronous Automata Network (or safe Petri nets).
+
+“The analysis of the dynamics of very large models is however a rather hard problem, which led us to previously introduce the Process Hitting framework (PH), which is a particular class of nondeterministic asynchronous automata network (or safe Petri nets).
 Its major advantage lies in the efficiency of several static analyses recently designed to assess dynamical properties, making it possible to tackle very large models.”

### 2014-11-12 main.tex
complete phage lambda example
todo--
fix \irB
phage lambda++
cleanup bib, make Fig 1 more visible, reference to K equivalence
plus de couleurs
minor corrections
last minor corrections

--- 2014-09-03 main.tex
@@
 \newrefformat{tb}{Table~\ref{#1}}
+\newrefformat{rem}{Remark~\ref{#1}}
 \def\pref{\prettyref}
@@
 \newdefinition{property}{Property}
+\newdefinition{remark}{Remark}
 \newdefinition{proposition}{Proposition}
@@
 \def\modOR#1{\textcolor{olive}{#1}}
+\def\modMF#1{#1} \def\modLP#1{#1} \def\modMM#1{#1} \def\modOR#1{#1}
 
-% Id est
-\newcommand{\ie}{i.e.~}
+% Abbréviations
+\newcommand{\ie}{i.e.,\ }
+\newcommand{\eg}{e.g.,\ }
 
@@
 \begin{abstract}
-\modMF{Maxime}
-\modLP{Lo\"ic}
-\modMM{Morgan}
-\modOR{Olivier}
-
-%\parindent 0.5cm
-The Process Hitting (PH) \modLP{is a particular class of Indeterministic Asynchronous Automata
-Network (or safe Petri nets) on which have been recently designed several static analyses for
-dynamical properties that are tractable on very large systems.}
+\modOR{Qualitative formalisms offer a well-established alternative to the more traditionally used differential equation models of Biological Regulatory Networks (BRNs).
+These formalisms led to numerous theoretical works and practical tools to understand emerging behaviors.} \modMM{The analysis of the dynamics of very large models is however a rather hard problem, which led us to previously introduce} \modMF{the Process Hitting framework (PH),} \modMM{which} \modLP{is a particular class of nondeterministic asynchronous automata
+network (or safe Petri nets).} \modMM{Its major advantage lies in the efficiency of several static analyses recently designed to assess dynamical properties, making it possible to tackle very large models.}
 
 \modLP{In this paper, we address the formal identification of qualitative
-models of Biological Regulatory Networks (BRNs) from PH models.}
-First, the inference of the Interaction Graph (IG) from a PH model
+models of BRNs from PH models.}
+First, the inference of the Interaction Graph from a PH model
 \modLP{summarizes the signed influences between the components} that are effective for the dynamics.
@@
 that are compatible with a given PH.
-As the PH allows the specification of indeterministic interactions between
+As the PH allows the specification of nondeterministic interactions between
 components, our inference emphasizes the ability of PH to deal with large BRNs with incomplete knowledge
@@
 
-The inference of corresponding Thomas models is provided using Answer Set Programming,
-which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
-
-%\todo{Rather name Kaufman as the creator of BRNs.}
-%On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
-%formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
-%In addition to be able to use the existing results on this formalism,
-%obtaining the Thomas representation of a PH model allows to iteratively refine a model.
-%Linking this representation to PH would not only allow to use the existing results on this formalism,
-%but also to iteratively refine the model of a system.
-
-%This paper establishes formal links between these two formalisms, permitting to appreciate their
-%complementary features.
+The inference of corresponding Thomas models is \modLP{implemented} using Answer Set Programming,
+which allows \modMF{in particular} an efficient enumeration of (possibly numerous) compatible parametrizations.
 \end{abstract}
@@
 
-%\input{parts/contributions}
+\appendix
+\input{annex-phage-lambda}
 

### 2014-11-12 parts-asp.tex
last minor corrections

@@
 \modMF{
-Furthermore, both the lower and upper bounds are $1$,
+Furthermore, the lower and upper bounds are both $1$,
 which forces each enumerated parameter to have exactly one value.
@@
 The previous cardinality rule allows to produce 9 parametrizations,
-in which these three parameters can take all possible values:
+in which these two parameters can take all possible values:
 %(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):

### 2014-11-12 parts-discussion.tex
section 4.1
corrections
last minor corrections

@@
 The enumeration of such parametrizations is efficiently tackled using Answer Set Programming.
-We illustrate the overall method with several results on large biological models.
+We illustrate the overall method with several results
+on \modMF{both small and} large biological models.
 
@@
 the study and formalization of such process would allow to predict behavioral changes and lead to more accurate models.
-Such work would also require further study on the inference of the IG in order to better understand and possibly avoid
-the presence of some of the current spurious auto-edges.
 Second, the inference of BRN multiplexes \cite{BernotMultiplexes} may be of practical interest 
@@
 the French National Agency for Research (ANR-10-BLANC-0218 BioTempo project).
-The work of LP was supported by the Swiss SystemsX.ch project.

### 2014-11-12 parts-example.tex
last minor corrections

@@
 where \texttt{model.ig} is an IG model file in \textsc{Pint} format.
-The admissible parametrizations enumeration is performed by adding the \texttt{-{}-enumerate}
+The admissible parametrizations enumeration is performed by appending the \texttt{-{}-enumerate}
 option to the command.
@@
 The composition of all models and the results of the inferences are summarized in \pref{tb:computation}.
+\modMF{%
+In every case, all parameters could be inferred and it was not necessary to
+enumerate compatible parametrizations.
+}%
 %We note that due to a very high number of parameters, no parameters inference could be performed on the $94$ and $104$ component models.
@@
   \begin{tabular}{r|c|c|c|c|c}
-    \textbf{Model} & $\mathbf{|\Gamma|}$ & $\mathbf{|\PHs \setminus \Gamma|}$ & \textbf{$\Delta t$ IG} & \textbf{$\Delta t$ K} & $\mathbf{|K|}$
+    \textbf{Model} & $\mathbf{|\Gamma|}$ & $\mathbf{|\Delta|}$ & \textbf{IG inference} & \textbf{K inference} & $\mathbf{|K|}$
 \\\hline\hline

### 2014-11-12 parts-grn.tex
corrections mineures
last minor corrections

@@
 \modMM{
-The principle of qualitative modeling was introduced as synchronous Boolean networks by Stuart Kauffman on the one hand \cite{kauffman1969metabolic}, asynchronous Ren\'e Thomas’ networks on the other \cite{Thomas73}.
+The principle of qualitative modeling was introduced as synchronous Boolean networks by Stuart Kauffman on the one hand \cite{kauffman1969metabolic}, and asynchronous Ren\'e Thomas' networks on the other \cite{Thomas73}.
 Both models have been considered of interest and led to numerous works.
@@
 
-\modMM{In this paragraph, we} concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
+\modMM{In this section, we} concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
 \cite{thomas1990biological, Richard06,BernotSemBRN}.
@@
 In particular, a component may be regulated by both a positive and negative interaction derived from two different components.
-To refine the dynamics, a parametrization had been added to the model \cite{snoussi1989qualitative}.
+To refine the dynamics, a parametrization has been added to the model \cite{snoussi1989qualitative}.
 This parametrization allows to define focal points, \ie it clarifies the level towards which a component evolves depending on the expression of its regulators.
-Some of the information included in the IG and the parametrization may appear as slighltly redundant.
+Some of the information included in the IG and the parametrization may appear as slightly redundant.
 Especially, if the single parametrization were enriched with additional constraints (\eg monotonicity requiring an ordering on the value of parameters), then it may directly capture the signs.
@@
 the same.
-It is notably the case with auto-regulations:
+It is notably the case with self-regulations:
 if $\GRNedge{a}{s}{t}{a}$,
@@
 $f^a(s\{b_i\}) < \text{(resp. $>$) } f^a(s\{b_{i+1}\})$, 
-where $i < l_b$ and $s\{b_i\}$ denotes the state $s$ where the component $b$ is assigned to $b_i$.
+where $i < l_b$ and $s\{b_i\}$ denotes the state $s$ where the component $b$ is assigned to $i$.
 \modLP{If all the interactions from $b$ to $a$ are positive (resp. negative), there is a positive

### 2014-11-12 parts-infer-ig.tex
last minor corrections

@@
 \modLP{The inference relies on finding configurations $\sigma$ of groups $g$ of regulators of $a$
-such that the increase of one of the regulator $b\in g$ from $b_t$ to $b_{t+1}$
+such that the increase of one of the regulators $b\in g$ from $b_t$ to $b_{t+1}$
 changes the direction of the evolution of $a$.
@@
 \end{align}
-\modMF{where $\irB_a^g(\sigma) \subset \PHl_a$ is defined as follows:}
+\modMF{where $\irB_a^g(\sigma) \subseteq \PHl_a$ is defined as follows:}
 \begin{align}
@@
 In the end, the threshold of each edge is the minimum threshold for which an influence has been found.
-\modLP{Note that as regulations matches with sign changes of the evolution of the
+\modLP{Note that as regulations match with sign changes in the evolution of the
 regulated components,
@@
 necessarily, a third regulation of opposite sign has been inferred in between.
-The only exception is when $a$ is never hit ($\PHpredecgene{a} = \emptyset$), in
-such a case, we arbitrarily pick the minimum threshold (1) for the
+The only exception is when $a$ is never hit ($\PHpredecgene{a} = \emptyset$); in
+such a case, we arbitrarily pick the minimum threshold (that is, 1) for the
 self-activation.}

### 2014-11-12 parts-infer-param.tex
last minor corrections

@@
 well-formed for parameter inference (characterized in the next subsection by \pref{pro:wf-ph-K}),
-there may be several focal processes (non-deterministic behaviour) or none (cyclic behaviour).  Such
+there may be several focal processes (nondeterministic behaviour) or none (cyclic behaviour).  Such
 a case occurs notably when the PH encodes the union of several Boolean or discrete networks, as
@@
 Hence, we propose in \pref{ssec:admissible-K} the notion of \emph{compatible} parametrization with respect
-ot the PH dynamics: a Thomas model is compatible with a PH if Thomas dynamics is included in the PH
+ot the PH dynamics: a Thomas model is compatible with a PH if its dynamics is included in the PH
 dynamics, \ie all the transitions in the Thomas model are possible transitions in the PH.
@@
 compatible with a configuration $\sigma$: if $a$ is specified in $\sigma$, then $S_a(\sigma) = \{a_i\}$
-where $a_i\in\sigma$; otherwise $S_a(\sigma)$ contains all the processes of sort $a$, $L_a$.}
+where $a_i\in\sigma$; otherwise $S_a(\sigma)$ \modMF{is the set of all processes of sort $a$, that is,} $L_a$.}
 \begin{align}
@@
 and using the IG of \pref{fig:runningBRN}(left)
-which is the result of the IG inference on this PH model,
-as explained in \pref{ssec:infer-ig-examples},
+which is the result of the IG inference on this PH model
+(as explained in \pref{ssec:infer-ig-examples}),
 \pref{pps:param_K} infers the parametrization of \pref{fig:runningBRN}(right).
-The inference of parameters finds back all parameters
+The inference of parameters thus finds back all parameters
 of the original model when the cooperations are fully defined.

### 2014-11-12 parts-intro.tex
last minor corrections

--- 2014-11-10 parts-intro.tex
@@
 With the purpose of analyzing these systems, they are often modeled as graphs which make it possible to determine the possible evolutions of all the interacting components of the system.
-\modOR{Besides continuous models of physicists, generally designed through systems of ordinary differential equations, modeling regulatory networks by means of Boolean networks has become popular in the wake of Kauffman work \cite{kauffman1969metabolic}.
+\modOR{Besides continuous models of physicists, generally designed through systems of ordinary differential equations, modeling regulatory networks by means of Boolean networks has become popular in the wake of Kauffman's work \cite{kauffman1969metabolic}.
 We based our work on a logical formalism developed by Ren\'e Thomas from 1973 \cite{Thomas73} that generalizes upon Boolean networks in the sense that it allows variables to have more than two values and transitions between states to occur asynchronously.}
@@
 specify Thomas' parameters, especially for large networks.
-\modOR{Other works have also been carried out with the aim of dealing with large systems (\eg \cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generaly focus on some restricted properties,} \modOR{as finding out steady states.}
+\modOR{Other works have also been carried out with the aim of dealing with large systems (\eg \cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generally focus on some restricted properties,} \modOR{such as finding steady states.}
 %\todo{The reviewer proposes to nuance this with several papers:
@@
 In other words, it captures the logical functions stating how} \modOR{various elements coalesce and act together upon an other element among the network.}
-\modLP{In particular, PH allows to specify partial (non-deterministic) cooperations that superpose the dynamics of 
+\modLP{In particular, PH allows to specify partial (nondeterministic) cooperations that superpose the dynamics of 
 candidate fully-specified (deterministic) cooperations.
@@
 In a more general framework and in an interesting way,
-we should be able to synthesize families of BRN having some featuring behavioral properties.
+we should be able to synthesize families of BRNs having some featuring behavioral properties.
 
@@
 Programming (ASP) \cite{Baral03},  which turns out to be effective for these enumerative searches.
-The whole method is also applied to the biological model of ERBB receptor-regulated G1/S transition,
-containing 20 components,
-which allows to detail how joint actions can be defined inside a PH model in order
-to refine its dynamics, and study a more precise class of underlying Thomas models.
+The whole method is also applied to
+the model of the bacteriophage lambda immunity control with 4 components,
+in order to give a detailed application of our method,
+and to the biological model of the ERBB receptor-regulated G1/S transition,
+which contains 20 components
+and allows to show how joint actions can be defined inside a PH model in order
+to refine its dynamics and study a more precise class of underlying Thomas models.
 %\todo{describe application}
 
-Our work is related to the approach of \cite{Khalis09} which relies on temporal logic, and \cite{20646302,DBLP:conf/ipcat/CorblinFTCT12} which also uses constraint programming.
-Both aim at determining a class of models which are consistent with available partial data on the regulatory structure and dynamical properties.
+Our work is related to the approach of \cite{Khalis09} which relies on temporal logic, and to \cite{20646302,DBLP:conf/ipcat/CorblinFTCT12} that use constraint programming.
+All three of them aim at determining a class of models which are consistent with available partial data on the regulatory structure and dynamical properties.
 %These classes are built in order to infer properties common to some studied models.

### 2014-11-12 parts-ph.tex
last minor corrections

@@
 This PH example actually models a BRN where the component $a$ has three qualitative
-levels and components $b$ and $c$ are boolean.
+levels and components $b$ and $c$ are Boolean.
 In this BRN, $b$ and $c$ activate $a$, while $a$ inhibits $b$.
@@
 As described in \cite{PMR10-TCSB}, such cooperations can be encoded in PH with
-the use of some intermediate sorts, so-called \emph{cooperative sorts}, that
+the use of some intermediate sorts, called \emph{cooperative sorts}, that
 merge information on the state of cooperating sorts.}
@@
 Hence, $bc_{01}$ represents the sub-state $\state{b_0,c_1}$, and so on.
-Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the state of the sorts $b$
+Each process of sorts $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the state of the sorts $b$
 and $c$ (\eg $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).

### 2014-11-12 parts-phage-lambda.tex
fix figures
corrections mineures
last minor corrections

@@
+% vi:spell spelllang=en:
+
 \modMF{
@@
   \underbrace{\cIcrocII, \cIcroN, \cIcro, \crocII}_{\Delta} \}\]
-The four components have respectively 3, 4 2 and 2 expression levels,
+The four components have respectively 3, 4, 2 and 2 expression levels,
 which we denote:
@@
 
-\begin{figure}[p]
+\begin{figure}
 \centering
@@
 
-\begin{figure}[t]
+\begin{figure}
 \centering
@@
   \item Our inference returns $K_{\cro, \{\cI;\cro\}} = 2$, \modLP{which is a value
-    equivalent to the original $0$ value according to \pref{rem:K-equiv}:
+    equivalent to the original value $0$ according to \pref{rem:K-equiv}:
     the resulting dynamics is the same, and $1$ would also have been an

### 2014-11-12 parts-tr2global.tex
last minor corrections

@@
 Such a function is always deterministic (whereas the associated asynchronous dynamics can be
-indeterministic due to concurrent evolution of components).
+nondeterministic due to concurrent evolution of components).
 In contrast, 
@@
 applied in a given configuration into a single process that correspond to the level toward which the
-component evolve.
+component evolves.
 This step is detailed in \pref{ssec:focal} in which is defined the $\focals$ function associating 
@@
 for a given configuration.
-It is possible that several different focal processes are identified, indicating an indeterministic
+It is possible that several different focal processes are identified, indicating an nondeterministic
 evolution of the sort;

### 2014-11-12 parts-intro.tex
corrections from Olivier

@@
 candidate fully-specified (deterministic) cooperations.
-This permit to model BRNs with a partial knowledge on precise evolution functions for components
+This permits to model BRNs with a partial knowledge on precise evolution functions for components
 by capturing the largest (the most general) dynamics, without a combinatorics enumeration of

### 2014-11-12 parts-tr2global.tex
corrections from Olivier

@@
 Thomas parametrization defines a global function
-which associates to each component and each resource configuration a level toward which the component
+which associates to each component and each resource configuration a level towards which the component
 eventually evolves in the given configuration.
@@
 Hence, identifying Thomas models from PH requires to lump a set of transitions that can be
-applied in a given configuration into a single process that correspond to the level toward which the
+applied in a given configuration into a single process that correspond to the level towards which the
 component evolves.
@@
 Finally, \pref{ssec:split-sorts} addresses the separation of PH sorts into
-components and intermediate cooperate sorts, the latter having to be masked in Thomas models.
+components and intermediate cooperative sorts, the latter having to be masked in Thomas models.
 It notably defines the predecessors, the (groups of) regulators of a
@@
 (\pref{pro:wf-ph}), one can recursively
-evaluate the focal process of each cooperative sort that hit $a$ with respect to
+evaluate the focal process of each cooperative sort that hits $a$ with respect to
 the configuration $\sigma$ of the regulators $g$.

### 2014-12-03 main.tex
typos on Thomas's name after the editor's review

@@
 \modLP{summarizes the signed influences between the components} that are effective for the dynamics.
-Second, we provide the inference of all Ren\'e Thomas models of BRNs 
+Second, we provide the inference of all Ren\'e-Thomas models of BRNs 
 that are compatible with a given PH.
@@
 components, our inference emphasizes the ability of PH to deal with large BRNs with incomplete knowledge
-on interactions, where Thomas' approach fails because of the combinatorics of parameters.
+on interactions, where Thomas's approach fails because of the combinatorics of parameters.
 
@@
 interaction graph \sep
-René Thomas' parameters \sep
+René Thomas's parameters \sep
 answer set programming

### 2014-12-03 parts-discussion.tex
typos on Thomas's name after the editor's review

@@
 
-This work establishes the abstraction relationship between PH and Thomas' approaches for
+This work establishes the abstraction relationship between PH and Thomas's approaches for
 qualitative BRN modeling.
 The PH allows an abstract representation of BRNs dynamics (allowing incomplete knowledge on the
-cooperation between components) that cannot be exactly represented in Ren\'e Thomas' formalism by a
+cooperation between components) that cannot be exactly represented in Ren\'e Thomas's formalism by a
 single instance of BRN parametrization.
-This motivates the concretization of PH models into a set of compatible Thomas' models in order to benefit
+This motivates the concretization of PH models into a set of compatible Thomas models in order to benefit
 of the complementary advantages of these two formal frameworks.
@@
 others.
-Then, based on a prior inference of Ren\'e Thomas' parametrization for BRNs from a PH model, we
-delimit the set of admissible Thomas' parametrizations that are compatible with the PH dynamics,
+Then, based on a prior inference of Ren\'e Thomas's parametrization for BRNs from a PH model, we
+delimit the set of admissible Thomas's parametrizations that are compatible with the PH dynamics,
 and give arguments for their correctness.

### 2014-12-03 parts-grn.tex
typos on Thomas's name after the editor's review

@@
 % vim:spell spelllang=en
-\subsection{Thomas' modeling}
+\subsection{Thomas's modeling}
 \label{ssec:thomas}
@@
 \modMM{
-The principle of qualitative modeling was introduced as synchronous Boolean networks by Stuart Kauffman on the one hand \cite{kauffman1969metabolic}, and asynchronous Ren\'e Thomas' networks on the other \cite{Thomas73}.
+The principle of qualitative modeling was introduced as synchronous Boolean networks by Stuart Kauffman on the one hand \cite{kauffman1969metabolic}, and asynchronous Ren\'e-Thomas networks on the other \cite{Thomas73}.
 Both models have been considered of interest and led to numerous works.
-In the following, we choose to focus on  Thomas' modelling, with its asynchronous, unitary and multi-level semantics.}
+In the following, we choose to focus on  Thomas's modeling, with its asynchronous, unitary and multi-level semantics.}
 
-\modMM{In this section, we} concisely present Thomas' modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
+\modMM{In this section, we} concisely present Thomas's modeling of a Biological Regulatory Network (BRN) and its dynamics, merely inspired by
 \cite{thomas1990biological, Richard06,BernotSemBRN}.
@@
 
-Thomas' formalism lies on two complementary descriptions of the system. First, the
+Thomas's formalism lies on two complementary descriptions of the system. First, the
 \emph{Interaction Graph} (IG) models the structure of the system by defining the components'
@@
 %referred to as the \emph{resources} of $a$ in this state (\pref{def:resources}).
-Relying on this observation, the specificity of Thomas' approach lies in the use of discrete \emph{parameters} to represent the
+Relying on this observation, the specificity of Thomas's approach lies in the use of discrete \emph{parameters} to represent the
 focal level interval towards which the component will evolve in every configuration of the resources (\pref{def:param}).
@@
 \begin{definition}[Asynchronous dynamics ($\GRNtrans{}{}$)]\label{def:dynamics}
-The dynamics of a BRN using Thomas' parameters is given by the transition relation $\GRNtrans{}{}\ \in \GRNstates \times \GRNstates$ defined by:
+The dynamics of a BRN using Thomas's parameters is given by the transition relation $\GRNtrans{}{}\ \in \GRNstates \times \GRNstates$ defined by:
 \begin{align*}

### 2014-12-03 parts-infer-ig.tex
typos on Thomas's name after the editor's review

@@
 This is due to the fact that the actions $\PHfrappe{a_2}{b_1}{b_0}$ and $\PHfrappe{a_2}{b_0}{b_1}$
-introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
+introduce an oscillation only caused by $a$, which cannot be represented in Thomas's modeling.
 }

### 2014-12-03 parts-infer-param.tex
typos on Thomas's name after the editor's review

@@
 When there is a unique focal process of a component for a given configuration of its regulators,
-the focal process matches with the value of the associated Thomas parameter.}
+the focal process matches with the value of the associated Thomas's parameter.}
 
@@
 well-formed for parameter inference (characterized in the next subsection by \pref{pro:wf-ph-K}),
-there may be several focal processes (nondeterministic behaviour) or none (cyclic behaviour).  Such
+there may be several focal processes (nondeterministic behavior) or none (cyclic behavior).  Such
 a case occurs notably when the PH encodes the union of several Boolean or discrete networks, as
@@
 
-Let $K_{a,\omega}$ be a Thomas parameter for a given component $a \in \Gamma$ 
+Let $K_{a,\omega}$ be a Thomas's parameter for a given component $a \in \Gamma$ 
 with $\omega \subset \GRNreg{a}$ a set of its regulators.
@@
 \modLP{$K_{a,\omega} = \focals(a,S_a(\sigma_{a,\omega}),\sigma_{a,\omega})$ if this latter is
-a singleton, denoting a deterministic behaviour (\pref{pps:param_K})}.
+a singleton, denoting a deterministic behavior (\pref{pps:param_K})}.
 
@@
 
-Then, we use some additional biological constraints on Thomas' parameters given in
+Then, we use some additional biological constraints on Thomas's parameters given in
 \cite{BernotSemBRN}, that we sum up in the following three properties:

### 2014-12-03 parts-intro.tex
typos on Thomas's name after the editor's review

@@
 Nevertheless, these dynamics can be precisely established only with regard to some discrete parameters,
-hereafter called ``Thomas' parameters'',
+hereafter called ``Thomas's parameters'',
 which stand for kinds of ``focal points'', \ie the evolutionary tendency from each state and depending on the set of the other currently interacting components.
@@
 
-Thomas' modeling has motivated numerous works around the link between the Interaction Graph (IG)
+Thomas's modeling has motivated numerous works around the link between the Interaction Graph (IG)
 (summarizing the global influences between components) and the possible dynamics (\eg
@@
 state graph explosion, an other major drawback of this framework is the difficulty to
-specify Thomas' parameters, especially for large networks.
+specify Thomas's parameters, especially for large networks.
 \modOR{Other works have also been carried out with the aim of dealing with large systems (\eg \cite{Corblin200991,Bollman2007,Delaplace2012, 10.1371/journal.pone.0007992}).} \modMM{Because of the complexity of such systems, these works generally focus on some restricted properties,} \modOR{such as finding steady states.}
@@
 various degrees of knowledge for joint regulations;
-and Thomas' parameters, which completely specify the functions governing the evolution of
+and Thomas's parameters, which completely specify the functions governing the evolution of
 components.
@@
 of the system dynamics (see \cite{PR11-SASB} for a short survey).
-The most known are the Thomas' conjectures (that have been proved since,
+The most known are the Thomas's conjectures (that have been proved since,
 \eg \cite{RRT08,RiCo07,Richard2010378}
@@
 
-Second, we tackle the inference of Thomas' parameters that are compatible with a given PH
+Second, we tackle the inference of Thomas's parameters that are compatible with a given PH
 model, \ie that fully specify the evolution of components while respecting the cooperations
@@
 transition in the (asynchronous) Thomas model exists in the PH model.
-In practice, the number of compatible Thomas parameters can be huge for large networks where only a
+In practice, the number of compatible Thomas's parameters can be huge for large networks where only a
 few cooperations are specified.
@@
 \pref{sec:infer-IG} defines the IG inference from PH;
-\pref{sec:infer-K} details the parameters inference and the enumeration of Thomas parametrizations compatible with a PH;
+\pref{sec:infer-K} details the parameters inference and the enumeration of Thomas's parametrizations compatible with a PH;
 \pref{sec:impl} discusses the implementation of the latter in ASP.

### 2014-12-03 parts-tr2global.tex
typos on Thomas's name after the editor's review

@@
 
-Process Hitting (PH) and Thomas modelling rely on two different paradigms for the specification of
+Process Hitting (PH) and Thomas modeling rely on two different paradigms for the specification of
 network dynamics:
-Thomas parametrization defines a global function
+Thomas's parametrization defines a global function
 which associates to each component and each resource configuration a level towards which the component
@@
 The \pref{ssec:dfocals} discusses on constraints for obtaining deterministic
-(local) behaviours in PH.
+(local) behaviors in PH.
 
@@
 
-For the inference of the interaction graph and Thomas parameters from a PH model, the following
+For the inference of the interaction graph and Thomas's parameters from a PH model, the following
 sections assume that \emph{all the cooperative sorts are deterministic}.
